### 2022-07-30 README.Rmd
branding

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nhanesTrends <img src="man/figures/logo.png" align="right" height="139" />

<!-- badges: start -->
<!-- badges: end -->

Our goal is to provide a platform for exploration of the National Health and Nutrition Examination Survey (NHANES) data.

## Installation

You can install the development version of nhanesTrends from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
remotes::install_github("jhs-hwg/nhanesShinyBP")
```

## Example

The shiny application is hosted on a public server, but you can also run it locally:

```{r, eval=FALSE}

library(nhanesShinyBP)
app_run()

```

### 2022-07-30 logo.R

@@ -1,0 +1,57 @@
+
+devtools::load_all()
+
+library(ggplot2)
+
+ds <- nhanes_bp %>%
+ .[svy_subpop_htn == 1] %>%
+ .[, svy_weight := svy_weight_mec] %>%
+ svy_design_new(
+  n_exposure_group = numeric(),
+  years = levels(nhanes_bp$svy_year),
+  pool = 'no'
+ ) %>%
+ subset(demo_pregnant == 'No' | is.na(demo_pregnant)) %>%
+ subset(bp_med_use == 'Yes')
+
+smry <- ds %>%
+ svy_design_summarize(outcome = 'bp_control_jnc7',
+                      key = nhanes_key,
+                      age_standardize = TRUE,
+                      user_calls = c('percentage')) %>%
+ filter(bp_control_jnc7 == 'Yes')
+
+
+trend_1 <-  c("1999-2000", "2007-2008")
+
+trend_2 <- c("2007-2008", "2013-2014")
+
+trend_3 <- c("2013-2014",  "2017-2020")
+
+p <- ggplot(smry, aes(x = svy_year, y = estimate)) +
+ geom_point(size=1.2, color = '#8a0303') +
+ geom_line(data = filter(smry, svy_year %in% trend_1),
+           aes(group = 1),
+           color = '#8a0303') +
+ geom_line(data = filter(smry, svy_year %in% trend_2),
+           aes(group = 1),
+           color = '#8a0303') +
+ geom_line(data = filter(smry, svy_year %in% trend_3),
+           aes(group = 1),
+           color = '#8a0303') +
+ theme_void()
+
+
+hexSticker::sticker(
+ p,
+ package = "nhanesTrends",
+ p_color = 'black',
+ h_fill = 'white',
+ h_color = 'grey',
+ p_size = 18,
+ s_x = 1,
+ s_y = 0.85,
+ s_width = 1.7,
+ s_height = 0.75,
+ filename = "logo.png"
+)

### 2022-09-18 README.Rmd
remotes for install
more accessible for blogposts

@@
 ``` r
-# install.packages("devtools")
+# install.packages("remotes")
 remotes::install_github("jhs-hwg/nhanesShinyBP")
@@
 ```
+
+## Validation of results from the shiny application
+
+We have replicated results from prior NHANES publications to verify that routines used to conduct analyses in this app are valid.
+Based on these tests, we have identified some cases where our app does not exactly match previously published articles, and we note those cases here.
+
+TO BE WRITTEN: notes on tests showing exact match with [Muntner et al](https://jamanetwork.com/journals/jama/fullarticle/2770254)
+
+TO BE WRITTEN: notes on discrepancies with results in [Carey et al.](https://www.ahajournals.org/doi/10.1161/HYPERTENSIONAHA.118.12191)

### 2022-09-18 R/nhanes_load.R

@@ -1,5 +1,5 @@
 
-#' @title Load the nhanes data
+#' Load the nhanes data
 #'
 #' No re-coding is applied.
 #'
@@ -9,7 +9,6 @@
 #' @return data set inheriting from `as`
 #'
 #' @noRd
-#'
 nhanes_load <- function(fpath = file.path(here::here(), 'data-raw'),
                         as = 'tibble'){
 

### 2022-09-24 tests/testthat/test-nhanes_calibrate.R
weight checking

@@ -1,0 +1,62 @@
+
+# benchmark the nhanes_calibrate function
+# takes about ~30-35 milliseconds on average
+# microbenchmark::microbenchmark(
+#  bcj_calib = nhanes_calibrate(nhanes_full = nhanes_bp,
+#                               nhanes_sub = nhanes_bp[svy_subpop_htn == 1])
+# )
+
+# validate calibration in subgroups defined by:
+variables <- c('svy_year',
+               'demo_age_cat',
+               'demo_gender')
+
+nhanes_htn <- nhanes_calibrate(nhanes_full = nhanes_bp,
+                               nhanes_sub = nhanes_bp[svy_subpop_htn == 1])
+
+test_that(
+ desc = "sum of weights in the calibrated data matches the original",
+ code = {
+
+  # overall
+  expect_equal(
+   sum(nhanes_htn$svy_weight_cal),
+   sum(nhanes_bp$svy_weight_mec)
+  )
+
+
+  for(v in variables){
+   for(i in levels(nhanes_bp[[v]])){
+    expect_equal(
+     sum(nhanes_htn$svy_weight_cal[nhanes_htn[[v]]==i]),
+     sum(nhanes_bp$svy_weight_mec[nhanes_bp[[v]]==i])
+    )
+   }
+  }
+ }
+)
+
+nhanes_chol <- nhanes_calibrate(nhanes_full = nhanes_bp,
+                                nhanes_sub = nhanes_bp[svy_subpop_chol == 1])
+
+test_that(
+ desc = "sum of weights in the calibrated data matches the original",
+ code = {
+
+  # overall
+  expect_equal(
+   sum(nhanes_chol$svy_weight_cal),
+   sum(nhanes_bp$svy_weight_mec)
+  )
+
+  for(v in variables){
+   for(i in levels(nhanes_bp[[v]])){
+    expect_equal(
+     sum(nhanes_chol$svy_weight_cal[nhanes_chol[[v]]==i]),
+     sum(nhanes_bp$svy_weight_mec[nhanes_bp[[v]]==i])
+    )
+   }
+  }
+ }
+)
+

### 2022-09-26 README.Rmd
weight checking
actions!

@@
 <!-- badges: start -->
+[![Codecov test coverage](https://codecov.io/gh/jhs-hwg/nhanes-shiny-bp/branch/main/graph/badge.svg)](https://app.codecov.io/gh/jhs-hwg/nhanes-shiny-bp?branch=main)
+[![R-CMD-check](https://github.com/jhs-hwg/nhanes-shiny-bp/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhs-hwg/nhanes-shiny-bp/actions/workflows/R-CMD-check.yaml)
 <!-- badges: end -->
@@
 
+## Weights
+
+if counts are requested, we re-calibrate weights.
+
 ## Validation of results from the shiny application

### 2022-09-26 R/nhanes_bp.R

@@ -1,49 +1,9 @@
 #' NHANES blood pressure data, 1999-2020
 #'
 #' A dataset containing variables that pertain to blood pressure from the
-#'   NHANES continuous examination cycles.
+#'   NHANES continuous examination cycles. The documenation of the variables
+#'   in this data are available online (TODO: add link)
 #'
-#' @format A data frame with 59799 rows and 35 variables:
-#' \describe{
-#'   \item{price}{price, in US dollars}
-#'   \item{carat}{weight of the diamond, in carats}
-#'   \item{svy_id}{survey id}
-#'   \item{svy_weight}{survey weight}
-#'   \item{svy_psu}{survey psu}
-#'   \item{svy_strata}{survey strata}
-#'   \item{svy_year}{survey year}
-#'   \item{demo_age_cat}{age category}
-#'   \item{demo_race}{race/ethnicity}
-#'   \item{demo_age_years}{age, years}
-#'   \item{demo_pregnant}{pregnancy, yes or no}
-#'   \item{demo_gender}{gender, male or female}
-#'   \item{bp_sys_mean}{systolic blood pressure, mm Hg}
-#'   \item{bp_dia_mean}{diastolic blood pressure, mm Hg}
-#'   \item{bp_cat_meds_excluded}{blood pressure category, excluding those using antihypertensive medication}
-#'   \item{bp_cat_meds_included}{blood pressure category, including those using antihypertensive medication}
-#'   \item{bp_med_use}{using antihypertensive medication, yes or no}
-#'   \item{bp_med_recommended_jnc7}{antihypertensive medication recommended by JNC7 guideline, yes or no}
-#'   \item{bp_med_recommended_accaha}{antihypertensive medication recommended by ACC/AHA 2017 guideline, yes or no}
-#'   \item{bp_med_n_class}{number of antihypertensive medication classes being used}
-#'   \item{bp_control_jnc7}{blood pressure control according to JNC7}
-#'   \item{bp_control_accaha}{blood pressure control according to ACC/AHA 2017 guideline}
-#'   \item{htn_jnc7}{hypertension according to JNC7}
-#'   \item{htn_accaha}{hypertension according to ACC/AHA 2017 guideline}
-#'   \item{htn_aware}{awareness of hypertenison}
[…]

### 2022-09-26 R/reliability_mean.R

@@ -1,18 +1,23 @@
 
+# reliability of a mean
+
 is_reliable_mean <- function(smry){
-   output <- smry %>%
-   #Calculate realtive standard error:
-   mutate(rse=std_error/mean,
 
-          # if rse>=.3 then p_reliable=0;
-          p_reliable=ifelse(rse>=0.3, 0, 1),
+ output <- smry %>%
+  #Calculate realtive standard error:
+  mutate(
+   rse=std_error/mean,
 
-          #create means with confidence intervals
-          means=format(round((mean), digits=1), nsmall=1),
-          se=format(round((std_error), digits=1), nsmall=1),
+   # if rse>=.3 then p_reliable=0;
+   p_reliable=ifelse(rse>=0.3, 0, 1),
 
-          CI_mean=glue("{means} ({se})"),
+   #create means with confidence intervals
+   means=format(round((mean), digits=1), nsmall=1),
+   se=format(round((std_error), digits=1), nsmall=1),
 
-          #create the finalized means with confidence intervals, with the value suppressed if p_reliable==0
-          CI_final=ifelse(p_reliable==0, NA, CI_mean))
- }
+   CI_mean=glue("{means} ({se})"),
+
+   #create the finalized means with confidence intervals, with the value suppressed if p_reliable==0
+   CI_final=ifelse(p_reliable==0, NA, CI_mean)
+  )
+}

### 2022-09-26 R/reliability_prop.R

@@ -1,9 +1,8 @@
 
 
-###################################Function for reliability of proportion################################
-#########################################################################################################
-is_reliable_prop <- function(data, design, smry)
- {
+# reliability of proportions
+
+is_reliable_prop <- function(data, design, smry) {
  Nobs <- get_obs_count(data)
  Npsu <- get_psu_count(design)
  Nstrata <- get_strata_count(design)

### 2022-09-26 R/svy_design_new.R

@@ -2,7 +2,7 @@
 
 #' Make a new design object
 #'
-#' @param data nhanes data - should be recoded (see [nhanes_recode])
+#' @param data nhanes data - should be recoded
 #' @param years the years to be included in the analysis
 #' @param exposure the exposure variable
 #' @param n_exposure_group how many exposure groups? only applies to
@@ -12,7 +12,7 @@
 #' @param pool are the results pooled across multiple NHANES cycles?
 #'  if so, say 'yes', o.w. say 'no'
 #'
-#' @export
+#' @noRd
 #'
 #' @return a `svydesign` object
 #'
@@ -61,6 +61,9 @@ svy_design_new <- function(data,
                      weights = ~ svy_weight,
                      data = data_design,
                      nest = TRUE)
+
+ attr(design, 'pool') <- pool
+ attr(design, 'exposure') <- exposure
 
  design
 

### 2022-09-26 tests/testthat/test-nhanes_variables.R

@@ -2,16 +2,6 @@
 library(testthat)
 library(haven)
 
-# pm_23sep2022 <- haven::read_sas('data-raw/final9920_0923022.sas7bdat')
-#
-# pm_23jul2022 <- haven::read_sas('data-raw/nhanes_bp-raw.sas7bdat')
-#
-# length(names(pm_23jul2022))
-#
-# setdiff(names(pm_23jul2022), names(pm_23sep2022))
-
-nhanes_sas <- nhanes_load(as = 'tibble') %>%
- nhanes_rename()
 
 key <- nhanes_key$data
 
@@ -40,44 +30,56 @@ test_that(
 
 )
 
-test_that(
- desc = 'Number of NAs and category counts match',
- code = {
-  for(i in names(nhanes_bp)){
-
-   type <- key[variable == i, type]
-
-   expect_equal(sum(is.na(nhanes_sas[[i]])),
-                sum(is.na(nhanes_bp[[i]])))
-
-   if(!grepl('^svy_weight', x = i) && i %in% names(nhanes_sas)){
-
-    if(type %in% c('ctns', 'svy') ){
-
-     # these variable should be exactly the same
-     expect_equal(nhanes_sas[[i]], nhanes_bp[[i]], ignore_attr = TRUE)
-
-    } else if(type %in% c('catg', 'bnry', 'time')){
-
-     # categorical variables that were recoded should have same counts
-     counts_sas   <- sort(table(nhanes_sas[[i]]))
-     counts_shiny <- sort(table(nhanes_bp[[i]]))
-
-     # don't test the counts if I manually lumped categories together
-     if(length(counts_sas) == length(counts_shiny))
-      expect_equal(counts_sas, counts_shiny, ignore_attr = TRUE)
-
-    } else {
-
-     stop("unrecognized variable type: ", type)
-
-    }
-
-   }
-
-  }
- }
-)
+# LOCAL tests
+# pm_23sep2022 <- haven::read_sas('data-raw/final9920_0923022.sas7bdat')
+#
+# pm_23jul2022 <- haven::read_sas('data-raw/nhanes_bp-raw.sas7bdat')
+#
+# length(names(pm_23jul2022))
+#
+# setdiff(names(pm_23jul2022), names(pm_23sep2022))
+#
+# nhanes_sas <- nhanes_load(as = 'tibble') %>%
+#  nhanes_rename()
+#
+# test_that(
+#  desc = 'Number of NAs and category counts match',
+#  code = {
[…]

### 2022-09-26 tests/testthat/test-svy_stat.R

@@ -474,4 +474,24 @@ test_that(
  }
 )
 
+#
+# testServer(
+#  app = server, {
+#   session$setInputs(do = 'data',
+#                     years = c(
+#                      "1999-2000",
+#                      "2001-2002",
+#                      "2003-2004",
+#                      "2005-2006",
+#                      "2007-2008",
+#                      "2009-2010",
+#                      "2011-2012",
+#                      "2013-2014",
+#                      "2015-2016",
+#                      "2017-2020"
+#                     ),
+#                     pool = 'no')
+#  }
+#
+# )
 

### 2022-09-26 README.Rmd
name changed!

@@
 
-# nhanesTrends <img src="man/figures/logo.png" align="right" height="139" />
+# cardioStats.USA <img src="man/figures/logo.png" align="right" height="139" />
 
@@
 # install.packages("remotes")
-remotes::install_github("jhs-hwg/nhanesShinyBP")
+remotes::install_github("jhs-hwg/cardioStats.USA")
 ```
@@
 
-library(nhanesShinyBP)
+library(cardioStats.USA)
 app_run()

### 2022-09-26 R/nhanes_summarize.R

@@ -45,8 +45,8 @@ nhanes_summarize <- function(outcome,
                              age_wts = NULL){
 
  # back-end objects
- key          <- nhanesShinyBP::nhanes_key
- nhanes       <- nhanesShinyBP::nhanes_bp
+ key          <- cardioStats.USA::nhanes_key
+ nhanes       <- cardioStats.USA::nhanes_bp
 
  # TODO check inputs
 

### 2022-09-26 R/nhanes_visualize.R

@@ -31,7 +31,7 @@ nhanes_visualize <- function(outcome,
                              width = NULL,
                              height = NULL){
 
- nhanes <- nhanesShinyBP::nhanes_bp
+ nhanes <- cardioStats.USA::nhanes_bp
 
  years <- switch(
   years[1],
@@ -50,10 +50,10 @@ nhanes_visualize <- function(outcome,
  if(is.null(statistic_primary)){
 
   outcome_type <-
-   nhanesShinyBP::nhanes_key$variables[[outcome]]$type
+   cardioStats.USA::nhanes_key$variables[[outcome]]$type
 
   statistic_primary <-
-   nhanesShinyBP::nhanes_key$svy_calls[[outcome_type]][1]
+   cardioStats.USA::nhanes_key$svy_calls[[outcome_type]][1]
 
  }
 

### 2022-09-26 R/svy_design_summarize.R

@@ -31,7 +31,7 @@ svy_design_summarize <- function(
   age_wts = c(0.155, 0.454, 0.215, 0.177)
 ){
 
- key <- nhanesShinyBP::nhanes_key
+ key <- cardioStats.USA::nhanes_key
  pool_svy_years = attr(design, 'pool') == 'yes'
  time_var <- ifelse(pool_svy_years, 'None', key$time_var)
  outcome_type <- key$variables[[outcome]]$type

### 2022-09-26 README.Rmd
drop the .

--- 2022-09-26 README.Rmd
@@
 
-# nhanesTrends <img src="man/figures/logo.png" align="right" height="139" />
+# cardioStatsUSA <img src="man/figures/logo.png" align="right" height="139" />
 
@@
 # install.packages("remotes")
-remotes::install_github("jhs-hwg/nhanesShinyBP")
+remotes::install_github("jhs-hwg/cardioStatsUSA")
 ```
@@
 
-library(nhanesShinyBP)
+library(cardioStatsUSA)
 app_run()

### 2022-09-26 R/is_discrete.R

@@ -2,7 +2,7 @@
 
 is_var_type <- function(var_name, var_types){
  if(is.null(var_name)) return(FALSE)
- cardioStats.USA::nhanes_key$variables[[var_name]]$type %in% var_types
+ cardioStatsUSA::nhanes_key$variables[[var_name]]$type %in% var_types
 }
 
 is_binary <- function(variable_name) {

### 2022-09-26 R/nhanes_summarize.R

@@ -45,8 +45,8 @@ nhanes_summarize <- function(outcome,
                              age_wts = NULL){
 
  # back-end objects
- key          <- cardioStats.USA::nhanes_key
- nhanes       <- cardioStats.USA::nhanes_bp
+ key          <- cardioStatsUSA::nhanes_key
+ nhanes       <- cardioStatsUSA::nhanes_bp
 
  # TODO check inputs
 

### 2022-09-26 tests/testthat.R

@@ -1,4 +1,4 @@
 library(testthat)
-library(cardioStats.USA)
+library(cardioStatsUSA)
 
-test_check("cardioStats.USA")
+test_check("cardioStatsUSA")

### 2022-09-27 R/svy_stat_tidy.R
pkgdown

@@ -7,6 +7,8 @@
 #'
 #' @return a data frame
 #'
+#' @noRd
+#'
 svy_stat_tidy <- function(x, outcome, by_vars) {
 
  tidy_fun <-  paste('svy', get_svy_stat_fun(x),

### 2022-09-27 README.Rmd
nhaneTrends no more
put table 1 in readme?

@@
 
-You can install the development version of nhanesTrends from [GitHub](https://github.com/) with:
+You can install the development version of cardioStatsUSA from [GitHub](https://github.com/) with:
 
@@
 
-## Weights
+## Why we made this
+
+## Methods
+
+### Weights
 
@@
 
-## Validation of results from the shiny application
+### Validation of results from the shiny application
 
@@
 TO BE WRITTEN: notes on discrepancies with results in [Carey et al.](https://www.ahajournals.org/doi/10.1161/HYPERTENSIONAHA.118.12191)
+
+## Results
+
+### Participant characteristics
+
+```{r}
+
+suppressPackageStartupMessages({
+ library(cardioStatsUSA)
+ library(magrittr)
+ library(gtsummary)
+ library(tidyverse)
+ library(survey)
+ library(gt)
+})
+
+tbl_summary_data <- as_tibble(nhanes_bp) %>% 
+ filter(svy_subpop_htn == 1) %>% 
+ mutate(nobs = 1) %>% 
+ select(
+  svy_psu,
+  svy_strata,
+  svy_weight_mec,
+  svy_year,
+  nobs,
+  starts_with('demo'),
+  starts_with('bp'), - bp_cat_meds_excluded,
+  starts_with('htn'),
+  starts_with('cc')
+ )
+
+for(i in names(tbl_summary_data)){
+ attr(tbl_summary_data[[i]], 'label') <- nhanes_key$variables[[i]]$label
+}
+
+tbl_summary_design <- tbl_summary_data %>% 
+ svydesign(ids = ~ svy_psu,
+           strata = ~ svy_strata,
+           weights = ~ svy_weight_mec,
+           data = .,
+           nest = TRUE)
+
+tbl_summary_design %>% 
+ tbl_svysummary(
+  by = svy_year,
+  include = c(nobs, 
+              starts_with("demo"),
+              starts_with('bp'), 
+              starts_with('htn'),
+              starts_with('cc')),
+  missing = 'no',
+  label = list(nobs ~ "No. of Adults"),
+  statistic = list(
+   all_categorical() ~ "{p}",
+   all_continuous() ~ "{mean} ({sd})",
+   nobs ~ "{N_unweighted}"
+  )
+ ) %>%
+ modify_header(all_stat_cols() ~ "**{level}**") %>%
+ modify_footnote(all_stat_cols() ~ NA) %>%
+ as_gt() %>%
+ tab_source_note("Table values are mean (standard deviation) or percent.")
+
+```

### 2022-09-27 README.Rmd
tweaks

@@
  modify_header(all_stat_cols() ~ "**{level}**") %>%
- modify_footnote(all_stat_cols() ~ NA) %>%
- as_gt() %>%
- tab_source_note("Table values are mean (standard deviation) or percent.")
+ modify_footnote(all_stat_cols() ~ NA)
+
+# %>%
+#  as_gt() %>%
+#  tab_source_note("Table values are mean (standard deviation) or percent.")
 

### 2022-09-28 paper/results.Rmd
beginning to look like a paper

---
title: "Tables and Figures"
always_allow_html: true
output: 
  officedown::rdocx_document:
    keep_md: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
 echo = FALSE,
 message = FALSE,
 warning = FALSE,
 fig.width = 7,
 fig.height = 4, 
 dpi = 300
)

suppressPackageStartupMessages({
 library(officer)
 library(cardioStatsUSA)
 library(magrittr)
 library(gtsummary)
 library(tidyverse)
 library(survey)
 library(flextable)
 library(plotly)
})

```

__Table 1__ Participant characteristics

```{r, cache=TRUE}

tbl_summary_data <- as_tibble(nhanes_bp) %>% 
 filter(svy_subpop_htn == 1) %>% 
 mutate(nobs = 1) %>% 
 select(
  svy_psu,
  svy_strata,
  svy_weight_mec,
  svy_year,
  nobs,
  starts_with('demo'),
  starts_with('bp'), - bp_cat_meds_excluded,
  starts_with('htn'),
  starts_with('cc')
 )

for(i in names(tbl_summary_data)){
 attr(tbl_summary_data[[i]], 'label') <- nhanes_key$variables[[i]]$label
}

tbl_summary_design <- tbl_summary_data %>% 
 svydesign(ids = ~ svy_psu,
           strata = ~ svy_strata,
           weights = ~ svy_weight_mec,
           data = .,
           nest = TRUE)

tbl_gtsummary <- tbl_summary_design %>% 
 tbl_svysummary(
  by = svy_year,
  include = c(nobs, 
              starts_with("demo"),
              starts_with('bp'), 
              starts_with('htn'),
              starts_with('cc')),
  missing = 'no',
  label = list(nobs ~ "No. of Adults"),
  statistic = list(
   all_categorical() ~ "{p}",
   all_continuous() ~ "{mean} ({sd})",
   nobs ~ "{N_unweighted}"
  )
 ) %>%
 modify_header(all_stat_cols() ~ "**{level}**") %>%
 modify_footnote(all_stat_cols() ~ NA)

as_flex_table(tbl_gtsummary) %>% 
 theme_box() %>% 
 width(width = 0.90) %>% 
 width(j = 1, width = 2) %>% 
 align(align = 'center') %>% 
 align(j = 1, align = 'left')

```

```{r}
block_section(
  prop_section(
    page_size = page_size(orient = "landscape"),
    page_margins = page_mar(bottom = 1/12, top = 1/12, 
                            right = 1/8, left = 1/8),
    type = "continuous"
  )
)
```

__Figure 1__ Prevalence of hypertension by calendar year

```{r}

subset_variables <- list("demo_pregnant")
subset_values <- list(c("No", "Missing"))

nhanes_visualize(outcome = 'htn_accaha', 
                 subset_variables = subset_variables,
                 subset_values = subset_values,
                 years = 'all')[[1]] %>% 
 layout(font = list(size = 40))

```

\newpage

__Figure 2__ Antihypertensive medication use by calendar year among adults with hypertension

```{r}

nhanes_visualize(
 outcome = 'bp_med_use', 
 subset_variables = c(subset_variables, 'htn_accaha'),
 subset_values = c(subset_values, 'Yes'),
 years = 'all'
)[[1]] %>% 
 layout(font = list(size = 40))

```

\newpage

__Figure 3__ Prevalence of resistant hypertension among adults with hypertension taking antihypertensive medication

```{r}

nhanes_visualize(
 outcome = 'htn_resistant_accaha', 
 subset_variables = c(subset_variables, 'htn_accaha', 'bp_med_use'),
 subset_values = c(subset_values, 'Yes', 'Yes'),
 years = 'all'
)[[1]] %>% 
 layout(font = list(size = 40))

```

\newpage

__Figure S1__: Participants included in the current analysis

\newpage

__Figure S2__ Prevalence of hypertension by age group

```{r}

nhanes_visualize(outcome = 'htn_accaha', 
                 exposure = 'demo_age_cat',
                 subset_variables = subset_variables,
                 subset_values = subset_values,
                 years = 'most_recent')[[1]] %>% 
 layout(font = list(size = 40))

```

\newpage

__Figure S3__ Prevalence of hypertension by sex

```{r}

nhanes_visualize(outcome = 'htn_accaha', 
                 exposure = 'demo_gender',
                 subset_variables = subset_variables,
                 subset_values = subset_values,
                 years = 'most_recent')[[1]] %>% 
 layout(font = list(size = 40))

```

\newpage

__Figure S4__ Prevalence of hypertension by race/ethnicity

```{r}

nhanes_visualize(outcome = 'htn_accaha', 
                 exposure = 'demo_race',
                 subset_variables = subset_variables,
                 subset_values = subset_values,
                 years = 'most_recent')[[1]] %>% 
 layout(font = list(size = 40))

```

### 2022-09-29 paper/results.Rmd
results first draft

@@
 ```{r setup, include = FALSE}
+
 knitr::opts_chunk$set(
@@
 
-__Table 1__ Participant characteristics
+**RESULTS**
+
+A total of `r sum(nhanes_bp$svy_subpop_htn)` survey participants (mean age 46 years, 52% women, 68% non-Hispanic white) were included in the BP and hypertension cohort (**Table 1**).
+The estimated prevalence of hypertension was 45.9% (95% CI 43.1, 48.8) in 1999-2000, 43.8% (95% CI 41.1, 46.5) in 2009-2010, and 47.0% (95% CI 44.5, 49.5) in 2017-2020 (**Figure 1**).
+For US adults in 2017-2020, hypertension was estimated to be more prevalent among older versus younger adults (**Figure S2**), more prevalent among men versus women (**Figure S3**), and more prevalent among adults are non-Hispanic Black versus other race/ethnicity subgroups (**Figure S4**).
+
+**Self-reported antihypertensive medication use**
+
+Among US adults with hypertension, the estimated percentage who self-reported using antihypertensive medication was 35.0% (95% CI 30.7, 39.7) in 1999-2000, 52.8% (95% CI 48.3, 57.2) in 2009-2010, and 51.3% (95% CI 48.7, 54.0) in 2017-2020 (**Figure 2**).
+For US adults with hypertension in 2017-2020, adults aged 65 to 74 years had a higher estimated percentage who self-report using antihypertensive medication versus other age groups (**Figure S5**), women had a higher estimated percentage who self-report using antihypertensive medication compared to men (**Figure S6**), and Hispanic adults had a lower estimated percentage who self-report using antihypertensive medication versus other race/ethnicity subgroups (**Figure S7**).
+
+**Classes of antihypertensive medication**
+
+Among US adults with hypertension, the estimated percentage using one or more classes of antihypertensive medication was 38.5% (95% CI 34.5, 42.4) in 1999-2000, 54.3% (95% CI 50.1, 58.5) in 2009-2010, and 55.4% (95% CI 54.0, 57.8) in 2017-2020 (**Figure 3**).
+For US adults with hypertension in 2017-2020, adults aged 75+ had a higher estimated percentage using one or more antihypertensive medication class versus other age groups (**Figure S8**), women a higher estimated percentage using one or more antihypertensive medication class versus men (**Figure S9**), and non-Hispanic white adults a higher estimated percentage using one or more classes of antihypertensive medication versus other race/ethnicity groups (**Figure S10**).
+
+**Resistant hypertension**
+
+Among US adults with hypertension using three or more classes of antihypertensive medication, the estimated prevalence of resistant hypertension was 70.6 (95% CI 62.8, 78.3) in 1999-2000, 62.7 (95% CI 58.6, 66.7) in 2009-2010, and 66.0 (95% CI 61.9, 70.1) in 2017-2020 (**Figure 4**).
+For US adults with hypertension using three or more classes of antihypertensive medication in 2017-2020, adults aged 18 to 44 years had a higher estimated prevalence of resistant hypertension versus other age groups (**Figure S11**), men a higher estimated prevalence versus women (**Figure S12**), and non-Hispanic Black adults a higher estimated prevalence versus other race/ethnicity groups (**Figure S13**).
+
+```{r}
+block_section(
+  prop_section(
+    page_size = page_size(orient = "portrait"),
+    type = "continuous"
+  )
+)
+```
+
+__Table 1__ Characteristics of participants included in the blood pressure and hypertension cohort of the online platform.
 
@@
 for(i in names(tbl_summary_data)){
- attr(tbl_summary_data[[i]], 'label') <- nhanes_key$variables[[i]]$label
+ attr(tbl_summary_data[[i]], 'label') <-
+  nhanes_key$variables[[i]]$label
 }
@@
  modify_header(all_stat_cols() ~ "**{level}**") %>%
+ add_overall(last = FALSE, col_label = "**Overall**") %>% 
  modify_footnote(all_stat_cols() ~ NA)
@@
  theme_box() %>% 
- width(width = 0.90) %>% 
+ width(width = 0.8) %>% 
  width(j = 1, width = 2) %>% 
@@
 
-__Figure 1__ Prevalence of hypertension by calendar year
+__Figure 1__ Prevalence of hypertension by calendar year.
 
@@
 
-__Figure 2__ Antihypertensive medication use by calendar year among adults with hypertension
+__Figure 2__ Antihypertensive medication use by calendar year among adults with hypertension.
 
@@
 
-__Figure 3__ Prevalence of resistant hypertension among adults with hypertension taking antihypertensive medication
+__Figure 3__ Number of antihypertensive medication classes by calendar period among adults with hypertension.
 
@@
 nhanes_visualize(
- outcome = 'htn_resistant_accaha', 
- subset_variables = c(subset_variables, 'htn_accaha', 'bp_med_use'),
- subset_values = c(subset_values, 'Yes', 'Yes'),
+ outcome = 'bp_med_n_class', 
+ subset_variables = c(subset_variables, 'htn_accaha'),
+ subset_values = c(subset_values, 'Yes'),
  years = 'all'
@@
 
-__Figure S1__: Participants included in the current analysis
+__Figure 4__ Prevalence of resistant hypertension among adults with hypertension taking three or more classes of antihypertensive medication.
+
+```{r}
+
+nhanes_visualize(
+ outcome = 'htn_resistant_accaha',
+ subset_variables = c(subset_variables, 
+                      'htn_accaha', 
+                      'bp_med_n_class'),
+ subset_values = c(subset_values, 
+                   'Yes', 
+                   list(c('Three', 'Four or more'))),
+ years = 'all'
+)[[1]] %>%
+ layout(font = list(size = 40))
+
+```
 
@@
 
-__Figure S2__ Prevalence of hypertension by age group
+__Figure S1__: Participants included in the current analysis.
+
+\newpage
+
+__Figure S2__ Prevalence of hypertension by age group in 2017-2020.
 
@@
 
-__Figure S3__ Prevalence of hypertension by sex
+__Figure S3__ Prevalence of hypertension by sex in 2017-2020.
 
@@
 
-__Figure S4__ Prevalence of hypertension by race/ethnicity
+__Figure S4__ Prevalence of hypertension by race/ethnicity in 2017-2020.
 
@@
 ```
+
+\newpage
+
+__Figure S5__ Antihypertensive medication use by age category in 2017-2020 among adults with hypertension.
+
+```{r}
+
+nhanes_visualize(
+ outcome = 'bp_med_use', 
+ exposure = 'demo_age_cat',
+ subset_variables = c(subset_variables, 'htn_accaha'),
+ subset_values = c(subset_values, 'Yes'),
+ years = 'most_recent'
+)[[1]] %>% 
+ layout(font = list(size = 40))
+
+```
+
+\newpage
+
+__Figure S6__ Antihypertensive medication use by sex in 2017-2020 among adults with hypertension.
+
+```{r}
+
+nhanes_visualize(
+ outcome = 'bp_med_use', 
+ exposure = 'demo_gender',
+ subset_variables = c(subset_variables, 'htn_accaha'),
+ subset_values = c(subset_values, 'Yes'),
+ years = 'most_recent'
+)[[1]] %>% 
+ layout(font = list(size = 40))
+
+```
+
+\newpage
+
+__Figure S7__ Antihypertensive medication use by race/ethnicity in 2017-2020 among adults with hypertension.
+
+```{r}
+
+nhanes_visualize(
+ outcome = 'bp_med_use', 
+ exposure = 'demo_race',
+ subset_variables = c(subset_variables, 'htn_accaha'),
+ subset_values = c(subset_values, 'Yes'),
+ years = 'most_recent'
+)[[1]] %>% 
+ layout(font = list(size = 40))
+
+```
+
+\newpage
+
+__Figure S8__ Number of antihypertensive medication classes by age category in 2017-2020 among adults with hypertension.
+
+```{r, fig.height = 5}
+
+margin <- list(l = 50, r = 50, b = 50, t = 300)
+
+nhanes_visualize(
+ outcome = 'bp_med_n_class', 
+ exposure = 'demo_age_cat',
+ subset_variables = c(subset_variables, 'htn_accaha'),
+ subset_values = c(subset_values, 'Yes'),
+ years = 'most_recent',
+ pool = TRUE
+)[[1]] %>% 
+ layout(font = list(size = 40), margin = margin)
+
+```
+
+\newpage
+
+__Figure S9__ Number of antihypertensive medication classes by sex in 2017-2020 among adults with hypertension.
+
+```{r, fig.height = 5}
+
+nhanes_visualize(
+ outcome = 'bp_med_n_class', 
+ exposure = 'demo_gender',
+ subset_variables = c(subset_variables, 'htn_accaha'),
+ subset_values = c(subset_values, 'Yes'),
+ years = 'most_recent',
+ pool = TRUE
+)[[1]] %>% 
+ layout(font = list(size = 40), margin = margin)
+
+```
+
+\newpage
+
+__Figure S10__ Number of antihypertensive medication classes by race/ethnicity in 2017-2020 among adults with hypertension.
+
+```{r, fig.height = 5}
+
+nhanes_visualize(
+ outcome = 'bp_med_n_class', 
+ exposure = 'demo_race',
+ subset_variables = c(subset_variables, 'htn_accaha'),
+ subset_values = c(subset_values, 'Yes'),
+ years = 'most_recent',
+ pool = TRUE
+)[[1]] %>% 
+ layout(font = list(size = 40), margin = margin)
+
+```
+
+\newpage
+
+__Figure S11__ Prevalence of resistant hypertension by age category in 2017-2020 among adults with hypertension taking three or more classes of antihypertensive medication.
+
+```{r}
+
+nhanes_visualize(
+ outcome = 'htn_resistant_accaha',
+ exposure = 'demo_age_cat',
+ subset_variables = c(subset_variables, 
+                      'htn_accaha', 
+                      'bp_med_n_class'),
+ subset_values = c(subset_values, 
+                   'Yes', 
+                   list(c('Three', 'Four or more'))),
+ years = 'most_recent'
+)[[1]] %>%
+ layout(font = list(size = 40))
+
+```
+
+\newpage
+
+__Figure S12__ Prevalence of resistant hypertension by sex in 2017-2020 among adults with hypertension taking three or more classes of antihypertensive medication.
+
+```{r}
+
+nhanes_visualize(
+ outcome = 'htn_resistant_accaha',
+ exposure = 'demo_gender',
+ subset_variables = c(subset_variables, 
+                      'htn_accaha', 
+                      'bp_med_n_class'),
+ subset_values = c(subset_values, 
+                   'Yes', 
+                   list(c('Three', 'Four or more'))),
+ years = 'most_recent'
+)[[1]] %>%
+ layout(font = list(size = 40))
+
+```
+
+\newpage
+
+__Figure S13__ Prevalence of resistant hypertension by race/ethnicity in 2017-2020 among adults with hypertension taking three or more classes of antihypertensive medication.
+
+```{r}
+
+nhanes_visualize(
+ outcome = 'htn_resistant_accaha',
+ exposure = 'demo_race',
+ subset_variables = c(subset_variables, 
+                      'htn_accaha', 
+                      'bp_med_n_class'),
+ subset_values = c(subset_values, 
+                   'Yes', 
+                   list(c('Three', 'Four or more'))),
+ years = 'most_recent'
+)[[1]] %>%
+ layout(font = list(size = 40))
+
+```

### 2022-09-29 R/nhanes_calibrate.R

@@ -8,7 +8,8 @@ nhanes_calibrate <- function(nhanes_full,
                              nhanes_sub,
                              calib_by = c('svy_year',
                                           'demo_age_cat',
-                                          'demo_gender')){
+                                          'demo_gender',
+                                          'demo_race_black')){
 
 
  # find total sum of weights across all combinations of the by groups

### 2022-09-29 data-raw/nhanes_recode.R

@@ -53,6 +53,11 @@ nhanes_recode <- function(data = NULL){
     )
    ),
 
+   race_black = factor(
+    if_else(race_wbaho == "Non-Hispanic Black", "Yes", "No"),
+    levels = c("No", "Yes")
+   ),
+
 
    # number of BP medication classes
    num_htn_class = factor(

### 2022-09-29 data-raw/nhanes_rename.R

@@ -26,11 +26,12 @@ nhanes_rename <- function(data){
   svy_subpop_chol = subpopldl,
 
   # Demographics ------------------------------------------------------------
-  demo_age_cat   = agecat4,
-  demo_race      = race_wbaho,
-  demo_age_years = ridageyr,
-  demo_pregnant  = pregnant,
-  demo_gender    = riagendr,
+  demo_age_cat    = agecat4,
+  demo_race       = race_wbaho,
+  demo_race_black = race_black,
+  demo_age_years  = ridageyr,
+  demo_pregnant   = pregnant,
+  demo_gender     = riagendr,
 
   # Blood pressure ----------------------------------------------------------
   bp_sys_mean               = avgsbp,

### 2022-09-30 README.Rmd
beginning to look like a paper
keep the shiny link

--- 2022-09-26 README.Rmd
@@
 Our goal is to provide a platform for exploration of the National Health and Nutrition Examination Survey (NHANES) data.
+You can find the online application for exploratory data here: https://bcjaeger.shinyapps.io/nhanesShinyBP/
 

### 2022-09-30 paper/results.Rmd
keep the shiny link
about to overhaul results

@@
 ```
+
+**INTRODUCTION**
+
+The National Health and Nutrition Examination Survey (NHANES) is a program conducted by the US National Center for Health Statistics (NCHS) of the Centers for Disease Control and Prevention (CDC) and is designed to assess the health and nutritional status of the US population.
+The NHANES program was initiated in the early 1960s and beginning in 1999 has been conducted continuously, in two-year cycles.
+NHANES includes participant interviews and physical examinations, each conducted following rigorous protocols.
+The NHANES interview includes questions about prior diagnoses of hypertension and self-reported antihypertensive medication use and a medication inventory where containers for prescription medications are reviewed and recorded.
+The physical examination includes blood pressure (BP) measurements by trained and certified physicians.
+
+NHANES data have been used to provide estimates of the proportion of US adults with hypertension, aware they had hypertension, taking antihypertensive medication and with controlled BP.
+These data have been informative for public health, informing policy making, identifying gaps in guideline-recommended care and for scientific discovery.
+The 2017 American College of Cardiology/American Heart Association (ACC/AHA) BP guideline included analyses of NHANES data estimating the percentage of US adults with hypertension based on its updated definition.
+Additionally, NHANES data have been used to track the proportion of US adults with controlled BP, which were included in a Call-to-Action to Control BP from the US Surgeon General.
+
+NHANES data are publically available and easily accessible through the CDC website.
+However, there are barriers to analyzing NHANES data.
+The NHANES program uses a multi-stage sampling design to select participants for enrollment and NHANES data require specific programming commands to provide valid statistical estimates.
+Also, the NHANES protocols are detailed and each two-year NHANES cycle includes over 50 data files, each of which has detailed informational material on the included data.
+To address the barriers to analyzing NHANES data, we developed an open-access online platform that allows users to obtain visual and tabular summaries of the NHANES BP and hypertension data.
+The goal of this platform is to increase the usage of NHANES data for BP and hypertension research and policy decision making with a focus on ensuring results are valid and transparent.
+In this manuscript, we review the design of the platform, we show some results of NHANES analyses using the platform, and we provide a tutorial on how people can use the platform.
+
+**METHODS**
+
+The open-access online platform that we describe in this manuscript contains data from 10 NHANES cycles – NHANES 1999-2000 through NHANES 2017-2020.
+In each NHANES cycle, potential participants are identified using a multi-stage selection process.
+First, every county across the US is placed into 1 of 15 groups based on their characteristics.
+Next, one county is selected from each group.
+Each of the 15 selected counties are divided into neighborhoods and 20 to 24 neighborhoods are selected and 30 households are chosen from all houses and apartments in the neighborhood.
+Within the selected household, a random process is used to select none, some or all members for inclusion in NHANES.
+The participation rate in NHANES has declined from 76% of individuals who completed the interview and examination in the 1999-2000 cycle to 47% in the 2017-2020 cycle.
+
+NHANES data were collected through in-home interviews and study examinations conducted at mobile examination centers.
+Data elements included on the platform are listed, along with variable definitions, in Supplemental Table 1.
+Information on age, sex, race/ethnicity, cigarette smoking, a prior diagnosis of diabetes, use of insulin or oral glucose-lowering medication, a prior diagnosis of hypertension, self-reported antihypertensive medication use, history of coronary heart disease, myocardial infarction, stroke, heart failure were obtained by self-report during the in-home interview.
+
+Details on prescription medications taken in the last 30 days were obtained by a pill bottle review conducted as part of the in-home interview.
+Specifically, the labels of medications were reviewed and the complete drug names were recorded into a computer-assisted personal interview system.
+For the data platform, we categorized antihypertensive medications by class: aldosterone receptor antagonist, α-blocker, angiotensin converting enzyme inhibitor (ACEI), angiotensin receptor blocker (ARB), beta-blocker, calcium channel blocker (CCB), centrally acting agents (e.g., clonidine), direct vasodilator, loop diuretic, potassium sparing diuretic, thiazide diuretic, and renin inhibitor.
+Data on medication dosages are not available in NHANES.
+
+Following the in-home interview, participants complete a study examination during which height and weight are measured and used to calculate body mass index (BMI).
+Systolic and diastolic blood pressure
+are measured three times, and blood and spot urine samples are collected.
+Of relevance to the data platform, the collected blood was used to measure total and HDL cholesterol, glycated hemoglobin and serum creatinine.
+The spot urine was used to measure albumin and creatinine.
+
+The current data platform included participants from NHANES 1999-2000 to 2017-2020 (Figure S1; n=107,622).
+As statistics for hypertension and SBP and DBP levels in children and adolescents are markedly different than for adults, we restricted the data in the platform to adults ≥ 18 years of age.
+We further restricted the platform to participants who completed the in-home study visit and study examination, with one or more SBP and DBP measurements and who had data on self-reported antihypertensive medication use.
+After these exclusions were applied, the data set that we used included 56,035 participants (**Figure S1**).
+
+Statistical analysis
+All statistical analyses account for the non-random selection of NHANES participants using survey software in R Version 4.1.3.
+Additionally, to provide estimates for the non-institutionalized US population, we applied the NHANES examination weights to all calculations.
+For estimating population counts, we recalibrated the weights to account for participants being excluded due to missing information on SBP, DBP or self-reported antihypertensive medication use.
+The weights were calibrated within age, sex, race/ethnicity and NHANES cycle.
 
@@
  layout(font = list(size = 40))
+
+```
+
+\newpage
+
+__Table S1__ Variable definitions for the online platform.
+
+```{r}
+
+library(tidyverse)
+library(flextable)
+
+nhanes_key$data %>% 
+ as_tibble() %>%
+ select(class, label, source, description) %>% 
+ as_grouped_data(groups = 'class') %>% 
+ as_flextable(hide_grouplabel = TRUE) %>% 
+ theme_box() %>% 
+ width(width = 1) %>% 
+ width(j = 1, width = 2.5)
 

### 2022-10-01 README.Rmd
logo correct
jhs retreat

@@
 
-# cardioStatsUSA <img src="man/figures/logo.png" align="right" height="139" />
+# cardioStatsUSA <a href="https://jhs-hwg.github.io/cardioStatsUSA/"><img src="man/figures/logo.png" align="right" height="139" /></a>
 
@@
 
-## Why we made this
+## TODO list
 
-## Methods
+- Write vignettes for tests of the app
+- Write vignette for medication data
+- Replication with AB's group
+- Incorporate suppression of unreliable estimates
+- Documentation/demos for R functions used in the app
+- Definitions/notes for all BP variables (table S1 in manuscript)
+- Tutorial videos
+- Do we need to use AF weights?
 
-### Weights
-
-if counts are requested, we re-calibrate weights.
-
-### Validation of results from the shiny application
+## Validation of results from the shiny application
 

### 2022-10-01 paper/manuscript.Rmd
jhs retreat

---
title: "Cardiometabolic statistics for US adults: an open source web application for exploratory analysis"
always_allow_html: true
output: 
  officedown::rdocx_document:
    reference_docx: style_manuscript_times_new_roman.docx
    keep_md: true
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
 echo = FALSE,
 message = FALSE,
 warning = FALSE,
 fig.width = 10.5,
 fig.height = 4, 
 dpi = 300
)

suppressPackageStartupMessages({
 library(officer)
 library(cardioStatsUSA)
 library(magrittr)
 library(gtsummary)
 library(tidyverse)
 library(survey)
 library(flextable)
 library(plotly)
})

# default is double space 
# single spacing for title page, abstract, captions
single_par <- fp_par(line_spacing=1)

```

Byron C. Jaeger, PhD,^1^ 
Ligong Chen, PhD,^2^
Adam P. Bress, PharmD, MS,^3,4^ 
YOUR NAME, DEGREE,^x^
Bharat Poudel, MSPH,^2^
Ashley Christenson, BS,^2^ 
Lisandro D. Colantonio, PhD,^2^
Paul Muntner, PhD,^2^

^1^Department of Biostatistics and Data Science, Wake Forest School of Medicine, Winston-Salem, NC. `r single_par`
^2^Department of Epidemiology, University of Alabama at Birmingham. `r single_par`
^3^Informatics, Decision-Enhancement, and Analytic Sciences (IDEAS) Center, Veterans Affairs, Salt Lake City Health Care System, Salt Lake City, UT. `r single_par`
^4^Department of Population Health Sciences, University of Utah School of Medicine, Salt Lake City, UT. `r single_par`
^x^ YOUR AFFILIATION

\newpage

**ABSTRACT**

\newpage

The National Health and Nutrition Examination Survey (NHANES) is a program conducted by the US National Center for Health Statistics (NCHS) of the Centers for Disease Control and Prevention (CDC) and is designed to assess the health and nutritional status of the US population. The NHANES program was initiated in the early 1960s and beginning in 1999 has been conducted continuously, in two-year cycles. 
NHANES data are collected through participant interviews and physical examinations, each conducted following standardized protocols. The NHANES interview includes questions about prior diagnoses of hypertension and self-reported antihypertensive medication use and a medication inventory where containers for prescription medications are reviewed and recorded. The physical examination includes blood pressure (BP) measurements by trained and certified physicians.

NHANES data have been used to provide estimates of the proportion of US adults with hypertension, aware they had hypertension, taking antihypertensive medication and with controlled BP. Data from NHANES have been used to estimate the impact of the lower BP levels used to define hypertension and controlled BP in the 2017 American College of Cardiology/American Heart Association (ACC/AHA) BP guideline versus the Seventh Joint National Committee (JNC7) BP guideline on the proportion and number of US adults with hypertension, recommended antihypertensive medication and with uncontrolled BP.  Additionally, NHANES data have been used to track the proportion of US adults with controlled BP, which were included in a Call-to-Action to Control BP from the US Surgeon General.

NHANES data are publically available and easily accessible through the CDC website. However, analyzing NHANES data may be challenging for many people due to the complexity of its study design. The NHANES program uses a multi-stage sampling design to select participants for enrollment and data analysis requires specific programming commands to provide valid statistical estimates. Also, each NHANES cycle includes over 50 data files, each of which has detailed informational material. NHANES data collection protocols have changed over time for certain variables, and these differences should be accounted for when analyzing data. To address the barriers to analyzing NHANES data, we developed an open-access web-based application that allows users to obtain BP and hypertension statistics for US adults using NHANES data. The goal of the application is to increase the usage of NHANES data for BP and hypertension research and policy decision making with a focus on ensuring results are valid and transparent. In this manuscript, we review the design of the application, show some results of NHANES analyses using the application, and provide a tutorial on how people can use the application.

**METHODS**

The open-access web-based application uses data from 10 NHANES cycles, specifically NHANES 1999-2000 through NHANES 2017- March 2020, prior to the COVID-19 pandemic.  In each NHANES cycle, potential participants were identified using a multi-stage sampling process.  The protocols for each NHANES cycle were approved by the NCHS Institutional Review Board. Written informed consent was obtained from each participant.

NHANES data were collected through in-home interviews and study examinations conducted at mobile examination centers. Data elements included in the application, along with their definitions, are listed in Supplemental Table 1. Information on age, sex, race/ethnicity, pregnancy status for women, cigarette smoking, a prior diagnosis of diabetes, use of insulin or oral glucose-lowering medication, a prior diagnosis of hypertension, self-reported antihypertensive medication use, and history of coronary heart disease, myocardial infarction, stroke, and heart failure were obtained by self-report during the in-home interview.  During the in-home interview, the labels of medications that participants reported taking in the preceding 30 days were recorded. For the application, we categorized antihypertensive medications by class as defined in the 2017 ACC/AHA BP guideline. Data on medication dosages are not available in NHANES.

During the study examination, height and weight were measured and used to calculate body mass index. Systolic and diastolic BP (SBP and DBP, respectively) were measured three times, and blood and spot urine samples were collected.  BP was measured using a mercury sphygmomanometer from 1999-2000 through 2015-2016 and using an oscillometric device in 2017-2020.  The oscillometric BP values were calibrated to the mercury device as described previously. Blood was used to measure total and high-density lipoprotein (HDL) cholesterol, glycated hemoglobin and serum creatinine.  Spot urine was used to measure albumin and creatinine and to conduct a pregnancy test.

Overall, 107,622 US individuals participated in NHANES 1999-2000 to 2017- March 2020.  As statistics for hypertension and SBP and DBP levels in children and adolescents are markedly different than for adults, we restricted the data in the application to adults ≥ 18 years of age.  We further restricted the application to participants who completed the in-home interview and study examination, with one or more SBP and DBP measurement and who had data on self-reported antihypertensive medication use.  After these exclusions were applied, the application included data from 56,035 participants (Figure S1).

*Statistical analysis*

The application was created using Shiny, an open-source software package for creating web applications using the R programming language. Shiny translates R programming code into HTML, CSS, or JavaScript commands and creates a website interface. The statistical analyses that are used on the application are conducted with R programming language packages that account for the non-random selection of NHANES participants.  All statistical analyses on the application are weighted to provide estimates for the non-institutionalized US population. Following CDC recommendations, the application automatically suppresses statistics which are unreliable due to being based on a small sample size.

The application can be used to estimate characteristics of the population which are relevant when investigating BP and hypertension data (Table 1, top panel).  The application also allows users to estimate statistics for a number of hypertension-related variables (Table 1, bottom panel).

The application was designed with a number of options so that analyses can be customized to the user’s specifications. Results can be presented as figures or in tables. Figures can be made with the data represented as bars or points.  The 10 NHANES cycles can be analyzed separately or multiple cycles can be pooled together, which provides a larger sample size leading to more stable statistical estimates. The analyses can be done crude or with age-adjustment using direct standardization in four age groups (18-44, 45-64, 65-74 and ≥75 years of age).  Users can select the age distribution to be used as the standard population.  Analyses can be done in sub-groups and stratified by participant characteristics.

To demonstrate the features and utility of the application, we conducted a series of statistical analyses. We estimated the crude and age-adjusted mean SBP and prevalence of hypertension for US adults by NHANES cycle from 1999-2000 through 2017-2020. We estimated the age-adjusted prevalence of hypertension by NHANES cycle for individuals with and without chronic kidney disease, separately.  For age adjustment in these analyses, we used the estimated age distribution of US adults from 1999 to 2020 as the standard (49.3%, 33.6%, 10.1% and 7.0% being 18 to 44, 45 to 64, 65 to 74 and ≥75 years of age, respectively).  Also, we estimated the absolute number of US adults with hypertension from 1999-2000 through 2017-2020. For the estimation of absolute counts of US adults, the application recalibrates the statistical weights to account for participants being excluded due to missing information on SBP, DBP or self-reported antihypertensive medication use.  Weights are recalibrated within age, sex, race/ethnicity groups and NHANES cycle.

The age-adjusted distribution of the number of antihypertensive medication classes being taken was estimated for US adults with hypertension in each NHANES cycle, overall and among those who self-reported taking antihypertensive medication and had at least one antihypertensive drug class identified in their medication inventory. The age-adjusted proportion of US adults with controlled BP, defined using the 2017 ACC/AHA BP guideline, was estimated by NHANES cycle among US adults with hypertension and who self-reported taking antihypertensive medication.   The age-adjusted prevalence of resistant hypertension was estimated by NHANES cycle for US adults with hypertension taking antihypertensive medication and among those with hypertension taking ≥ 3 classes of antihypertensive medication.  For the analyses of number of antihypertensive medication classes being taken, resistant hypertension and BP control, we set the age distribution for the standard population to represent US adults with hypertension from 1999 to 2020: 26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥75 years of age, respectively.  Using the stratification option in the application, we estimated the prevalence of hypertension in NHANES 2017-2020 by age group, sex, and race-ethnicity. Using the data pooling option in the application, we grouped the NHANES 2011-2020 cycles to estimate the proportion of pregnant women who had uncontrolled BP according to the 2017 ACC/AHA BP guideline, SBP ≥ 130 mm Hg or DBP ≥ 80 mm Hg.

**RESULTS**

The estimated crude mean SBP for all US adults was 123 mm Hg in 1999-2000, 120 mm Hg in 2009-2010, and 123 mm Hg in 2017-2020 (Figure 1; panel A). The estimated age-adjusted mean SBP was 124, 120, and 122 mm Hg in 1999-2000, 2009-2010 and 2017-2020, respectively (Figure 1; panel B). The estimated crude prevalence of hypertension among US adults was lowest in 2009-2010 (43.3%) and highest in 2017-2020 (46.6%) (Figure 2; panel A). After age-adjustment, the prevalence of hypertension was highest in 1999-2000 (47.9%), lowest in 2009-2010 and 2013-2014 (43.0%) and was 44.7% in 2017-2020. (Figure 2; panel B). The number of US adults with hypertension increased from 89.8 million in 1999-2000 to 115.3 million US adults in 2017-2020 (Figure 3). The age adjusted mean SBP was 140 in 1999-2000, 134 in 2009-2010, and 136 in 2017-2020 among US adults with hypertension and 117 in 1999-2000, 115 in 2009-2010, and 117 in 2017-2020 among US adults without hypertension (Figure S2). In each NHANES cycle, the age-adjusted estimated mean SBP was 18 to 23 mm Hg higher among US adults with versus without hypertension.  In 2017-2020, a higher percentage of US adults with versus without hypertension were non-Hispanic Black (13.5% versus 9.3%) while a lower percentage of US adults with versus without hypertension were Hispanic (12.3% versus 18.1%) (Table 1). In each NHANES cycle, the estimated age-adjusted prevalence of hypertension among US adults with CKD was higher versus their counterparts without CKD (Figure S3).

Among US adults with hypertension, 61.5% were not taking antihypertensive medication in 1999-2020 compared with 44.6% in 2017-2020. (Figure 4; panel A). Among US adults who self-reported taking antihypertensive medication, the percentage taking four or more classes of antihypertensive medication increased from 3.5% in 1999-2000 to 7.6% in 2017-2020 (Figure 4; panel B). The age-adjusted estimated prevalence of BP control among US adults with hypertension was 10.2%, 25.8%, and 23.8% in 1999-2000, 2009-2010, and 2017-2020, respectively (Figure 5; panel A). For adults who self-report taking antihypertensive medication, 26.2%, 44.1%, and 43.3% had controlled BP in 1999-2000, 2009-2010, and 2017-2020, respectively (Figure 5; panel B).

Among US adults with hypertension who self-reported taking antihypertensive medication, the prevalence of resistant hypertension was 13.7% in 1999-2000, 17.7% in 2009-2010, and 16.2% in 2017-2020 (Figure S4; panel A). Among those who were taking three or more classes of antihypertensive medication, the prevalence of resistant hypertension was 73.5%, 64.1%, and 67.9% in 1999-2000, 2009-2010, and 2017-2020, respectively (Figure S4; panel B).

In 2017-2020, the prevalence of hypertension was higher at progressively older age (Figure S5),  among men versus women (Figure S6), and among non-Hispanic Black adults versus non-Hispanic White adults, non-Hispanic Asian adults, Hispanic adults and adults from other race/ethnic groups (Figure S7).

**DISCUSSION**

In the current manuscript, we describe the development of an open-access web-based application that allows the calculation of nationally representative estimates for a number of BP-related outcomes using publicly available data from NHANES. Using this app, we generated statistics for some of these BP-related outcomes over time, both crude and adjusted by age. We also generated statistics stratified by characteristics of US adults (e.g., CKD), and by pooling multiple NHANES cycles to obtain more precise estimates when working with small group sizes. Researchers, stakeholders, and the public in general interested in BP-related outcomes can now use the current app to accomplish a variety of purposes.

NHANES is a program specifically designed to obtain nationally representative estimates of the health and nutrition status of non-institutionalized US adults. Therefore, NHANES represents the ideal data source to obtain valid and reliable statistic on BP-related outcomes among US adults. Although NHANES data are publicly available, working with these data is not exempt of challenges. Examples of challenges that may face working with NHANES data include the need to download multiple files, even for a single NHANES cycle; combining multiple NHANES variables to create the measurement of interest, which may require dealing with missing data and questionnaire skip patterns; harmonizing definitions across multiple NHANES cycles; and the analyses of complex survey design data. NHANES has been conducted over many cycles and some of the methods have been updated over time. For example, NHANES switched from using sphygmomanometer devices to oscillometric devices to measure BP in 2017-2018. NHANES data users need to be careful when analyzing data across multiple cycles. Reducing barriers to using NHANES data will facilitate the use of these data to inform BP-related public health decisions and future research.

Several design decisions have been incorporated into the app.  We required participants to have only a single SBP and DBP to be included. This approach is consistent with several analyses conducted by CDC investigators (cite pmid 26633197, 29155682). However, mean BP and the prevalence of high BP would be lower if we required multiple BP measurements (cite pmid 23126346). Any bias resulting from this decision is likely to be small as over 95% of adult NHANES participants within one SBP and DBP measurement had three SBP and DBP measurements.  We re-calibrated the NHANES weights for the calculation of population counts (cite pmid 12500213). When estimating population counts, participants missing data on SBP, DBP or antihypertensive medication use cannot have BP or hypertension-related outcomes.  Weights were not recalibrated when estimating portions as participants missing data are removed from the numerator and denominator.  Medication classes were coded using generic names and the drug classes in the 2017 ACC/AHA BP guideline.  We recognize the NCHS recommends using Lexicon Plus®, a proprietary database, to categorize medication classes.  While the categorization of most medications is identical using generic drug names and Lexicon Plus®, some differences exist.  Many additional decisions were made regarding the definitions of variables, inclusion of study participants, and analytic approach.  We sought to make decisions that would be acceptable to people who use the app and transparent to ensure the results could be described accurately.

We were able to replicate several prior manuscripts using the app.  However, some results could not be replicated.  For example, a prior manuscript reported the prevalence of resistant hypertension defined by the 2017 ACC/AHA BP guideline to be 19.7% in 2009-2014.  When estimated by the app, the prevalence of resistant hypertension was 17.7%, 16.5% and 17.2% in 2009-2010, 2011-2012, and 2013-2014, respectively (cite pmid 30580690).  The difference in the prevalence estimates can be attributed to the approach used to categorize medication into classes. For example, the prior publication used Lexicon Plus®, which counted spironolactone as two drug classes, a potassium-sparing diuretic and an aldosterone antagonist.  Additionally, Lexicon Plus®includes Sotalol as a beta blocker and nitroglycerine as a direct vasodilator and we did not include these drugs as antihypertensive medications as they are not listed in the 2017 ACC/AHA BP guideline.  The differences in results between the app and this previously published manuscript emphasize two important points. First, while we encourage users to check results when using the app with prior publications, it may not be possible to always generate results that are identical to prior publications. Second, users should be aware of the choices made in defining variables.

The web application has a number of complementary features that allow users to complete various analysis tasks. In particular, users can generate descriptive summaries of the US population spanning over 20 years, including summaries of characteristics and cardiometabolic outcomes. The flexibility of inclusion criteria paired with the large set of outcomes and stratifying variables available allow an enormous number of hypotheses to be explored. Outputs from the app can be included in scientific proposals, presentations, or articles, citing the application's publicly available documentation to contextualize results. With the ability to download summary data from the application, users can visualize or tabulate results with aesthetics and software of their choosing. The web application can also be used for educational purposes, as teachers can use it to analyze health outcomes for US adults interactively with students.

With an open-source software license and publicly available documentation and data, the web application we developed is designed to be extended. The Shiny interface and R codes supporting this application can support general sub-populations of NHANES participants, such as a sub-population of participants with cholesterol data or a sub-population of pediatric survey participants. As the current analysis demonstrates creation and analysis of a sub-population of participants with BP measurements, others can use it as a template for developing extensions of the web application that engage with new sub-populations. In addition to including new sub-populations of NHANES participants, the application may be extended to include inferential statistics and hypothesis testing, and may also incorporate new cycles of NHANES data as they become available.

\newpage

```{r}
block_section(
  prop_section(
    page_size = page_size(orient = "portrait"),
    type = "continuous"
  )
)
```

**Table 1**:	Race/ethnicity distribution of the population with and without hypertension in 2017-2020

```{r}

library(table.glue)

nhanes_summarize(outcome = 'demo_race', 
                 years = 'most_recent', 
                 group = 'htn_accaha') %>%
 mutate(across(where(is.numeric), table_value)) %>% 
 filter(statistic == 'percentage') %>% 
 flextable() %>% 
 theme_box() %>% 
 width(width = 1) %>% 
 align(align = 'center', part = 'all') %>% 
 align(j = 1, align = 'center', part = 'all')

```

\newpage

**Figure 1** Mean systolic blood pressure for US adults by calendar year

A. Without age adjustment
```{r, fig.height=3.75}

fig_print <- function(x){
 x[[1]] %>% 
  layout(font = list(size = 40))
}

p1 <- nhanes_visualize(outcome = 'bp_sys_mean',
                       years = 'all', 
                       geom = 'bar')

age_wts_overall <- c(49.3, 33.6, 10.1, 7)

p2 <- nhanes_visualize(outcome = 'bp_sys_mean',
                       years = 'all', 
                       age_wts = age_wts_overall, 
                       geom = 'bar')

fig_print(p1)

```

B. With age adjustment through direct standardization.
```{r, fig.height=3.75}
fig_print(p2)
```

\newpage

**Figure 2** Prevalence of hypertension for US adults by calendar year

A. Without age adjustment
```{r, fig.height=3.75}

p1 <- nhanes_visualize(outcome = 'htn_accaha',
                       years = 'all', 
                       geom = 'bar')

p2 <- nhanes_visualize(outcome = 'htn_accaha',
                       years = 'all', 
                       age_wts = age_wts_overall, 
                       geom = 'bar')

fig_print(p1)

```

B. With age adjustment through direct standardization.
```{r, fig.height=3.75}
fig_print(p2)
```

```{r}

nhanes_visualize(outcome = 'htn_accaha',
                 years = 'all',
                 exposure = 'cc_ckd',
                 age_wts = age_wts_overall, 
                 geom = 'point') %>% 
 fig_print()

```

\newpage

**Figure 3** Number of US adults with hypertension

```{r}

nhanes_visualize(outcome = 'htn_accaha',
                 years = 'all',
                 statistic_primary = 'count',
                 geom = 'bar') %>% 
 fig_print()

```

\newpage

**Figure 4** Distribution of the number of antihypertensive medication classes being taken among those with hypertension by calendar period

A. Among those with hypertension
```{r, fig.height=3.75}

p1 <- nhanes_visualize(outcome = 'bp_med_n_class',
                       subset_variables = list('htn_accaha'),
                       subset_values = list("Yes"),
                       years = 'all', 
                       geom = 'bar')

p2 <- nhanes_visualize(outcome = 'bp_med_n_class',
                       subset_variables = list('htn_accaha', 'bp_med_use'),
                       subset_values = list("Yes", "Yes"),
                       years = 'all', 
                       geom = 'bar')

fig_print(p1)

```

B. Among those with hypertension who self-report taking antihypertensive medication 
```{r, fig.height=3.75}
fig_print(p2)
```

\newpage

**Figure 5** Blood pressure control by calendar year

A. Age-adjusted among those with hypertension
```{r, fig.height=3.75}

age_wts_htn <- c(26.4, 43.5, 17.0, 13.2)

p1 <- nhanes_visualize(outcome = 'bp_control_accaha',
                       subset_variables = list('htn_accaha'),
                       subset_values = list("Yes"),
                       age_wts = age_wts_htn,
                       years = 'all', 
                       geom = 'bar')

p2 <- nhanes_visualize(outcome = 'bp_control_accaha',
                       subset_variables = list('htn_accaha',
                                               'bp_med_use'),
                       subset_values = list("Yes", 
                                            "Yes"),
                       age_wts = age_wts_htn,
                       years = 'all', 
                       geom = 'bar')

fig_print(p1)

```

B. Age-adjusted among those who self-report taking antihypertensive medication 
```{r, fig.height=3.75}
fig_print(p2)
```

\newpage

**Figure S1** Application of inclusion and exclusion criteria for the application of NHANES hypertension data.

```{r}

data.frame(
 check.names = FALSE,
 ` ` = c(
  "Participants",
  "\u2265 18 years old",
  "Completed interview and examination",
  "Had SBP and DBP measurements",
  "Had self-reported information on antihypertensive medication",
  "Final Sample size"
 ),
 Overall = c(107622, 63041, 59799, 56286, 56035, 56035),
 `1999-2000` = c(9965, 5448, 4976, 4755, 4694,   4694),
 `2001-2002` = c(11039, 5993, 5592, 5251, 5184,  5184),
 `2003-2004` = c(10122, 5620, 5303, 4902, 4838,  4838),
 `2005-2006` = c(10348, 5563, 5334, 5028, 5015,  5015),
 `2007-2008` = c(10149, 6228, 5995, 5670, 5665,  5665),
 `2009-2010` = c(10537, 6527, 6360, 6053, 6043,  6043),
 `2011-2012` = c(9756, 5864, 5615, 5436, 5337,   5337),
 `2013-2014` = c(10175, 6113, 5924, 5700, 5694,  5694),
 `2015-2016` = c(9971, 5992, 5735, 5557, 5552,   5552),
 `2017-2020` = c(15560, 9693, 8965, 8024, 8013,  8013)
) %>% 
 flextable() %>% 
 theme_box() %>% 
 width(width = 0.8) %>% 
 width(j = 1, width = 1) %>% 
 align(align = 'center', part = 'all') %>% 
 align(j = 1, part = 'all')

```

\newpage

**Figure S2**:	Age-adjusted mean systolic blood pressure among people with and without hypertension

```{r}

nhanes_visualize(outcome = 'bp_sys_mean',
                 years = 'all',
                 exposure = 'htn_jnc7',
                 age_wts = age_wts_overall, 
                 geom = 'bar') %>% 
fig_print()

```

\newpage

**Figure S3**: Age-adjusted prevalence of hypertension among US adults with and without chronic kidney disease
\newpage

**Figure S4**: Prevalence of resistant hypertension by calendar year

A. Among those with hypertension who self report taking antihypertensive medication 
```{r, fig.height=3.75}

p1 <- nhanes_visualize(outcome = 'htn_resistant_accaha',
                       subset_variables = list('htn_accaha', 'bp_med_use'),
                       subset_values = list("Yes", "Yes"),
                       years = 'all', 
                       geom = 'bar')

p2 <- nhanes_visualize(outcome = 'htn_resistant_accaha',
                       subset_variables = list('htn_accaha', 
                                               'bp_med_use',
                                               'bp_med_n_class'),
                       subset_values = list("Yes", 
                                            "Yes",
                                            c('Three', 'Four or more')),
                       years = 'all', 
                       geom = 'bar')

fig_print(p1)

```

B. Among those with hypertension who self report and are taking 3+ classes of antihypertensive medication
```{r, fig.height=3.75}
fig_print(p2)
```
\newpage

**Figure S5** Prevalence of hypertension by age categories in 2017-2020

```{r}
nhanes_visualize(outcome = 'htn_accaha',
                 exposure = 'demo_age_cat',
                 years = 'most_recent',
                 geom = 'bar') %>% 
 fig_print()
```
\newpage

**Figure S6** Prevalence of hypertension by sex in 2017-2020

```{r}
nhanes_visualize(outcome = 'htn_accaha',
                 exposure = 'demo_gender',
                 years = 'most_recent',
                 geom = 'bar') %>% 
 fig_print()
```

\newpage

**Figure S7** Prevalence of hypertension by race/ethnicity in 2017-2020

```{r}
nhanes_visualize(outcome = 'htn_accaha',
                 exposure = 'demo_race',
                 years = 'most_recent',
                 geom = 'bar') %>% 
 fig_print()
```

\newpage

**Figure S8** Prevalence of uncontrolled blood pressure among pregnant women from 2011-2012 through 2017-2020.

```{r}

nhanes_visualize(outcome = 'bp_uncontrolled_accaha',
                 years = c("2011-2012", 
                           "2013-2014",
                           "2015-2016",
                           "2017-2020"),
                 pool = TRUE,
                 subset_variables = list('demo_pregnant'),
                 subset_values = list("Yes"),
                 geom = 'bar') %>% 
 fig_print()

```

```{r}
block_section(
  prop_section(
    page_size = page_size(orient = "landscape"),
    page_margins = page_mar(bottom = 1/8, top = 1/8, 
                            right = 1/4, left = 1/4),
    type = "continuous"
  )
)
```

### 2022-10-01 R/svy_design_new.R

@@ -17,13 +17,18 @@
 #' @return a `svydesign` object
 #'
 svy_design_new <- function(data,
-                           exposure,
+                           exposure = NULL,
                            n_exposure_group = NULL,
                            exposure_cut_type = NULL,
-                           years,
+                           years = NULL,
                            pool = 'no'){
 
  stopifnot(is.data.table(data))
+
+ if(is.null(years)) years <- levels(data$svy_year)
+ if(is.null(years)) years <- unique(data$svy_year)
+
+ if(is.null(exposure)) exposure <- "None"
 
  fctrs <- data %>%
   sapply(is.factor) %>%
@@ -33,6 +38,12 @@ svy_design_new <- function(data,
  divide_by <- length(years)
 
  if(pool == 'no') divide_by <- 1
+
+ # this happens if sub-population was not considered
+ # prior to making the design object
+ if(!("svy_weight" %in% names(data))){
+  data[, svy_weight := svy_weight_mec]
+ }
 
  data_design <- data[svy_year %in% years] %>%
   .[, svy_weight := svy_weight / divide_by]

### 2022-10-01 tests/testthat/test-reliability_prop.R

@@ -1,3 +1,16 @@
+
+
+ds <- nhanes_bp %>%
+ filter(svy_subpop_htn==1) %>%
+ nhanes_calibrate(nhanes_full = nhanes_bp, nhanes_sub = .) %>%
+ .[, svy_weight := svy_weight_cal] %>%
+ svy_design_new()
+
+smry <- svy_design_summarize(ds, outcome = 'htn_jnc7') %>%
+ reliability_prop(design = ds)
+
+
+
 # TODO
 # smry <- tribble(
 #  ~estimate,	~std_error, ~ci_lower, 	~ci_upper,	~n_eff_SAS,	~kg_wdth_SAS,	~kg_relw_p_SAS,	~kg_relw_q_SAS,	~p_reliable_SAS,	~q_reliable_SAS,	~p_staistical_SAS,

### 2022-10-07 README.Rmd
api improvements

@@
 
-## TODO list
+<!-- ## TODO list -->
 
-- Write vignettes for tests of the app
-- Write vignette for medication data
-- Replication with AB's group
-- Incorporate suppression of unreliable estimates
-- Documentation/demos for R functions used in the app
-- Definitions/notes for all BP variables (table S1 in manuscript)
-- Tutorial videos
-- Do we need to use AF weights?
+<!-- - Write vignettes for tests of the app -->
+<!-- - Write vignette for medication data -->
+<!-- - Replication with AB's group -->
+<!-- - Incorporate suppression of unreliable estimates -->
+<!-- - Documentation/demos for R functions used in the app -->
+<!-- - Definitions/notes for all BP variables (table S1 in manuscript) -->
+<!-- - Tutorial videos -->
+<!-- - Do we need to use AF weights? -->
+
+## Features of the application
+
+Using the web application, we estimated characteristics of the population which are relevant when investigating BP and hypertension data.
+As the application was designed with a number of options so that analyses can be customized to the user’s specifications.
+Results can be presented as figures or in tables.
+Figures can be made with the data represented as bars or points.
+The 10 NHANES cycles can be analyzed separately or multiple cycles can be pooled together, which provides a larger sample size leading to more stable statistical estimates.
+The analyses can be done crude or with age-adjustment using direct standardization in four age groups (18-44, 45-64, 65-74 and ≥75 years of age).
+Users can select the age distribution to be used as the standard population.
+Analyses can be done in sub-groups and stratified by participant characteristics.
+
+Full summaries of this application's features and associated tutorials are publicly available.
+Briefly, selected NHANES cycles can be analyzed separately, and contiguous cycles can be pooled together to increase reliability of statistical estimates.
+Users may select statistics to estimate, and based on the type of outcome they are analyzing.
+For example, a mean may be computed for a continuous variable, whereas a prevalence or an estimated count of US adults may be computed for a categorical variable.
+When a user requests to estimate a count, the application automatically calibrates survey weights to ensure the sum of the counts matches the estimated total number of US adults.
+Weights are recalibrated within age, sex, race/ethnicity groups and NHANES cycle.
+All statistical analyses on the application are weighted to provide representative estimates for the non-institutionalized US population.
+Analyses may incorporate age-adjustment using direct standardization, and users can select the age distribution to be used as the standard population.
+Analyses can be done in sub-groups and stratified by participant characteristics.
+Following CDC recommendations, the application automatically suppresses statistics which are unreliable, e.g., due to being based on a small sample size.
 

### 2022-10-07 paper/manuscript.Rmd
api improvements

@@
 
+conflicted::conflict_prefer('as_flextable', 'flextable')
+conflicted::conflict_prefer('filter', 'dplyr')
+conflicted::conflict_prefer('layout', 'plotly')
+
 # default is double space 
@@
 single_par <- fp_par(line_spacing=1)
+
+abbrvs <- list(
+  "CI"   = 'confidence interval',
+  "BP"   = "blood pressure",
+  "SBP"  = "systolic blood pressure",
+  "DBP"  = "diastolic blood pressure",
+  "SD"   = 'standard deviation',
+  "IQR"  = 'interquartile range',
+  "JNC7" = "Seventh Joint National Committee",
+  "ACC"  = "American College of Cardiology",
+  "AHA"  = "American Heart Association",
+  "CVD"  = 'cardiovascular disease'
+)
+
+abbrvs_paste <- function(strings){
+
+  front <- 'Abbreviations'
+
+  matter <- glue::glue_collapse(strings, sep = '; ', last = '; and ')
+
+  paste(front, matter, sep = ': ')
+
+}
+
+abbrvs_write <- function(abbr){
+
+  sorted <- abbr[sort(names(abbr))]
+
+  strings <- map2_chr(names(sorted), sorted, paste, sep = ' = ')
+
+  abbrvs_paste(strings)
+
+}
 
@@
 ^4^Department of Population Health Sciences, University of Utah School of Medicine, Salt Lake City, UT. `r single_par`
-^x^ YOUR AFFILIATION
+^x^ YOUR AFFILIATION IF IT ISN'T ALREADY DEFINED ABOVE
 
@@
 
+<!-- Abstract: Structured with the following headings and no more than 250 words: Background, Objectives, Methods, Results, Conclusions.
+The abstract should present essential data in 5 paragraphs.
+Use complete sentences.
+All data in the abstract also must appear in the manuscript text or tables.
+-->
+
+**Background**: Data from the National Health and Nutrition Examination Survey (NHANES) program are freely available. However, analysis of these data may be time-intensive and challenging.
+
+**Objectives**: Develop, validate, and share an open-source web application that allows for interactive, exploratory analysis of NHANES data that can be used and extended upon by others.
+
+**Methods**: We developed the open-source web application using the R programming language and several R packages.
+We validated the web application by using it to reproduce results from two prior studies and a report from the Centers for Disease Control.
+To demonstrate the application's capabilities, we estimated characteristics of adults in the US, the prevalence of hypertension, distribution of the number of antihypertensive medication classes being used, prevalence of controlled blood pressure (BP) among those with hypertension, and the prevalence of resistant hypertension.
+
+**Results**: The estimated mean systolic BP for all US adults was 123 mm Hg in 1999-2000, 120 mm Hg in 2009-2010, and 123 mm Hg in 2017-2020.
+The estimated prevalence of hypertension among US adults was lowest in 2009-2010 (43.3%) and highest in 2017-2020 (46.6%), with the estimated number of US adults with hypertension increasing from 89.8 million in 1999-2000 to 115.3 million in 2017-2020 .
+The age adjusted mean systolic BP was 140 in 1999-2000, 134 in 2009-2010, and 136 in 2017-2020 among US adults with hypertension.
+Among US adults who self-reported taking antihypertensive medication, the percentage taking four or more classes of antihypertensive medication increased from 3.5% in 1999-2000 to 7.6% in 2017-2020.
+The age-adjusted estimated prevalence of BP control among US adults with hypertension was 10.2%, 25.8%, and 23.8% in 1999-2000, 2009-2010, and 2017-2020, respectively.
+Among US adults with hypertension who self-reported taking antihypertensive medication, the prevalence of resistant hypertension was 13.7% in 1999-2000, 17.7% in 2009-2010, and 16.2% in 2017-2020.
+
+**Conclusions.**: The web application developed in the current analysis may increase the usage of NHANES data for research and policy decision making, ensuring results are valid, transparent, and reproducible.
+
 \newpage
@@
 The National Health and Nutrition Examination Survey (NHANES) is a program conducted by the US National Center for Health Statistics (NCHS) of the Centers for Disease Control and Prevention (CDC) and is designed to assess the health and nutritional status of the US population.
-The NHANES program was initiated in the early 1960s and beginning in 1999 has been conducted continuously, in two-year cycles.
-NHANES data are collected through participant interviews and physical examinations, each conducted following standardized protocols.
-The NHANES interview includes questions about prior diagnoses of hypertension and self-reported antihypertensive medication use and a medication inventory where containers for prescription medications are reviewed and recorded.
-The physical examination includes blood pressure (BP) measurements by trained and certified physicians.
-
-NHANES data have been used to provide estimates of the proportion of US adults with hypertension, aware they had hypertension, taking antihypertensive medication and with controlled BP.
-Data from NHANES have been used to estimate the impact of the lower BP levels used to define hypertension and controlled BP in the 2017 American College of Cardiology/American Heart Association (ACC/AHA) BP guideline versus the Seventh Joint National Committee (JNC7) BP guideline on the proportion and number of US adults with hypertension, recommended antihypertensive medication and with uncontrolled BP.
+NHANES data have been used to provide nationally representative estimates for adults living in the United States (US) in multiple studies of cardiometabolic outcomes.
+For example, NHANES data have been used to estimate the impact of the lower blood pressure (BP) levels used to define hypertension and controlled BP in the 2017 American College of Cardiology/American Heart Association (ACC/AHA) BP guideline versus the Seventh Joint National Committee (JNC7) BP guideline on the proportion and number of US adults with hypertension, recommended antihypertensive medication and with uncontrolled BP.
 Additionally, NHANES data have been used to track the proportion of US adults with controlled BP, which were included in a Call-to-Action to Control BP from the US Surgeon General.
 
-NHANES data are publically available and easily accessible through the CDC website.
-However, analyzing NHANES data may be challenging for many people due to the complexity of its study design.
-The NHANES program uses a multi-stage sampling design to select participants for enrollment and data analysis requires specific programming commands to provide valid statistical estimates.
+NHANES data are publicly available and accessible through the CDC website.
+However, analyzing NHANES data may be challenging for a number of reasons.
+In particular, specific statistical techniques are required to account for the multi-stage sampling design of NHANES data and produce nationally representative estimates.
 Also, each NHANES cycle includes over 50 data files, each of which has detailed informational material.
-NHANES data collection protocols have changed over time for certain variables, and these differences should be accounted for when analyzing data.
-To address the barriers to analyzing NHANES data, we developed an open-access web-based application that allows users to obtain BP and hypertension statistics for US adults using NHANES data.
-The goal of the application is to increase the usage of NHANES data for BP and hypertension research and policy decision making with a focus on ensuring results are valid and transparent.
-In this manuscript, we review the design of the application, show some results of NHANES analyses using the application, and provide a tutorial on how people can use the application.
+NHANES data collection protocols have changed over time for certain variables, and these differences should be accounted for when harmonizing or pooling data from multiple cycles.
+Another potential challenge is implementation of recommended standards from the CDC, such as their guidelines for suppression of unreliable results.
+
+To address the challenges of analyzing NHANES data, we developed an open-source web-based application that allows users to obtain nationally representative summaries of cardiometabolic outcomes for US adults using NHANES data.
+The goal of the application is to increase the usage of NHANES data for research and policy decision making with a focus on ensuring results are valid, transparent, and reproducible.
+The application is designed to be extended with the addition of "modules", which we define as a set of variables measured among a sub-population of NHANES participants.
+In this manuscript, we review the design and development of the application, present nationally representative summaries of outcomes from the "BP and Hypertension module", and describe online tutorials and resources that may help users engage with the application.
 
@@
 
-The open-access web-based application uses data from 10 NHANES cycles, specifically NHANES 1999-2000 through NHANES 2017- March 2020, prior to the COVID-19 pandemic.
-In each NHANES cycle, potential participants were identified using a multi-stage sampling process.
-The protocols for each NHANES cycle were approved by the NCHS Institutional Review Board.
+The NHANES program was initiated in the early 1960s and beginning in 1999 has been conducted continuously, in two-year cycles.
+In each cycle, potential participants were identified using a multi-stage sampling process.
+The protocols for each cycle were approved by the NCHS Institutional Review Board.
 Written informed consent was obtained from each participant.
 
-NHANES data were collected through in-home interviews and study examinations conducted at mobile examination centers.
-Data elements included in the application, along with their definitions, are listed in Supplemental Table 1.
-Information on age, sex, race/ethnicity, pregnancy status for women, cigarette smoking, a prior diagnosis of diabetes, use of insulin or oral glucose-lowering medication, a prior diagnosis of hypertension, self-reported antihypertensive medication use, and history of coronary heart disease, myocardial infarction, stroke, and heart failure were obtained by self-report during the in-home interview.
-During the in-home interview, the labels of medications that participants reported taking in the preceding 30 days were recorded.
-For the application, we categorized antihypertensive medications by class as defined in the 2017 ACC/AHA BP guideline.
-Data on medication dosages are not available in NHANES.
+NHANES data include in-home interviews and study examinations conducted at mobile examination centers.
+The interview includes questions about demographics, health behaviors, prior diagnoses, medication use, and medical history.
+During the interview, the labels of medications that participants reported taking in the preceding 30 days were recorded.
+During the study examination, height and weight were measured, blood samples were used to measure cholesterol, glycated hemoglobin and serum creatinine, and spot urine was used to measure albumin and creatinine and to conduct a pregnancy test.
+Systolic and diastolic BP (SBP and DBP, respectively) were measured three times.
+BP was measured using a mercury sphygmomanometer from 1999-2000 through 2015-2016 and using an oscillometric device in 2017-2020.
 
-During the study examination, height and weight were measured and used to calculate body mass index.
-Systolic and diastolic BP (SBP and DBP, respectively) were measured three times, and blood and spot urine samples were collected.
-BP was measured using a mercury sphygmomanometer from 1999-2000 through 2015-2016 and using an oscillometric device in 2017-2020.
-The oscillometric BP values were calibrated to the mercury device as described previously.
-Blood was used to measure total and high-density lipoprotein (HDL) cholesterol, glycated hemoglobin and serum creatinine.
-Spot urine was used to measure albumin and creatinine and to conduct a pregnancy test.
+*Features of the Open-Source Web-Based Application*
 
-Overall, 107,622 US individuals participated in NHANES 1999-2000 to 2017- March 2020.
-As statistics for hypertension and SBP and DBP levels in children and adolescents are markedly different than for adults, we restricted the data in the application to adults ≥ 18 years of age.
-We further restricted the application to participants who completed the in-home interview and study examination, with one or more SBP and DBP measurement and who had data on self-reported antihypertensive medication use.
-After these exclusions were applied, the application included data from 56,035 participants (Figure S1).
+Full summaries of this application's features and associated tutorials are available.
+Briefly, users can select NHANES cycles to be analyzed.
+Contiguous cycles can be combined to increase precision.
+Estimates are weighted to represent the non-institutionalized US population and may incorporate age-adjustment using direct standardization.
+Count estimates are calibrated to match the estimated total number of US adults.
+Users may visualize summaries of outcomes in groups (i.e., prevalence of hypertension by sex) with all groups in the same figure or with one figure per group, and may additionally specify subsets to exclude from their analyses.
+All outputs created with the web application can be downloaded and saved.
+Following CDC recommendations, unreliable statistical estimates are automatically suppressed.
+
+*Design of the Open-Source Web-Based Application*
+
+The web application was created using Shiny, an open-source software package that translates code from the R programming language into HTML, CSS, or JavaScript commands that create a website interface.
+The "cardioStatsUSA" R package provides additional details on the web application's design and comprehensive documentation of its components.
+
+*The Blood Pressure and Hypertension Module*
+
+We define a module as a subpopulation of NHANES participants paired with a collection of variables.
+Each variable in the collection has designated roles.
+For example, hypertension can be an outcome and can also be used to define groups when analyzing another outcome.
+This module definition is general enough to allow for diverse extensions of the web application, but may not be specific enough to illustrate how the web application works.
+Therefore, in the current analysis, we define and present results from the "BP and hypertension" module, defined as follows.
+
+**Subpopulation:** Beginning with 107,622 US individuals who participated in NHANES 1999-2000 to 2017- March 2020, we restricted the subpopulation to adults ≥ 18 years of age.
+This exclusion was applied because statistics for hypertension and BP levels in children and adolescents are markedly different than for adults.
+We further restricted the subpopulation to participants who completed the in-home interview and study examination, with one or more SBP and DBP measurement, and who had data on self-reported antihypertensive medication use.
+After these exclusions were applied, the sub-population included data from 56,035 participants (Figure S1).
+
+**Variables** The primary variables from the BP and hypertension module are listed in Table 1, with full definitions provided in Supplemental Table 1.
+Briefly, mean SBP and DBP levels were computed over all available measurements for each participant.
+Oscillometric BP values were calibrated to the mercury device as described previously.
+Antihypertensive medication classes were defined using recommendations from the 2017 ACC/AHA BP guideline.
+
+*Validation of the web application*
+
+We validated the web application by using it in combination with our BP and hypertension module to reproduce results from two prior studies and one CDC report.
+A full description of the validation procedure is available.
 
@@
 
-The application was created using Shiny, an open-source software package for creating web applications using the R programming language.
-Shiny translates R programming code into HTML, CSS, or JavaScript commands and creates a website interface.
-The statistical analyses that are used on the application are conducted with R programming language packages that account for the non-random selection of NHANES participants.
-All statistical analyses on the application are weighted to provide estimates for the non-institutionalized US population.
-Following CDC recommendations, the application automatically suppresses statistics which are unreliable due to being based on a small sample size.
-
-The application can be used to estimate characteristics of the population which are relevant when investigating BP and hypertension data (Table 1, top panel).
-The application also allows users to estimate statistics for a number of hypertension-related variables (Table 1, bottom panel).
-
-The application was designed with a number of options so that analyses can be customized to the user’s specifications.
-Results can be presented as figures or in tables.
-Figures can be made with the data represented as bars or points.
-The 10 NHANES cycles can be analyzed separately or multiple cycles can be pooled together, which provides a larger sample size leading to more stable statistical estimates.
-The analyses can be done crude or with age-adjustment using direct standardization in four age groups (18-44, 45-64, 65-74 and ≥75 years of age).
-Users can select the age distribution to be used as the standard population.
-Analyses can be done in sub-groups and stratified by participant characteristics.
-
-To demonstrate the features and utility of the application, we conducted a series of statistical analyses.
-We estimated the crude and age-adjusted mean SBP and prevalence of hypertension for US adults by NHANES cycle from 1999-2000 through 2017-2020.
-We estimated the age-adjusted prevalence of hypertension by NHANES cycle for individuals with and without chronic kidney disease, separately.
+We conducted a series of statistical analyses using the BP and hypertension module defined above.
+We estimated characteristics of the population, crude and age-adjusted mean SBP, and prevalence of hypertension for US adults by NHANES cycle.
+To demonstrate computation of results by group, we computed the age-adjusted prevalence of hypertension by NHANES cycle for individuals with versus without chronic kidney disease.
 For age adjustment in these analyses, we used the estimated age distribution of US adults from 1999 to 2020 as the standard (49.3%, 33.6%, 10.1% and 7.0% being 18 to 44, 45 to 64, 65 to 74 and ≥75 years of age, respectively).
-Also, we estimated the absolute number of US adults with hypertension from 1999-2000 through 2017-2020.
-For the estimation of absolute counts of US adults, the application recalibrates the statistical weights to account for participants being excluded due to missing information on SBP, DBP or self-reported antihypertensive medication use.
-Weights are recalibrated within age, sex, race/ethnicity groups and NHANES cycle.
+We also estimated the number of US adults with hypertension by NHANES cycle from 1999-2000 through 2017-2020.
 
 The age-adjusted distribution of the number of antihypertensive medication classes being taken was estimated for US adults with hypertension in each NHANES cycle, overall and among those who self-reported taking antihypertensive medication and had at least one antihypertensive drug class identified in their medication inventory.
-The age-adjusted proportion of US adults with controlled BP, defined using the 2017 ACC/AHA BP guideline, was estimated by NHANES cycle among US adults with hypertension and who self-reported taking antihypertensive medication.
+The age-adjusted proportion of US adults with controlled BP (i.e., SBP < 130 mm Hg and DBP < 80 mm Hg) was estimated by NHANES cycle among US adults with hypertension and who self-reported taking antihypertensive medication.
 The age-adjusted prevalence of resistant hypertension was estimated by NHANES cycle for US adults with hypertension taking antihypertensive medication and among those with hypertension taking ≥ 3 classes of antihypertensive medication.
-For the analyses of number of antihypertensive medication classes being taken, resistant hypertension and BP control, we set the age distribution for the standard population to represent US adults with hypertension from 1999 to 2020: 26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥75 years of age, respectively.
-Using the stratification option in the application, we estimated the prevalence of hypertension in NHANES 2017-2020 by age group, sex, and race-ethnicity.
-Using the data pooling option in the application, we grouped the NHANES 2011-2020 cycles to estimate the proportion of pregnant women who had uncontrolled BP according to the 2017 ACC/AHA BP guideline, SBP ≥ 130 mm Hg or DBP ≥ 80 mm Hg.
+For age adjustment in these analyses, we set the age distribution for the standard population to represent US adults with hypertension from 1999 to 2020: 26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥75 years of age, respectively.
+We estimated the prevalence of hypertension in NHANES 2017-2020 by age group, sex, and race-ethnicity.
+Using the data pooling option in the application, we grouped the NHANES 2011-2020 cycles to estimate the proportion of pregnant women who had uncontrolled BP (i.e., SBP ≥ 130 mm Hg or DBP ≥ 80 mm Hg).
 
@@
 
+**Table 1**: Variables included in the blood pressure and hypertension module.
+
+```{r}
+
+bp_med_classes <- c(
+ "bp_med_ace",
+ "bp_med_aldo",
+ "bp_med_alpha",
+ "bp_med_angioten",
+ "bp_med_beta",
+ "bp_med_central",
+ "bp_med_ccb",
+ "bp_med_diur_Ksparing",
+ "bp_med_diur_loop",
+ "bp_med_diur_thz",
+ "bp_med_renin_inhibitors",
+ "bp_med_vasod"
+)
+
+classes <- c("Blood pressure",
+             "Hypertension", 
+             "Antihypertensive medication")
+
+ft_data <- nhanes_key %>% 
+ .[module == 'htn'] %>% 
+ .[class %in% classes] %>% 
+ .[, class := factor(class, levels = classes)] %>% 
+ .[order(class)] %>% 
+ .[variable %in% bp_med_classes,
+   class := "Antihypertensive medication classes"] %>% 
+ select(class, variable, label) %>% 
+ as_grouped_data(groups = 'class')
+
+indent_meds_spanner <- 
+ which(ft_data$class == "Antihypertensive medication classes")
+indent_meds <- 
+ which(ft_data$variable %in% bp_med_classes)
+
+ft <- ft_data %>% 
+ as_flextable(hide_grouplabel = TRUE,
+              col_keys = c('class', 'label')) %>% 
+ set_header_labels(label = "Variables included in module") %>% 
+ theme_box() %>% 
+ bg(i = ~ !is.na(class), bg = 'grey80') %>%
+ bg(i = indent_meds_spanner, bg = 'white') %>%
+ italic(i = ~ !is.na(class)) %>%
+ padding(i = ~ is.na(class), padding.left = 10) %>% 
+ padding(i = indent_meds_spanner, padding.left = 10) %>% 
+ padding(i = indent_meds, padding.left = 20) %>% 
+ width(j = 1, width = 6) %>% 
+ add_footer_lines(
+  values = abbrvs_write(
+   abbrvs[c('BP', 'SBP', 'DBP', 'ACC', 'AHA', 'JNC7')]
+  )
+ )
+
+ft
+
+```
+
 ```{r}
@@
 
-**Table 1**:	Race/ethnicity distribution of the population with and without hypertension in 2017-2020
+**Table 2**:	Race/ethnicity distribution of the population with and without hypertension in 2017-2020
 
@@
 
-nhanes_summarize(outcome = 'demo_race', 
-                 years = 'most_recent', 
-                 group = 'htn_accaha') %>%
+nhanes_data %>% 
+ nhanes_summarize(key = nhanes_key,
+                  outcome_variable = 'demo_race',
+                  group_variable = 'htn_accaha',
+                  time_values = 'most_recent') %>%
  mutate(across(where(is.numeric), table_value)) %>% 
  filter(statistic == 'percentage') %>% 
+ select(-statistic, 
+        -ends_with("reason"),
+        -unreliable_status,
+        -review_needed) %>% 
  flextable() %>% 
@@
 
-p1 <- nhanes_visualize(outcome = 'bp_sys_mean',
-                       years = 'all', 
+p1 <- nhanes_visualize(data = nhanes_data,
+                       key = nhanes_key,
+                       outcome_variable = 'bp_sys_mean',
+                       time_values = 'all', 
                        geom = 'bar')
@@
 
-p2 <- nhanes_visualize(outcome = 'bp_sys_mean',
-                       years = 'all', 
+p2 <- nhanes_visualize(data = nhanes_data,
+                       key = nhanes_key,
+                       outcome_variable = 'bp_sys_mean',
+                       time_values = 'all', 
                        age_wts = age_wts_overall, 
@@
 
-p1 <- nhanes_visualize(outcome = 'htn_accaha',
-                       years = 'all', 
+p1 <- nhanes_visualize(data = nhanes_data,
+                       key = nhanes_key,
+                       outcome_variable = 'htn_accaha',
+                       time_values = 'all', 
                        geom = 'bar')
 
-p2 <- nhanes_visualize(outcome = 'htn_accaha',
-                       years = 'all', 
+p2 <- nhanes_visualize(data = nhanes_data,
+                       key = nhanes_key,
+                       outcome_variable = 'htn_accaha',
+                       time_values = 'all', 
                        age_wts = age_wts_overall, 
@@
 
-nhanes_visualize(outcome = 'htn_accaha',
-                 years = 'all',
-                 exposure = 'cc_ckd',
+nhanes_visualize(data = nhanes_data,
+                 key = nhanes_key,
+                 outcome_variable = 'htn_accaha',
+                 time_values = 'all',
+                 group_variable = 'cc_ckd',
                  age_wts = age_wts_overall, 
-                 geom = 'point') %>% 
+                 geom = 'point',
+                 size_point = 22,
+                 size_error = 4) %>% 
  fig_print()
@@
 
-nhanes_visualize(outcome = 'htn_accaha',
-                 years = 'all',
+nhanes_visualize(data = nhanes_data,
+                 key = nhanes_key,
+                 outcome_variable = 'htn_accaha',
+                 time_values = 'all',
                  statistic_primary = 'count',
@@
 
-p1 <- nhanes_visualize(outcome = 'bp_med_n_class',
+p1 <- nhanes_visualize(data = nhanes_data,
+                       key = nhanes_key,
+                       outcome_variable = 'bp_med_n_class',
                        subset_variables = list('htn_accaha'),
                        subset_values = list("Yes"),
-                       years = 'all', 
+                       time_values = 'all', 
                        geom = 'bar')
 
-p2 <- nhanes_visualize(outcome = 'bp_med_n_class',
+p2 <- nhanes_visualize(data = nhanes_data,
+                       key = nhanes_key,
+                       outcome_variable = 'bp_med_n_class',
                        subset_variables = list('htn_accaha', 'bp_med_use'),
                        subset_values = list("Yes", "Yes"),
-                       years = 'all', 
+                       time_values = 'all', 
                        geom = 'bar')
@@
 
-p1 <- nhanes_visualize(outcome = 'bp_control_accaha',
+p1 <- nhanes_visualize(data = nhanes_data,
+                       key = nhanes_key,
+                       outcome_variable = 'bp_control_accaha',
                        subset_variables = list('htn_accaha'),
@@
                        age_wts = age_wts_htn,
-                       years = 'all', 
+                       time_values = 'all', 
                        geom = 'bar')
 
-p2 <- nhanes_visualize(outcome = 'bp_control_accaha',
+p2 <- nhanes_visualize(data = nhanes_data,
+                       key = nhanes_key,
+                       outcome_variable = 'bp_control_accaha',
                        subset_variables = list('htn_accaha',
@@
                        age_wts = age_wts_htn,
-                       years = 'all', 
+                       time_values = 'all', 
                        geom = 'bar')
@@
 fig_print(p2)
+```
+
+\newpage
+
+**Table S1**
+
+```{r}
+
+classes <-
+ c("Survey",
+   "Demographics",
+   "Blood pressure",
+   "Hypertension",
+   "Antihypertensive medication",
+   "Comorbidities")
+
+ft_data <- nhanes_key %>% 
+ .[class %in% classes] %>%
+ .[variable != 'svy_subpop_chol'] %>% 
+ .[, class := factor(class, levels = classes)] %>% 
+ .[order(class)] %>% 
+ .[variable %in% bp_med_classes,
+   class := "Antihypertensive medication classes"] %>% 
+ select(class, variable, label, description) %>% 
+ as_grouped_data(groups = 'class')
+
+indent_meds_spanner <- 
+ which(ft_data$class == "Antihypertensive medication classes")
+
+indent_meds <- 
+ which(ft_data$variable %in% bp_med_classes)
+
+ft <- ft_data %>% 
+ as_flextable(hide_grouplabel = TRUE,
+              col_keys = c('class', 'label', 'description')) %>% 
+ set_header_labels(label = "All variables included in module",
+                   description = "Variable definition") %>% 
+ theme_box() %>% 
+ bg(i = ~ !is.na(class), bg = 'grey80') %>%
+ bg(i = indent_meds_spanner, bg = 'white') %>%
+ italic(i = ~ !is.na(class)) %>%
+ padding(i = ~ is.na(class), j = 1, padding.left = 10) %>% 
+ padding(i = indent_meds_spanner, j = 1, padding.left = 10) %>% 
+ padding(i = indent_meds, j = 1, padding.left = 20) %>% 
+ width(j = 1, width = 2) %>%
+ width(j = 2, width = 7) %>% 
+ add_footer_lines(
+  values = abbrvs_write(
+   abbrvs[c('BP', 'SBP', 'DBP', 'ACC', 'AHA', 'JNC7')]
+  )
+ )
+
+ft
+
 ```
@@
 
-nhanes_visualize(outcome = 'bp_sys_mean',
-                 years = 'all',
-                 exposure = 'htn_jnc7',
+nhanes_visualize(data = nhanes_data,
+                 key = nhanes_key,
+                 outcome_variable = 'bp_sys_mean',
+                 time_values = 'all',
+                 group_variable = 'htn_jnc7',
                  age_wts = age_wts_overall, 
@@
 
-p1 <- nhanes_visualize(outcome = 'htn_resistant_accaha',
+p1 <- nhanes_visualize(data = nhanes_data,
+                       key = nhanes_key,
+                       outcome_variable = 'htn_resistant_accaha',
                        subset_variables = list('htn_accaha', 'bp_med_use'),
                        subset_values = list("Yes", "Yes"),
-                       years = 'all', 
+                       time_values = 'all', 
                        geom = 'bar')
 
-p2 <- nhanes_visualize(outcome = 'htn_resistant_accaha',
+p2 <- nhanes_visualize(data = nhanes_data,
+                       key = nhanes_key,
+                       outcome_variable = 'htn_resistant_accaha',
                        subset_variables = list('htn_accaha', 
@@
                                             c('Three', 'Four or more')),
-                       years = 'all', 
+                       time_values = 'all', 
                        geom = 'bar')
@@
 ```{r}
-nhanes_visualize(outcome = 'htn_accaha',
-                 exposure = 'demo_age_cat',
-                 years = 'most_recent',
+nhanes_visualize(data = nhanes_data,
+                 key = nhanes_key,
+                 outcome_variable = 'htn_accaha',
+                 group_variable =  'demo_age_cat',
+                 time_values = 'most_recent',
                  geom = 'bar') %>% 
@@
 ```{r}
-nhanes_visualize(outcome = 'htn_accaha',
-                 exposure = 'demo_gender',
-                 years = 'most_recent',
+nhanes_visualize(data = nhanes_data,
+                 key = nhanes_key,
+                 outcome_variable = 'htn_accaha',
+                 group_variable = 'demo_gender',
+                 time_values = 'most_recent',
                  geom = 'bar') %>% 
@@
 ```{r}
-nhanes_visualize(outcome = 'htn_accaha',
-                 exposure = 'demo_race',
-                 years = 'most_recent',
+nhanes_visualize(data = nhanes_data,
+                 key = nhanes_key,
+                 outcome_variable = 'htn_accaha',
+                 group_variable = 'demo_race',
+                 time_values = 'most_recent',
                  geom = 'bar') %>% 
@@
 
-nhanes_visualize(outcome = 'bp_uncontrolled_accaha',
-                 years = c("2011-2012", 
+nhanes_visualize(data = nhanes_data,
+                 key = nhanes_key,
+                 outcome_variable = 'bp_uncontrolled_accaha',
+                 time_values = c("2011-2012", 
                                  "2013-2014",

### 2022-10-07 R/app_run.R

@@ -9,9 +9,40 @@ library(shiny)
 #' @export
 #'
 
-# TODO: fix bug with pooled results: 2017-2020 through 2017-2020
+app_run <- function(nhanes_data = cardioStatsUSA::nhanes_data,
+                    nhanes_key = cardioStatsUSA::nhanes_key) {
 
-app_run <- function(...) {
+ time_variable <- nhanes_key$variable[nhanes_key$type == 'time']
+ time_values <- levels(nhanes_data[[time_variable]])
+
+ variable_choices <- list(
+  outcome  = nhanes_key[outcome  == TRUE, .(class, label, variable)],
+  group    = nhanes_key[group    == TRUE, .(class, label, variable)],
+  subset   = nhanes_key[subset   == TRUE, .(class, label, variable)],
+  stratify = nhanes_key[stratify == TRUE, .(class, label, variable)]
+ ) %>%
+  purrr::map(
+  .f = ~ .x %>%
+   split(by = 'class', keep.by = FALSE) %>%
+   purrr::map(deframe)
+ )
+
+ ctns_variables <- nhanes_key[type == 'ctns', variable]
+
+ boundaries <- ctns_variables %>%
+  purrr::set_names(ctns_variables) %>%
+  purrr::map(
+   ~ nhanes_data[[.x]] %>%
+    range(na.rm = TRUE) %>%
+    as.list() %>%
+    purrr::set_names(nm = c('min', 'max'))
+  )
+
+ ctns_group_variables <- ctns_variables %>%
+  intersect(unlist(variable_choices$group))
+
+ ctns_subset_variables <- ctns_variables %>%
+  intersect(unlist(variable_choices$subset))
 
  # User interface (UI) -----------------------------------------------------
 
@@ -23,12 +54,8 @@ app_run <- function(...) {
    titlePanel("Cardiometabolic statistics for US adults"),
    data.step = 1,
    data.intro = paste(
-    "This application analyzes blood pressure from",
-    "1999 through 2020 using data from the National Health and",
-    "Nutrition Examination Survey (NHANES). It includes survey",
-    "participants who attended their NHANES interview and examination,",
-    "who provided information on antihypertensive medication use, and",
-    "who had complete data on systolic and diastolic blood pressure."
+    "This application analyzes data from the",
+    "National Health and Nutrition Examination Survey (NHANES)."
    )
   ),
 
@@ -218,7 +245,7 @@ app_run <- function(...) {
[…]

### 2022-10-07 R/cardioStatsUSA-package.R

@@ -15,7 +15,7 @@
 #' @importFrom rlang parse_expr %||%
 #' @importFrom plotly plot_ly add_trace layout
 #' @importFrom arules discretize
-#' @importFrom purrr map map2 map_dfr map2_chr is_empty discard
+#' @importFrom purrr map map2 map_dfr map2_chr map2_dfr is_empty discard
 #' @importFrom dplyr filter mutate select pull across
 #' @importFrom dplyr if_else all_of .data recode left_join
 #' @importFrom tidyr pivot_wider pivot_longer unite

### 2022-10-07 R/get_outcome_stats.R

@@ -1,0 +1,12 @@
+
+get_outcome_stats <- function(outcome_type) {
+
+ switch(
+  outcome_type,
+  'ctns' = c(Mean = 'mean', Quantiles = 'quantile'),
+  'catg' = c(Percentage = 'percentage', Count = 'count'),
+  'bnry' = c(Percentage = 'percentage', Count = 'count'),
+  'intg' = c(Percentage = 'percentage', Count = 'count')
+ )
+
+}

### 2022-10-07 R/is_discrete.R

@@ -1,25 +1,0 @@
-
-
-is_var_type <- function(var_name, var_types){
- if(is.null(var_name)) return(FALSE)
- cardioStatsUSA::nhanes_key$variables[[var_name]]$type %in% var_types
-}
-
-is_binary <- function(variable_name) {
-
- is_var_type(variable_name, 'bnry')
-
-}
-
-is_discrete <- function(variable_name){
-
- is_var_type(variable_name, c('bnry', 'catg', 'intg'))
-
-}
-
-
-is_continuous <- function(variable_name) {
-
- is_var_type(variable_name, 'ctns')
-
-}

### 2022-10-07 R/jsc_write_subset_ctns.R

@@ -1,0 +1,15 @@
+
+# this is used to identify when a continuous variable
+# has been selected. It's tricky b/c you have to write
+# the logic in JavaScript (hence the pre-fix jsc)
+jsc_write_subset_ctns <- function(input_id, ctns_variables){
+
+ x <- paste('input', input_id, sep = '.')
+
+ ctns_variables_quoted <- paste0("\'",ctns_variables, "\'")
+
+ paste(x, ctns_variables_quoted, sep = ' == ', collapse = ' | ')
+
+ # glue("{x}.length > 0 & ({var_is_ctns})")
+
+}

### 2022-10-07 R/nhanes_design.R

@@ -1,0 +1,333 @@
+
+#' Title
+#'
+#' @param data
+#' @param key
+#' @param outcome
+#' @param group
+#' @param group_cut_n
+#' @param group_cut_type
+#' @param stratify
+#' @param time_values
+#' @param pool
+#' @param run_checks
+#'
+#' @return
+#' @export
+#'
+#' @examples
+#'
+#' x <- nhanes_design(data = nhanes_data,
+#'                    key = nhanes_key,
+#'                    outcome_variable = 'bp_sys_mean')
+#'
+nhanes_design <- function(data,
+                          key,
+                          outcome_variable,
+                          outcome_quantiles = NULL,
+                          group_variable = NULL,
+                          group_cut_n = NULL,
+                          group_cut_type = NULL,
+                          stratify_variable = NULL,
+                          time_variable = 'svy_year',
+                          time_values = NULL,
+                          pool = FALSE,
+                          run_checks = TRUE){
+
+ info_cols <- c('class',
+                'variable',
+                'label',
+                'source',
+                'type',
+                'module',
+                'description')
+
+ dt_data <- as.data.table(data)
+ dt_key <- as.data.table(key)
+
+ # outcome variable ----
+
+ outcome_info <- as.list(dt_key[variable == outcome_variable])[info_cols]
+
+ if(outcome_info$type == 'ctns'){
+  if(is.null(outcome_quantiles)) outcome_quantiles <- c(0.25, 0.5, 0.75)
+  outcome_info$quantiles <- outcome_quantiles
+ }
+
+ # statistics to compute
+ outcome_stats <- get_outcome_stats(outcome_info$type)
+
+ if(is.null(time_values)) time_values <- 'all'
+
+ # time_values ----
+ time_values <- switch(
+  time_values[1],
+  'all' = levels(data[[time_variable]]),
+  'last_5' = last(levels(data[[time_variable]]), n = 5),
+  'most_recent' = last(levels(data[[time_variable]]), n = 1),
+  time_values
+ )
+
+ time_info <- list(variable = time_variable,
+                   values = time_values,
+                   label = key[variable==time_variable, label])
+
+
+ # subset and pool the weights (if needed) ----
+
[…]

### 2022-10-07 R/nhanes_design_standardize.R

@@ -1,0 +1,45 @@
+
+
+#' Title
+#'
+#' @param x
+#' @param standard_variable
+#' @param standard_weights
+#'
+#' @return
+#' @export
+#'
+#' @examples
+nhanes_design_standardize <- function(x,
+                                      standard_variable = 'demo_age_cat',
+                                      standard_weights = c(0.155,
+                                                       0.454,
+                                                       0.215,
+                                                       0.177)){
+
+ by_vars <- c(x$time$variable,
+              x$group$variable,
+              x$stratify$variable)
+
+ over <- by_vars %>%
+  setdiff(standard_variable) %>%
+  as_svy_formula()
+
+ exclude_miss <- c(x$outcome$variable, by_vars) %>%
+  setdiff(standard_variable) %>%
+  as_svy_formula()
+
+ x$design <- survey::svystandardize(by = as_svy_formula(standard_variable),
+                                    over = over,
+                                    design = x$design,
+                                    population = standard_weights,
+                                    excluding.missing = exclude_miss)
+
+ x$standard <- list(variable = standard_variable,
+                    weights = standard_weights)
+
+ x$operations <- c(x$operations, 'standardize')
+
+ x
+
+}

### 2022-10-07 R/nhanes_design_summarize.R

@@ -1,0 +1,93 @@
+
+nhanes_design_summarize <- function(x,
+                                    stats = x$stats,
+                                    simplify_output = FALSE){
+
+ use_statby <- !is.null(x$by_variables)
+
+ smry_output <- purrr::map_dfr(
+
+  .x = x$stats[x$stats %in% stats],
+
+  .f = ~ {
+
+   stat_string <- if(use_statby) 'statby' else 'stat'
+
+   # identify the function to call
+   # if you want to use svy_stat_proportion instead of svy_stat_percentage,
+   # uncomment the line below
+   # if(.x == 'percentage') .x <- 'proportion'
+
+   svy_stat_fun <- paste("svy", stat_string, .x,  sep = '_')
+   svy_suppress_fun <- paste('svy_stat_suppress', x$outcome$type, sep = '_')
+
+   .out <- do.call(what = svy_stat_fun,
+                   args = list(outcome = x$outcome$variable,
+                               by_vars = x$by_variables,
+                               design = x$design,
+                               quantiles = x$outcome$quantiles)) %>%
+    svy_stat_tidy(outcome = x$outcome$variable,
+                  by_vars = x$by_variables) %>%
+    svy_stat_suppress(type = x$outcome$type,
+                      stat = .x,
+                      n_obs = get_obs_count(x$design),
+                      n_psu = get_psu_count(x$design),
+                      n_strata = get_strata_count(x$design))
+
+    .out[, outcome := NULL]
+
+
+   if(outcome_is_discrete(x) && 'level' %in% names(.out)){
+
+    setnames(.out, old = 'level', new = x$outcome$variable)
+
+    lvls <- levels(x$design$variables[[x$outcome$variable]])
+    .out[[x$outcome$variable]] %<>% factor(levels = lvls)
+
+   }
+
+   if(has_group(x)){
+
+    lvls <- levels(x$design$variables[[x$group$variable]])
+
+    .out[[x$group$variable]] %<>% factor(levels = lvls)
+
+   }
+
+   if(has_stratify(x)){
+
+    lvls <- levels(x$design$variables[[x$stratify$variable]])
+
+    .out[[x$stratify$variable]] %<>% factor(levels = lvls)
+
+   }
+
+
+   neworder <- c(x$time$variable,
+                 x$group$variable,
+                 x$stratify$variable,
+                 x$outcome$variable,
+                 'statistic',
+                 'estimate',
[…]

### 2022-10-07 R/nhanes_key.R

@@ -1,6 +1,6 @@
 #' One-stop shop for variable names, roles, labels, and functions
 #'
-#' A list with various elements pertaining the `nhanes_bp`
+#' A list with various elements pertaining the `nhanes_data`
 #'
 #' @format A data frame with 59799 rows and 35 variables:
 #' \describe{

### 2022-10-07 R/nhanes_visualize.R

@@ -17,10 +17,16 @@
 #'
 #' @export
 #'
-nhanes_visualize <- function(outcome,
-                             exposure = NULL,
-                             group = NULL,
-                             years = 'last_5',
+nhanes_visualize <- function(data,
+                             key,
+                             outcome_variable,
+                             outcome_quantiles = NULL,
+                             group_variable = NULL,
+                             group_cut_n = NULL,
+                             group_cut_type = NULL,
+                             stratify_variable = NULL,
+                             time_variable = 'svy_year',
+                             time_values = NULL,
                              pool = FALSE,
                              subset_variables = list(),
                              subset_values = list(),
@@ -29,45 +35,38 @@ nhanes_visualize <- function(outcome,
                              geom = 'bar',
                              reorder_cats = FALSE,
                              width = NULL,
-                             height = NULL){
+                             height = NULL,
+                             size_point = NULL,
+                             size_error = NULL){
 
- nhanes <- cardioStatsUSA::nhanes_bp
+ if(geom %in% c('point', 'points')) geom <- 'scatter'
 
- years <- switch(
-  years[1],
-  'all' = levels(nhanes$svy_year),
-  'last_5' = last(levels(nhanes$svy_year), n = 5),
-  'most_recent' = last(levels(nhanes$svy_year), n = 1),
-  years
- )
+ smry <- nhanes_summarize(data = data,
+                          key = key,
+                          outcome_variable = outcome_variable,
+                          outcome_quantiles = outcome_quantiles,
+                          group_variable = group_variable,
+                          group_cut_n = group_cut_n,
+                          group_cut_type = group_cut_type,
+                          stratify_variable = stratify_variable,
+                          time_variable = time_variable,
+                          time_values = time_values,
+                          pool = pool,
+                          subset_variables = subset_variables,
+                          subset_values = subset_values,
+                          age_wts = age_wts,
+                          simplify_output = FALSE)
 
- # silly conversion to make pool consistent w/the svy functions
- if(is.logical(pool)) pool <- ifelse(pool, yes = 'yes', no = 'no')
 
- smry <- nhanes_summarize(outcome, exposure, group, years, pool,
-                          subset_variables, subset_values, age_wts)
-
- if(is.null(statistic_primary)){
-
[…]

### 2022-10-07 R/svy_stat_count.R

@@ -1,11 +1,23 @@
 
-svy_stat_count <- function(outcome, design, key, ...) {
 
- if(key$variables[[outcome]]$type == 'intg'){
+design_prep_count <- function(design, outcome){
 
-  design$variables[[outcome]] %<>% as.factor()
+ # ensure categorical inputs
+ design$variables[[outcome]] %<>% as.factor()
 
+ # use calibrated weights if they are present
+ if('svy_weight_cal' %in% names(design$variables)){
+  design$variables$svy_weight <- design$variables$svy_weight_cal
+  design$prob <- 1/design$variables$svy_weight
  }
+
+ design
+
+}
+
+svy_stat_count <- function(outcome, design, ...) {
+
+ design <- design_prep_count(design, outcome = outcome)
 
  svytotal(x = as_svy_formula(outcome),
           na.rm = TRUE,
@@ -15,6 +27,20 @@ svy_stat_count <- function(outcome, design, key, ...) {
 
 
 }
+
+svy_statby_count <- function(outcome, by_vars, design, ...){
+
+ design <- design_prep_count(design, outcome = outcome)
+
+ svy_statby(outcome = outcome,
+            by_vars = by_vars,
+            design = design,
+            svy_stat_fun = svytotal) %>%
+  svy_stat_adorn(stat_type = 'count',
+                 stat_fun = 'statby')
+
+}
+
 
 
 

### 2022-10-07 R/svy_stat_prop.R

@@ -2,6 +2,7 @@
 
 svy_stat_proportion <- function(outcome, design, ...){
 
+ # ensure categorical inputs
  design$variables[[outcome]] %<>% as.factor()
 
  lvls <- levels(design$variables[[outcome]])
@@ -9,14 +10,17 @@ svy_stat_proportion <- function(outcome, design, ...){
  glue::glue("I({outcome} == '{lvls}')") %>%
   set_names(lvls) %>%
   map(as_svy_formula) %>%
-  map(svyciprop, design = design) %>%
+  map(svyciprop, design = design, method = 'beta') %>%
   svy_stat_adorn(stat_type = 'proportion',
                  stat_fun = 'stat')
 
 }
 
+
+
 svy_statby_proportion <- function(outcome, by_vars, design, ...){
 
+ # ensure categorical inputs
  design$variables[[outcome]] %<>% as.factor()
 
  lvls <- levels(design$variables[[outcome]])
@@ -27,8 +31,9 @@ svy_statby_proportion <- function(outcome, by_vars, design, ...){
   map(svy_statby,
       by_vars = by_vars,
       design = design,
-      svy_stat_fun = svyciprop) %>%
- svy_stat_adorn(stat_type = 'proportion',
+      svy_stat_fun = svyciprop,
+      method = 'beta') %>%
+  svy_stat_adorn(stat_type = 'proportion',
                  stat_fun = 'statby')
 
 }

### 2022-10-07 R/svy_stat_suppress.R

@@ -1,0 +1,180 @@
+#' .. content for \description{} (no empty lines) ..
+#'
+#' .. content for \details{} ..
+#'
+#' @title
+#' @param n_obs
+#' @param n_psu
+#' @param n_strata
+svy_stat_suppress <- function(dt, type, stat, n_obs, n_psu, n_strata) {
+
+ suppress_fun <- paste('svy_stat_suppress', type, sep = '_')
+
+ dt$unreliable_status <- rep(FALSE, nrow(dt))
+ dt$unreliable_reason <- rep(NA_character_, nrow(dt))
+ dt$review_needed <- rep(FALSE, nrow(dt))
+ dt$review_reason <- rep(NA_character_, nrow(dt))
+
+ if(stat %in% c('count', 'quantile', 'mean')) return(dt)
+
+ out <- do.call(suppress_fun,
+                args = list(dt = dt,
+                            n_obs = n_obs,
+                            n_psu = n_psu,
+                            n_strata = n_strata))
+
+}
+
+svy_stat_suppress_ctns <- function(dt, n_obs, n_psu, n_strata){
+ dt
+}
+
+svy_stat_suppress_intg <- function(dt, n_obs, n_psu, n_strata){
+ svy_stat_suppress_bnry(dt, n_obs, n_psu, n_strata)
+}
+
+svy_stat_suppress_catg <- function(dt, n_obs, n_psu, n_strata){
+ svy_stat_suppress_bnry(dt, n_obs, n_psu, n_strata)
+}
+
+svy_stat_suppress_bnry <- function(dt, n_obs, n_psu, n_strata){
+
+ n_df <- n_psu - n_strata
+
+ # When applicable for complex surveys, if the sample size and confidence
+ # interval criteria are met for presentation and the degrees of freedom
+ # are fewer than 8, then the proportion should be flagged for statistical
+ # review by the clearance official. This review may result in either the
+ # presentation or the suppression of the proportion.
+
+ if(n_df < 8){
+
+  dt$review_needed[] <- TRUE
+  dt$review_reason[] <- "degrees of freedom < 8"
+
+
+ }
+
+ p  <- dt$estimate  / 100
+ q  <- 1 - p
+ se <- dt$std_error / 100
+
+ if(any(p < .Machine$double.eps)){
+
+  dt$review_needed[p < .Machine$double.eps] <- TRUE
+  dt$review_reason %<>% insert_text(which(p < .Machine$double.eps),
[…]

### 2022-10-07 R/svy_statby_count.R

@@ -1,20 +1,0 @@
-
-
-svy_statby_count <- function(outcome, by_vars, design, key, ...){
-
- if(key$variables[[outcome]]$type == 'intg'){
-
-  design$variables[[outcome]] %<>% as.factor()
-
- }
-
- svy_statby(outcome = outcome,
-            by_vars = by_vars,
-            design = design,
-            svy_stat_fun = svytotal) %>%
-  svy_stat_adorn(stat_type = 'count',
-                 stat_fun = 'statby')
-
-}
-
-

### 2022-10-07 R/svy_statby_percentage.R

@@ -1,17 +1,0 @@
-
-svy_statby_percentage <- function(outcome, by_vars, design, ...){
-
- if(is_discrete(outcome)){
-  design$variables[[outcome]] <- as.factor(design$variables[[outcome]])
- }
-
- svy_statby(
-  outcome = outcome,
-  by_vars = by_vars,
-  design = design,
-  svy_stat_fun = svymean
- ) %>%
-  svy_stat_adorn(stat_type = 'percentage',
-                 stat_fun = 'statby')
-
-}

### 2022-10-07 R/svy_statby_quantile.R

@@ -1,13 +1,0 @@
-
-
-svy_statby_quantile <- function(outcome, by_vars, design, quantiles, ...){
-
- svy_statby(outcome = outcome,
-            by_vars = by_vars,
-            design = design,
-            svy_stat_fun = svyquantile,
-            quantiles = quantiles) %>%
-  svy_stat_adorn(stat_type = 'quantile',
-                 stat_fun = 'statby')
-
-}

### 2022-10-07 R/zzz.R

@@ -1,8 +1,9 @@
 
 .onLoad <- function(libname, pkgname){
 
- rspec <- round_spec() %>%
-  round_using_magnitude(digits = c(2, 1, 0), breaks = c(10, 100, Inf))
+ rspec <- table.glue::round_spec() %>%
+  table.glue::round_using_magnitude(digits = c(2, 1, 0),
+                                    breaks = c(10, 100, Inf))
 
  names(rspec) <- paste('table.glue.', names(rspec), sep = '')
 

### 2022-10-07 logo.R

@@ -3,13 +3,13 @@ devtools::load_all()
 
 library(ggplot2)
 
-ds <- nhanes_bp %>%
+ds <- nhanes_data %>%
  .[svy_subpop_htn == 1] %>%
  .[, svy_weight := svy_weight_mec] %>%
  svy_design_new(
   exposure = NULL,
   n_exposure_group = numeric(),
-  years = levels(nhanes_bp$svy_year),
+  years = levels(nhanes_data$svy_year),
   pool = 'no'
  ) %>%
  subset(demo_pregnant == 'No' | is.na(demo_pregnant)) %>%

### 2022-10-07 tests/testthat/test-nhanes_calibrate.R

@@ -2,8 +2,8 @@
 # benchmark the nhanes_calibrate function
 # takes about ~30-35 milliseconds on average
 # microbenchmark::microbenchmark(
-#  bcj_calib = nhanes_calibrate(nhanes_full = nhanes_bp,
-#                               nhanes_sub = nhanes_bp[svy_subpop_htn == 1])
+#  bcj_calib = nhanes_calibrate(nhanes_full = nhanes_data,
+#                               nhanes_sub = nhanes_data[svy_subpop_htn == 1])
 # )
 
 # validate calibration in subgroups defined by:
@@ -12,8 +12,8 @@ variables <- c('svy_year',
                'demo_gender',
                'demo_race_black')
 
-nhanes_htn <- nhanes_calibrate(nhanes_full = nhanes_bp,
-                               nhanes_sub = nhanes_bp[svy_subpop_htn == 1])
+nhanes_htn <- nhanes_calibrate(nhanes_full = nhanes_data,
+                               nhanes_sub = nhanes_data[svy_subpop_htn == 1])
 
 test_that(
  desc = "sum of weights in the calibrated data matches the original",
@@ -22,23 +22,23 @@ test_that(
   # overall
   expect_equal(
    sum(nhanes_htn$svy_weight_cal),
-   sum(nhanes_bp$svy_weight_mec)
+   sum(nhanes_data$svy_weight_mec)
   )
 
 
   for(v in variables){
-   for(i in levels(nhanes_bp[[v]])){
+   for(i in levels(nhanes_data[[v]])){
     expect_equal(
      sum(nhanes_htn$svy_weight_cal[nhanes_htn[[v]]==i]),
-     sum(nhanes_bp$svy_weight_mec[nhanes_bp[[v]]==i])
+     sum(nhanes_data$svy_weight_mec[nhanes_data[[v]]==i])
     )
    }
   }
  }
 )
 
-nhanes_chol <- nhanes_calibrate(nhanes_full = nhanes_bp,
-                                nhanes_sub = nhanes_bp[svy_subpop_chol == 1])
+nhanes_chol <- nhanes_calibrate(nhanes_full = nhanes_data,
+                                nhanes_sub = nhanes_data[svy_subpop_chol == 1])
 
 test_that(
  desc = "sum of weights in the calibrated data matches the original",
@@ -47,14 +47,14 @@ test_that(
   # overall
[…]

### 2022-10-07 tests/testthat/test-nhanes_summarize.R

@@ -1,53 +1,51 @@
 
-
-
 outcomes <- c('bp_sys_mean',          # continuous outcome
               'bp_cat_meds_included', # categorical outcome
+              'bp_med_n_class',       # integer outcome
               'chol_hdl_low')         # binary outcome
 
 exposures <- c('demo_race',      # a categorical variable
                'demo_age_years', # a continuous variable
                'htn_jnc7')       # a binary variable
 
-subset_variables <- list("demo_gender")
-subset_values <- list("Women")
-years <- 'last_5'
+subset_calls <- list("demo_gender" = "Women")
+years <- 'most_recent'
 
 test_that(
  desc = "each outcome can be summarized overall and in exposure groups",
  code = {
   for(o in outcomes){
 
-   expect_true(is.data.table(nhanes_summarize(o, pool = TRUE)))
-   expect_true(is.data.table(nhanes_summarize(o, pool = FALSE)))
+   smry_overall <- nhanes_summarize(outcome_variable = o,
+                                    pool = TRUE,
+                                    simplify_output = TRUE)
+
+   smry_by_cycle <- nhanes_summarize(outcome_variable = o,
+                                     pool = FALSE,
+                                     simplify_output = TRUE)
+
+
+   expect_true(is.data.table(smry_overall))
+   expect_true(is.data.table(smry_by_cycle))
+
+   expect_true(
+    all(names(smry_overall) %in% names(smry_by_cycle))
+   )
 
    for(e in setdiff(exposures, o)){
-
-    expect_true(
-     is.data.table(
-      nhanes_summarize(
-       outcome = o,
-       exposure = e,
-       years = years,
-       subset_variables = subset_variables,
-       subset_values = subset_values,
-       pool = TRUE
+    for (p in c(TRUE, FALSE)){
+     expect_true(
+      is.data.table(
+       nhanes_summarize(
+        outcome_variable = o,
+        group_variable = e,
+        time_values = years,
+        subset_calls = subset_calls,
+        pool = p
+       )
       )
      )
-    )
-    expect_true(
-     is.data.table(
-      nhanes_summarize(
-       outcome = o,
-       exposure = e,
-       years = years,
-       subset_variables = subset_variables,
-       subset_values = subset_values,
-       pool = FALSE
-      )
-     )
-    )
-
+    }
    }
   }
  }

### 2022-10-07 tests/testthat/test-nhanes_variables.R

@@ -9,7 +9,7 @@ test_that(
  desc = 'inverse variables have total discordance',
  code = {
 
-  concordant_accaha <- with(nhanes_bp,
+  concordant_accaha <- with(nhanes_data,
                             table(bp_control_accaha,
                                   bp_uncontrolled_accaha)) %>%
    diag() %>%
@@ -17,7 +17,7 @@ test_that(
 
   expect_equal(concordant_accaha, 0)
 
-  concordant_jnc7 <- with(nhanes_bp,
+  concordant_jnc7 <- with(nhanes_data,
                           table(bp_control_jnc7,
                                 bp_uncontrolled_jnc7)) %>%
    diag() %>%
@@ -33,7 +33,7 @@ test_that(
 # LOCAL tests
 # pm_23sep2022 <- haven::read_sas('data-raw/final9920_0923022.sas7bdat')
 #
-# pm_23jul2022 <- haven::read_sas('data-raw/nhanes_bp-raw.sas7bdat')
+# pm_23jul2022 <- haven::read_sas('data-raw/nhanes_data-raw.sas7bdat')
 #
 # length(names(pm_23jul2022))
 #
@@ -45,25 +45,25 @@ test_that(
 # test_that(
 #  desc = 'Number of NAs and category counts match',
 #  code = {
-#   for(i in names(nhanes_bp)){
+#   for(i in names(nhanes_data)){
 #
 #    type <- key[variable == i, type]
 #
 #    expect_equal(sum(is.na(nhanes_sas[[i]])),
-#                 sum(is.na(nhanes_bp[[i]])))
+#                 sum(is.na(nhanes_data[[i]])))
 #
 #    if(!grepl('^svy_weight', x = i) && i %in% names(nhanes_sas)){
 #
 #     if(type %in% c('ctns', 'svy') ){
 #
 #      # these variable should be exactly the same
-#      expect_equal(nhanes_sas[[i]], nhanes_bp[[i]], ignore_attr = TRUE)
+#      expect_equal(nhanes_sas[[i]], nhanes_data[[i]], ignore_attr = TRUE)
 #
 #     } else if(type %in% c('catg', 'bnry', 'time')){
 #
 #      # categorical variables that were recoded should have same counts
 #      counts_sas   <- sort(table(nhanes_sas[[i]]))
-#      counts_shiny <- sort(table(nhanes_bp[[i]]))
+#      counts_shiny <- sort(table(nhanes_data[[i]]))
 #
[…]

### 2022-10-07 tests/testthat/test-reliability_prop.R

@@ -1,86 +1,0 @@
-
-
-ds <- nhanes_bp %>%
- filter(svy_subpop_htn==1) %>%
- nhanes_calibrate(nhanes_full = nhanes_bp, nhanes_sub = .) %>%
- .[, svy_weight := svy_weight_cal] %>%
- svy_design_new()
-
-smry <- svy_design_summarize(ds, outcome = 'htn_jnc7') %>%
- reliability_prop(design = ds)
-
-
-
-# TODO
-# smry <- tribble(
-#  ~estimate,	~std_error, ~ci_lower, 	~ci_upper,	~n_eff_SAS,	~kg_wdth_SAS,	~kg_relw_p_SAS,	~kg_relw_q_SAS,	~p_reliable_SAS,	~q_reliable_SAS,	~p_staistical_SAS,
-#  26.7192,	    1.3762,	   24.0241,	   29.5494,	   1033.87,	   0.055253,	     20.6793,	       7.53994,	            1,	              1,	            0,
-#  26.3107,	    1.237,	    23.8868,	   28.8466,	   1266.99,	   0.049597,	     18.8507,	       6.73061,	            1,	              1,	            0,
-#  29.7108,	    1.2419,	   27.2687,	   32.2426,	   1353.97,	   0.049739,	      16.741,	       7.07633,	            1,	              1,	            0,
-#  29.27,	      1.0892,	   27.1277,	   31.4833,	   1744.97,	   0.043556,	     14.8808,	       6.15808,	            1,	              1,	            0,
-#  29.8917,	    0.9171,	   28.0858,	   31.7457,	   2491.58,	   0.036599,	      12.244,	       5.22041,	            1,	              1,	            0,
-#  29.4645,	    1.2401,	   27.0264,	   31.9935,	   1351.43,	   0.049671,	     16.8579,	       7.04196,	            1,	              1,	            0,
-#  30.1448,	    1.5356,	   27.1274,	   33.2958,	   892.98,	    0.061684,	     20.4627,	       8.83032,	            1,	              1,	            0,
-#  31.3705,	    0.9034,	   29.5897,	   33.1934,	  2637.86,	    0.036037,	     11.4876,	       5.25096,	            1,	              1,	            0,
-#  31.3212,	    1.1973,	   28.9629,	   33.7539,	  1500.57,	     0.04791,	     15.2962,	        6.9759,            	1,              	1,	            0,
[…]

### 2022-10-09 README.Rmd
reliability!

@@
 <!-- - Documentation/demos for R functions used in the app -->
-<!-- - Definitions/notes for all BP variables (table S1 in manuscript) -->
 <!-- - Tutorial videos -->

### 2022-10-09 R/app_run.R

@@ -9,6 +9,8 @@ library(shiny)
 #' @export
 #'
 
+# TODO: drop lipids, fix pool, and drop grps with few obs
+
 app_run <- function(nhanes_data = cardioStatsUSA::nhanes_data,
                     nhanes_key = cardioStatsUSA::nhanes_key) {
 
@@ -142,7 +144,7 @@ app_run <- function(nhanes_data = cardioStatsUSA::nhanes_data,
      "input.do == 'figure'",
      radioGroupButtons(
       inputId = 'geom',
-      label = 'Select a plotting geometry',
+      label = 'Select figure type',
       choices = c("Bars" = "bar", "Points" = "scatter"),
       selected = "bar",
       status = "primary",
@@ -440,7 +442,7 @@ app_run <- function(nhanes_data = cardioStatsUSA::nhanes_data,
        6,
        style='padding-right: 2px;',
        pickerInput(
-        inputId = 'group_class',
+        inputId = 'stratify_class',
         label = 'Select stratify type',
         choices = names(variable_choices$stratify),
         selected = NULL,
@@ -675,40 +677,41 @@ app_run <- function(nhanes_data = cardioStatsUSA::nhanes_data,
     ss_val_ctns <- paste('subset_value', i, 'ctns', sep = '_')
     ss_val_catg <- paste('subset_value', i, 'catg', sep = '_')
 
-    if(!is.null(input[[ss_var]])){
-
-     if(input$outcome == input[[ss_var]]){
-
-      updatePickerInput(
-       session = session,
-       inputId = ss_var,
-       choices = nhanes_key$variable_choices$group,
-       selected = character(0)
-      )
-
-      updatePickerInput(
-       session = session,
-       inputId = ss_var,
-       choices = nhanes_key$variable_choices$group,
-       selected = character(0)
-      )
-
-      updatePrettyCheckboxGroup(
-       inputId = ss_val_catg,
-       selected = character(0)
-      )
-
-      updateSliderInput(
-       inputId = ss_val_ctns,
-       value = c(0,0)
-       # TODO: fix
-       # value = c(min(nhanes_data[[input[[ss_var]]]], na.rm = TRUE),
-       #           max(nhanes_data[[input[[ss_var]]]], na.rm = TRUE))
-      )
-
-     }
-
-    }
+    # I don't think this is necessary anymore
+    # if(!is.null(input[[ss_var]])){
+    #
+    #  if(input$outcome == input[[ss_var]]){
+    #
+    #   updatePickerInput(
[…]

### 2022-10-09 R/nhanes_design.R

@@ -75,11 +75,23 @@ nhanes_design <- function(data,
 
  # subset and pool the weights (if needed) ----
 
- divide_by <- if(pool) length(time_values) else 1
+ design_data <- dt_data[dt_data[[time_variable]] %in% time_values]
 
- design_data <- dt_data %>%
-  .[.[[time_variable]] %in% time_values] %>%
-  .[, svy_weight_mec := svy_weight_mec / divide_by]
+ if(pool){
+
+  divide_by <- length(time_values)
+
+  design_data[, svy_weight_mec := svy_weight_mec / divide_by]
+
+  if('svy_weight_cal' %in% names(design_data)){
+
+   design_data[, svy_weight_cal := svy_weight_cal / divide_by]
+
+  }
+
+ }
+
+
 
  # group variable ----
 
@@ -98,7 +110,7 @@ nhanes_design <- function(data,
        j = group_variable,
        value = discretize(design_data[[group_variable]],
                           method = group_cut_type,
-                          breaks = group_cut_n))
+                          breaks = as.numeric(group_cut_n)))
 
    group_info$cut_n <- group_cut_n
    group_info$cut_type <- group_cut_type
@@ -301,6 +313,29 @@ print.nhanes_design <- function(x, ...){
 
  }
 
+
+ cat_pretty(
+  "\n",
+  paste_pretty("N observations"),
+  "\n"
+ )
+
+ cat_pretty(
+  paste_pretty(
+   "- Unweighted: ",
+   table.glue::table_value(
+    get_obs_count(x$design, weighted = FALSE)
+   )
+  ),
+  "\n",
+  paste_pretty(
+   "- Weighted: ",
+   table.glue::table_value(
+    get_obs_count(x$design, weighted = TRUE)
+   ), "\n"
+  )
+ )
+
  cat(footer)
 
  invisible(x)
@@ -331,3 +366,6 @@ is_subsetted <- function(x){
  !is.null(x$subset_rows)
 }
 
+is_nhanes_design <- function(x){
+ inherits(x, 'nhanes_design')
+}

### 2022-10-09 R/nhanes_summarize.R

@@ -39,6 +39,7 @@ nhanes_summarize <- function(data,
                              key,
                              outcome_variable,
                              outcome_quantiles = NULL,
+                             outcome_stats = NULL,
                              group_variable = NULL,
                              group_cut_n = NULL,
                              group_cut_type = NULL,
@@ -49,7 +50,6 @@ nhanes_summarize <- function(data,
                              subset_calls = list(),
                              age_wts = NULL,
                              simplify_output = TRUE){
-
 
  if(missing(data)) data <- cardioStatsUSA::nhanes_data
  if(missing(key)) key <- cardioStatsUSA::nhanes_key
@@ -72,6 +72,7 @@ nhanes_summarize <- function(data,
   # re-calibrate weights to match the total in nhanes_data
   nhanes_calibrate(nhanes_full = dt_data)
 
+ # browser()
 
  ds <- nhanes_design(data = dt_sub,
                      key = dt_key,
@@ -93,6 +94,7 @@ nhanes_summarize <- function(data,
   )
  }
 
+
  if(!is.null(age_wts)){
   ds %<>% nhanes_design_standardize(
    standard_variable = 'demo_age_cat',
@@ -100,6 +102,8 @@ nhanes_summarize <- function(data,
   )
  }
 
- nhanes_design_summarize(ds, simplify_output = simplify_output)
+ nhanes_design_summarize(ds,
+                         stats = outcome_stats %||% ds$stats,
+                         simplify_output = simplify_output)
 
 }

### 2022-10-09 R/nhanes_visualize.R

@@ -28,8 +28,7 @@ nhanes_visualize <- function(data,
                              time_variable = 'svy_year',
                              time_values = NULL,
                              pool = FALSE,
-                             subset_variables = list(),
-                             subset_values = list(),
+                             subset_calls = list(),
                              age_wts = NULL,
                              statistic_primary = NULL,
                              geom = 'bar',
@@ -52,8 +51,7 @@ nhanes_visualize <- function(data,
                           time_variable = time_variable,
                           time_values = time_values,
                           pool = pool,
-                          subset_variables = subset_variables,
-                          subset_values = subset_values,
+                          subset_calls = subset_calls,
                           age_wts = age_wts,
                           simplify_output = FALSE)
 

### 2022-10-09 R/plotly_viz_make_hover.R

@@ -8,7 +8,7 @@ plotly_viz_make_hover <- function(data, stat_all, group, group_label) {
  data$group_label <- ""
  label_sep <- ""
 
- if(is_used(group)){
+ if(is_used(group) && group != 'fake_._group'){
   data$group_label <- group_label %>%
    paste(data[[group]], sep = ': ')
   label_sep <- "\n"

### 2022-10-09 app.R

@@ -1,2 +1,2 @@
 pkgload::load_all(".")
-app_run()
+app_run(nhanes_key = nhanes_key[module != 'chol'])

### 2022-10-09 data-raw/nhanes_load.R

@@ -13,11 +13,11 @@ nhanes_load <- function(fpath = file.path(here::here(), 'data-raw'),
                         as = 'tibble'){
 
  data_bp <- haven::read_sas(
-  file.path(fpath, 'nhanes_data-raw.sas7bdat')
+  file.path(fpath, 'htn-raw.sas7bdat')
  )
 
  data_lipids <- haven::read_sas(
-  file.path(fpath, 'nhanes_lipids-raw.sas7bdat')
+  file.path(fpath, 'lipids-raw.sas7bdat')
  )
 
  out <- left_join(data_bp, data_lipids, by = c("SEQN", "WTSAF2YR")) %>%

### 2022-10-09 tests/testthat/test-nhanes_design_subset.R

@@ -1,4 +1,11 @@
 
+# what happens if the subset is empty?
+
+ds <- nhanes_design(data = nhanes_data,
+                    key = nhanes_key,
+                    time_values = 'most_recent',
+                    outcome_variable = 'bp_sys_mean')
+
 
 # subsets -----------------------------------------------------------------
 

### 2022-10-09 tests/testthat/test-reliability_mean.R

@@ -1,1 +1,13 @@
 
+
+smry <- nhanes_summarize(
+ data = nhanes_data,
+ key = nhanes_key,
+ outcome_variable = 'bp_sys_mean',
+ outcome_stats = 'mean',
+ subset_calls = list(bp_med_ace = "Yes",
+                     bp_med_aldo = "Yes"),
+ stratify_variable = "demo_race"
+)
+
+smry[unreliable_status==TRUE]

### 2022-10-12 rmd/group_versus_stratify.Rmd
more docs/tests

The group variable and the stratify variable are both used to summarize an outcome within subgroups. If the summary is returned as a table, then the group and stratify variables can be considered interchangeable. If the summary is returned as a plot, then the subgroups defined by the group variable are shown on the same graph, while the subgroups defined by the stratify variable are shown on different graphs (i.e., one graph per strata). For example, if you wanted to estimate the race and sex specific prevalence of hypertension with one graph for men and one graphs for women, you would use `group = 'demo_race'` and `stratify = 'demo_gender'`

### 2022-10-12 rmd/nhanes_data.Rmd
more docs/tests

```{r child = 'aaa.Rmd'}

```

```{r, echo=FALSE, results='asis'}

classes <-
 list("Survey" = "In each NHANES cycle, potential participants were identified using a multi-stage sampling process. The variables below are based on this process.",
      "Demographics" = "The NHANES in-home interview included questions about participant demographics, including age, sex, race/ethnicity, and pregnancy status",
      "Blood pressure" = "Systolic and diastolic BP (SBP and DBP, respectively) were measured three times by trained and certified physicians. BP was measured using a mercury sphygmomanometer from 1999-2000 through 2015-2016 and using an oscillometric device in 2017-2020. The oscillometric SBP and DBP values were calibrated to the mercury device. Mean SBP and DBP levels were computed over all available",
      "Hypertension" = "These variables are derived according to the 2017 American College of Cardiology/American Heart Association (ACC/AHA) BP guideline and the Seventh Joint National Committee (JNC7) BP guideline." ,
      "Antihypertensive medication" = "During the interview, the labels of medications that participants reported taking in the preceding 30 days were recorded.",
      "Comorbidities" = "These variables describe comorbid conditions.")

data_to_doc <- nhanes_key %>% 
 .[class %in% names(classes)] %>%
 .[variable != 'svy_subpop_chol'] %>% 
 .[variable != 'svy_weight_cal'] %>% 
 .[, class := factor(class, levels = names(classes))] %>% 
 .[order(class)] %>% 
 droplevels()

data_to_doc$description %<>% stringr::str_replace_all('\u2265', '>=')
data_to_doc$label %<>% stringr::str_replace_all('\u2265', '>=')

key_classes <- split(data_to_doc, f = data_to_doc$class)

for(i in seq_along(key_classes)){

 cat("#", names(key_classes)[i], "variables", "\n")

 cat(classes[[names(key_classes)[i]]], "\n\n")

 for(j in seq_along(key_classes[[i]]$variable)){
  cat("##", key_classes[[i]]$variable[j], "\n")
  cat("- Label: ", key_classes[[i]]$label[j], "\n")
  cat("- Description: ", key_classes[[i]]$description[j], "\n\n")
 }

}

```

### 2022-10-12 R/discretize.R

@@ -1,0 +1,39 @@
+
+discretize <- function(x, method, n_group){
+
+ breaks <-
+  switch(method,
+         'frequency' = get_breaks_frequency(x, n_group),
+         'interval' = get_breaks_interval(x, n_group))
+
+ labels <- character(n_group)
+
+ for(i in seq_along(labels)){
+  labels[i] <- table_glue("{breaks[i]} to <{breaks[i+1]}")
+ }
+
+ cut(x,
+     breaks = breaks,
+     include.lowest = TRUE,
+     right = FALSE,
+     labels = labels)
+
+}
+
+get_breaks_frequency <- function(x, n_group){
+
+ probs <- seq(0, 1, length.out = n_group+1)
+
+ quantile(x, probs = probs, na.rm = TRUE)
+
+
+
+
+}
+get_breaks_interval <- function(x, n_group){
+
+ bounds <- range(x, na.rm = TRUE)
+
+ seq(bounds[1], bounds[2], length.out = n_group+1)
+
+}

### 2022-10-12 R/nhanes_bp.R

@@ -1,10 +1,0 @@
-#' NHANES blood pressure data, 1999-2020
-#'
-#' A dataset containing variables that pertain to blood pressure from the
-#'   NHANES continuous examination cycles. The documenation of the variables
-#'   in this data are available online (TODO: add link)
-#'
-#' @source \url{https://wwwn.cdc.gov/nchs/nhanes/continuousnhanes/default.aspx}
-"nhanes_data"
-
-

### 2022-10-12 R/nhanes_data.R

@@ -1,0 +1,12 @@
+#' NHANES data, 1999-2020
+#'
+#' `r describe_nhanes_program()`
+#'
+#' `r describe_nhanes_data()`
+#'
+#' @includeRmd rmd/nhanes_data.Rmd
+#'
+#' @source \url{https://wwwn.cdc.gov/nchs/nhanes/continuousnhanes/default.aspx}
+"nhanes_data"
+
+

### 2022-10-12 R/nhanes_design.R

@@ -1,25 +1,98 @@
 
-#' Title
+#' Survey Design for NHANES Data
 #'
-#' @param data
-#' @param key
-#' @param outcome
-#' @param group
-#' @param group_cut_n
-#' @param group_cut_type
-#' @param stratify
-#' @param time_values
-#' @param pool
-#' @param run_checks
+#' @description NHANES design objects are the data structure used in
+#'   the `cardioStatsUSA` package for analysis of NHANES data.
 #'
-#' @return
+#' @param data \[data.frame\]
+#' `r describe_nhanes_data_input()`
+#'  See Details for specific requirements.
+#'  See [nhanes_data] for an example.
+#'
+#' @param key \[data.frame\]
+#' `r describe_nhanes_key_input()`
+#'  See Details for specific requirements.
+#'  See [nhanes_key] for an example.
+#'
+#' @param outcome_variable \[character(1)\]
+#' The name of the outcome variable to be summarized.
+#'
+#' @param outcome_quantiles \[numeric(1+)\]
+#' The quantiles to be summarized for a continuous outcome. The default is
+#'   `c(0.25, 0.50, 0.75)`. For example,
+#'
+#' - `outcome_quantiles = c(.5)` will compute the 50th percentile
+#'   (i.e., the median)
+#'
+#' - `outcome_quantiles = c(.25, .5, .75)` will compute the 25th, 50th,
+#'    and 75th percentile.
+#'
+#' - `outcome_quantiles = seq(.1, .9, by = .1)` will compute every 10th
+#'   percentile, except for the 0th and 100th
+#'
+#' @param group_variable \[character(1)\]
+#' The name of the group variable. See Details for a description of
+#'   the group variable and the stratify variable.
+#'
+#' @param group_cut_n \[integer(1)\]
+#'  The number of groups to form using the group variable. This is only
+#'    relevant if the group variable is continuous, and can be omitted.
+#'    Default is 3
+#'
+#' @param group_cut_type \[character(1)\]
+#'  The method used to create groups with the grouping variable. This is only
+#'    relevant i fthe group variable is continuous, and can be omitted.
+#'    Valid options are:
+#'
+#'    - "interval": equal interval width, e.g., three groups with ages of
+#'      0 to <10, 10 to <20, and 20 to < 30 years.
+#'
+#'    - "frequency": equal frequency, e.g., three groups with ages of
+#'      0 to <q, q to <p, and p to <r, where q, p, and r are selected
[…]

### 2022-10-12 R/nhanes_design_viz.R

@@ -21,8 +21,6 @@
 #  nhanes_design_summarize(simplify_output = FALSE) %>%
 #  nhanes_design_viz()
 
-#'
-#'
 nhanes_design_viz <- function(x,
                               statistic_primary = NULL,
                               geom = 'bar',
@@ -134,11 +132,18 @@ nhanes_design_viz <- function(x,
   stat_all <- x$stats
 
   if('quantile' %in% stat_all){
-   stat_all <- c(stat_all, 'q25', 'q50', 'q75')
+   stats_in_x <- unique(data$statistic)
+   stats_not_quantile <- setdiff(get_outcome_stats('ctns'), 'quantile')
+   stats_quantile <- setdiff(stats_in_x, stats_not_quantile)
+   stat_all <- c(stat_all, stats_quantile)
   }
 
   if(statistic_primary == 'quantile'){
-   statistic_primary <- 'q50'
+   n_quantiles <- length(stats_quantile)
+   if(n_quantiles %% 2 == 0)
+    statistic_primary <- stats_quantile[n_quantiles/2]
+   else
+    statistic_primary <- stats_quantile[(n_quantiles+1)/2]
   }
 
   stacked_and_pooled <- outcome_type == 'catg' && pool
@@ -382,7 +387,6 @@ nhanes_design_viz <- function(x,
  output
 
 }
-
 
 
 

### 2022-10-12 tests/testthat/test-nhanes_visualize.R

@@ -1,0 +1,242 @@
+
+# continuous outcome ----
+
+test_that(
+ desc = 'one group blots, continuous outcome mean',
+ code = {
+  expect_snapshot(nhanes_visualize(outcome_variable = 'bp_sys_mean'),
+                  cran = FALSE)
+
+  expect_snapshot(nhanes_visualize(outcome_variable = 'bp_sys_mean',
+                                   geom = 'point'),
+                  cran = FALSE)
+
+ }
+)
+
+test_that(
+ desc = 'one group bar blots, continuous outcome quantile',
+ code = {
+  expect_snapshot(nhanes_visualize(outcome_variable = 'bp_sys_mean',
+                                   outcome_quantiles = seq(1,10)/10,
+                                   statistic_primary = 'quantile'),
+                  cran = FALSE)
+  expect_snapshot(nhanes_visualize(outcome_variable = 'bp_sys_mean',
+                                   outcome_quantiles = seq(1,10)/10,
+                                   statistic_primary = 'quantile',
+                                   geom = 'point'),
+                  cran = FALSE)
+ }
+)
+
+test_that(
+ desc = 'multiple group bar blots, continuous outcome mean',
+ code = {
+  expect_snapshot(nhanes_visualize(outcome_variable = 'bp_sys_mean',
+                                   group_variable = 'demo_gender'),
+                  cran = FALSE)
+  expect_snapshot(nhanes_visualize(outcome_variable = 'bp_sys_mean',
+                                   group_variable = 'demo_gender',
+                                   geom = 'point'),
+                  cran = FALSE)
+ }
+)
+
+test_that(
+ desc = 'multiple group plots, continuous outcome quantile',
+ code = {
+  expect_snapshot(nhanes_visualize(outcome_variable = 'bp_sys_mean',
+                                   outcome_quantiles = seq(1,10)/10,
+                                   statistic_primary = 'quantile',
+                                   group_variable = 'demo_gender'),
+                  cran = FALSE)
+  expect_snapshot(nhanes_visualize(outcome_variable = 'bp_sys_mean',
+                                   outcome_quantiles = seq(1,10)/10,
+                                   statistic_primary = 'quantile',
+                                   group_variable = 'demo_gender',
+                                   geom = 'point'),
+                  cran = FALSE)
+ }
+)
+
+test_that(
+ desc = 'stratified group plots, continuous outcome quantile',
+ code = {
[…]

### 2022-10-21 paper/manuscript.Rmd
big update
added CI options for percentage, tweaked api, fixed many small things

@@
 ---
-title: "Cardiometabolic statistics for US adults: an open source web application for exploratory analysis"
+title: "Hypertension statistics for US adults: an open-source web application for analysis and visualization of US NHANES data"
 always_allow_html: true
@@
 conflicted::conflict_prefer('layout', 'plotly')
+
+# define age weights for standardization
+age_wts_overall <- c(49.3, 33.6, 10.1, 7)
+age_wts_htn <- c(26.4, 43.5, 17.0, 13.2)
 
@@
 
-**ABSTRACT**
+# ABSTRACT
 
@@
 
-**Background**: Data from the National Health and Nutrition Examination Survey (NHANES) program are freely available. However, analysis of these data may be time-intensive and challenging.
+**Background**: Data from the US National Health and Nutrition Examination Survey (NHANES) are freely available and can be analyzed to produce hypertension statistics for the non-institutionalized US population.
+However, the analysis of these data requires statistical programming knowledge.
 
-**Objectives**: Develop, validate, and share an open-source web application that allows for interactive, exploratory analysis of NHANES data that can be used and extended upon by others.
+**Methods**: We developed and validated a web-based application that provides hypertension statistics for US adults using 10 cycles of NHANES data, 1999-2000 through 2017-2020 without requiring users to conduct statistical programming.
+The application’s interface allows users to estimate crude and age-adjusted means, quantiles, and proportions.
+Population counts can also be estimated.
+To demonstrate the application’s capabilities, we estimated hypertension statistics for US adults.
 
-**Methods**: We developed the open-source web application using the R programming language and several R packages.
-We validated the web application by using it to reproduce results from two prior studies and a report from the Centers for Disease Control.
-To demonstrate the application's capabilities, we estimated characteristics of adults in the US, the prevalence of hypertension, distribution of the number of antihypertensive medication classes being used, prevalence of controlled blood pressure (BP) among those with hypertension, and the prevalence of resistant hypertension.
+**Results**: The estimated mean systolic blood pressure (BP) declined from 123 mmHg in 1999-2000 to 120 mmHg in 2009-2010 and increased to 123 mmHg in 2017-2020.
+The age-adjusted prevalence of hypertension (i.e., systolic BP ≥130 mmHg, diastolic BP ≥80 mmHg or self-reported antihypertensive medication use) was 47.9% in 1999-2000, 43.0% in 2009-2010, and 44.7% in 2017-2020.
+In 2017-2020, 115.3 million US adults had hypertension.
+The age-adjusted prevalence of controlled BP, defined according to the 2017 American College of Cardiology/American Heart Association BP guideline, among US adults with hypertension was 9.6% in 1999-2000, 24.6% in 2013-2014, and 21.7% in 2017-2020.
+After age-adjustment and among US adults who self-report taking antihypertensive medication, 27.4%, 48.5%, and 42.8% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively.
+The application was validated by replicating statistics from three publications.
 
-**Results**: The estimated mean systolic BP for all US adults was 123 mm Hg in 1999-2000, 120 mm Hg in 2009-2010, and 123 mm Hg in 2017-2020.
-The estimated prevalence of hypertension among US adults was lowest in 2009-2010 (43.3%) and highest in 2017-2020 (46.6%), with the estimated number of US adults with hypertension increasing from 89.8 million in 1999-2000 to 115.3 million in 2017-2020 .
-The age adjusted mean systolic BP was 140 in 1999-2000, 134 in 2009-2010, and 136 in 2017-2020 among US adults with hypertension.
-Among US adults who self-reported taking antihypertensive medication, the percentage taking four or more classes of antihypertensive medication increased from 3.5% in 1999-2000 to 7.6% in 2017-2020.
-The age-adjusted estimated prevalence of BP control among US adults with hypertension was 10.2%, 25.8%, and 23.8% in 1999-2000, 2009-2010, and 2017-2020, respectively.
-Among US adults with hypertension who self-reported taking antihypertensive medication, the prevalence of resistant hypertension was 13.7% in 1999-2000, 17.7% in 2009-2010, and 16.2% in 2017-2020.
-
-**Conclusions.**: The web application developed in the current analysis may increase the usage of NHANES data for research and policy decision making, ensuring results are valid, transparent, and reproducible.
+**Conclusions**: The application developed and validated is publicly available and produced valid, transparent, and reproducible results.
 
@@
 
-The National Health and Nutrition Examination Survey (NHANES) is a program conducted by the US National Center for Health Statistics (NCHS) of the Centers for Disease Control and Prevention (CDC) and is designed to assess the health and nutritional status of the US population.
-NHANES data have been used to provide nationally representative estimates for adults living in the United States (US) in multiple studies of cardiometabolic outcomes.
-For example, NHANES data have been used to estimate the impact of the lower blood pressure (BP) levels used to define hypertension and controlled BP in the 2017 American College of Cardiology/American Heart Association (ACC/AHA) BP guideline versus the Seventh Joint National Committee (JNC7) BP guideline on the proportion and number of US adults with hypertension, recommended antihypertensive medication and with uncontrolled BP.
-Additionally, NHANES data have been used to track the proportion of US adults with controlled BP, which were included in a Call-to-Action to Control BP from the US Surgeon General.
+The National Health and Nutrition Examination Survey (NHANES) is a program conducted by the US National Center for Health Statistics (NCHS) of the Centers for Disease Control and Prevention (CDC) and is designed to assess the health and nutritional status of the non-institutionalized US population.
+NHANES data have been analyzed to provide hypertension statistics for non-institutionalized US adults with important policy and public health implications.
+For example, NHANES data have been used to estimate the impact of the lower blood pressure (BP) levels that define hypertension and controlled BP in the 2017 American College of Cardiology/American Heart Association (ACC/AHA) BP guideline versus the Seventh Joint National Committee (JNC7) BP guideline.
+Additionally, NHANES data have been used to track the proportion of US adults with hypertension that have controlled BP, which were included in a Call-to-Action to Control BP from the US Surgeon General.
 
@@
 However, analyzing NHANES data may be challenging for a number of reasons.
-In particular, specific statistical techniques are required to account for the multi-stage sampling design of NHANES data and produce nationally representative estimates.
-Also, each NHANES cycle includes over 50 data files, each of which has detailed informational material.
-NHANES data collection protocols have changed over time for certain variables, and these differences should be accounted for when harmonizing or pooling data from multiple cycles.
-Another potential challenge is implementation of recommended standards from the CDC, such as their guidelines for suppression of unreliable results.
+Specific statistical techniques are required to account for the multi-stage sampling design used to select NHANES participants and analyses need to be weighted to produce nationally representative estimates.
+Each NHANES cycle includes over 50 data files, each of which has detailed informational material.
+NHANES data collection protocols have changed over time for some variables, and these differences should be taken into account when comparing or aggregating data from multiple cycles.
+Also, tests need to be performed to ensure the statistical estimates are reliable, and unstable estimates should be suppressed.
 
-To address the challenges of analyzing NHANES data, we developed an open-source web-based application that allows users to obtain nationally representative summaries of cardiometabolic outcomes for US adults using NHANES data.
-The goal of the application is to increase the usage of NHANES data for research and policy decision making with a focus on ensuring results are valid, transparent, and reproducible.
-The application is designed to be extended with the addition of "modules", which we define as a set of variables measured among a sub-population of NHANES participants.
-In this manuscript, we review the design and development of the application, present nationally representative summaries of outcomes from the "BP and Hypertension module", and describe online tutorials and resources that may help users engage with the application.
+To address these challenges, we developed an open-source web-based application that provides nationally representative BP and hypertension statistics for non-institutionalized US adults using NHANES data without requiring users to conduct statistical programming.
+The goal of the application is to increase the use of NHANES data for research and policy decision making with a focus on ensuring results are valid, transparent, and reproducible.
+In this manuscript, we review the design, development and validation of the application, present BP and hypertension statistics for US adults, and describe online tutorials and resources that may help users engage with the application.
 
-**METHODS**
+# METHODS
 
@@
 
-NHANES data include in-home interviews and study examinations conducted at mobile examination centers.
-The interview includes questions about demographics, health behaviors, prior diagnoses, medication use, and medical history.
+NHANES data are collected through in-home interviews and study examinations conducted at mobile examination centers.
+The interview included questions about demographics, health behaviors, medical history, and medication use.
 During the interview, the labels of medications that participants reported taking in the preceding 30 days were recorded.
-During the study examination, height and weight were measured, blood samples were used to measure cholesterol, glycated hemoglobin and serum creatinine, and spot urine was used to measure albumin and creatinine and to conduct a pregnancy test.
-Systolic and diastolic BP (SBP and DBP, respectively) were measured three times.
+During the study examination, height, weight and BP were measured and blood and spot urine samples were collected.
+Of relevance to the application, blood samples were used to measure total and HDL cholesterol, glycated hemoglobin and serum creatinine, and the urine sample was used to measure albumin and creatinine and to conduct a pregnancy test.
+The protocol for the BP measurement is available online.
+In brief, systolic and diastolic BP (SBP and DBP, respectively) were measured three times by trained and certified physicians.
 BP was measured using a mercury sphygmomanometer from 1999-2000 through 2015-2016 and using an oscillometric device in 2017-2020.
+The mean SBP and DBP levels were computed over all available measurements for each participant.
+The oscillometric SBP and DBP values were calibrated to the mercury device.
+Antihypertensive medication classes were defined using recommendations from the 2017 ACC/AHA BP guideline.
+For the current analysis, we defined hypertension, BP control, and resistant hypertension according to the 2017 ACC/AHA BP guideline.
+The application also has these variables defined according to the Seventh Report of the Joint National Committee on Prevention, Detection, Evaluation, and Treatment of High Blood Pressure definitions.
+A complete list of the variables available in the data set is provided in Table 1 with full definitions provided in Supplemental Table 1.
 
-*Features of the Open-Source Web-Based Application*
+There were 107,622 NHANES participants in the 10 cycles from 1999-2000 to 2017-2020.
+We restricted the dataset to adults ≥ 18 years of age.
+This exclusion was applied because statistics for BP levels and hypertension in children and adolescents are markedly different than for adults.
+We further restricted the population to participants who completed the in-home interview and study examination, with one or more SBP and DBP measurement, and who had data on self-reported antihypertensive medication use.
+After these exclusions were applied, the population in this application included 56,035 participants (Figure S1).
 
-Full summaries of this application's features and associated tutorials are available.
-Briefly, users can select NHANES cycles to be analyzed.
-Contiguous cycles can be combined to increase precision.
-Estimates are weighted to represent the non-institutionalized US population and may incorporate age-adjustment using direct standardization.
-Count estimates are calibrated to match the estimated total number of US adults.
-Users may visualize summaries of outcomes in groups (i.e., prevalence of hypertension by sex) with all groups in the same figure or with one figure per group, and may additionally specify subsets to exclude from their analyses.
-All outputs created with the web application can be downloaded and saved.
+## Features of the Open-Source Web Application
+
+A full summary of this application’s features and associated tutorials are available online.
+Briefly, users can select NHANES cycles from 1999-2000 to 2017-2020 to be analyzed.
+Estimates are weighted to represent the non-institutionalized US population and users may incorporate age-adjustment through direct standardization.
+Users can restrict analyses to subsets of participants (e.g., participants who self-reported taking antihypertensive medication).
+When population count estimates are requested, survey weights are calibrated to account for participants missing information on SBP, DBP or self-reported antihypertensive medication use.
+Users may tabulate or visualize summaries, and can present results for the overall population or in subgroups defined by a stratifying variable.
+All figures and datasets created with the web application can be downloaded and saved.
 Following CDC recommendations, unreliable statistical estimates are automatically suppressed.
+To increase precision and reliability of estimates, contiguous NHANES cycles can be combined.
 
-*Design of the Open-Source Web-Based Application*
+## Development and validation of the Open-Source Web Application
 
 The web application was created using Shiny, an open-source software package that translates code from the R programming language into HTML, CSS, or JavaScript commands that create a website interface.
-The "cardioStatsUSA" R package provides additional details on the web application's design and comprehensive documentation of its components.
+We created the “cardioStatsUSA” R package to provide additional details on the web application’s design and comprehensive documentation of its components.
+We validated the web application by using it to reproduce statistics reported in two prior studies and one CDC report.
 
-*The Blood Pressure and Hypertension Module*
+## Statistical analysis
 
-We define a module as a subpopulation of NHANES participants paired with a collection of variables.
-Each variable in the collection has designated roles.
-For example, hypertension can be an outcome and can also be used to define groups when analyzing another outcome.
-This module definition is general enough to allow for diverse extensions of the web application, but may not be specific enough to illustrate how the web application works.
-Therefore, in the current analysis, we define and present results from the "BP and hypertension" module, defined as follows.
+We performed statistical analyses to demonstrate core features of the application.
+We estimated the mean SBP for US adults by NHANES cycle, 1999-2000 through 2017-2020, with points showing the estimated means and error bars showing the 95% confidence intervals.
+We made bar charts presenting the age-adjusted prevalence of hypertension and the estimated number of US adults with hypertension.
+For age adjustment, we used the estimated age distribution of US adults from 1999 to 2020 as the standard (49.3%, 33.6%, 10.1% and 7.0% being 18 to 44, 45 to 64, 65 to 74 and ≥75 years of age, respectively).
+To illustrate how results from the application appear when saved in various formats (e.g., comma separated, portable document format, Microsoft Excel), we tabulated the estimated race/ethnicity distribution of US adults with and without hypertension, separately.
+We demonstrated stratification by estimating the prevalence of hypertension by NHANES cycle for US adults with and without chronic kidney disease (CKD).
+We showed the application’s ability to suppress output when statistical estimates are unstable by attempting to estimate the distribution of BP categories (SBP/DBP < 120/80 mm Hg, 120-129/<80 mm Hg, 130-139/80-89 mm Hg, 140-159/90-99 mm Hg and ≥ 160/100 mm Hg) among pregnant women in 2017-2020.
+We then showed that reliable estimates can be obtained for the distribution of BP categories among pregnant women by pooling NHANES cycles from 2009-2010 through 2017-2020.
 
-**Subpopulation:** Beginning with 107,622 US individuals who participated in NHANES 1999-2000 to 2017- March 2020, we restricted the subpopulation to adults ≥ 18 years of age.
-This exclusion was applied because statistics for hypertension and BP levels in children and adolescents are markedly different than for adults.
-We further restricted the subpopulation to participants who completed the in-home interview and study examination, with one or more SBP and DBP measurement, and who had data on self-reported antihypertensive medication use.
-After these exclusions were applied, the sub-population included data from 56,035 participants (Figure S1).
+We also illustrated how core features of the application can be combined to perform customized analyses.
+Specifically, we estimated the age-adjusted proportion of US adults with controlled BP by NHANES cycle among non-pregnant US adults with hypertension, overall and among those who self-reported taking antihypertensive medication.
+We also estimated the age-adjusted prevalence of resistant hypertension by NHANES cycle for US adults with hypertension taking antihypertensive medication and among US adults with hypertension taking ≥ 3 classes of antihypertensive medication.
+For age adjustment in the analysis of resistant hypertension, we set the age distribution for the standard population to represent US adults with hypertension from 1999 to 2020: 26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥75 years of age, respectively.
 
-**Variables** The primary variables from the BP and hypertension module are listed in Table 1, with full definitions provided in Supplemental Table 1.
-Briefly, mean SBP and DBP levels were computed over all available measurements for each participant.
-Oscillometric BP values were calibrated to the mercury device as described previously.
-Antihypertensive medication classes were defined using recommendations from the 2017 ACC/AHA BP guideline.
+# RESULTS
 
-*Validation of the web application*
+The web application successfully replicated results from prior studies.
+Full details on how to use the application to replicate these studies are available online.
+Following its validation, version 0.0.1 of the application was released and deployed in a publicly available server on **DATE TBD**.(add a citation to version 0.0.1 of the cardioStatsUSA package and a citation to the url where the shiny app lives)
 
-We validated the web application by using it in combination with our BP and hypertension module to reproduce results from two prior studies and one CDC report.
-A full description of the validation procedure is available.
+## Demonstration of core features
 
-*Statistical analysis*
+Among non-institutionalized US adults ≥ 18 years of age, the estimated mean SBP was XXX mm Hg (95% CI ) in 1999-2000, XXX mm Hg (95% CI) in 2009-2010 and XXX mm Hg (95% CI ) in 2017-2020 (**Figure 1**).
+The age-adjusted prevalence of hypertension was highest in 1999-2000 (47.9%), lowest in 2009-2010 (43.0%), and 44.7% in 2017-2020 (**Figure 2**).
+In 1999-2000, there were an estimated 89.8 million US adults with hypertension (**Figure 3**).
+The number of US adults with hypertension increased to 115.3 million in 2017-2020.
+In each NHANES cycle, the estimated age-adjusted prevalence of hypertension was higher among US adults with versus without CKD (**Figure S2**).
+In 2017-2020, a higher percentage of US adults with versus without hypertension were non-Hispanic Black (13.5% versus 9.3%) while a lower percentage of US adults with versus without hypertension were Hispanic (12.3% versus 18.1%) (**Table 2**).
+Among pregnant women, the distribution of BP categories could not be estimated reliably in 2017-2020 (**Figure S3; Panel A**), but was stable after pooling NHANES cycles from 2009-2010 through 2017-2020 (**Figure S3; Panel B**)
 
-We conducted a series of statistical analyses using the BP and hypertension module defined above.
-We estimated characteristics of the population, crude and age-adjusted mean SBP, and prevalence of hypertension for US adults by NHANES cycle.
-To demonstrate computation of results by group, we computed the age-adjusted prevalence of hypertension by NHANES cycle for individuals with versus without chronic kidney disease.
-For age adjustment in these analyses, we used the estimated age distribution of US adults from 1999 to 2020 as the standard (49.3%, 33.6%, 10.1% and 7.0% being 18 to 44, 45 to 64, 65 to 74 and ≥75 years of age, respectively).
-We also estimated the number of US adults with hypertension by NHANES cycle from 1999-2000 through 2017-2020.
+## Demonstration of customized analyses
 
-The age-adjusted distribution of the number of antihypertensive medication classes being taken was estimated for US adults with hypertension in each NHANES cycle, overall and among those who self-reported taking antihypertensive medication and had at least one antihypertensive drug class identified in their medication inventory.
-The age-adjusted proportion of US adults with controlled BP (i.e., SBP < 130 mm Hg and DBP < 80 mm Hg) was estimated by NHANES cycle among US adults with hypertension and who self-reported taking antihypertensive medication.
-The age-adjusted prevalence of resistant hypertension was estimated by NHANES cycle for US adults with hypertension taking antihypertensive medication and among those with hypertension taking ≥ 3 classes of antihypertensive medication.
-For age adjustment in these analyses, we set the age distribution for the standard population to represent US adults with hypertension from 1999 to 2020: 26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥75 years of age, respectively.
-We estimated the prevalence of hypertension in NHANES 2017-2020 by age group, sex, and race-ethnicity.
-Using the data pooling option in the application, we grouped the NHANES 2011-2020 cycles to estimate the proportion of pregnant women who had uncontrolled BP (i.e., SBP ≥ 130 mm Hg or DBP ≥ 80 mm Hg).
+The age-adjusted prevalence of BP control among US adults with hypertension was lowest in 1999-2000 (9.6%), highest in 2013-2014 (24.6%), and 21.7% in 2017-2020 (**Figure 4; panel A**).
+Among US adults who self-reported taking antihypertensive medication, 27.4%, 48.5%, and 42.8% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively (**Figure 4; panel B**).
+The age-adjusted prevalence of resistant hypertension among US adults who self-reported taking antihypertensive medication was lowest in 1999-2000 (13.7%), highest in 2005-2006 (20.4), and 16.2% in 2017-2020 (**Figure S4; panel A**).
+Among US adults who were taking three or more classes of antihypertensive medication, the prevalence of resistant hypertension was lowest in 2009-2010 (64.1%), highest in 2003-2004 (78.3%), and 67.9% in 2017-2020 (**Figure S4; panel B**).
 
-**RESULTS**
-
-The estimated crude mean SBP for all US adults was 123 mm Hg in 1999-2000, 120 mm Hg in 2009-2010, and 123 mm Hg in 2017-2020 (Figure 1; panel A).
-The estimated age-adjusted mean SBP was 124, 120, and 122 mm Hg in 1999-2000, 2009-2010 and 2017-2020, respectively (Figure 1; panel B).
-The estimated crude prevalence of hypertension among US adults was lowest in 2009-2010 (43.3%) and highest in 2017-2020 (46.6%) (Figure 2; panel A).
-After age-adjustment, the prevalence of hypertension was highest in 1999-2000 (47.9%), lowest in 2009-2010 and 2013-2014 (43.0%) and was 44.7% in 2017-2020.
-(Figure 2; panel B).
-The number of US adults with hypertension increased from 89.8 million in 1999-2000 to 115.3 million US adults in 2017-2020 (Figure 3).
-The age adjusted mean SBP was 140 in 1999-2000, 134 in 2009-2010, and 136 in 2017-2020 among US adults with hypertension and 117 in 1999-2000, 115 in 2009-2010, and 117 in 2017-2020 among US adults without hypertension (Figure S2).
-In each NHANES cycle, the age-adjusted estimated mean SBP was 18 to 23 mm Hg higher among US adults with versus without hypertension.
-In 2017-2020, a higher percentage of US adults with versus without hypertension were non-Hispanic Black (13.5% versus 9.3%) while a lower percentage of US adults with versus without hypertension were Hispanic (12.3% versus 18.1%) (Table 1).
-In each NHANES cycle, the estimated age-adjusted prevalence of hypertension among US adults with CKD was higher versus their counterparts without CKD (Figure S3).
-
-Among US adults with hypertension, 61.5% were not taking antihypertensive medication in 1999-2020 compared with 44.6% in 2017-2020.
+<!-- Among US adults with hypertension, 61.5% were not taking antihypertensive medication in 1999-2000 compared with 44.6% not taking antihypertensive medication in 2017-2020.
 (Figure 4; panel A).
 Among US adults who self-reported taking antihypertensive medication, the percentage taking four or more classes of antihypertensive medication increased from 3.5% in 1999-2000 to 7.6% in 2017-2020 (Figure 4; panel B).
-The age-adjusted estimated prevalence of BP control among US adults with hypertension was 10.2%, 25.8%, and 23.8% in 1999-2000, 2009-2010, and 2017-2020, respectively (Figure 5; panel A).
-For adults who self-report taking antihypertensive medication, 26.2%, 44.1%, and 43.3% had controlled BP in 1999-2000, 2009-2010, and 2017-2020, respectively (Figure 5; panel B).
+-->
 
-Among US adults with hypertension who self-reported taking antihypertensive medication, the prevalence of resistant hypertension was 13.7% in 1999-2000, 17.7% in 2009-2010, and 16.2% in 2017-2020 (Figure S4; panel A).
-Among those who were taking three or more classes of antihypertensive medication, the prevalence of resistant hypertension was 73.5%, 64.1%, and 67.9% in 1999-2000, 2009-2010, and 2017-2020, respectively (Figure S4; panel B).
+# DISCUSSION
 
-In 2017-2020, the prevalence of hypertension was higher at progressively older age (Figure S5),  among men versus women (Figure S6), and among non-Hispanic Black adults versus non-Hispanic White adults, non-Hispanic Asian adults, Hispanic adults and adults from other race/ethnic groups (Figure S7).
+In the current manuscript, we present a web application that allows the calculation of nationally representative estimates for BP, hypertension and antihypertensive medication-related outcomes using publicly available data from NHANES.
+Using this application, we generated crude and age-adjusted BP and hypertension statistics.
+Also, we generated statistics stratified by characteristics of US adults (e.g., CKD), and pooling multiple NHANES cycles to obtain more precise estimates when working with the small sub-group of pregnant women.
+Researchers, clinicians, and the general public can use the application to generate customized BP and hypertension statistics for US adults.
 
-**DISCUSSION**
+NHANES was designed to obtain nationally representative estimates of the health and nutrition status of non-institutionalized US adults and is an ideal data source to obtain statistics related to hypertension.
+SBP and DBP were measured following a standardized protocol by trained and certified staff.
+NHANES data are publicly available to download but working with these data requires understanding variable definitions and advanced programming and statistical knowledge.
+Challenges that users encounter when analyzing with NHANES include the need to download and merge multiple data files, even for a single NHANES cycle; combining multiple variables to create outcome definitions, which may require dealing with missing data and questionnaire skip patterns; harmonizing definitions across multiple NHANES cycles; and the analyses of complex survey design data.
+The web application that we present in the current manuscript was designed to reduce barriers to using NHANES so that these data can more fully inform public health decisions and future research related to hypertension.
 
-In the current manuscript, we describe the development of an open-access web-based application that allows the calculation of nationally representative estimates for a number of BP-related outcomes using publicly available data from NHANES.
-Using this app, we generated statistics for some of these BP-related outcomes over time, both crude and adjusted by age.
-We also generated statistics stratified by characteristics of US adults (e.g., CKD), and by pooling multiple NHANES cycles to obtain more precise estimates when working with small group sizes.
-Researchers, stakeholders, and the public in general interested in BP-related outcomes can now use the current app to accomplish a variety of purposes.
-
-NHANES is a program specifically designed to obtain nationally representative estimates of the health and nutrition status of non-institutionalized US adults.
-Therefore, NHANES represents the ideal data source to obtain valid and reliable statistic on BP-related outcomes among US adults.
-Although NHANES data are publicly available, working with these data is not exempt of challenges.
-Examples of challenges that may face working with NHANES data include the need to download multiple files, even for a single NHANES cycle; combining multiple NHANES variables to create the measurement of interest, which may require dealing with missing data and questionnaire skip patterns; harmonizing definitions across multiple NHANES cycles; and the analyses of complex survey design data.
-NHANES has been conducted over many cycles and some of the methods have been updated over time.
-For example, NHANES switched from using sphygmomanometer devices to oscillometric devices to measure BP in 2017-2018.
-NHANES data users need to be careful when analyzing data across multiple cycles.
-Reducing barriers to using NHANES data will facilitate the use of these data to inform BP-related public health decisions and future research.
-
-Several design decisions have been incorporated into the app.
-We required participants to have only a single SBP and DBP to be included.
-This approach is consistent with several analyses conducted by CDC investigators (cite pmid 26633197, 29155682).
+Several design decisions have been incorporated into the application.
+We required participants to have a single SBP and DBP to be included while some prior analyses required three SBP and DBP measurements.
+We chose an approach that is consistent with several analyses conducted by CDC investigators (cite pmid 26633197, 29155682).
 However, mean BP and the prevalence of high BP would be lower if we required multiple BP measurements (cite pmid 23126346).
-Any bias resulting from this decision is likely to be small as over 95% of adult NHANES participants within one SBP and DBP measurement had three SBP and DBP measurements.
-We re-calibrated the NHANES weights for the calculation of population counts (cite pmid 12500213).
+Any bias resulting from this decision is likely to be small as over 95% of adult NHANES participants with at least one SBP and DBP measurement had three SBP and DBP measurements.
+The application re-calibrates the NHANES weights for the estimation of population counts (cite pmid 12500213).
 When estimating population counts, participants missing data on SBP, DBP or antihypertensive medication use cannot have BP or hypertension-related outcomes.
-Weights were not recalibrated when estimating portions as participants missing data are removed from the numerator and denominator.
+Weights were not re-calibrated when estimating proportions as participants missing data are removed from the numerator and denominator.
 Medication classes were coded using generic names and the drug classes in the 2017 ACC/AHA BP guideline.
@@
 Many additional decisions were made regarding the definitions of variables, inclusion of study participants, and analytic approach.
-We sought to make decisions that would be acceptable to people who use the app and transparent to ensure the results could be described accurately.
+We sought to make decisions that would be widely acceptable and transparent to ensure the results could be described accurately.
 
-We were able to replicate several prior manuscripts using the app.
-However, some results could not be replicated.
-For example, a prior manuscript reported the prevalence of resistant hypertension defined by the 2017 ACC/AHA BP guideline to be 19.7% in 2009-2014.
-When estimated by the app, the prevalence of resistant hypertension was 17.7%, 16.5% and 17.2% in 2009-2010, 2011-2012, and 2013-2014, respectively (cite pmid 30580690).
+We were able to replicate several prior manuscripts using the application.
+However, we were not able to replicate some results.
+For example, a prior manuscript reported the prevalence of resistant hypertension defined by the 2017 ACC/AHA BP guideline to be 19.7% in 2009-2014(cite pmid 30580690).
+When estimated by the application, the prevalence of resistant hypertension over this time period was X%.
 The difference in the prevalence estimates can be attributed to the approach used to categorize medication into classes.
-For example, the prior publication used Lexicon Plus®, which counted spironolactone as two drug classes, a potassium-sparing diuretic and an aldosterone antagonist.
-Additionally, Lexicon Plus®includes Sotalol as a beta blocker and nitroglycerine as a direct vasodilator and we did not include these drugs as antihypertensive medications as they are not listed in the 2017 ACC/AHA BP guideline.
-The differences in results between the app and this previously published manuscript emphasize two important points.
-First, while we encourage users to check results when using the app with prior publications, it may not be possible to always generate results that are identical to prior publications.
-Second, users should be aware of the choices made in defining variables.
+The prior publication used Lexicon Plus®, which counted spironolactone as two drug classes, a potassium-sparing diuretic and an aldosterone antagonist.
+We included spironolactone as a single drug class, an aldosterone antagonist.
+Additionally, Lexicon Plus® includes Sotalol as a beta blocker and nitroglycerine as a direct vasodilator and we did not include these drugs as antihypertensive medications as they are not listed in the 2017 ACC/AHA BP guideline.
+The differences in results between the application and this previously published manuscript emphasize that it may not be possible to always generate results that are identical to prior publications.
+Users should be aware of the choices made in defining variables as this may affect the statistical estimates generated using the application.
 
-The web application has a number of complementary features that allow users to complete various analysis tasks.
-In particular, users can generate descriptive summaries of the US population spanning over 20 years, including summaries of characteristics and cardiometabolic outcomes.
-The flexibility of inclusion criteria paired with the large set of outcomes and stratifying variables available allow an enormous number of hypotheses to be explored.
-Outputs from the app can be included in scientific proposals, presentations, or articles, citing the application's publicly available documentation to contextualize results.
-With the ability to download summary data from the application, users can visualize or tabulate results with aesthetics and software of their choosing.
+The application has a number of features that can be used in combination to create highly customized statistics.
+Variables can be analyzed as outcomes or used to stratify results or restrict analyses to subsets of participants.
+Combining the variables, tens of thousands of unique statistics can be estimated.
+In addition, users can generate statistics for US adults over 22 years, pooling results or stratifying the data.
+All results from the application can be saved as an image and included in scientific proposals or presentations.
+In addition, the ability to download summary data from the application allows users to further customize their results in tabular or graphical formats.
 The web application can also be used for educational purposes, as teachers can use it to analyze health outcomes for US adults interactively with students.
 
-With an open-source software license and publicly available documentation and data, the web application we developed is designed to be extended.
-The Shiny interface and R codes supporting this application can support general sub-populations of NHANES participants, such as a sub-population of participants with cholesterol data or a sub-population of pediatric survey participants.
-As the current analysis demonstrates creation and analysis of a sub-population of participants with BP measurements, others can use it as a template for developing extensions of the web application that engage with new sub-populations.
-In addition to including new sub-populations of NHANES participants, the application may be extended to include inferential statistics and hypothesis testing, and may also incorporate new cycles of NHANES data as they become available.
+With an open-source software license and publicly available documentation and data, the web application we developed can be extended.
+The current manuscript described a set of hypertension variables and a sub-population of NHANES participants with data on BP and antihypertensive medication use.
+The application’s interface and its supporting R package allows additional variables to be added.
+In addition, the application may be extended to include data from additional NHANES cycles as they become publically available.
+
+## Strengths/limitations
+
+This study has a number of strengths.
+We used NHANES data, which are publicly available, rigorously collected, and allow for estimation of nationally representative statistics.
+Also, we leveraged open-source software to ensure that the application developed in the current study is transparent, able to be extended, and freely available.
+In addition to the application, we developed online resources with tutorials for new users and extensive documentation to contextualize results.
+This study also has several limitations.
+First, NHANES participants had their mean BP measured during a single visit, and BP guidelines recommend obtaining mean BP using at least two BP values measured on separate days.
+Second, the response rate for NHANES has declined from 1999-2000 through 2017-2020.
+Effects of this decline are unclear.
+
+## Conclusions
+
+We developed a web-based application for analysis of hypertension outcomes among non-institutionalized adults living in the US from 1999-2000 through 2017-2020.
+The application is publicly available and produces valid, transparent, and reproducible results.
 
@@
 
-**Table 1**: Variables included in the blood pressure and hypertension module.
+### Table 1: Blood pressure, hypertension and antihypertensive medication variables that may be analyzed using the web application.
 
@@
               col_keys = c('class', 'label')) %>% 
- set_header_labels(label = "Variables included in module") %>% 
+ set_header_labels(label = "Variable") %>% 
  theme_box() %>% 
@@
 
-**Table 2**:	Race/ethnicity distribution of the population with and without hypertension in 2017-2020
+### Table 2:	Race/ethnicity distribution of US adults with and without hypertension in 2017-2020
 
@@
 library(table.glue)
+
+rspec <- round_spec() %>% 
+ round_using_decimal(digits = 1)
 
@@
                   time_values = 'most_recent') %>%
- mutate(across(where(is.numeric), table_value)) %>% 
+ mutate(
+  across(where(is.numeric), 
+         .fns = table_value, 
+         rspec = rspec)
+ ) %>% 
  filter(statistic == 'percentage') %>% 
@@
  flextable() %>% 
+ set_header_labels(
+  svy_year = 'NHANES cycle',
+  htn_accaha = 'Hypertension',
+  demo_race = 'Race/Ethnicity',
+  estimate = 'Estimate',
+  std_error = 'Standard error',
+  ci_lower = 'CI lower',
+  ci_upper = 'CI upper',
+  n_obs = 'N observations'
+ ) %>% 
  theme_box() %>% 
@@
 
-**Figure 1** Mean systolic blood pressure for US adults by calendar year
+### Figure 1 Mean systolic blood pressure for US adults by calendar year
 
-A. Without age adjustment
 ```{r, fig.height=3.75}
@@
                        time_values = 'all', 
-                       geom = 'bar')
-
-age_wts_overall <- c(49.3, 33.6, 10.1, 7)
-
-p2 <- nhanes_visualize(data = nhanes_data,
-                       key = nhanes_key,
-                       outcome_variable = 'bp_sys_mean',
-                       time_values = 'all', 
-                       age_wts = age_wts_overall, 
-                       geom = 'bar')
+                       geom = 'points',
+                       size_point = 25,
+                       size_error = 1)
 
@@
 ```
-
-B. With age adjustment through direct standardization.
-```{r, fig.height=3.75}
-fig_print(p2)
-```
+Dots represent mean systolic blood pressure. Vertical lines represent the 95% confidence interval.
 
@@
 
-**Figure 2** Prevalence of hypertension for US adults by calendar year
+### Figure 2 Age-adjusted prevalence of hypertension for US adults by calendar year
 
-A. Without age adjustment
 ```{r, fig.height=3.75}
-
-p1 <- nhanes_visualize(data = nhanes_data,
-                       key = nhanes_key,
-                       outcome_variable = 'htn_accaha',
-                       time_values = 'all', 
-                       geom = 'bar')
-
-p2 <- nhanes_visualize(data = nhanes_data,
-                       key = nhanes_key,
-                       outcome_variable = 'htn_accaha',
-                       time_values = 'all', 
-                       age_wts = age_wts_overall, 
-                       geom = 'bar')
-
-fig_print(p1)
-
-```
-
-B. With age adjustment through direct standardization.
-```{r, fig.height=3.75}
-fig_print(p2)
-```
-
-```{r}
 
@@
                  time_values = 'all', 
-                 group_variable = 'cc_ckd',
                  age_wts = age_wts_overall, 
-                 geom = 'point',
-                 size_point = 22,
-                 size_error = 4) %>% 
+                 geom = 'bar') %>% 
  fig_print()
@@
 ```
+Age adjustment was performed through direct standardization
 
@@
 
-**Figure 3** Number of US adults with hypertension
+### Figure 3 Number of US adults with hypertension
 
@@
 
-**Figure 4** Distribution of the number of antihypertensive medication classes being taken among those with hypertension by calendar period
+### Figure 4 Age-adjusted prevalence of blood pressure control by calendar year.
 
@@
                        key = nhanes_key,
-                       outcome_variable = 'bp_med_n_class',
-                       subset_variables = list('htn_accaha'),
-                       subset_values = list("Yes"),
-                       time_values = 'all', 
-                       geom = 'bar')
-
-p2 <- nhanes_visualize(data = nhanes_data,
-                       key = nhanes_key,
-                       outcome_variable = 'bp_med_n_class',
-                       subset_variables = list('htn_accaha', 'bp_med_use'),
-                       subset_values = list("Yes", "Yes"),
-                       time_values = 'all', 
-                       geom = 'bar')
-
-fig_print(p1)
-
-```
-
-B. Among those with hypertension who self-report taking antihypertensive medication 
-```{r, fig.height=3.75}
-fig_print(p2)
-```
-
-\newpage
-
-**Figure 5** Blood pressure control by calendar year
-
-A. Age-adjusted among those with hypertension
-```{r, fig.height=3.75}
-
-age_wts_htn <- c(26.4, 43.5, 17.0, 13.2)
-
-p1 <- nhanes_visualize(data = nhanes_data,
-                       key = nhanes_key,
                        outcome_variable = 'bp_control_accaha',
-                       subset_variables = list('htn_accaha'),
-                       subset_values = list("Yes"),
+                       subset_calls = list("htn_accaha" = "Yes"),
                        age_wts = age_wts_htn,
@@
                        outcome_variable = 'bp_control_accaha',
-                       subset_variables = list('htn_accaha',
-                                               'bp_med_use'),
-                       subset_values = list("Yes", 
-                                            "Yes"),
+                       subset_calls = list("htn_accaha" = "Yes",
+                                           "bp_med_use" = "Yes"),
                        age_wts = age_wts_htn,
@@
 
-B. Age-adjusted among those who self-report taking antihypertensive medication 
+B. Among those who self-report taking antihypertensive medication 
 ```{r, fig.height=3.75}
@@
 ```
+Age adjustment was performed through direct standardization
+
+<!-- \newpage -->
+
+<!-- **Figure 4** Distribution of the number of antihypertensive medication classes being taken among those with hypertension by calendar period -->
+
+<!-- A. Among those with hypertension -->
+<!-- ```{r, fig.height=3.75} -->
+
+<!-- p1 <- nhanes_visualize(data = nhanes_data, -->
+<!--                        key = nhanes_key, -->
+<!--                        outcome_variable = 'bp_med_n_class', -->
+<!--                        subset_calls = list("htn_accaha" = "Yes"), -->
+<!--                        time_values = 'all', -->
+<!--                        geom = 'bar') -->
+
+<!-- p2 <- nhanes_visualize(data = nhanes_data, -->
+<!--                        key = nhanes_key, -->
+<!--                        outcome_variable = 'bp_med_n_class', -->
+<!--                        subset_calls = list("htn_accaha" = "Yes", -->
+<!--                                            "bp_med_use" = "Yes"), -->
+<!--                        time_values = 'all', -->
+<!--                        geom = 'bar') -->
+
+<!-- fig_print(p1) -->
+
+<!-- ``` -->
+
+<!-- B. Among those with hypertension who self-report taking antihypertensive medication -->
+<!-- ```{r, fig.height=3.75} -->
+<!-- fig_print(p2) -->
+<!-- ``` -->
 
@@
 
-**Table S1**
+### Table S1: All variables included in the web-based application.
 
@@
 
-**Figure S1** Application of inclusion and exclusion criteria for the application of NHANES hypertension data.
+### Figure S1 Application of inclusion and exclusion criteria for the application of NHANES hypertension data.
 
@@
  theme_box() %>% 
+ add_header_row(values = c("", "NHANES cycle"), colwidths = c(2, 10)) %>% 
  width(width = 0.8) %>% 
@@
 
-**Figure S2**:	Age-adjusted mean systolic blood pressure among people with and without hypertension
+### Figure S2: Prevalence of hypertension among US adults with and without chronic kidney disease
 
@@
                  key = nhanes_key,
-                 outcome_variable = 'bp_sys_mean',
+                 outcome_variable = 'htn_accaha',
                  time_values = 'all',
-                 group_variable = 'htn_jnc7',
-                 age_wts = age_wts_overall, 
+                 group_variable = 'cc_ckd',
                  geom = 'bar') %>% 
@@
 
-**Figure S3**: Age-adjusted prevalence of hypertension among US adults with and without chronic kidney disease
+### Figure S3: Distribution of blood pressure categories among pregnant women.
+
+A. In 2017-2020
+```{r}
+
+nhanes_visualize(data = nhanes_data,
+                 key = nhanes_key,
+                 outcome_variable = 'bp_cat_meds_excluded',
+                 time_values = 'most_recent',
+                 subset_calls = list("demo_pregnant" = "Yes"),
+                 geom = 'bar') %>% 
+ fig_print()
+
+```
+
+B. In 2009-2010 through 2017-2020
+```{r}
+
+nhanes_visualize(data = nhanes_data,
+                 key = nhanes_key,
+                 outcome_variable = 'bp_cat_meds_excluded',
+                 time_values = 'last_5',
+                 pool = TRUE,
+                 subset_calls = list("demo_pregnant" = "Yes"),
+                 geom = 'bar') %>% 
+ fig_print()
+
+```
+
 \newpage
 
-**Figure S4**: Prevalence of resistant hypertension by calendar year
+### Figure S4: Age-adjusted prevalence of resistant hypertension by calendar year.
 
-A. Among those with hypertension who self report taking antihypertensive medication 
+A. Among those who self report taking antihypertensive medication 
 ```{r, fig.height=3.75}
@@
                        outcome_variable = 'htn_resistant_accaha',
-                       subset_variables = list('htn_accaha', 'bp_med_use'),
-                       subset_values = list("Yes", "Yes"),
+                       subset_calls = list("bp_med_use" = "Yes"),
                        time_values = 'all', 
@@
 
-p2 <- nhanes_visualize(data = nhanes_data,
+p2 <- nhanes_visualize(
+ data = nhanes_data,
  key = nhanes_key,
  outcome_variable = 'htn_resistant_accaha',
-                       subset_variables = list('htn_accaha', 
-                                               'bp_med_use',
-                                               'bp_med_n_class'),
-                       subset_values = list("Yes", 
-                                            "Yes",
-                                            c('Three', 'Four or more')),
+ subset_calls = list("bp_med_use" = "Yes",
+                     "bp_med_n_class" = c('Three', 'Four or more')),
  time_values = 'all', 
-                       geom = 'bar')
+ geom = 'bar'
+)
 
@@
 
-B. Among those with hypertension who self report and are taking 3+ classes of antihypertensive medication
+B. Among those who self report taking antihypertensive medication and are taking 3 or more classes.
 ```{r, fig.height=3.75}
@@
 ```
-\newpage
-
-**Figure S5** Prevalence of hypertension by age categories in 2017-2020
-
-```{r}
-nhanes_visualize(data = nhanes_data,
-                 key = nhanes_key,
-                 outcome_variable = 'htn_accaha',
-                 group_variable =  'demo_age_cat',
-                 time_values = 'most_recent',
-                 geom = 'bar') %>% 
- fig_print()
-```
-\newpage
-
-**Figure S6** Prevalence of hypertension by sex in 2017-2020
-
-```{r}
-nhanes_visualize(data = nhanes_data,
-                 key = nhanes_key,
-                 outcome_variable = 'htn_accaha',
-                 group_variable = 'demo_gender',
-                 time_values = 'most_recent',
-                 geom = 'bar') %>% 
- fig_print()
-```
+Age adjustment was performed through direct standardization
 
@@
 
-**Figure S7** Prevalence of hypertension by race/ethnicity in 2017-2020
+<!-- **Figure S5** Prevalence of hypertension by age categories in 2017-2020 -->
 
-```{r}
-nhanes_visualize(data = nhanes_data,
-                 key = nhanes_key,
-                 outcome_variable = 'htn_accaha',
-                 group_variable = 'demo_race',
-                 time_values = 'most_recent',
-                 geom = 'bar') %>% 
- fig_print()
-```
+<!-- ```{r} -->
+<!-- nhanes_visualize(data = nhanes_data, -->
+<!--                  key = nhanes_key, -->
+<!--                  outcome_variable = 'htn_accaha', -->
+<!--                  group_variable =  'demo_age_cat', -->
+<!--                  time_values = 'most_recent', -->
+<!--                  geom = 'bar') %>%  -->
+<!--  fig_print() -->
+<!-- ``` -->
+<!-- \newpage -->
 
-\newpage
+<!-- **Figure S6** Prevalence of hypertension by sex in 2017-2020 -->
 
-**Figure S8** Prevalence of uncontrolled blood pressure among pregnant women from 2011-2012 through 2017-2020.
+<!-- ```{r} -->
+<!-- nhanes_visualize(data = nhanes_data, -->
+<!--                  key = nhanes_key, -->
+<!--                  outcome_variable = 'htn_accaha', -->
+<!--                  group_variable = 'demo_gender', -->
+<!--                  time_values = 'most_recent', -->
+<!--                  geom = 'bar') %>%  -->
+<!--  fig_print() -->
+<!-- ``` -->
 
-```{r}
+<!-- \newpage -->
 
-nhanes_visualize(data = nhanes_data,
-                 key = nhanes_key,
-                 outcome_variable = 'bp_uncontrolled_accaha',
-                 time_values = c("2011-2012", 
-                                 "2013-2014",
-                                 "2015-2016",
-                                 "2017-2020"),
-                 pool = TRUE,
-                 subset_variables = list('demo_pregnant'),
-                 subset_values = list("Yes"),
-                 geom = 'bar') %>% 
- fig_print()
+<!-- **Figure S7** Prevalence of hypertension by race/ethnicity in 2017-2020 -->
 
-```
+<!-- ```{r} -->
+<!-- nhanes_visualize(data = nhanes_data, -->
+<!--                  key = nhanes_key, -->
+<!--                  outcome_variable = 'htn_accaha', -->
+<!--                  group_variable = 'demo_race', -->
+<!--                  time_values = 'most_recent', -->
+<!--                  geom = 'bar') %>%  -->
+<!--  fig_print() -->
+<!-- ``` -->
+
+<!-- \newpage -->
+
+<!-- **Figure S8** Prevalence of uncontrolled blood pressure among pregnant women from 2011-2012 through 2017-2020. -->
+
+<!-- ```{r} -->
+
+<!-- nhanes_visualize(data = nhanes_data, -->
+<!--                  key = nhanes_key, -->
+<!--                  outcome_variable = 'bp_cat_meds_excluded', -->
+<!--                  time_values = c("last_5"), -->
+<!--                  pool = TRUE, -->
+<!--                  subset_calls = list("demo_pregnant" = "Yes"), -->
+<!--                  geom = 'bar') %>%  -->
+<!--  fig_print() -->
+
+<!-- ``` -->
 

### 2022-10-21 rmd/nhanes_design_subset.Rmd
big update
added CI options for percentage, tweaked api, fixed many small things

```{r child = 'aaa.Rmd'}

```

# Examples

First we create an NHANES design object

```{r child='make_simple_nhanes_design.Rmd'}

```

Subset calls work as they do in many other `subset` functions:

```{r}

ds_men <- nhanes_design_subset(ds, demo_gender == 'Men')

```

To verify `ds_men` has been subsetted, you can do something like this:

```{r}
table(ds_men$design$variables$demo_gender)
```

and you can also find the subsetted indices here:

```{r}
ds_men$subset_rows
```

Why does the design object save this information? In case you update it in the future and subsetting operations need to be repeated.

## Only one subset is allowed

Due to the way `nhanes_design` objects are handled, it is not a good idea to subset them more than once. In fact, it is currently not allowed. If this becomes problematic, please open an issue on GitHub and let us know!

```{r, error=TRUE}
nhanes_design_subset(ds_men, demo_age_years > 40)
```

### 2022-10-21 R/cardioStatsUSA-package.R

@@ -13,7 +13,6 @@
 #' @importFrom DT datatable DTOutput formatRound
 #' @importFrom rlang parse_expr %||%
 #' @importFrom plotly plot_ly add_trace layout
-#' @importFrom arules discretize
 #' @importFrom purrr map map2 map_dfr map2_chr map2_dfr is_empty discard
 #' @importFrom dplyr filter mutate select pull across
 #' @importFrom dplyr if_else all_of .data recode left_join

### 2022-10-21 R/check.R

@@ -10,6 +10,32 @@ check_nobs <- function(x, min_size = 30){
 
 }
 
+check_vars_in_data <- function(..., data_names, data_label) {
+
+ .dots <- rlang::dots_list(...)
+
+ for(i in seq_along(.dots)){
+
+  if(!is.null(.dots[[i]])){
+
+   variable_role <- names(.dots)[i]
+   variable_name <- .dots[[i]]
+
+   if( !(variable_name %in% data_names) ){
+
+    stop(
+     variable_role, " ", variable_name,
+     " was not found in ", data_label,
+     call. = FALSE
+    )
+
+   }
+
+  }
+
+ }
+
+}
 
 # nocov start
 check_group_counts <- function(x, min_size = 5){

### 2022-10-21 R/get_outcome_stats.R

@@ -3,10 +3,17 @@ get_outcome_stats <- function(outcome_type) {
 
  switch(
   outcome_type,
-  'ctns' = c(Mean = 'mean', Quantiles = 'quantile'),
-  'catg' = c(Percentage = 'percentage', Count = 'count'),
-  'bnry' = c(Percentage = 'percentage', Count = 'count'),
-  'intg' = c(Percentage = 'percentage', Count = 'count')
+  'ctns' = c(Mean = 'mean',
+             Quantiles = 'quantile'),
+  'catg' = c(Percentage = 'percentage',
+             `Percentage (Korn and Graubard CI)` = 'percentage_kg',
+             Count = 'count'),
+  'bnry' = c(Percentage = 'percentage',
+             `Percentage (Korn and Graubard CI)` = 'percentage_kg',
+             Count = 'count'),
+  'intg' = c(Percentage = 'percentage',
+             `Percentage (Korn and Graubard CI)` = 'percentage_kg',
+             Count = 'count')
  )
 
 }

### 2022-10-21 R/get_var_info.R

@@ -3,14 +3,35 @@
 
 #' Title
 #'
-#' @param key
-#' @param x
-#' @param field
+#' @param x \[character(1)\]
 #'
-#' @return
+#' a variable name that is present in the `variable` column of `key`
+#'
+#' @param key \[data.frame\]
+#'
+#' `r describe_nhanes_key_input()`
+#'
+#' @param field \[character(1)\]
+#'
+#' the column name in `nhanes_key` to get information from.
+#'
+#' @return a character value
+#'
 #' @export
 #'
 #' @examples
+#'
+#' # the low-level working function
+#' get_var_info('bp_sys_mean', key = nhanes_key, field = 'class')
+#'
+#' # the convenience functions
+#' get_variable_class('bp_sys_mean')
+#' get_variable_label('bp_sys_mean')
+#' get_variable_source('bp_sys_mean')
+#' get_variable_description('bp_sys_mean')
+#' get_variable_module('bp_sys_mean')
+#'
+
 get_var_info <- function(x, key, field){
  if(missing(key)) key <- cardioStatsUSA::nhanes_key
  key[[field]][key$variable == x]
@@ -33,4 +54,4 @@ get_variable_description <- function(x, key) get_var_info(x, key, 'description')
 get_variable_type <- function(x, key) get_var_info(x, key, 'type')
 #' @rdname get_var_info
 #' @export
-get_variable_subpop <- function(x, key) get_var_info(x, key, 'subpop')
+get_variable_module <- function(x, key) get_var_info(x, key, 'module')

### 2022-10-21 R/jsc_write_cpanel.R

@@ -14,8 +14,8 @@
 jsc_write_cpanel <- function(key_data, var_type, var_role){
 
  variable_names <- key_data %>%
-  filter(type == var_type, .data[[var_role]]) %>%
-  pull(variable)
+  .[type == var_type & .[[var_role]] == TRUE] %>%
+  .[['variable']]
 
  if(is_empty(variable_names))
   return("false")

### 2022-10-21 R/nhanes_design_standardize.R

@@ -1,21 +1,33 @@
 
 
-#' Title
+#' Standardize an NHANES design
 #'
-#' @param x
-#' @param standard_variable
-#' @param standard_weights
+#' @param x \[nhanes_design\]
 #'
-#' @return
+#' `r document_nhanes_design()`
+#'
+#' @param standard_variable \[character(1)\]
+#'
+#' The name of the variable used to create standardization groups.
+#'   The default is to use `demo_age_cat`, which leads to age
+#'   standardization.
+#'
+#' @param standard_weights \[numeric(n)\]
+#'
+#' The proportionate weights for each group defined by the standard variable.
+#'   The number of weights should equal the number of groups defined by
+#'   `standard_variable` and all weights must be >0.
+#'
+#' @return an [nhanes_design] object.
+#'
 #' @export
 #'
-#' @examples
 nhanes_design_standardize <- function(x,
                                       standard_variable = 'demo_age_cat',
                                       standard_weights = c(0.155,
-                                                       0.454,
-                                                       0.215,
-                                                       0.177)){
+                                                           0.454,
+                                                           0.215,
+                                                           0.177)){
 
  by_vars <- c(x$time$variable,
               x$group$variable,

### 2022-10-21 R/nhanes_design_subset.R

@@ -1,13 +1,25 @@
-#' Title
+#' Subset an NHANES design object
 #'
-#' @param x
-#' @param subset
+#' @param x \[nhanes_design\]
+#'
+#' `r document_nhanes_design()`
+#'
+#' @param subset \[expression\]
+#'
+#' An expression indicating what rows to keep.
+#'   Missing values are taken as false.
+#'
 #' @param ...
 #'
-#' @return
+#' Currently not used.
+#'
+#' @return an [nhanes_design] object.
+#'
 #' @export
 #'
-#' @examples
+#' @includeRmd rmd/nhanes_design_subset.Rmd
+
+
 nhanes_design_subset <- function(x, subset){
 
  if(missing(subset)) return(x)

### 2022-10-21 R/nhanes_design_viz.R

@@ -3,7 +3,10 @@
 #' Analyze the prevalence, mean, or quantiles of an outcome
 #'   over time.
 #'
-#' @param x an `nhanes_design` object
+#' @param x \[nhanes_design\]
+#'
+#' `r document_nhanes_design()`
+#'
 #' @param statistic_primary (character) the statistic that defines the
 #'   geometric objects in the plot.
 #' @param geom (character) can be 'bar' or 'points'
@@ -23,6 +26,7 @@
 
 nhanes_design_viz <- function(x,
                               statistic_primary = NULL,
+                              title = NULL,
                               geom = 'bar',
                               reorder_cats=FALSE,
                               width = NULL,
@@ -323,9 +327,16 @@ nhanes_design_viz <- function(x,
 
   if(stacked_and_pooled) {
 
+   if(outcome_type == 'catg'){
+    # to make a little more room for the second row of the title
+    yaxis$range <- list(0, 110)
+   }
+
    # this occurs when there are no title addons to worry about
    if(title_addon[1] == ""){
-    title_addon <- levels(data_fig[[time_variable]])
+    # data_fig is always a list of data frames at this point,
+    # so just access the first one.
+    title_addon <- levels(data_fig[[1]][[time_variable]])
    } else {
     # if there are title addons, split them across each year
     title_addon <- c(levels(data_fig[[time_variable]]),title_addon) %>%
@@ -343,7 +354,23 @@ nhanes_design_viz <- function(x,
    xaxis$tickmode <- 'array'
   }
 
-  fig_title <-  paste(c(outcome_label, title_addon), collapse = '<br>')
+  fig_title <- NULL
+
+  if(is.null(title)){
+
+   fig_title <- list(
+    text = paste(c(outcome_label, title_addon), collapse = '<br>'),
+    x = 0.5
+   )
+
+  } else if (title != ""){
+
+   fig_title <- list(
+    text = title,
+    x = 0.5
+   )
+
+  }
 
   xaxis$title = ifelse(stacked_and_pooled, group_label %||% "", time_label)
 
@@ -359,10 +386,7 @@ nhanes_design_viz <- function(x,
 
   output[[j]] <- fig |>
    layout(
-    title = list(
-     text = fig_title,
-     x = 0.5
-    ),
+    title = fig_title,
     font = list(size = 16),
     margin = list(
      t = 100,

### 2022-10-21 R/nhanes_visualize.R

@@ -4,23 +4,55 @@
 #'
 #' @inheritParams nhanes_summarize
 #'
-#' @param statistic_primary (character) the statistic that defines the
-#'   geometric objects in the plot.
-#' @param geom (character) can be 'bar' or 'points'
-#' @param reorder_cats (logical) whether to re-order the categorical exposure
+#' @param statistic_primary \[character(1)\]
+#'
+#'   the statistic that defines the geometric objects in the plot.
+#'
+#' @param geom \[character(1)\]
+#'
+#' What type of plot to create. Valid options are
+#'
+#' - `'bar'` A bar plot will be created
+#'
+#' - `'point'` A scatter plot will be created
+#'
+#' @param reorder_cats \[logical(1)\]
+#'
+#' whether to re-order the categorical exposure
 #'   variable so that its levels are shown in increasing order by the expected
 #'   outcome.
-#' @param width width of the plot, in pixels
-#' @param height height of the plot, in pixels
+#'
+#' @param width
+#'
+#' width of the plot, in pixels
+#'
+#' @param height
+#'
+#' height of the plot, in pixels
+#'
+#' @param size_point
+#'
+#' the size of points in the plot (only relevant if `geom = 'point'`)
+#'
+#' @param size_error
+#'
+#' the size of error bars in the plot (only relevant if `geom = 'point'`)
 #'
 #' @return a plotly visualization
 #'
 #' @export
 #'
+#' @examples
+#'
+#' # Plotly objects do not render for example R code. Please see vignettes
+#' # for examples of nhanes_visualize. TODO: ADD LINK
+#'
+#'
 nhanes_visualize <- function(data,
                              key,
                              outcome_variable,
                              outcome_quantiles = NULL,
+                             outcome_stats = NULL,
                              group_variable = NULL,
                              group_cut_n = NULL,
                              group_cut_type = NULL,
@@ -31,6 +63,7 @@ nhanes_visualize <- function(data,
                              subset_calls = list(),
                              age_wts = NULL,
                              statistic_primary = NULL,
+                             title = NULL,
                              geom = 'bar',
                              reorder_cats = FALSE,
                              width = NULL,
@@ -40,10 +73,13 @@ nhanes_visualize <- function(data,
 
  if(geom %in% c('point', 'points')) geom <- 'scatter'
 
+ outcome_stats <- outcome_stats %||% statistic_primary
+
  smry <- nhanes_summarize(data = data,
                           key = key,
                           outcome_variable = outcome_variable,
[…]

### 2022-10-21 R/plotly_viz_make_hover.R

@@ -34,7 +34,7 @@ plotly_viz_make_hover <- function(data, stat_all, group, group_label) {
     stat_show
    ),
    stat_label = glue(
-    "{str_to_title(statistic)}: {stat_show}"
+    "{stat_recode(statistic)}: {stat_show}"
    )
   ) %>%
   select(-all_of(numeric_cols),
@@ -47,5 +47,15 @@ plotly_viz_make_hover <- function(data, stat_all, group, group_label) {
   unite(col = 'hover', !!!stat_cols, sep = '<br>') %>%
   mutate(hover = glue("{group_label}{label_sep}{hover}")) %>%
   as.data.table()
+
+}
+
+
+stat_recode <- function(stat){
+
+ fcase(
+  stat == 'percentage_kg', 'Percentage (Korn & Graubard CI)',
+  rep(TRUE, length(stat)), stringr::str_to_title(stat)
+ )
 
 }

### 2022-10-21 R/roxygen_document.R

@@ -1,19 +1,108 @@
 
 # nocov start
 
+describe_age_wts <- function(age_wts, age_lvls=NULL){
+
+ if(is.null(age_lvls))
+  age_lvls <- levels(cardioStatsUSA::nhanes_data$demo_age_cat)
+
+ .age_wts <- age_wts %>%
+  paste0("%") %>%
+  glue::glue_collapse(sep = ', ', last = ', and ')
+
+ .age_lvls <- glue::glue_collapse(age_lvls, sep = ', ', last = ', and ')
+
+ glue::glue("{.age_wts} being {.age_lvls} years of age, respectively")
+
+}
+
+describe_module_sample <- function(type = 'htn'){
+ switch(
+  type,
+  "htn" = "There were 107,622 NHANES 1999-2000 to 2017-March 2020 participants. We restricted the the dataset to adults >=18 years of age. This exclusion was applied because statistics for BP levels and hypertension in children and adolescents are markedly different than for adults. We further restricted the population to participants who completed the in-home interview and study examination, with one or more SBP and DBP measurement, and who had data on self-reported antihypertensive medication use."
+ )
+}
+
+tabulate_module_sample <- function(type = 'htn',
+                                   subset_calls = NULL,
+                                   subset_labs = NULL){
+
+ kable_data <- switch(
+  type,
+  "htn" = data.frame(
+   check.names = FALSE,
+   Inclusion = c(
+    "Participants",
+    ">=18 years old",
+    "Completed interview and examination",
+    "Had SBP and DBP measurements",
+    "Had self-reported information on antihypertensive medication"
+   ),
+    Overall = c(107622, 63041, 59799, 56286, 56035),
+   `1999-2000` = c(9965, 5448, 4976, 4755, 4694),
+   `2001-2002` = c(11039, 5993, 5592, 5251, 5184),
+   `2003-2004` = c(10122, 5620, 5303, 4902, 4838),
+   `2005-2006` = c(10348, 5563, 5334, 5028, 5015),
+   `2007-2008` = c(10149, 6228, 5995, 5670, 5665),
+   `2009-2010` = c(10537, 6527, 6360, 6053, 6043),
+   `2011-2012` = c(9756, 5864, 5615, 5436, 5337),
+   `2013-2014` = c(10175, 6113, 5924, 5700, 5694),
[…]

### 2022-10-21 R/svy_stat_suppress.R

@@ -1,11 +1,19 @@
-#' .. content for \description{} (no empty lines) ..
+
+#' FILL IN
 #'
-#' .. content for \details{} ..
+#' @param dt FILL IN
+#' @param type FILL IN
+#' @param stat FILL IN
+#' @param n_obs FILL IN
+#' @param n_psu FILL IN
+#' @param n_strata FILL IN
 #'
-#' @title
-#' @param n_obs
-#' @param n_psu
-#' @param n_strata
+#' @return FILL IN
+#'
+#' @export
+#'
+#' @noRd
+#'
 svy_stat_suppress <- function(dt, type, stat, n_obs, n_psu, n_strata) {
 
  # fails <- is.nan(dt$std_error) | dt$std_error == 0
@@ -138,7 +146,7 @@ svy_stat_suppress_bnry <- function(dt, n_obs, n_psu, n_strata){
 
  ci_width <- with(dt, (ci_upper - ci_lower) / 100)
 
- if( any(is.nan(ci_width)) ) rse[is.nan(ci_width)] <- Inf
+ if( any(is.nan(ci_width)) ) ci_width[is.nan(ci_width)] <- Inf
 
  # Large absolute confidence interval width
  # If the absolute confidence interval width is greater than or

### 2022-10-21 R/svy_stat_tidy_prop.R

@@ -1,61 +1,0 @@
-
-svy_stat_tidy_proportion <- function(x, outcome, ...){
-
- map2_dfr(
-  x,
-  names(x),
-  ~ {
-
-   x_ci <- confint(.x)
-
-   data.table(outcome = outcome,
-              level = .y,
-              statistic = c('proportion'),
-              estimate = as.numeric(.x),
-              ci_lower = x_ci[, 1],
-              ci_upper = x_ci[, 2]) %>%
-    .[, estimate := estimate * 100] %>%
-    .[, std_error := std_error * 100] %>%
-    .[, ci_lower := pmax(ci_lower, 0) * 100] %>%
-    .[, ci_upper := pmin(ci_upper, 1) * 100]
-  }
- )
-
-}
-
-
-svy_statby_tidy_proportion <- function(x, outcome, by_vars, ...){
-
- map2_dfr(
-  x,
-  names(x),
-  ~ {
-
-   outcome_expression <- as.character(glue("I({outcome} == \"{.y}\")"))
-   se_expression <- paste0('se.as.numeric(', outcome_expression, ')')
-
-   as.data.table(.x) %>%
-    setnames(old = c(outcome_expression, se_expression, 'ci_l', 'ci_u'),
-             new = c('estimate', 'std_error', 'ci_lower', 'ci_upper')) %>%
-    .[, outcome := outcome] %>%
-    .[, level := .y] %>%
-    .[, statistic := "proportion"] %>%
-    setcolorder(neworder = c('outcome',
-                             'level',
-                             by_vars,
-                             'statistic',
-                             'estimate',
-                             'std_error',
-                             'ci_lower',
-                             'ci_upper')) %>%
-    .[, estimate := estimate * 100] %>%
-    .[, std_error := std_error * 100] %>%
-    .[, ci_lower := pmax(ci_lower, 0) * 100] %>%
-    .[, ci_upper := pmin(ci_upper, 1) * 100]
-
-  }
- )
-
-
-
-}

### 2022-10-21 data-raw/nhanes_key.R

@@ -228,15 +228,15 @@ nhanes_variables <- data.table::data.table(
   "Gender", # demo_gender
   "Systolic blood pressure, mm Hg", # bp_sys_mean
   "Diastolic blood pressure, mm Hg", # bp_dia_mean
-  "Blood pressure category (not including antihypertensive medication use)", # bp_cat_meds_excluded
+  "Blood pressure category", # bp_cat_meds_excluded
   "Blood pressure category (including antihypertensive medication use)", # bp_cat_meds_included
-  "Blood pressure control (SBP < 140 mm Hg and DBP < 90 mm Hg)", # bp_control_jnc7
-  "Blood pressure control (SBP < 130 mm Hg and DBP < 80 mm Hg)", # bp_control_accaha
-  "Uncontrolled BP (SBP \u2265 140 mm Hg or DBP \u2265 90 mm Hg)", # bp_uncontrolled_jnc7
-  "Uncontrolled BP (SBP \u2265 130 mm Hg or DBP \u2265 80 mm Hg)", # bp_uncontrolled_accaha
+  "Blood pressure control (JNC7 guideline)", # bp_control_jnc7
+  "Blood pressure control (2017 ACC/AHA BP guideline)", # bp_control_accaha
+  "Uncontrolled BP (JNC7 guideline)", # bp_uncontrolled_jnc7
+  "Uncontrolled BP (2017 ACC/AHA BP guideline)", # bp_uncontrolled_accaha
   "Self-reported antihypertensive medication use", # bp_med_use
-  "Antihypertensive medications recommended by JNC7", # bp_med_recommended_jnc7
-  "Antihypertensive medications recommended by ACC/AHA 2017", # bp_med_recommended_accaha
+  "Antihypertensive medications recommended (JNC7 guideline)", # bp_med_recommended_jnc7
+  "Antihypertensive medications recommended (2017 ACC/AHA BP guideline)", # bp_med_recommended_accaha
   "Number of antihypertensive medication classes", # bp_med_n_class
   "ACE inhibitors", # bp_med_ace
   "Aldosterone antagonists", # bp_med_aldo
@@ -250,13 +250,13 @@ nhanes_variables <- data.table::data.table(
   "Thiazide or thiazide-type diuretics", # bp_med_diur_thz
   "Direct renin inhibitors", # bp_med_renin_inhibitors
   "Direct vasodilators", # bp_med_vasod
[…]

### 2022-10-21 tests/testthat/test-check_vars_in_data.R

@@ -1,0 +1,37 @@
+
+expect_invisible(
+ check_vars_in_data(outcome_variable = 'bp_sys_mean',
+                    group_variable = NULL,
+                    data_names = names(nhanes_data),
+                    data_label = 'nhanes_data')
+)
+
+expect_invisible(
+ check_vars_in_data(outcome_variable = 'bp_sys_mean',
+                    data_names = nhanes_key$variable,
+                    data_label = 'nhanes_key')
+)
+
+expect_error(
+ check_vars_in_data(outcome_variable = 'not_right',
+                    data_names = nhanes_key$variable,
+                    data_label = 'nhanes_key'),
+ regexp = 'outcome_variable not_right'
+)
+
+expect_error(
+ check_vars_in_data(outcome_variable = 'bp_sys_mean',
+                    group_variable = 'not_right',
+                    data_names = nhanes_key$variable,
+                    data_label = 'nhanes_key'),
+ regexp = 'group_variable not_right'
+)
+
+expect_error(
+ check_vars_in_data(outcome_variable = 'bp_sys_mean',
+                    stratify_variable = 'not_right',
+                    data_names = names(nhanes_data),
+                    data_label = 'nhanes_key'),
+ regexp = 'stratify_variable not_right'
+)
+

### 2022-10-23 paper/manuscript.Rmd
website updates
polished pages for results, starting to work on validation pages.

@@
  fig.height = 4, 
- dpi = 300
+ dpi = 300,
+ ft.keepnext = FALSE
 )
@@
 # define age weights for standardization
-age_wts_overall <- c(49.3, 33.6, 10.1, 7)
-age_wts_htn <- c(26.4, 43.5, 17.0, 13.2)
+standard_weights_overall <- c(49.3, 33.6, 10.1, 7)
+standard_weights_htn <- c(26.4, 43.5, 17.0, 13.2)
 
@@
 Adam P. Bress, PharmD, MS,^3,4^ 
-YOUR NAME, DEGREE,^x^
+Kathryn Foti,^2^
+Shakia T. Hardy,^2^
 Bharat Poudel, MSPH,^2^
@@
 ^4^Department of Population Health Sciences, University of Utah School of Medicine, Salt Lake City, UT. `r single_par`
-^x^ YOUR AFFILIATION IF IT ISN'T ALREADY DEFINED ABOVE
 
@@
 
-**Methods**: We developed and validated a web-based application that provides hypertension statistics for US adults using 10 cycles of NHANES data, 1999-2000 through 2017-2020 without requiring users to conduct statistical programming.
+**Methods**: We developed and validated a web-based application that provides hypertension statistics for US adults using 10 cycles of NHANES data, 1999-2000 through 2017-2020, without requiring users to conduct statistical programming.
 The application’s interface allows users to estimate crude and age-adjusted means, quantiles, and proportions.
@@
 The age-adjusted prevalence of controlled BP, defined according to the 2017 American College of Cardiology/American Heart Association BP guideline, among US adults with hypertension was 9.6% in 1999-2000, 24.6% in 2013-2014, and 21.7% in 2017-2020.
-After age-adjustment and among US adults who self-report taking antihypertensive medication, 27.4%, 48.5%, and 42.8% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively.
+After age-adjustment and among US adults who self-reported taking antihypertensive medication, 27.4%, 48.5%, and 42.8% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively.
 The application was validated by replicating statistics from three publications.
 
-**Conclusions**: The application developed and validated is publicly available and produced valid, transparent, and reproducible results.
+**Conclusions**: The application developed and validated in the current study produced valid, transparent, and reproducible results and is publicly available.
 
@@
 
-The National Health and Nutrition Examination Survey (NHANES) is a program conducted by the US National Center for Health Statistics (NCHS) of the Centers for Disease Control and Prevention (CDC) and is designed to assess the health and nutritional status of the non-institutionalized US population.
+The National Health and Nutrition Examination Survey (NHANES) is a program conducted by the US National Center for Health Statistics (NCHS) of the Centers for Disease Control and Prevention (CDC) and is designed to assess the ongoing health and nutritional status of the non-institutionalized US population.
 NHANES data have been analyzed to provide hypertension statistics for non-institutionalized US adults with important policy and public health implications.
-For example, NHANES data have been used to estimate the impact of the lower blood pressure (BP) levels that define hypertension and controlled BP in the 2017 American College of Cardiology/American Heart Association (ACC/AHA) BP guideline versus the Seventh Joint National Committee (JNC7) BP guideline.
+For example, NHANES data have been used to estimate the impact of the lower blood pressure (BP) levels that define hypertension and controlled BP in the 2017 American College of Cardiology/American Heart Association (ACC/AHA) BP guideline versus the Seventh Report of the Joint National Committee on Prevention, Detection, Evaluation, and Treatment of High BP (JNC7) guideline.
 Additionally, NHANES data have been used to track the proportion of US adults with hypertension that have controlled BP, which were included in a Call-to-Action to Control BP from the US Surgeon General.
@@
 Specific statistical techniques are required to account for the multi-stage sampling design used to select NHANES participants and analyses need to be weighted to produce nationally representative estimates.
-Each NHANES cycle includes over 50 data files, each of which has detailed informational material.
+NHANES data are currently collected in two-year periods, referred to as cycles, and each cycle  includes over 50 data files and accompanying informational material.
 NHANES data collection protocols have changed over time for some variables, and these differences should be taken into account when comparing or aggregating data from multiple cycles.
@@
 
-To address these challenges, we developed an open-source web-based application that provides nationally representative BP and hypertension statistics for non-institutionalized US adults using NHANES data without requiring users to conduct statistical programming.
+We developed an open-source web-based application that provides nationally representative BP and hypertension statistics for non-institutionalized US adults using NHANES data without requiring users to conduct statistical programming.
 The goal of the application is to increase the use of NHANES data for research and policy decision making with a focus on ensuring results are valid, transparent, and reproducible.
@@
 
-NHANES data are collected through in-home interviews and study examinations conducted at mobile examination centers.
+NHANES data were collected through in-home interviews and study examinations conducted at mobile examination centers.
 The interview included questions about demographics, health behaviors, medical history, and medication use.
@@
 Of relevance to the application, blood samples were used to measure total and HDL cholesterol, glycated hemoglobin and serum creatinine, and the urine sample was used to measure albumin and creatinine and to conduct a pregnancy test.
-The protocol for the BP measurement is available online.
+The protocol for obtaining BP measurements is available online.
 In brief, systolic and diastolic BP (SBP and DBP, respectively) were measured three times by trained and certified physicians.
@@
 The oscillometric SBP and DBP values were calibrated to the mercury device.
-Antihypertensive medication classes were defined using recommendations from the 2017 ACC/AHA BP guideline.
+Antihypertensive medication classes were defined using those listed in the 2017 ACC/AHA BP guideline.
 For the current analysis, we defined hypertension, BP control, and resistant hypertension according to the 2017 ACC/AHA BP guideline.
-The application also has these variables defined according to the Seventh Report of the Joint National Committee on Prevention, Detection, Evaluation, and Treatment of High Blood Pressure definitions.
-A complete list of the variables available in the data set is provided in Table 1 with full definitions provided in Supplemental Table 1.
+The application also has these variables defined according to the JNC7 definitions.
+A complete list of the variables available in the data set is provided in **Table 1** with full definitions provided in **Table S1**.
 
@@
 We further restricted the population to participants who completed the in-home interview and study examination, with one or more SBP and DBP measurement, and who had data on self-reported antihypertensive medication use.
-After these exclusions were applied, the population in this application included 56,035 participants (Figure S1).
+After these exclusions were applied, the population in this application included 56,035 participants (**Figure S1**).
 
-## Features of the Open-Source Web Application
+## Features of the web application
 
@@
 Users can restrict analyses to subsets of participants (e.g., participants who self-reported taking antihypertensive medication).
-When population count estimates are requested, survey weights are calibrated to account for participants missing information on SBP, DBP or self-reported antihypertensive medication use.
+When population count estimates are requested, survey weights are calibrated to account for missing information on SBP, DBP or self-reported antihypertensive medication use.
 Users may tabulate or visualize summaries, and can present results for the overall population or in subgroups defined by a stratifying variable.
@@
 
-## Development and validation of the Open-Source Web Application
+## Development and validation of the web application
 
@@
 
-The web application successfully replicated results from prior studies.
-Full details on how to use the application to replicate these studies are available online.
-Following its validation, version 0.0.1 of the application was released and deployed in a publicly available server on **DATE TBD**.(add a citation to version 0.0.1 of the cardioStatsUSA package and a citation to the url where the shiny app lives)
-
-## Demonstration of core features
-
-Among non-institutionalized US adults ≥ 18 years of age, the estimated mean SBP was XXX mm Hg (95% CI ) in 1999-2000, XXX mm Hg (95% CI) in 2009-2010 and XXX mm Hg (95% CI ) in 2017-2020 (**Figure 1**).
+Among non-institutionalized US adults ≥ 18 years of age, the estimated mean SBP was 123 (95% CI 121, 124) mm Hg in 1999-2000, 120 (95% CI 120, 121) mm Hg in 2009-2010 and 123 (95% CI 122, 124) mm Hg in 2017-2020 (**Figure 1**).
 The age-adjusted prevalence of hypertension was highest in 1999-2000 (47.9%), lowest in 2009-2010 (43.0%), and 44.7% in 2017-2020 (**Figure 2**).
@@
 The number of US adults with hypertension increased to 115.3 million in 2017-2020.
-In each NHANES cycle, the estimated age-adjusted prevalence of hypertension was higher among US adults with versus without CKD (**Figure S2**).
+In each NHANES cycle, the estimated prevalence of hypertension was higher among US adults with versus without CKD (**Figure S2**).
 In 2017-2020, a higher percentage of US adults with versus without hypertension were non-Hispanic Black (13.5% versus 9.3%) while a lower percentage of US adults with versus without hypertension were Hispanic (12.3% versus 18.1%) (**Table 2**).
-Among pregnant women, the distribution of BP categories could not be estimated reliably in 2017-2020 (**Figure S3; Panel A**), but was stable after pooling NHANES cycles from 2009-2010 through 2017-2020 (**Figure S3; Panel B**)
-
-## Demonstration of customized analyses
+Among pregnant women, the distribution of BP categories could not be estimated reliably in 2017-2020 (**Figure S3; Panel A**) except for the group with SBP of 130 to < 140 mm Hg or DBP of 80 to < 90 mm Hg, but was stable for all groups after pooling NHANES cycles from 2009-2010 through 2017-2020 (**Figure S3; Panel B**)
 
@@
 
-In the current manuscript, we present a web application that allows the calculation of nationally representative estimates for BP, hypertension and antihypertensive medication-related outcomes using publicly available data from NHANES.
+In the current manuscript, we present a web application that allows the calculation of nationally representative estimates for BP, hypertension and antihypertensive medication-related outcomes using publicly available NHANES data.
 Using this application, we generated crude and age-adjusted BP and hypertension statistics.
-Also, we generated statistics stratified by characteristics of US adults (e.g., CKD), and pooling multiple NHANES cycles to obtain more precise estimates when working with the small sub-group of pregnant women.
-Researchers, clinicians, and the general public can use the application to generate customized BP and hypertension statistics for US adults.
+Also, we generated statistics stratified by characteristics of US adults and pooling multiple NHANES cycles to obtain more precise estimates when working with the small sub-group of pregnant women.
+Following its validation, version 0.0.1 of the application was released and deployed in a publicly available server on **DATE TBD**.(add a citation to version 0.0.1 of the cardioStatsUSA package and a citation to the url where the shiny app lives) Researchers, clinicians, and the public can use the application to generate customized BP and hypertension statistics for US adults.
 
@@
 The application re-calibrates the NHANES weights for the estimation of population counts (cite pmid 12500213).
-When estimating population counts, participants missing data on SBP, DBP or antihypertensive medication use cannot have BP or hypertension-related outcomes.
+This was done because participants missing data on SBP, DBP or antihypertensive medication use cannot have BP or hypertension-related outcomes.
 Weights were not re-calibrated when estimating proportions as participants missing data are removed from the numerator and denominator.
@@
 For example, a prior manuscript reported the prevalence of resistant hypertension defined by the 2017 ACC/AHA BP guideline to be 19.7% in 2009-2014(cite pmid 30580690).
-When estimated by the application, the prevalence of resistant hypertension over this time period was X%.
+When estimated by the application, the prevalence of resistant hypertension over this time period was 17.1%.
 The difference in the prevalence estimates can be attributed to the approach used to categorize medication into classes.
@@
 The application’s interface and its supporting R package allows additional variables to be added.
-In addition, the application may be extended to include data from additional NHANES cycles as they become publically available.
+In addition, the application may be extended to include data from additional NHANES cycles as they become available.
 
@@
 We used NHANES data, which are publicly available, rigorously collected, and allow for estimation of nationally representative statistics.
-Also, we leveraged open-source software to ensure that the application developed in the current study is transparent, able to be extended, and freely available.
-In addition to the application, we developed online resources with tutorials for new users and extensive documentation to contextualize results.
+Also, we leveraged open-source software to ensure that our application is transparent, able to be extended, and freely available.
 This study also has several limitations.
@@
 
+<!-- TODO: make this identical to website output, including even the unreliability columns. Add footnotes to explain what column names mean -->
+
 ### Table 2:	Race/ethnicity distribution of US adults with and without hypertension in 2017-2020
@@
                  time_values = 'all', 
-                 age_wts = age_wts_overall, 
+                 standard_weights = standard_weights_overall, 
                  geom = 'bar') %>% 
@@
                        subset_calls = list("htn_accaha" = "Yes"),
-                       age_wts = age_wts_htn,
+                       standard_weights = standard_weights_htn,
                        time_values = 'all', 
@@
                                            "bp_med_use" = "Yes"),
-                       age_wts = age_wts_htn,
+                       standard_weights = standard_weights_htn,
                        time_values = 'all', 

### 2022-10-23 R/app_run.R

@@ -200,19 +200,19 @@ app_run <- function(nhanes_data = cardioStatsUSA::nhanes_data,
       h5("Weights for each age group (must be >0)"),
 
       splitLayout(
-       numericInput(inputId="age_wts_1",
+       numericInput(inputId="standard_weights_1",
                     label="18-44",
                     value = 49.3,
                     min = 5),
-       numericInput(inputId="age_wts_2",
+       numericInput(inputId="standard_weights_2",
                     label="45-64",
                     value = 33.6,
                     min = 5),
-       numericInput(inputId="age_wts_3",
+       numericInput(inputId="standard_weights_3",
                     label="65-74",
                     value = 10.1,
                     min = 5),
-       numericInput(inputId="age_wts_4",
+       numericInput(inputId="standard_weights_4",
                     label="75+",
                     value = 7.0,
                     min = 5),
@@ -798,7 +798,11 @@ app_run <- function(nhanes_data = cardioStatsUSA::nhanes_data,
        inputId = ss_val_catg,
        choices =
         levels(nhanes_data[[ input[[ss_var]] ]]) %||%
-        sort(unique(na.omit(nhanes_data[[ input[[ss_var]] ]]))) %>%
+        sort(
+         unique(
+          stats::na.omit(nhanes_data[[ input[[ss_var]] ]])
+         )
+        ) %>%
         c(miss_add),
        selected = character(0) #subset_value_selected
       )
@@ -880,13 +884,13 @@ app_run <- function(nhanes_data = cardioStatsUSA::nhanes_data,
 
     pool <- input$pool == 'yes'
 
-    age_wts <- NULL
+    standard_weights <- NULL
 
     if(input$age_standardize){
-     age_wts <- c(input$age_wts_1,
-                  input$age_wts_2,
-                  input$age_wts_3,
-                  input$age_wts_4)
+     standard_weights <- c(input$standard_weights_1,
+                  input$standard_weights_2,
+                  input$standard_weights_3,
+                  input$standard_weights_4)
     }
 
     outcome_variable  <- input_infer(input$outcome,  default = NULL)
@@ -962,7 +966,7 @@ app_run <- function(nhanes_data = cardioStatsUSA::nhanes_data,
        time_values = time_values_selected,
        pool = pool,
        subset_calls = subset_calls,
-       age_wts = age_wts,
+       standard_weights = standard_weights,
        simplify_output = FALSE
       ),
      silent = TRUE

### 2022-10-23 R/discretize.R

@@ -24,10 +24,7 @@ get_breaks_frequency <- function(x, n_group){
 
  probs <- seq(0, 1, length.out = n_group+1)
 
- quantile(x, probs = probs, na.rm = TRUE)
-
-
-
+ stats::quantile(x, probs = probs, na.rm = TRUE)
 
 }
 get_breaks_interval <- function(x, n_group){

### 2022-10-23 R/nhanes_summarize.R

@@ -7,33 +7,32 @@
 #' - Who is included in the data to be summarized
 #' - How is the summary computed
 #'
-#' @param outcome (*character*; length 1; no default)
-#'   The variable to be summarized.
-#' @param exposure (*character*; length 1; default is none)
+#' @inheritParams nhanes_design
 #'
-#' @param group (*character*; length 1; default is none)
+#' @inheritParams nhanes_design_summarize
 #'
-#' @param years (*character*; length may vary; default = 'all')
-#'   Valid options are
-#'   - 'all' (includes all NHANES cycles)
-#'   - 'last_5' (includes the last 5 most recent NHANES cycles)
-#'   - 'most_recent' (include the most recent NHANES cycle)
+#' @param subset_calls \[named list(n)\]
 #'
-#' @param pool (*logical*; length 1; default = `FALSE`)
+#'  the names of `subset_calls` are variable names, and the values are
+#'    values of the variable to include in the subsetted data. For example,
+#'    `subset_calls = list("demo_gender" = "Women")` will subset the data
+#'    to include rows where `demo_gender` is equal to `"Women"`. Multiple
+#'    entries are allowed and collapsed with the logical `&` operator.
+#'    For example, `subset_calls = list(demo_gender = "Women", bp_med_use = "Yes")`
+#'    will subset the data to include rows where `demo_gender` is equal
+#'    to `'Women'` AND `bp_med_use` is equal to `"Yes"`
 #'
-#' @param subset_variables (*list*; length may vary)
+#' @inheritParams nhanes_design_standardize
 #'
-#' @param subset_values (*list*; length must match `subset_variables`)
-#'
-#' @param age_wts (*numeric*; length 4; default is none)
-#'
-#' @return a `data.table` with summarized values
+#' @inherit nhanes_design_summarize return
 #'
 #' @export
 #'
 #' @examples
 #'
-#' nhanes_summarize(nhanes_data, nhanes_key, outcome_variable = "bp_sys_mean")
+#' nhanes_summarize(data = nhanes_data,
+#'                  key = nhanes_key,
+#'                  outcome_variable = "bp_sys_mean")
 #'
 nhanes_summarize <- function(data,
                              key,
@@ -48,7 +47,8 @@ nhanes_summarize <- function(data,
                              time_values = NULL,
                              pool = FALSE,
                              subset_calls = list(),
-                             age_wts = NULL,
+                             standard_variable = 'demo_age_cat',
[…]

### 2022-10-23 tests/testthat/test-nhanes_variables.R

@@ -1,7 +1,5 @@
 
 library(testthat)
-library(haven)
-
 
 key <- nhanes_key$data
 

### 2022-10-23 tests/testthat/test-replicate_muntner_jama_2020.R

@@ -1,70 +1,130 @@
 
+# TODO: do you need to subset before standardize? should make an error if so.
+suppressPackageStartupMessages({
+ library(dplyr)
+ library(data.table)
+ library(tidyr)
+ library(testthat)
+})
+
 # JAMA Trends in BP tests -------------------------------------------------
 
-# TODO: do you need to subset before standardize? should make an error if so.
+#' ## Who is included in this test
+#'
+#' For these tests, we include the same cohort as Muntner et al. Full details
+#' on the inclusion criteria are given in the paper, but here all that
+#' we need to do is subset the data to include rows where `svy_subpop_htn`
+#' is equal to 1.
 
-wts <- c(15.5, 45.4, 21.5, 17.7)
+#+ echo = TRUE
+nhanes_subpop_htn <- nhanes_data[svy_subpop_htn == 1]
 
+#' after these exclusions,
+{{nrow(nhanes_subpop_htn)}}
+#' participants are included.
+#'
+#' ## Age adjustment
+#'
+#' We use direct standardization with the same standard weights as in Munter
+#' et al.
+#'
+#+ echo = TRUE
+standard_weights <- c(15.5, 45.4, 21.5, 17.7)
+#'
+#' ## NHANES design objects
+#'
+#' In most cases, we use `nhanes_summarize()` or `nhanes_visualize` when
+#' generating results with `cardioStatsUSA`. In this case, we find it's
+#' easier to use the `nhanes_design` family of functions, which are the lower
+#' level functions that support `nhanes_summarize` and `nhanes_visualize`.
+#' Full details on this family can be found at the help page for
+#' [nhanes_design]. For this analysis, we create three designs:
+#'
+#' - an initial design that encapsulates the two others:
+#'
+#+ echo = TRUE
 ds_init <- nhanes_design(
  data = nhanes_data[svy_subpop_htn == 1],
  key = nhanes_key,
  outcome_variable = 'bp_control_jnc7'
 )
-
+#' - a design for non-pregnant adults with hypertension
+#+ echo = TRUE
 ds_all <- ds_init %>%
  nhanes_design_subset(
[…]

### 2022-10-23 paper/manuscript.Rmd
PM edits, fpath fix attempt

@@
 
+<br>
+
 NHANES data are publicly available and accessible through the CDC website.
@@
 
+<br>
+
 We developed an open-source web-based application that provides nationally representative BP and hypertension statistics for non-institutionalized US adults using NHANES data without requiring users to conduct statistical programming.
@@
 In this manuscript, we review the design, development and validation of the application, present BP and hypertension statistics for US adults, and describe online tutorials and resources that may help users engage with the application.
+
+<br>
 
@@
 Written informed consent was obtained from each participant.
+
+<br>
 
@@
 
+<br>
+
 There were 107,622 NHANES participants in the 10 cycles from 1999-2000 to 2017-2020.
@@
 After these exclusions were applied, the population in this application included 56,035 participants (**Figure S1**).
+
+<br>
 
@@
 
+<br>
+
 ## Development and validation of the web application
@@
 We validated the web application by using it to reproduce statistics reported in two prior studies and one CDC report.
+
+<br>
 
@@
 
+<br>
+
 We also illustrated how core features of the application can be combined to perform customized analyses.
@@
 For age adjustment in the analysis of resistant hypertension, we set the age distribution for the standard population to represent US adults with hypertension from 1999 to 2020: 26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥75 years of age, respectively.
+
+<br>
 
@@
 
+<br>
+
 The age-adjusted prevalence of BP control among US adults with hypertension was lowest in 1999-2000 (9.6%), highest in 2013-2014 (24.6%), and 21.7% in 2017-2020 (**Figure 4; panel A**).
@@
 Among US adults who were taking three or more classes of antihypertensive medication, the prevalence of resistant hypertension was lowest in 2009-2010 (64.1%), highest in 2003-2004 (78.3%), and 67.9% in 2017-2020 (**Figure S4; panel B**).
+
+<br>
 
@@
 
+<br>
+
 NHANES was designed to obtain nationally representative estimates of the health and nutrition status of non-institutionalized US adults and is an ideal data source to obtain statistics related to hypertension.
@@
 The web application that we present in the current manuscript was designed to reduce barriers to using NHANES so that these data can more fully inform public health decisions and future research related to hypertension.
+
+<br>
 
@@
 
+<br>
+
 We were able to replicate several prior manuscripts using the application.
@@
 
+<br>
+
 The application has a number of features that can be used in combination to create highly customized statistics.
@@
 
+<br>
+
 With an open-source software license and publicly available documentation and data, the web application we developed can be extended.
@@
 The application’s interface and its supporting R package allows additional variables to be added.
-In addition, the application may be extended to include data from additional NHANES cycles as they become available.
+In addition, the application can be extended to include data from additional NHANES cycles as they become available.
+
+<br>
 
@@
 This study also has several limitations.
-First, NHANES participants had their mean BP measured during a single visit, and BP guidelines recommend obtaining mean BP using at least two BP values measured on separate days.
-Second, the response rate for NHANES has declined from 1999-2000 through 2017-2020.
+NHANES participants had their mean BP measured during a single visit, and BP guidelines recommend obtaining mean BP using at least two BP values measured on separate days.
+The response rate for NHANES has declined from 1999-2000 through 2017-2020.
 Effects of this decline are unclear.
+
+<br>
 

### 2022-10-23 README.Rmd
badges were behind on repo name

@@
 <!-- badges: start -->
-[![Codecov test coverage](https://codecov.io/gh/jhs-hwg/nhanes-shiny-bp/branch/main/graph/badge.svg)](https://app.codecov.io/gh/jhs-hwg/nhanes-shiny-bp?branch=main)
-[![R-CMD-check](https://github.com/jhs-hwg/nhanes-shiny-bp/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhs-hwg/nhanes-shiny-bp/actions/workflows/R-CMD-check.yaml)
+[![Codecov test coverage](https://codecov.io/gh/jhs-hwg/cardioStatsUSA/branch/main/graph/badge.svg)](https://app.codecov.io/gh/jhs-hwg/cardioStatsUSA?branch=main)
+[![R-CMD-check](https://github.com/jhs-hwg/cardioStatsUSA/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhs-hwg/cardioStatsUSA/actions/workflows/R-CMD-check.yaml)
 <!-- badges: end -->

### 2022-10-23 paper/manuscript.Rmd
badges were behind on repo name

@@
 
-<!-- TODO: make this identical to website output, including even the unreliability columns. Add footnotes to explain what column names mean -->
-
 ### Table 2:	Race/ethnicity distribution of US adults with and without hypertension in 2017-2020
@@
  filter(statistic == 'percentage') %>% 
- select(-statistic, 
-        -ends_with("reason"),
-        -unreliable_status,
-        -review_needed) %>% 
  flextable() %>% 
- set_header_labels(
-  svy_year = 'NHANES cycle',
-  htn_accaha = 'Hypertension',
-  demo_race = 'Race/Ethnicity',
-  estimate = 'Estimate',
-  std_error = 'Standard error',
-  ci_lower = 'CI lower',
-  ci_upper = 'CI upper',
-  n_obs = 'N observations'
+ theme_box() %>% 
+ width(width = 0.5) %>% 
+ align(align = 'center', part = 'all') %>% 
+ align(j = 1, align = 'center', part = 'all') %>% 
+ add_footer_lines(
+  values = "Table format is identical to the web application's output."
  ) %>% 
- theme_box() %>% 
- width(width = 1) %>% 
- align(align = 'center', part = 'all') %>% 
- align(j = 1, align = 'center', part = 'all')
+ add_footer_lines(
+  values = "svy_year indicates the NHANES cycle"
+ ) %>% 
+ add_footer_lines(
+  values = "htn_accaha indicates hypertension status." 
+ ) %>% 
+ add_footer_lines(
+  values = "demo_race indicates the race subgroup."
+ ) %>% 
+ add_footer_lines(
+  values = "statistic indicates the type of statistic computed."
+ ) %>% 
+ add_footer_lines(
+  values = "estimate is the point estimate of the given statistic"
+ ) %>% 
+ add_footer_lines(
+  values = "std_error is the standard error for estimate"
+ ) %>% 
+ add_footer_lines(
+  values = "ci_lower is the lower bound of a 95% confidence interval for estimate"
+ ) %>% 
+ add_footer_lines(
+  values = "ci_upper is the upper bound of a 95% confidence interval for estimate"
+ ) %>% 
+ add_footer_lines(
+  values = "n_obs is the number of observations (unweighted) used for computations"
+ ) %>% 
+ add_footer_lines(
+  values = "unreliable_status is TRUE if the result is unreliable, FALSE otherwise"
+ ) %>% 
+ add_footer_lines(
+  values = "unreliable_reason is blank if the result is reliable, and lists reason(s) for unreliability otherwise"
+ ) %>% 
+ add_footer_lines(
+  values = "review_needed is TRUE if the result should be reviewed, FALSE otherwise"
+ ) %>% 
+ add_footer_lines(
+  values = "review_reason is blank if the result does not require review, and lists reason(s) for requiring review otherwise"
+ )
 

### 2022-10-24 paper/manuscript.Rmd
refs added to paper

--- 2022-10-23 paper/manuscript.Rmd
@@
 always_allow_html: true
+bibliography: refs.bib
+csl: jama.csl
 output: 
@@
 
-<!-- Abstract: Structured with the following headings and no more than 250 words: Background, Objectives, Methods, Results, Conclusions.
-The abstract should present essential data in 5 paragraphs.
-Use complete sentences.
-All data in the abstract also must appear in the manuscript text or tables.
--->
-
 **Background**: Data from the US National Health and Nutrition Examination Survey (NHANES) are freely available and can be analyzed to produce hypertension statistics for the non-institutionalized US population.
@@
 
-**Conclusions**: The application developed and validated in the current study produced valid, transparent, and reproducible results and is publicly available.
+**Conclusions**: The application developed and validated in the current study is publicly available and produced valid, transparent, and reproducible results.
 
@@
 NHANES data have been analyzed to provide hypertension statistics for non-institutionalized US adults with important policy and public health implications.
-For example, NHANES data have been used to estimate the impact of the lower blood pressure (BP) levels that define hypertension and controlled BP in the 2017 American College of Cardiology/American Heart Association (ACC/AHA) BP guideline versus the Seventh Report of the Joint National Committee on Prevention, Detection, Evaluation, and Treatment of High BP (JNC7) guideline.
-Additionally, NHANES data have been used to track the proportion of US adults with hypertension that have controlled BP, which were included in a Call-to-Action to Control BP from the US Surgeon General.
+For example, NHANES data have been used to estimate the impact of the lower blood pressure (BP) levels that define hypertension and controlled BP in the 2017 American College of Cardiology/American Heart Association (ACC/AHA) BP guideline versus the Seventh Report of the Joint National Committee on Prevention, Detection, Evaluation, and Treatment of High BP (JNC7) guideline.[@whelton20182017; @chobanian2003seventh] Additionally, NHANES data have been used to track the proportion of US adults with hypertension that have controlled BP, which were included in a Call-to-Action to Control BP from the US Surgeon General.[@us_surgeon_2020]
 
@@
 
-NHANES data are publicly available and accessible through the CDC website.
-However, analyzing NHANES data may be challenging for a number of reasons.
+NHANES data are publicly available and accessible through the CDC website.[@nhanes_home] However, analyzing NHANES data may be challenging for a number of reasons.
 Specific statistical techniques are required to account for the multi-stage sampling design used to select NHANES participants and analyses need to be weighted to produce nationally representative estimates.
@@
 NHANES data collection protocols have changed over time for some variables, and these differences should be taken into account when comparing or aggregating data from multiple cycles.
-Also, tests need to be performed to ensure the statistical estimates are reliable, and unstable estimates should be suppressed.
+Also, tests need to be performed to ensure the statistical estimates are reliable, and unstable estimates should be suppressed.[@parker2017national]
 
@@
 During the study examination, height, weight and BP were measured and blood and spot urine samples were collected.
-Of relevance to the application, blood samples were used to measure total and HDL cholesterol, glycated hemoglobin and serum creatinine, and the urine sample was used to measure albumin and creatinine and to conduct a pregnancy test.
+Of relevance to the application, blood samples were used to measure total and high-density lipoprotein cholesterol, glycated hemoglobin and serum creatinine, and the urine sample was used to measure albumin and creatinine and to conduct a pregnancy test.
 The protocol for obtaining BP measurements is available online.
@@
 The mean SBP and DBP levels were computed over all available measurements for each participant.
-The oscillometric SBP and DBP values were calibrated to the mercury device.
-Antihypertensive medication classes were defined using those listed in the 2017 ACC/AHA BP guideline.
-For the current analysis, we defined hypertension, BP control, and resistant hypertension according to the 2017 ACC/AHA BP guideline.
+The oscillometric SBP and DBP values were calibrated to the mercury device.[@muntner2022blood] Antihypertensive medication classes were defined using those listed in the 2017 ACC/AHA BP guideline.[@whelton20182017] For the current analysis, we defined hypertension, BP control, and resistant hypertension according to the 2017 ACC/AHA BP guideline.
 The application also has these variables defined according to the JNC7 definitions.
-A complete list of the variables available in the data set is provided in **Table 1** with full definitions provided in **Table S1**.
+Chronic kidney disease (CKD) was defined by estimated glomerular filtration rate < 60 ml/min/1.73 m^2^ calculated using the 2021 serum creatinine-based equation or albumin-to-creatinine ratio ≥ 30 mg/g.[@inker_new_2021] A list of BP, hypertension, and antihypertensive medication variables is provided in **Table 1** with full definitions for all variables provided in **Table S1**.
 
@@
 We restricted the dataset to adults ≥ 18 years of age.
-This exclusion was applied because statistics for BP levels and hypertension in children and adolescents are markedly different than for adults.
+This exclusion was applied because statistics for BP levels and hypertension in children and adolescents are markedly different than for adults (**What should we cite here?** Perhaps this?
+doi:10.1001/jama.291.17.2107).
 We further restricted the population to participants who completed the in-home interview and study examination, with one or more SBP and DBP measurement, and who had data on self-reported antihypertensive medication use.
@@
 
-A full summary of this application’s features and associated tutorials are available online.
-Briefly, users can select NHANES cycles from 1999-2000 to 2017-2020 to be analyzed.
+A full summary of this application’s features and associated tutorials are available online.[@rpack_cardioStatsUSA] Briefly, users can select NHANES cycles from 1999-2000 to 2017-2020 to be analyzed.
 Estimates are weighted to represent the non-institutionalized US population and users may incorporate age-adjustment through direct standardization.
@@
 Following CDC recommendations, unreliable statistical estimates are automatically suppressed.
-To increase precision and reliability of estimates, contiguous NHANES cycles can be combined.
+To increase precision and reliability of estimates, contiguous NHANES cycles can be combined.[@nhanes_tutorial_weights]
 
@@
 
-The web application was created using Shiny, an open-source software package that translates code from the R programming language into HTML, CSS, or JavaScript commands that create a website interface.
-We created the “cardioStatsUSA” R package to provide additional details on the web application’s design and comprehensive documentation of its components.
-We validated the web application by using it to reproduce statistics reported in two prior studies and one CDC report.
+The web application was created using Shiny,[@rpack_shiny; @rpack_shinyWidgets; @rpack_shinyalert] an open-source software package that translates code from the R programming language into HTML, CSS, or JavaScript commands that create a website interface.[@r_language] We created the “cardioStatsUSA” R package to provide additional details on the web application’s design and comprehensive documentation of its components.[@rpack_cardioStatsUSA] We validated the web application by using it to reproduce statistics reported in two prior studies and one CDC report.[@muntner2020htn; @muntner2022blood; @fryar_hypertension_2017]
 
@@
 We performed statistical analyses to demonstrate core features of the application.
-We estimated the mean SBP for US adults by NHANES cycle, 1999-2000 through 2017-2020, with points showing the estimated means and error bars showing the 95% confidence intervals.
+We estimated the mean SBP for US adults by NHANES cycle, 1999-2000 through 2017-2020, with points showing the estimated means and error bars showing the 95% confidence interval (CI) for the estimate.
 We made bar charts presenting the age-adjusted prevalence of hypertension and the estimated number of US adults with hypertension.
-For age adjustment, we used the estimated age distribution of US adults from 1999 to 2020 as the standard (49.3%, 33.6%, 10.1% and 7.0% being 18 to 44, 45 to 64, 65 to 74 and ≥75 years of age, respectively).
+For age adjustment, we used the estimated age distribution of US adults from 1999 to 2020 as the standard (49.3%, 33.6%, 10.1% and 7.0% being 18 to 44, 45 to 64, 65 to 74 and ≥ 75 years of age, respectively).
 To illustrate how results from the application appear when saved in various formats (e.g., comma separated, portable document format, Microsoft Excel), we tabulated the estimated race/ethnicity distribution of US adults with and without hypertension, separately.
-We demonstrated stratification by estimating the prevalence of hypertension by NHANES cycle for US adults with and without chronic kidney disease (CKD).
+We demonstrated stratification by estimating the prevalence of hypertension by NHANES cycle for US adults with and without CKD.
 We showed the application’s ability to suppress output when statistical estimates are unstable by attempting to estimate the distribution of BP categories (SBP/DBP < 120/80 mm Hg, 120-129/<80 mm Hg, 130-139/80-89 mm Hg, 140-159/90-99 mm Hg and ≥ 160/100 mm Hg) among pregnant women in 2017-2020.
@@
 We also estimated the age-adjusted prevalence of resistant hypertension by NHANES cycle for US adults with hypertension taking antihypertensive medication and among US adults with hypertension taking ≥ 3 classes of antihypertensive medication.
-For age adjustment in the analysis of resistant hypertension, we set the age distribution for the standard population to represent US adults with hypertension from 1999 to 2020: 26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥75 years of age, respectively.
+For age adjustment in the analysis of resistant hypertension, we set the age distribution for the standard population to represent US adults with hypertension from 1999 to 2020: 26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥ 75 years of age, respectively.
 
@@
 
-<!-- Among US adults with hypertension, 61.5% were not taking antihypertensive medication in 1999-2000 compared with 44.6% not taking antihypertensive medication in 2017-2020.
-(Figure 4; panel A).
-Among US adults who self-reported taking antihypertensive medication, the percentage taking four or more classes of antihypertensive medication increased from 3.5% in 1999-2000 to 7.6% in 2017-2020 (Figure 4; panel B).
--->
-
 # DISCUSSION
@@
 Also, we generated statistics stratified by characteristics of US adults and pooling multiple NHANES cycles to obtain more precise estimates when working with the small sub-group of pregnant women.
-Following its validation, version 0.0.1 of the application was released and deployed in a publicly available server on **DATE TBD**.(add a citation to version 0.0.1 of the cardioStatsUSA package and a citation to the url where the shiny app lives) Researchers, clinicians, and the public can use the application to generate customized BP and hypertension statistics for US adults.
+Following its validation, version 0.0.1 of the application was released and deployed in a publicly available server on **DATE TBD**.[@app_cardioStatsUSA] Researchers, clinicians, and the public can use the application to generate customized BP and hypertension statistics for US adults.
 
@@
 We required participants to have a single SBP and DBP to be included while some prior analyses required three SBP and DBP measurements.
-We chose an approach that is consistent with several analyses conducted by CDC investigators (cite pmid 26633197, 29155682).
-However, mean BP and the prevalence of high BP would be lower if we required multiple BP measurements (cite pmid 23126346).
-Any bias resulting from this decision is likely to be small as over 95% of adult NHANES participants with at least one SBP and DBP measurement had three SBP and DBP measurements.
-The application re-calibrates the NHANES weights for the estimation of population counts (cite pmid 12500213).
-This was done because participants missing data on SBP, DBP or antihypertensive medication use cannot have BP or hypertension-related outcomes.
+We chose an approach that is consistent with several analyses conducted by CDC investigators.[@yoon_hypertension_2015; @fryar_hypertension_2017] However, mean BP and the prevalence of high BP may be lower if we required multiple BP measurements.[@handler_impact_2012] Any bias resulting from this decision is likely to be small as over 95% of adult NHANES participants with at least one SBP and DBP measurement had three SBP and DBP measurements.
+The application re-calibrates the NHANES weights for the estimation of population counts.[@coresh_prevalence_2003] This was done because participants missing data on SBP, DBP or antihypertensive medication use cannot have BP or hypertension-related outcomes.
 Weights were not re-calibrated when estimating proportions as participants missing data are removed from the numerator and denominator.
@@
 However, we were not able to replicate some results.
-For example, a prior manuscript reported the prevalence of resistant hypertension defined by the 2017 ACC/AHA BP guideline to be 19.7% in 2009-2014(cite pmid 30580690).
-When estimated by the application, the prevalence of resistant hypertension over this time period was 17.1%.
+For example, a prior manuscript reported the prevalence of resistant hypertension defined by the 2017 ACC/AHA BP guideline to be 19.7% in 2009-2014.[@carey_prevalence_2019] When estimated by the application, the prevalence of resistant hypertension over this time period was 17.1%.
 The difference in the prevalence estimates can be attributed to the approach used to categorize medication into classes.
@@
 All results from the application can be saved as an image and included in scientific proposals or presentations.
-In addition, the ability to download summary data from the application allows users to further customize their results in tabular or graphical formats.
+In addition, the ability to download results as a data set from the application allows users to further customize their results in tabular or graphical formats.
 The web application can also be used for educational purposes, as teachers can use it to analyze health outcomes for US adults interactively with students.
@@
 The application is publicly available and produces valid, transparent, and reproducible results.
+
+\newpage
+
+# REFERENCES
+
+<div id="refs"></div>
+
+\newpage
+
+# ACKNOWLEDGMENTS
+
+**this section is under construction**
+
+The views expressed in this paper are those of the authors and do not represent the official position of the National Institutes of Health (NIH), the National Heart, Lung, and Blood Institute, or the U.S.
+Government.
+
+# FUNDING/SUPPORT
+
+**this section is under construction and should be checked for correctness**
+
+Drs Muntner and Jaeger receive support through grant R01HL144773 from the National Heart, Lung, and Blood Institute.
+Drs Muntner and Hardy receive support through grant R01HL117323 from the National Heart, Lung, and Blood Institute.
+Dr Muntner receives support through grant R01HL139716 from the National Heart, Lung, and Blood Institute.
+
+# FINANCIAL DISCLOSURE
+
+**this section is under construction and should be checked for correctness**
+
+Drs Muntner reported receiving grant funding and consulting fees from Amgen Inc. Dr Colantonio reported receiving grant funding from Amgen Inc. No other disclosures were reported.
 
@@
  filter(statistic == 'percentage') %>% 
- select(-statistic, 
-        -ends_with("reason"),
-        -unreliable_status,
-        -review_needed) %>% 
  flextable() %>% 
- set_header_labels(
-  svy_year = 'NHANES cycle',
-  htn_accaha = 'Hypertension',
-  demo_race = 'Race/Ethnicity',
-  estimate = 'Estimate',
-  std_error = 'Standard error',
-  ci_lower = 'CI lower',
-  ci_upper = 'CI upper',
-  n_obs = 'N observations'
+ theme_box() %>% 
+ width(width = 0.875) %>% 
+ width(j = ~demo_race, width = 1) %>% 
+ width(j = ~statistic, width = 1) %>% 
+ width(j = ~unreliable_reason, width = 0.7) %>% 
+ width(j = ~review_reason, width = 0.7) %>% 
+ align(align = 'center', part = 'all') %>% 
+ align(j = 1, align = 'center', part = 'all') %>% 
+ add_footer_lines(
+  values = "Table format is identical to the web application's output."
  ) %>% 
- theme_box() %>% 
- width(width = 1) %>% 
- align(align = 'center', part = 'all') %>% 
- align(j = 1, align = 'center', part = 'all')
+ add_footer_lines(
+  values = "svy_year indicates the NHANES cycle"
+ ) %>% 
+ add_footer_lines(
+  values = "htn_accaha indicates hypertension status." 
+ ) %>% 
+ add_footer_lines(
+  values = "demo_race indicates the race subgroup."
+ ) %>% 
+ add_footer_lines(
+  values = "statistic indicates the type of statistic computed."
+ ) %>% 
+ add_footer_lines(
+  values = "estimate is the point estimate of the given statistic"
+ ) %>% 
+ add_footer_lines(
+  values = "std_error is the standard error for estimate"
+ ) %>% 
+ add_footer_lines(
+  values = "ci_lower is the lower bound of a 95% confidence interval for estimate"
+ ) %>% 
+ add_footer_lines(
+  values = "ci_upper is the upper bound of a 95% confidence interval for estimate"
+ ) %>% 
+ add_footer_lines(
+  values = "n_obs is the number of observations (unweighted) used for computations"
+ ) %>% 
+ add_footer_lines(
+  values = "unreliable_status is TRUE if the result is unreliable, FALSE otherwise"
+ ) %>% 
+ add_footer_lines(
+  values = "unreliable_reason is blank if the result is reliable, and lists reason(s) for unreliability otherwise"
+ ) %>% 
+ add_footer_lines(
+  values = "review_needed is TRUE if the result should be reviewed, FALSE otherwise"
+ ) %>% 
+ add_footer_lines(
+  values = "review_reason is blank if the result does not require review, and lists reason(s) for requiring review otherwise"
+ )
 
@@
 
-### Figure 1 Mean systolic blood pressure for US adults by calendar year
+### Figure 1: Mean systolic blood pressure for US adults by calendar year
 
@@
 
-### Figure 2 Age-adjusted prevalence of hypertension for US adults by calendar year
+### Figure 2: Age-adjusted prevalence of hypertension for US adults by calendar year
 
@@
 
-### Figure 3 Number of US adults with hypertension
+### Figure 3: Number of US adults with hypertension
 
@@
 
-### Figure 4 Age-adjusted prevalence of blood pressure control by calendar year.
+### Figure 4: Age-adjusted prevalence of blood pressure control by calendar year.
 

### 2022-10-28 README.Rmd
paper v7

@@
 Our goal is to provide a platform for exploration of the National Health and Nutrition Examination Survey (NHANES) data.
-You can find the online application for exploratory data here: https://bcjaeger.shinyapps.io/nhanesShinyBP/
+We created this website to give context and documentation for users of the application, which can be accessed here: https://bcjaeger.shinyapps.io/nhanesShinyBP/
 
-## Installation
+## Examples
 
-You can install the development version of cardioStatsUSA from [GitHub](https://github.com/) with:
+Below are a few graphs produced by the platform.
+Additional graphs can be found by selecting the “Results” tab above.
+Instructions on how to generate statistics using the application are available by clicking on the “Get Started” tab.
 
-``` r
-# install.packages("remotes")
-remotes::install_github("jhs-hwg/cardioStatsUSA")
+**Systolic blood pressure by age group in the last three NHANES cycles.**
+
+```{r, echo = FALSE, fig.height=10, fig.width=10}
+library(cardioStatsUSA)
+nhanes_visualize(outcome_variable = 'bp_sys_mean', 
+                 group_variable = 'demo_age_cat',
+                 outcome_stats = c("mean"),
+                 time_values = c('2013-2014', '2015-2016', '2017-2020'), 
+                 title = "",
+                 width = 800, 
+                 height = 600)[[1]]
+
 ```
 
-## Example
+**Prevalence of blood pressure control by sex among adults with hypertension.**
 
-The shiny application is hosted on a public server, but you can also run it locally:
-
-```{r, eval=FALSE}
-
-library(cardioStatsUSA)
-app_run()
+```{r, echo = FALSE, fig.height=10, fig.width=10}
+nhanes_visualize(outcome_variable = 'bp_control_accaha', 
+                 group_variable = 'demo_gender',
+                 outcome_stats = c("percentage"),
+                 subset_calls = c("htn_accaha" = "Yes"),
+                 time_values = 'all', 
+                 title = "",
+                 width = 800, 
+                 height = 600)[[1]]
 
 ```
+
+<!-- ## Installation -->
+
+<!-- You can install the development version of cardioStatsUSA from [GitHub](https://github.com/) with: -->
+
+<!-- ``` r -->
+<!-- # install.packages("remotes") -->
+<!-- remotes::install_github("jhs-hwg/cardioStatsUSA") -->
+<!-- ``` -->
+
+<!-- ## Example -->
+
+<!-- The shiny application is hosted on a public server, but you can also run it locally: -->
+
+<!-- ```{r, eval=FALSE} -->
+
+<!-- library(cardioStatsUSA) -->
+<!-- app_run() -->
+
+<!-- ``` -->
 
@@
 
-## Features of the application
+<!-- ## Features of the application -->
 
-Using the web application, we estimated characteristics of the population which are relevant when investigating BP and hypertension data.
+<!-- Using the web application, we estimated characteristics of the population which are relevant when investigating BP and hypertension data.
 As the application was designed with a number of options so that analyses can be customized to the user’s specifications.
@@
 Analyses can be done in sub-groups and stratified by participant characteristics.
+-->
 
-Full summaries of this application's features and associated tutorials are publicly available.
+<!-- Full summaries of this application's features and associated tutorials are publicly available.
 Briefly, selected NHANES cycles can be analyzed separately, and contiguous cycles can be pooled together to increase reliability of statistical estimates.
@@
 Following CDC recommendations, the application automatically suppresses statistics which are unreliable, e.g., due to being based on a small sample size.
+-->
 
-## Validation of results from the shiny application
+<!-- ## Validation of results from the shiny application -->
 
-We have replicated results from prior NHANES publications to verify that routines used to conduct analyses in this app are valid.
+<!-- We have replicated results from prior NHANES publications to verify that routines used to conduct analyses in this app are valid.
 Based on these tests, we have identified some cases where our app does not exactly match previously published articles, and we note those cases here.
+-->
 
-TO BE WRITTEN: notes on tests showing exact match with [Muntner et al](https://jamanetwork.com/journals/jama/fullarticle/2770254)
+<!-- TO BE WRITTEN: notes on tests showing exact match with [Muntner et al](https://jamanetwork.com/journals/jama/fullarticle/2770254) -->
 
-TO BE WRITTEN: notes on discrepancies with results in [Carey et al.](https://www.ahajournals.org/doi/10.1161/HYPERTENSIONAHA.118.12191)
+<!-- TO BE WRITTEN: notes on discrepancies with results in [Carey et al.](https://www.ahajournals.org/doi/10.1161/HYPERTENSIONAHA.118.12191)  -->

### 2022-10-28 paper/manuscript.Rmd
paper v7

@@
 
-The National Health and Nutrition Examination Survey (NHANES) is a program conducted by the US National Center for Health Statistics (NCHS) of the Centers for Disease Control and Prevention (CDC) and is designed to assess the ongoing health and nutritional status of the non-institutionalized US population.
-NHANES data have been analyzed to provide hypertension statistics for non-institutionalized US adults with important policy and public health implications.
-For example, NHANES data have been used to estimate the impact of the lower blood pressure (BP) levels that define hypertension and controlled BP in the 2017 American College of Cardiology/American Heart Association (ACC/AHA) BP guideline versus the Seventh Report of the Joint National Committee on Prevention, Detection, Evaluation, and Treatment of High BP (JNC7) guideline.[@whelton20182017; @chobanian2003seventh] Additionally, NHANES data have been used to track the proportion of US adults with hypertension that have controlled BP, which were included in a Call-to-Action to Control BP from the US Surgeon General.[@us_surgeon_2020]
+The National Health and Nutrition Examination Survey (NHANES) is a program conducted by the US National Center for Health Statistics (NCHS) of the Centers for Disease Control and Prevention (CDC) and is designed to assess the ongoing health and nutritional status of the non-institutionalized US population.[@nhanes_about] NHANES data have been analyzed to provide hypertension statistics for non-institutionalized US adults with important policy and public health implications.
+For example, NHANES data have been used to estimate the impact of the lower blood pressure (BP) levels that define hypertension and controlled BP in the 2017 American College of Cardiology/American Heart Association (ACC/AHA) BP guideline versus the Seventh Report of the Joint National Committee on Prevention, Detection, Evaluation, and Treatment of High BP (JNC7) guideline.[@muntner_potential_2018] Additionally, NHANES data have been used to track the proportion of US adults with hypertension that have controlled BP, which were included in a Call-to-Action to Control BP from the US Surgeon General.[@us_surgeon_2020]
 
@@
 
-NHANES data are publicly available and accessible through the CDC website.[@nhanes_home] However, analyzing NHANES data may be challenging for a number of reasons.
+NHANES data are publicly available and accessible through the CDC website.[@nhanes_home; @nhanes_data] However, analyzing NHANES data may be challenging for a number of reasons.
 Specific statistical techniques are required to account for the multi-stage sampling design used to select NHANES participants and analyses need to be weighted to produce nationally representative estimates.
@@
 Of relevance to the application, blood samples were used to measure total and high-density lipoprotein cholesterol, glycated hemoglobin and serum creatinine, and the urine sample was used to measure albumin and creatinine and to conduct a pregnancy test.
-The protocol for obtaining BP measurements is available online.
-In brief, systolic and diastolic BP (SBP and DBP, respectively) were measured three times by trained and certified physicians.
+The protocol for obtaining BP measurements is available online.[@ostchega_national_2003] In brief, systolic and diastolic BP (SBP and DBP, respectively) were measured three times by trained and certified physicians.
 BP was measured using a mercury sphygmomanometer from 1999-2000 through 2015-2016 and using an oscillometric device in 2017-2020.
 The mean SBP and DBP levels were computed over all available measurements for each participant.
-The oscillometric SBP and DBP values were calibrated to the mercury device.[@muntner2022blood] Antihypertensive medication classes were defined using those listed in the 2017 ACC/AHA BP guideline.[@whelton20182017] For the current analysis, we defined hypertension, BP control, and resistant hypertension according to the 2017 ACC/AHA BP guideline.
-The application also has these variables defined according to the JNC7 definitions.
-Chronic kidney disease (CKD) was defined by estimated glomerular filtration rate < 60 ml/min/1.73 m^2^ calculated using the 2021 serum creatinine-based equation or albumin-to-creatinine ratio ≥ 30 mg/g.[@inker_new_2021] A list of BP, hypertension, and antihypertensive medication variables is provided in **Table 1** with full definitions for all variables provided in **Table S1**.
+The oscillometric SBP and DBP values were calibrated to the mercury device.[@muntner2022blood; @ostchega2021comparing] Antihypertensive medication classes were defined using those listed in the 2017 ACC/AHA BP guideline.[@whelton20182017] For the current analysis, we defined hypertension, BP control, and resistant hypertension according to the 2017 ACC/AHA BP guideline.[@whelton20182017] The application also has these variables defined according to the JNC7 definitions.[@chobanian2003seventh] Chronic kidney disease (CKD) was defined by estimated glomerular filtration rate < 60 ml/min/1.73 m^2^ calculated using the 2021 serum creatinine-based equation or albumin-to-creatinine ratio ≥ 30 mg/g.[@inker_new_2021] A list of BP, hypertension, and antihypertensive medication variables is provided in **Table 1** with full definitions for all variables provided in **Table S1**.
 
@@
 We restricted the dataset to adults ≥ 18 years of age.
-This exclusion was applied because statistics for BP levels and hypertension in children and adolescents are markedly different than for adults (**What should we cite here?** Perhaps this?
-doi:10.1001/jama.291.17.2107).
-We further restricted the population to participants who completed the in-home interview and study examination, with one or more SBP and DBP measurement, and who had data on self-reported antihypertensive medication use.
+This exclusion was applied because statistics for BP levels and the prevalence of hypertension in children and adolescents are markedly different than for adults.[@muntner2004trends] We further restricted the population to participants who completed the in-home interview and study examination, with one or more SBP and DBP measurement, and who had data on self-reported antihypertensive medication use.
 After these exclusions were applied, the population in this application included 56,035 participants (**Figure S1**).
@@
 When population count estimates are requested, survey weights are calibrated to account for missing information on SBP, DBP or self-reported antihypertensive medication use.
-Users may tabulate or visualize summaries, and can present results for the overall population or in subgroups defined by a stratifying variable.
+Users may tabulate or visualize summaries, and can present results for the overall population or in subgroups.
 All figures and datasets created with the web application can be downloaded and saved.
@@
 For age adjustment, we used the estimated age distribution of US adults from 1999 to 2020 as the standard (49.3%, 33.6%, 10.1% and 7.0% being 18 to 44, 45 to 64, 65 to 74 and ≥ 75 years of age, respectively).
-To illustrate how results from the application appear when saved in various formats (e.g., comma separated, portable document format, Microsoft Excel), we tabulated the estimated race/ethnicity distribution of US adults with and without hypertension, separately.
+We created a table of the estimated race/ethnicity distribution of US adults with and without hypertension, separately.
 We demonstrated stratification by estimating the prevalence of hypertension by NHANES cycle for US adults with and without CKD.
@@
 Specifically, we estimated the age-adjusted proportion of US adults with controlled BP by NHANES cycle among non-pregnant US adults with hypertension, overall and among those who self-reported taking antihypertensive medication.
-We also estimated the age-adjusted prevalence of resistant hypertension by NHANES cycle for US adults with hypertension taking antihypertensive medication and among US adults with hypertension taking ≥ 3 classes of antihypertensive medication.
-For age adjustment in the analysis of resistant hypertension, we set the age distribution for the standard population to represent US adults with hypertension from 1999 to 2020: 26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥ 75 years of age, respectively.
+We also estimated the age-adjusted prevalence of resistant hypertension by NHANES cycle for non-pregnant US adults with hypertension who self-reported taking antihypertensive medication and had ≥ 1 classes of antihypertensive medication identified during the medication inventory and among those with ≥ 3 classes of antihypertensive medication identified during the medication inventory.
+For age adjustment in the analysis of BP control and resistant hypertension, we set the age distribution for the standard population to represent US adults with hypertension from 1999 to 2020: 26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥ 75 years of age, respectively.
 
@@
 In 2017-2020, a higher percentage of US adults with versus without hypertension were non-Hispanic Black (13.5% versus 9.3%) while a lower percentage of US adults with versus without hypertension were Hispanic (12.3% versus 18.1%) (**Table 2**).
-Among pregnant women, the distribution of BP categories could not be estimated reliably in 2017-2020 (**Figure S3; Panel A**) except for the group with SBP of 130 to < 140 mm Hg or DBP of 80 to < 90 mm Hg, but was stable for all groups after pooling NHANES cycles from 2009-2010 through 2017-2020 (**Figure S3; Panel B**)
+Among pregnant women, the distribution of BP categories could not be estimated reliably in 2017-2020 (**Figure S3; Panel A**), but it could be estimated reliably after pooling NHANES cycles from 2009-2010 through 2017-2020 (**Figure S3; Panel B**)
 
@@
 
-The age-adjusted prevalence of BP control among US adults with hypertension was lowest in 1999-2000 (9.6%), highest in 2013-2014 (24.6%), and 21.7% in 2017-2020 (**Figure 4; panel A**).
-Among US adults who self-reported taking antihypertensive medication, 27.4%, 48.5%, and 42.8% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively (**Figure 4; panel B**).
-The age-adjusted prevalence of resistant hypertension among US adults who self-reported taking antihypertensive medication was lowest in 1999-2000 (13.7%), highest in 2005-2006 (20.4), and 16.2% in 2017-2020 (**Figure S4; panel A**).
-Among US adults who were taking three or more classes of antihypertensive medication, the prevalence of resistant hypertension was lowest in 2009-2010 (64.1%), highest in 2003-2004 (78.3%), and 67.9% in 2017-2020 (**Figure S4; panel B**).
+The age-adjusted prevalence of BP control among non-pregnant US adults with hypertension was lowest in 1999-2000 (9.6%), highest in 2013-2014 (24.5%), and 21.7% in 2017-2020 (**Figure 4; panel A**).
+Among US adults who self-reported taking antihypertensive medication, 27.6%, 48.4%, and 42.8% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively (**Figure 4; panel B**).
+The age-adjusted prevalence of resistant hypertension among non-pregnant US adults taking antihypertensive medication was lowest in 1999-2000 (14.3%), highest in 2005-2006 (21.4), and 16.6% in 2017-2020 (**Figure S4; panel A**).
+Among non-pregnant US adults who were taking three or more classes of antihypertensive medication, the prevalence of resistant hypertension was lowest in 2009-2010 (64.1%), highest in 2003-2004 (78.3%), and 67.9% in 2017-2020 (**Figure S4; panel B**).
 
@@
 Medication classes were coded using generic names and the drug classes in the 2017 ACC/AHA BP guideline.
-We recognize the NCHS recommends using Lexicon Plus®, a proprietary database, to categorize medication classes.
-While the categorization of most medications is identical using generic drug names and Lexicon Plus®, some differences exist.
+We recognize the NCHS recommends using Lexicon Plus®, a proprietary database, to categorize medication classes.[@nhanes_rxq_drug] While the categorization of most medications is identical using generic drug names and Lexicon Plus®, some differences exist.
 Many additional decisions were made regarding the definitions of variables, inclusion of study participants, and analytic approach.
@@
 We included spironolactone as a single drug class, an aldosterone antagonist.
-Additionally, Lexicon Plus® includes Sotalol as a beta blocker and nitroglycerine as a direct vasodilator and we did not include these drugs as antihypertensive medications as they are not listed in the 2017 ACC/AHA BP guideline.
-The differences in results between the application and this previously published manuscript emphasize that it may not be possible to always generate results that are identical to prior publications.
+Additionally, Lexicon Plus® includes Sotalol as a beta blocker and nitroglycerine as a direct vasodilator and we did not include these drugs as antihypertensive medications as they are not listed in the 2017 ACC/AHA BP guideline.[@whelton20182017] The differences in results between the application and this previously published manuscript emphasize that it may not be possible to always generate results that are identical to prior publications.
 Users should be aware of the choices made in defining variables as this may affect the statistical estimates generated using the application.
@@
 This study also has several limitations.
-NHANES participants had their mean BP measured during a single visit, and BP guidelines recommend obtaining mean BP using at least two BP values measured on separate days.
-The response rate for NHANES has declined from 1999-2000 through 2017-2020.
+NHANES participants had their mean BP measured during a single visit, and BP guidelines recommend obtaining mean BP using at least two BP values measured on separate days.[@whelton20182017] The response rate for NHANES has declined from 1999-2000 through 2017-2020.
 Effects of this decline are unclear.
@@
 ```
-Age adjustment was performed through direct standardization
+Age adjustment was performed through direct standardization, using the estimated age distribution of US adults from 1999 to 2020 as the standard (49.3%, 33.6%, 10.1% and 7.0% being 18 to 44, 45 to 64, 65 to 74 and ≥ 75 years of age, respectively)
 
@@
 
-### Figure 3: Number of US adults with hypertension
+### Figure 3: Number of US adults with hypertension by calendar year
 
@@
 
-A. Among those with hypertension
+A. Among non-pregnant US adults with hypertension
 ```{r, fig.height=3.75}
 
-p1 <- nhanes_visualize(data = nhanes_data,
+p1 <- 
+ nhanes_visualize(
+  data = nhanes_data,
   key = nhanes_key,
   outcome_variable = 'bp_control_accaha',
-                       subset_calls = list("htn_accaha" = "Yes"),
+  subset_calls = list("htn_accaha" = "Yes",
+                      "demo_pregnant" = c("No", "Missing")),
   standard_weights = standard_weights_htn,
@@
 
-p2 <- nhanes_visualize(data = nhanes_data,
+p2 <- 
+ nhanes_visualize(
+  data = nhanes_data,
   key = nhanes_key,
@@
   subset_calls = list("htn_accaha" = "Yes",
-                                           "bp_med_use" = "Yes"),
+                      "bp_med_use" = "Yes",
+                      "demo_pregnant" = c("No", "Missing")),
   standard_weights = standard_weights_htn,
   time_values = 'all', 
-                       geom = 'bar')
+  geom = 'bar'
+ )
 
@@
 ```
+Age adjustment was performed through direct standardization, using the estimated age distribution of US adults with hypertension from 1999 to 2020 as the standard (26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥ 75 years of age, respectively)
 
-B. Among those who self-report taking antihypertensive medication 
+\newpage
+
+B. Among non-pregnant US adults who self-report taking antihypertensive medication 
 ```{r, fig.height=3.75}
@@
 ```
-Age adjustment was performed through direct standardization
-
-<!-- \newpage -->
-
-<!-- **Figure 4** Distribution of the number of antihypertensive medication classes being taken among those with hypertension by calendar period -->
-
-<!-- A. Among those with hypertension -->
-<!-- ```{r, fig.height=3.75} -->
-
-<!-- p1 <- nhanes_visualize(data = nhanes_data, -->
-<!--                        key = nhanes_key, -->
-<!--                        outcome_variable = 'bp_med_n_class', -->
-<!--                        subset_calls = list("htn_accaha" = "Yes"), -->
-<!--                        time_values = 'all', -->
-<!--                        geom = 'bar') -->
-
-<!-- p2 <- nhanes_visualize(data = nhanes_data, -->
-<!--                        key = nhanes_key, -->
-<!--                        outcome_variable = 'bp_med_n_class', -->
-<!--                        subset_calls = list("htn_accaha" = "Yes", -->
-<!--                                            "bp_med_use" = "Yes"), -->
-<!--                        time_values = 'all', -->
-<!--                        geom = 'bar') -->
-
-<!-- fig_print(p1) -->
-
-<!-- ``` -->
-
-<!-- B. Among those with hypertension who self-report taking antihypertensive medication -->
-<!-- ```{r, fig.height=3.75} -->
-<!-- fig_print(p2) -->
-<!-- ``` -->
+Age adjustment was performed through direct standardization, using the estimated age distribution of US adults with hypertension from 1999 to 2020 as the standard (26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥ 75 years of age, respectively)
 
@@
 
-### Figure S2: Prevalence of hypertension among US adults with and without chronic kidney disease
+### Figure S2: Prevalence of hypertension among US adults with and without chronic kidney disease by calendar year
 
@@
 
+\newpage
+
 B. In 2009-2010 through 2017-2020
@@
 ```
+The estimate for SBP of 140 to < 160 or 90 to < 100 mm Hg is 1.1%. This can be seen on the application by hovering over the red bar with your mouse.
 
@@
 
-A. Among those who self report taking antihypertensive medication 
+A. Among non-pregnant US adults who self report taking antihypertensive medication and are taking 1 or more antihypertensive medication classes 
 ```{r, fig.height=3.75}
 
-p1 <- nhanes_visualize(data = nhanes_data,
+p1 <- nhanes_visualize(
+ data = nhanes_data,
  key = nhanes_key,
  outcome_variable = 'htn_resistant_accaha',
-                       subset_calls = list("bp_med_use" = "Yes"),
+ subset_calls = list(
+  "demo_pregnant" = c("No", "Missing"),
+  "bp_med_use" = "Yes",
+  "bp_med_n_class" = c('One', 'Two', 'Three', 'Four or more')
+ ),
  time_values = 'all', 
-                       geom = 'bar')
+ geom = 'bar'
+)
 
@@
  outcome_variable = 'htn_resistant_accaha',
- subset_calls = list("bp_med_use" = "Yes",
-                     "bp_med_n_class" = c('Three', 'Four or more')),
+ subset_calls = list(
+  "demo_pregnant" = c("No", "Missing"),
+  "bp_med_use" = "Yes",
+  "bp_med_n_class" = c('Three', 'Four or more')
+ ),
  time_values = 'all', 
@@
 ```
+Age adjustment was performed through direct standardization, using the estimated age distribution of US adults with hypertension from 1999 to 2020 as the standard (26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥ 75 years of age, respectively)
 
-B. Among those who self report taking antihypertensive medication and are taking 3 or more classes.
+\newpage
+
+B. Among non-pregnant US adults who self report taking antihypertensive medication and are taking 3 or more antihypertensive medication classes
 ```{r, fig.height=3.75}
@@
 ```
-Age adjustment was performed through direct standardization
+Age adjustment was performed through direct standardization, using the estimated age distribution of US adults with hypertension from 1999 to 2020 as the standard (26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥ 75 years of age, respectively)
 

### 2022-10-28 R/nhanes_design_subset.R

@@ -30,10 +30,14 @@ nhanes_design_subset <- function(x, subset){
        " subset(B)")
  }
 
+
+
+
  e <- substitute(subset)
  r <- eval(e, x$design$variables, parent.frame())
  r <- r & !is.na(r)
  x$design <- x$design[r, ]
+ # TODO: make subset_rows a list so we can do multiple subsets
  x$subset_rows <- which(r)
  x$operations <- c(x$operations, 'subset')
  x

### 2022-10-28 R/plotly_viz_make_hover.R

@@ -15,38 +15,60 @@ plotly_viz_make_hover <- function(data, stat_all, group, group_label) {
  }
 
  data %>%
-  as_tibble() %>%
-  mutate(across(all_of(numeric_cols), table_value)) %>%
-  mutate(
-   stat_show = if_else(
-    review_needed,
-    paste0(" -Review needed-<br>  Reasons: ", review_reason),
-    ""
-   ),
-   stat_show = if_else(
-    unreliable_status,
-    paste0(" -Unreliable-<br>  Reasons: ", unreliable_reason),
-    stat_show
-   ),
-   stat_show = if_else(
-    stat_show == "",
-    as.character(glue("{estimate} (95% CI {ci_lower}, {ci_upper})")),
-    stat_show
-   ),
-   stat_label = glue(
-    "{stat_recode(statistic)}: {stat_show}"
-   )
-  ) %>%
-  select(-all_of(numeric_cols),
-         -c(unreliable_reason,
-            unreliable_status,
-            review_reason,
-            review_needed,
-            stat_show)) %>%
-  pivot_wider(names_from = statistic, values_from = stat_label) %>%
-  unite(col = 'hover', !!!stat_cols, sep = '<br>') %>%
-  mutate(hover = glue("{group_label}{label_sep}{hover}")) %>%
-  as.data.table()
+  .[, (numeric_cols) := lapply(.SD, table_value), .SDcols = numeric_cols] %>%
+  .[,
+    stat_show := fifelse(
+     unreliable_status,
+     yes = paste0(" -Unreliable-<br>  Reasons: ", unreliable_reason),
+     no = ""
+    )] %>%
+  .[stat_show == "",
+    stat_show := paste0(estimate, " (95% CI ", ci_lower, ", ", ci_upper, ")")
+  ] %>%
+  .[, stat_label := paste(stat_recode(statistic), stat_show, sep = ": ")] %>%
+  .[, c(numeric_cols,
+        "unreliable_reason",
+        "unreliable_status",
+        "review_reason",
+        "review_needed",
+        "stat_show") := NULL] %>%
+  dcast(data = ., formula = ... ~ statistic, value.var = "stat_label") %>%
+  .[, hover := do.call(paste, c(.SD, sep="<br>")), .SDcols = stat_cols] %>%
+  .[, hover := paste0(group_label, label_sep, hover)]
+
+ # data %>%
+ #  as_tibble() %>%
[…]

### 2022-10-28 data-raw/muntner_hypertension_2022_table_s2.R

@@ -1,0 +1,42 @@
+
+
+muntner_hypertension_2022_table_s2 <- tibble::tribble(
+ ~ svy_year, ~variable, ~group, ~estimate, ~ci_lower, ~ci_upper,
+ "2009-2010", "Overall", "Overall", 53.5, 50.2, 56.7,
+ "2011-2012", "Overall", "Overall", 52.2, 47.7, 56.7,
+ "2013-2014", "Overall", "Overall", 54.1, 48.9, 59.3,
+ "2015-2016", "Overall", "Overall", 48.6, 44.4, 52.7,
+ "2017-2020", "Overall", "Overall", 48.2, 45.7, 50.8,
+ "2009-2010", "Age group, years", "18 to 44", 35.5, 26.8, 44.2,
+ "2011-2012", "Age group, years", "18 to 44", 42.2, 31.5, 52.9,
+ "2013-2014", "Age group, years", "18 to 44", 44.6, 38.2, 50.9,
+ "2015-2016", "Age group, years", "18 to 44", 40.0, 33.9, 46.1,
+ "2017-2020", "Age group, years", "18 to 44", 40.3, 30.7, 49.9,
+ "2009-2010", "Age group, years", "45 to 64", 57.1, 52.6, 62.0,
+ "2011-2012", "Age group, years", "45 to 64", 56.2, 49.6, 62.9,
+ "2013-2014", "Age group, years", "45 to 64", 57.0, 49.8, 64.2,
+ "2015-2016", "Age group, years", "45 to 64", 53.7, 48.1, 59.2,
+ "2017-2020", "Age group, years", "45 to 64", 52.0, 47.9, 56.2,
+ "2009-2010", "Age group, years", "65 to 74", 59.2, 54.6, 63.7,
+ "2011-2012", "Age group, years", "65 to 74", 57.8, 51.1, 64.5,
+ "2013-2014", "Age group, years", "65 to 74", 59.2, 54.7, 63.6,
+ "2015-2016", "Age group, years", "65 to 74", 51.5, 44.4, 58.5,
+ "2017-2020", "Age group, years", "65 to 74", 54.2, 49.3, 59.0,
+ "2009-2010", "Age group, years", "75+", 50.3, 44.1, 56.5,
+ "2011-2012", "Age group, years", "75+", 42.0, 33.4, 50.6,
[…]

### 2022-10-28 data-raw/muntner_jama_2020_etable_1.R

@@ -1,0 +1,110 @@
+
+
+muntner_jama_2020_etable_1 <- tibble::tribble(
+ ~svy_year,   ~group              , ~estimate, ~ci_lower, ~ci_upper,
+ "1999-2000", "Overall"           ,  53.4    ,  49.0    ,  57.9    ,
+ "2001-2002", "Overall"           ,  58.6    ,  56.0    ,  61.2    ,
+ "2003-2004", "Overall"           ,  62.1    ,  57.9    ,  66.2    ,
+ "2005-2006", "Overall"           ,  65.5    ,  62.0    ,  69.0    ,
+ "2007-2008", "Overall"           ,  68.3    ,  65.9    ,  70.8    ,
+ "2009-2010", "Overall"           ,  69.7    ,  67.4    ,  72.1    ,
+ "2011-2012", "Overall"           ,  70.3    ,  66.8    ,  73.8    ,
+ "2013-2014", "Overall"           ,  72.2    ,  68.6    ,  75.8    ,
+ "2015-2016", "Overall"           ,  66.8    ,  63.0    ,  70.6    ,
+ "2017-2018", "Overall"           ,  64.8    ,  61.3    ,  68.3    ,
+ "1999-2000", "18 to 44"          ,  66.1    ,  54.7    ,  77.5    ,
+ "2001-2002", "18 to 44"          ,  70.8    ,  60.4    ,  81.1    ,
+ "2003-2004", "18 to 44"          ,  81.7    ,  73.2    ,  90.2    ,
+ "2005-2006", "18 to 44"          ,  77.2    ,  67.1    ,  87.4    ,
+ "2007-2008", "18 to 44"          ,  83.7    ,  78.6    ,  88.7    ,
+ "2009-2010", "18 to 44"          ,  65.5    ,  56.0    ,  74.9    ,
+ "2011-2012", "18 to 44"          ,  83.3    ,  74.2    ,  92.4    ,
+ "2013-2014", "18 to 44"          ,  81.0    ,  76.0    ,  85.9    ,
+ "2015-2016", "18 to 44"          ,  70.9    ,  63.2    ,  78.6    ,
+ "2017-2018", "18 to 44"          ,  73.3    ,  60.5    ,  86.0    ,
+ "1999-2000", "45 to 64"          ,  60.6    ,  55.1    ,  66.1    ,
+ "2001-2002", "45 to 64"          ,  63.3    ,  58.6    ,  68.1    ,
[…]

### 2022-10-28 data-raw/muntner_jama_2020_table_2.R

@@ -1,0 +1,56 @@
+
+muntner_jama_2020_table_2 <- tibble::tribble(
+ ~svy_year,   ~group,        ~estimate , ~ci_lower, ~ci_upper,
+ "1999-2000", "lt_120_80",    9.2      ,  7.4     ,  10.9    ,
+ "2001-2002", "lt_120_80",    12.5     ,  10.0    ,  15.0    ,
+ "2003-2004", "lt_120_80",    13.1     ,  11.2    ,  14.9    ,
+ "2005-2006", "lt_120_80",    14.9     ,  12.8    ,  17.0    ,
+ "2007-2008", "lt_120_80",    18.2     ,  15.8    ,  20.6    ,
+ "2009-2010", "lt_120_80",    21.4     ,  19.4    ,  23.4    ,
+ "2011-2012", "lt_120_80",    18.4     ,  14.7    ,  22.1    ,
+ "2013-2014", "lt_120_80",    20.2     ,  16.8    ,  23.6    ,
+ "2015-2016", "lt_120_80",    17.2     ,  13.5    ,  20.9    ,
+ "2017-2018", "lt_120_80",    15.8     ,  12.9    ,  18.6    ,
+ "1999-2000", "gt_120_lt_80", 6.5      ,  5.3     ,  7.6     ,
+ "2001-2002", "gt_120_lt_80", 7.1      ,  5.7     ,  8.5     ,
+ "2003-2004", "gt_120_lt_80", 9.6      ,  8.0     ,  11.3    ,
+ "2005-2006", "gt_120_lt_80", 10.7     ,  8.5     ,  12.8    ,
+ "2007-2008", "gt_120_lt_80", 12.3     ,  10.0    ,  14.5    ,
+ "2009-2010", "gt_120_lt_80", 12.5     ,  10.9    ,  14.0    ,
+ "2011-2012", "gt_120_lt_80", 15.8     ,  13.8    ,  17.8    ,
+ "2013-2014", "gt_120_lt_80", 14.6     ,  12.5    ,  16.7    ,
+ "2015-2016", "gt_120_lt_80", 14.3     ,  11.6    ,  16.9    ,
+ "2017-2018", "gt_120_lt_80", 11.3     ,  9.4     ,  13.2    ,
+ "1999-2000", "gt_130_80",    16.2     ,  12.4    ,  19.9    ,
+ "2001-2002", "gt_130_80",    15.3     ,  13.0    ,  17.5    ,
[…]

### 2022-10-28 tests/testthat/test-replicate_muntner_jama_2020.R

@@ -125,112 +125,8 @@ shiny_answers_etable_1
 #' match their answers. There isn't really an easy way to load these data
 #' programmatically, so they are written manually into a data set here.
 #+ echo = TRUE
-jama_etable_1 <- tribble(
- ~svy_year,   ~group              , ~estimate, ~ci_lower, ~ci_upper,
- "1999-2000", "Overall"           ,  53.4    ,  49.0    ,  57.9    ,
- "2001-2002", "Overall"           ,  58.6    ,  56.0    ,  61.2    ,
- "2003-2004", "Overall"           ,  62.1    ,  57.9    ,  66.2    ,
- "2005-2006", "Overall"           ,  65.5    ,  62.0    ,  69.0    ,
- "2007-2008", "Overall"           ,  68.3    ,  65.9    ,  70.8    ,
- "2009-2010", "Overall"           ,  69.7    ,  67.4    ,  72.1    ,
- "2011-2012", "Overall"           ,  70.3    ,  66.8    ,  73.8    ,
- "2013-2014", "Overall"           ,  72.2    ,  68.6    ,  75.8    ,
- "2015-2016", "Overall"           ,  66.8    ,  63.0    ,  70.6    ,
- "2017-2018", "Overall"           ,  64.8    ,  61.3    ,  68.3    ,
- "1999-2000", "18 to 44"          ,  66.1    ,  54.7    ,  77.5    ,
- "2001-2002", "18 to 44"          ,  70.8    ,  60.4    ,  81.1    ,
- "2003-2004", "18 to 44"          ,  81.7    ,  73.2    ,  90.2    ,
- "2005-2006", "18 to 44"          ,  77.2    ,  67.1    ,  87.4    ,
- "2007-2008", "18 to 44"          ,  83.7    ,  78.6    ,  88.7    ,
- "2009-2010", "18 to 44"          ,  65.5    ,  56.0    ,  74.9    ,
- "2011-2012", "18 to 44"          ,  83.3    ,  74.2    ,  92.4    ,
- "2013-2014", "18 to 44"          ,  81.0    ,  76.0    ,  85.9    ,
- "2015-2016", "18 to 44"          ,  70.9    ,  63.2    ,  78.6    ,
[…]

### 2022-11-05 paper/manuscript.Rmd
progressing validation + tutorials

@@
   "AHA"    = "American Heart Association",
-  "CVD"  = 'cardiovascular disease'
+  "ASCVD"  = "atherosclerotic cardiovascular disease",
+  "CVD"    = 'cardiovascular disease',
+  "HbA1c"  = "hemoglobin A1C",
+  "ACE"    = "angiotensin-converting enzyme",
+  "NHANES" = "National Health and Nutrition Examination Survey"
 )
@@
 Ligong Chen, PhD,^2^
-Adam P. Bress, PharmD, MS,^3,4^ 
 Kathryn Foti,^2^
 Shakia T. Hardy,^2^
+Adam P. Bress, PharmD, MS,^3,4^ 
+Sean P. Kane, PharmD, BCPS,^5^
+Jennifer S. Herrick, MS,^3,6^
+Catherine Derington, PharmD, MS,^7^ 
 Bharat Poudel, MSPH,^2^
@@
 
+**AUTHOR ORDER NOT FINALIZED**
+
 ^1^Department of Biostatistics and Data Science, Wake Forest School of Medicine, Winston-Salem, NC. `r single_par`
@@
 ^4^Department of Population Health Sciences, University of Utah School of Medicine, Salt Lake City, UT. `r single_par`
-
+^5^Department of Pharmacy Practice, Rosalind Franklin University of Medicine and Science, North Chicago, IL, USA
+^6^Department of Internal Medicine, University of Utah School of Medicine, Salt Lake City, UT
+^7^ University of Colorado Skaggs School of Pharmacy and Pharmaceutical Sciences, Aurora, CO, USA
 \newpage
@@
 **Background**: Data from the US National Health and Nutrition Examination Survey (NHANES) are freely available and can be analyzed to produce hypertension statistics for the non-institutionalized US population.
-However, the analysis of these data requires statistical programming knowledge.
+The analysis of these data requires statistical programming knowledge.
 
@@
 Population counts can also be estimated.
-To demonstrate the application’s capabilities, we estimated hypertension statistics for US adults.
+To demonstrate the application’s capabilities, we estimated hypertension statistics for US adults.The application was validated by replicating statistics from three publications.
 
@@
 In 2017-2020, 115.3 million US adults had hypertension.
-The age-adjusted prevalence of controlled BP, defined according to the 2017 American College of Cardiology/American Heart Association BP guideline, among US adults with hypertension was 9.6% in 1999-2000, 24.6% in 2013-2014, and 21.7% in 2017-2020.
-After age-adjustment and among US adults who self-reported taking antihypertensive medication, 27.4%, 48.5%, and 42.8% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively.
-The application was validated by replicating statistics from three publications.
+The age-adjusted prevalence of controlled BP, defined by the 2017 American College of Cardiology/American Heart Association BP guideline, among US adults with hypertension was 9.7% in 1999-2000, 25.0% in 2013-2014, and 21.9% in 2017-2020.
+After age-adjustment and among US adults who self-reported taking antihypertensive medication, 27.6%, 48.5%, and 43.0% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively.
 
-**Conclusions**: The application developed and validated in the current study is publicly available and produced valid, transparent, and reproducible results.
+**Conclusions**: The application developed in the current study is publicly available and produced valid, transparent, and reproducible results.
 
@@
 The goal of the application is to increase the use of NHANES data for research and policy decision making with a focus on ensuring results are valid, transparent, and reproducible.
-In this manuscript, we review the design, development and validation of the application, present BP and hypertension statistics for US adults, and describe online tutorials and resources that may help users engage with the application.
+In this manuscript, we review the design, development and validation of the application and present BP and hypertension statistics for US adults that were generated using the application.
 
@@
 
-NHANES data were collected through in-home interviews and study examinations conducted at mobile examination centers.
+NHANES data were collected through an in-home interview and study examination conducted at a mobile examination center.
 The interview included questions about demographics, health behaviors, medical history, and medication use.
 During the interview, the labels of medications that participants reported taking in the preceding 30 days were recorded.
-During the study examination, height, weight and BP were measured and blood and spot urine samples were collected.
+Antihypertensive medication classes were defined using those listed in the 2017 ACC/AHA BP guideline.[@whelton20182017] During the study examination, height, weight and BP were measured and blood and spot urine samples were collected.
 Of relevance to the application, blood samples were used to measure total and high-density lipoprotein cholesterol, glycated hemoglobin and serum creatinine, and the urine sample was used to measure albumin and creatinine and to conduct a pregnancy test.
-The protocol for obtaining BP measurements is available online.[@ostchega_national_2003] In brief, systolic and diastolic BP (SBP and DBP, respectively) were measured three times by trained and certified physicians.
+The protocol for measuring BP is available online.[@ostchega_national_2003] In brief, systolic and diastolic BP (SBP and DBP, respectively) were measured three times by trained and certified physicians.
 BP was measured using a mercury sphygmomanometer from 1999-2000 through 2015-2016 and using an oscillometric device in 2017-2020.
 The mean SBP and DBP levels were computed over all available measurements for each participant.
-The oscillometric SBP and DBP values were calibrated to the mercury device.[@muntner2022blood; @ostchega2021comparing] Antihypertensive medication classes were defined using those listed in the 2017 ACC/AHA BP guideline.[@whelton20182017] For the current analysis, we defined hypertension, BP control, and resistant hypertension according to the 2017 ACC/AHA BP guideline.[@whelton20182017] The application also has these variables defined according to the JNC7 definitions.[@chobanian2003seventh] Chronic kidney disease (CKD) was defined by estimated glomerular filtration rate < 60 ml/min/1.73 m^2^ calculated using the 2021 serum creatinine-based equation or albumin-to-creatinine ratio ≥ 30 mg/g.[@inker_new_2021] A list of BP, hypertension, and antihypertensive medication variables is provided in **Table 1** with full definitions for all variables provided in **Table S1**.
+The oscillometric SBP and DBP values were calibrated to the mercury device.[@muntner2022blood; @ostchega2021comparing] For the current analysis, we defined hypertension, BP control, and resistant hypertension according to the 2017 ACC/AHA BP guideline.[@whelton20182017] The application also has these variables defined according to the JNC7 definitions.[@chobanian2003seventh] Chronic kidney disease (CKD) was defined by estimated glomerular filtration rate < 60 ml/min/1.73 m^2^ calculated using the 2021 serum creatinine-based equation or albumin-to-creatinine ratio > 30 mg/g.[@inker_new_2021] A list of BP, hypertension, and antihypertensive medication variables is provided in **Table 1** with full definitions for all variables provided in **Table S1**.
 
 <br>
+
+```{r, echo=FALSE}
+n_included <- 
+ table.glue::table_value(sum(nhanes_data$svy_subpop_htn))
+```
 
@@
 This exclusion was applied because statistics for BP levels and the prevalence of hypertension in children and adolescents are markedly different than for adults.[@muntner2004trends] We further restricted the population to participants who completed the in-home interview and study examination, with one or more SBP and DBP measurement, and who had data on self-reported antihypertensive medication use.
-After these exclusions were applied, the population in this application included 56,035 participants (**Figure S1**).
+After these exclusions were applied, the population in this application included `r n_included` participants (**Figure S1**).
 
@@
 When population count estimates are requested, survey weights are calibrated to account for missing information on SBP, DBP or self-reported antihypertensive medication use.
-Users may tabulate or visualize summaries, and can present results for the overall population or in subgroups.
-All figures and datasets created with the web application can be downloaded and saved.
-Following CDC recommendations, unreliable statistical estimates are automatically suppressed.
-To increase precision and reliability of estimates, contiguous NHANES cycles can be combined.[@nhanes_tutorial_weights]
+The results can be presented in tables or figures and for the overall population or in subgroups.
+All tables and figures created with the web application can be downloaded and saved.
+Following CDC recommendations, unreliable statistical estimates are automatically suppressed.[@parker2017national] To increase precision and reliability of estimates, contiguous NHANES cycles can be combined.[@nhanes_tutorial_weights]
 
@@
 
-The web application was created using Shiny,[@rpack_shiny; @rpack_shinyWidgets; @rpack_shinyalert] an open-source software package that translates code from the R programming language into HTML, CSS, or JavaScript commands that create a website interface.[@r_language] We created the “cardioStatsUSA” R package to provide additional details on the web application’s design and comprehensive documentation of its components.[@rpack_cardioStatsUSA] We validated the web application by using it to reproduce statistics reported in two prior studies and one CDC report.[@muntner2020htn; @muntner2022blood; @fryar_hypertension_2017]
+The web application was created using Shiny,[@rpack_shiny; @rpack_shinyWidgets; @rpack_shinyalert] an open-source software package that translates code from the R programming language into HTML, CSS, or JavaScript commands and creates a website interface.[@r_language] We created the “cardioStatsUSA” R package to provide additional details on the web application’s design and comprehensive documentation of its components.[@rpack_cardioStatsUSA] We validated the web application by using it to reproduce statistics reported in two prior studies and one CDC report.[@muntner2020htn; @muntner2022blood; @fryar_hypertension_2017]
 
@@
 We performed statistical analyses to demonstrate core features of the application.
-We estimated the mean SBP for US adults by NHANES cycle, 1999-2000 through 2017-2020, with points showing the estimated means and error bars showing the 95% confidence interval (CI) for the estimate.
+We estimated the mean SBP for US adults by NHANES cycle, 1999-2000 through 2017-2020, with points and error bars representing the estimated means and 95% confidence intervals (CI), respectively.
 We made bar charts presenting the age-adjusted prevalence of hypertension and the estimated number of US adults with hypertension.
@@
 
-The age-adjusted prevalence of BP control among non-pregnant US adults with hypertension was lowest in 1999-2000 (9.6%), highest in 2013-2014 (24.5%), and 21.7% in 2017-2020 (**Figure 4; panel A**).
-Among US adults who self-reported taking antihypertensive medication, 27.6%, 48.4%, and 42.8% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively (**Figure 4; panel B**).
-The age-adjusted prevalence of resistant hypertension among non-pregnant US adults taking antihypertensive medication was lowest in 1999-2000 (14.3%), highest in 2005-2006 (21.4), and 16.6% in 2017-2020 (**Figure S4; panel A**).
-Among non-pregnant US adults who were taking three or more classes of antihypertensive medication, the prevalence of resistant hypertension was lowest in 2009-2010 (64.1%), highest in 2003-2004 (78.3%), and 67.9% in 2017-2020 (**Figure S4; panel B**).
+The age-adjusted prevalence of BP control among non-pregnant US adults with hypertension was lowest in 1999-2000 (9.7%), highest in 2013-2014 (25.0%), and 21.9% in 2017-2020 (**Figure 4; panel A**).
+Among US adults with hypertension who self-reported taking antihypertensive medication, 27.6%, 48.5%, and 43.0% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively (**Figure 4; panel B**).
+The age-adjusted prevalence of resistant hypertension among non-pregnant US adults with hypertension, who self-reported taking antihypertensive medication and had ≥1 antihypertensive medication class identified during the pill-bottle review was lowest in 1999-2000 (14.3%), highest in 2005-2006 (21.4%), and 16.6% in 2017-2020 (**Figure S4; panel A**).
+Further restricting this analysis to those who were taking three or more classes of antihypertensive medication, the prevalence of resistant hypertension was lowest in 2009-2010 (64.1%), highest in 2003-2004 (78.3%), and 67.9% in 2017-2020 (**Figure S4; panel B**).
 
@@
 Also, we generated statistics stratified by characteristics of US adults and pooling multiple NHANES cycles to obtain more precise estimates when working with the small sub-group of pregnant women.
-Following its validation, version 0.0.1 of the application was released and deployed in a publicly available server on **DATE TBD**.[@app_cardioStatsUSA] Researchers, clinicians, and the public can use the application to generate customized BP and hypertension statistics for US adults.
+Following its validation, version 0.0.1 of the application was released and deployed on a publicly available server on **DATE TBD (it's almost ready)**.[@app_cardioStatsUSA] Researchers, clinicians, and the public can use the application to generate customized BP and hypertension statistics for US adults.
 
@@
 NHANES was designed to obtain nationally representative estimates of the health and nutrition status of non-institutionalized US adults and is an ideal data source to obtain statistics related to hypertension.
-SBP and DBP were measured following a standardized protocol by trained and certified staff.
+SBP and DBP were measured following a standardized protocol by trained and certified physicians.
 NHANES data are publicly available to download but working with these data requires understanding variable definitions and advanced programming and statistical knowledge.
-Challenges that users encounter when analyzing with NHANES include the need to download and merge multiple data files, even for a single NHANES cycle; combining multiple variables to create outcome definitions, which may require dealing with missing data and questionnaire skip patterns; harmonizing definitions across multiple NHANES cycles; and the analyses of complex survey design data.
-The web application that we present in the current manuscript was designed to reduce barriers to using NHANES so that these data can more fully inform public health decisions and future research related to hypertension.
+Challenges that users encounter when analyzing NHANES data include the need to download and merge multiple data files, even for a single NHANES cycle; combining multiple variables to create outcome definitions, which may require dealing with missing data and questionnaire skip patterns; harmonizing variables across multiple NHANES cycles; and the analysis of complex survey design data.
+The web application that we present in the current manuscript addresses these challenges, allowing NHANES data to more fully inform public health decisions and future research related to hypertension.
 
@@
 Several design decisions have been incorporated into the application.
-We required participants to have a single SBP and DBP to be included while some prior analyses required three SBP and DBP measurements.
-We chose an approach that is consistent with several analyses conducted by CDC investigators.[@yoon_hypertension_2015; @fryar_hypertension_2017] However, mean BP and the prevalence of high BP may be lower if we required multiple BP measurements.[@handler_impact_2012] Any bias resulting from this decision is likely to be small as over 95% of adult NHANES participants with at least one SBP and DBP measurement had three SBP and DBP measurements.
+We required participants to have a single SBP and DBP to be included while some prior analyses required three SBP and DBP measurements.[@muntner_potential_2018]  We chose an approach that is consistent with several analyses conducted by CDC investigators.[@yoon_hypertension_2015; @fryar_hypertension_2017] However, mean BP and the prevalence of high BP may be lower if we required multiple BP measurements.[@handler_impact_2012] Any bias resulting from this decision is likely to be small as over 95% of adult NHANES participants with at least one SBP and DBP measurement had three SBP and DBP measurements.
 The application re-calibrates the NHANES weights for the estimation of population counts.[@coresh_prevalence_2003] This was done because participants missing data on SBP, DBP or antihypertensive medication use cannot have BP or hypertension-related outcomes.
@@
 
-We were able to replicate several prior manuscripts using the application.
-However, we were not able to replicate some results.
-For example, a prior manuscript reported the prevalence of resistant hypertension defined by the 2017 ACC/AHA BP guideline to be 19.7% in 2009-2014.[@carey_prevalence_2019] When estimated by the application, the prevalence of resistant hypertension over this time period was 17.1%.
+We were able to replicate several prior manuscripts using the application.[@muntner2020htn; @muntner2022blood; @fryar_hypertension_2017] However, we were not able to replicate some results.[@carey_prevalence_2019] For example, a prior manuscript reported the prevalence of resistant hypertension defined by the 2017 ACC/AHA BP guideline to be 19.7% in 2009-2014.[@carey_prevalence_2019] When estimated by the application, the prevalence of resistant hypertension over this time period was 17.1%.
 The difference in the prevalence estimates can be attributed to the approach used to categorize medication into classes.
@@
 Variables can be analyzed as outcomes or used to stratify results or restrict analyses to subsets of participants.
-Combining the variables, tens of thousands of unique statistics can be estimated.
-In addition, users can generate statistics for US adults over 22 years, pooling results or stratifying the data.
+Combining the variables, tens of thousands of unique statistics can be estimated using the application.
+In addition, users can generate statistics for US adults from 1999 to March 2020, pooling results or stratifying the data.
 All results from the application can be saved as an image and included in scientific proposals or presentations.
 In addition, the ability to download results as a data set from the application allows users to further customize their results in tabular or graphical formats.
-The web application can also be used for educational purposes, as teachers can use it to analyze health outcomes for US adults interactively with students.
-
-<br>
-
-With an open-source software license and publicly available documentation and data, the web application we developed can be extended.
-The current manuscript described a set of hypertension variables and a sub-population of NHANES participants with data on BP and antihypertensive medication use.
-The application’s interface and its supporting R package allows additional variables to be added.
-In addition, the application can be extended to include data from additional NHANES cycles as they become available.
 
@@
 We used NHANES data, which are publicly available, rigorously collected, and allow for estimation of nationally representative statistics.
-Also, we leveraged open-source software to ensure that our application is transparent, able to be extended, and freely available.
+Also, we leveraged open-source software to ensure that our application is transparent and freely available.
 This study also has several limitations.
@@
 
-# ACKNOWLEDGMENTS
-
-**this section is under construction**
-
-The views expressed in this paper are those of the authors and do not represent the official position of the National Institutes of Health (NIH), the National Heart, Lung, and Blood Institute, or the U.S.
-Government.
-
 # FUNDING/SUPPORT
@@
 
-### Table 1: Blood pressure, hypertension and antihypertensive medication variables that may be analyzed using the web application.
+### Table 1: Blood pressure, hypertension and antihypertensive medication variables that are available in the web application.
 
@@
  .[class %in% classes] %>% 
- .[, class := factor(class, levels = classes)] %>% 
+ .[, class := factor(class, 
+                     levels = classes,
+                     labels = paste(classes, 'domain'))] %>% 
  .[order(class)] %>% 
@@
  which(ft_data$class == "Antihypertensive medication classes")
+
 indent_meds <- 
@@
 
+```{r footer_functions}
+
+footer_hover <- function(parenthetical,
+                         hover_over){
+
+ glue::glue(
+ "The graph is identical to the web application's output. \\
+ All values ({parenthetical}) can be obtained in the app \\
+ by hovering over {hover_over}."
+ )
+
+}
+
+```
+
 ### Figure 1: Mean systolic blood pressure for US adults by calendar year
@@
 
-p1 <- nhanes_visualize(data = nhanes_data,
+nhanes_visualize(data = nhanes_data,
                  key = nhanes_key,
@@
                  size_point = 25,
-                       size_error = 1)
-
-fig_print(p1)
+                 size_error = 1) %>% 
+ fig_print()
 
 ```
-Dots represent mean systolic blood pressure. Vertical lines represent the 95% confidence interval.
+Dots represent mean systolic blood pressure. Vertical lines represent the 95% confidence interval. `r single_par`
+
+`r footer_hover(parenthetical = "estimated mean and the upper and lower limits of the 95% confidence interval", hover_over = "the mean systolic blood pressure")` `r single_par`
 
@@
 ```
-Age adjustment was performed through direct standardization, using the estimated age distribution of US adults from 1999 to 2020 as the standard (49.3%, 33.6%, 10.1% and 7.0% being 18 to 44, 45 to 64, 65 to 74 and ≥ 75 years of age, respectively)
+Age adjustment was performed through direct standardization, using the estimated age distribution of US adults from 1999 to 2020 as the standard (49.3%, 33.6%, 10.1% and 7.0% being 18 to 44, 45 to 64, 65 to 74 and ≥ 75 years of age, respectively).
+`r single_par`
+
+`r footer_hover(parenthetical = "estimated prevalence and the upper and lower limits of the 95% confidence interval", hover_over = "the bars")` `r single_par`
 
@@
 ```
+`r footer_hover(parenthetical = "estimated count and the upper and lower limits of the 95% confidence interval", hover_over = "the bars")` `r single_par`
 
@@
 
-A. Among non-pregnant US adults with hypertension
+A. Among non-pregnant US adults with hypertension `r single_par`
 ```{r, fig.height=3.75}
@@
 ```
-Age adjustment was performed through direct standardization, using the estimated age distribution of US adults with hypertension from 1999 to 2020 as the standard (26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥ 75 years of age, respectively)
+Age adjustment was performed through direct standardization, using the estimated age distribution of US adults with hypertension from 1999 to 2020 as the standard (26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥ 75 years of age, respectively) `r single_par`
+
+`r footer_hover(parenthetical = "estimated prevalence and the upper and lower limits of the 95% confidence interval", hover_over = "the bars")` `r single_par`
+
+Survey participants with missing values for pregnancy status were assumed to be non-pregnant `r single_par`
 
@@
 
-B. Among non-pregnant US adults who self-report taking antihypertensive medication 
+B. Among non-pregnant US adults who self-report taking antihypertensive medication `r single_par`
 ```{r, fig.height=3.75}
@@
 
-### Table S1: All variables included in the web-based application.
+### Table S1: Definitions of variables included in the web-based application.
 
@@
   values = abbrvs_write(
-   abbrvs[c('BP', 'SBP', 'DBP', 'ACC', 'AHA', 'JNC7')]
+   abbrvs[c("BP",
+         "SBP",
+         "DBP",
+         "JNC7",
+         "ACC",
+         "AHA",
+         "ASCVD",
+         "CVD",
+         "HbA1c",
+         "ACE",
+         "NHANES")]
   )
@@
 
-### Figure S1 Application of inclusion and exclusion criteria for the application of NHANES hypertension data.
+### Figure S1: Sample size after inclusion and exclusion criteria were applied for the NHANES hypertension application
 
@@
  ),
- Overall = c(107622, 63041, 59799, 56286, 56035, 56035),
+ Overall = c(107622, 63041, 59799, 56286, 56017, 56017),
  `1999-2000` = c(9965, 5448, 4976, 4755, 4694, 4694),
- `2001-2002` = c(11039, 5993, 5592, 5251, 5184,  5184),
- `2003-2004` = c(10122, 5620, 5303, 4902, 4838,  4838),
- `2005-2006` = c(10348, 5563, 5334, 5028, 5015,  5015),
- `2007-2008` = c(10149, 6228, 5995, 5670, 5665,  5665),
+ `2001-2002` = c(11039, 5993, 5592, 5251, 5181, 5181),
+ `2003-2004` = c(10122, 5620, 5303, 4902, 4836, 4836),
+ `2005-2006` = c(10348, 5563, 5334, 5028, 5012, 5012),
+ `2007-2008` = c(10149, 6228, 5995, 5670, 5664, 5664),
  `2009-2010` = c(10537, 6527, 6360, 6053, 6043, 6043),
- `2011-2012` = c(9756, 5864, 5615, 5436, 5337,   5337),
- `2013-2014` = c(10175, 6113, 5924, 5700, 5694,  5694),
- `2015-2016` = c(9971, 5992, 5735, 5557, 5552,   5552),
- `2017-2020` = c(15560, 9693, 8965, 8024, 8013,  8013)
+ `2011-2012` = c(9756, 5864, 5615, 5436, 5334, 5334),
+ `2013-2014` = c(10175, 6113, 5924, 5700, 5692, 5692),
+ `2015-2016` = c(9971, 5992, 5735, 5557, 5551, 5551),
+ `2017-2020` = c(15560, 9693, 8965, 8024, 8010, 8010)
 ) %>% 
@@
  align(align = 'center', part = 'all') %>% 
- align(j = 1, part = 'all')
+ align(j = 1, part = 'all') %>% 
+ add_footer_lines(
+  values = abbrvs_write(
+   abbrvs[c("SBP",
+            "DBP",
+            "NHANES")]
+  )
+ )
 
@@
 
-### Figure S2: Prevalence of hypertension among US adults with and without chronic kidney disease by calendar year
+### Figure S2: Prevalence of hypertension among US adults with and without chronic kidney disease by calendar year.
 
@@
 ```
+`r footer_hover(parenthetical = "estimated prevalence and the upper and lower limits of the 95% confidence interval", hover_over = "the bars")` `r single_par`
 
@@
 
-A. In 2017-2020
+A. In 2017-2020 `r single_par`
 ```{r}
@@
 ```
+The `r paste("--")` symbol indicates that an estimate is not reliable. `r single_par`
+
+2.26 represents the estimated proportion of pregnant women with systolic blood pressure of 130 to < 140 mm Hg or diastolic blood pressure of 80 to < 90 mm Hg. `r single_par`
+
+`r footer_hover(parenthetical = "estimated prevalence, upper and lower limits of the 95% confidence interval, or reasons for the suppression of data if applicable", hover_over = "the bars")` `r single_par`
+
+There were no pregnant women in the two highest blood pressure categories, so the figure does not include segments for these categories. `r single_par`
 
@@
 
-B. In 2009-2010 through 2017-2020
+B. In 2009-2010 through 2017-2020 `r single_par`
 ```{r}
@@
 ```
-The estimate for SBP of 140 to < 160 or 90 to < 100 mm Hg is 1.1%. This can be seen on the application by hovering over the red bar with your mouse.
+`r footer_hover(parenthetical = "estimated prevalence and the upper and lower limits of the 95% confidence interval", hover_over = "the bars")` `r single_par`
+
+The estimate for systolic blood pressure of 140 to < 160 or diastolic blood pressure of 90 to < 100 mm Hg is 1.1%.
+This can be seen on the application by hovering over the red bar with your mouse.
+`r single_par`
 
@@
 
-A. Among non-pregnant US adults who self report taking antihypertensive medication and are taking 1 or more antihypertensive medication classes 
+A. Among non-pregnant US adults who self report taking antihypertensive medication and are taking 1 or more antihypertensive medication classes `r single_par`
 ```{r, fig.height=3.75}
@@
 
-B. Among non-pregnant US adults who self report taking antihypertensive medication and are taking 3 or more antihypertensive medication classes
+B. Among non-pregnant US adults who self report taking antihypertensive medication and are taking 3 or more antihypertensive medication classes `r single_par`
 ```{r, fig.height=3.75}
@@
 ```
-Age adjustment was performed through direct standardization, using the estimated age distribution of US adults with hypertension from 1999 to 2020 as the standard (26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥ 75 years of age, respectively)
+Age adjustment was performed through direct standardization, using the estimated age distribution of US adults with hypertension from 1999 to 2020 as the standard (26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥ 75 years of age, respectively) `r single_par`
+
+`r footer_hover(parenthetical = "estimated prevalence and the upper and lower limits of the 95% confidence interval", hover_over = "the bars")` `r single_par`
+
+Survey participants with missing values for pregnancy status were assumed to be non-pregnant `r single_par`
 
 \newpage
-
-<!-- **Figure S5** Prevalence of hypertension by age categories in 2017-2020 -->
-
-<!-- ```{r} -->
-<!-- nhanes_visualize(data = nhanes_data, -->
-<!--                  key = nhanes_key, -->
-<!--                  outcome_variable = 'htn_accaha', -->
-<!--                  group_variable =  'demo_age_cat', -->
-<!--                  time_values = 'most_recent', -->
-<!--                  geom = 'bar') %>%  -->
-<!--  fig_print() -->
-<!-- ``` -->
-<!-- \newpage -->
-
-<!-- **Figure S6** Prevalence of hypertension by sex in 2017-2020 -->
-
-<!-- ```{r} -->
-<!-- nhanes_visualize(data = nhanes_data, -->
-<!--                  key = nhanes_key, -->
-<!--                  outcome_variable = 'htn_accaha', -->
-<!--                  group_variable = 'demo_gender', -->
-<!--                  time_values = 'most_recent', -->
-<!--                  geom = 'bar') %>%  -->
-<!--  fig_print() -->
-<!-- ``` -->
-
-<!-- \newpage -->
-
-<!-- **Figure S7** Prevalence of hypertension by race/ethnicity in 2017-2020 -->
-
-<!-- ```{r} -->
-<!-- nhanes_visualize(data = nhanes_data, -->
-<!--                  key = nhanes_key, -->
-<!--                  outcome_variable = 'htn_accaha', -->
-<!--                  group_variable = 'demo_race', -->
-<!--                  time_values = 'most_recent', -->
-<!--                  geom = 'bar') %>%  -->
-<!--  fig_print() -->
-<!-- ``` -->
-
-<!-- \newpage -->
-
-<!-- **Figure S8** Prevalence of uncontrolled blood pressure among pregnant women from 2011-2012 through 2017-2020. -->
-
-<!-- ```{r} -->
-
-<!-- nhanes_visualize(data = nhanes_data, -->
-<!--                  key = nhanes_key, -->
-<!--                  outcome_variable = 'bp_cat_meds_excluded', -->
-<!--                  time_values = c("last_5"), -->
-<!--                  pool = TRUE, -->
-<!--                  subset_calls = list("demo_pregnant" = "Yes"), -->
-<!--                  geom = 'bar') %>%  -->
-<!--  fig_print() -->
-
-<!-- ``` -->
 

### 2022-11-05 R/app_run.R

@@ -25,7 +25,7 @@ app_run <- function(nhanes_data = cardioStatsUSA::nhanes_data,
 
  # coerce inputs to data.tables but don't copy them.
  if(!is.data.table(nhanes_data)) setDT(nhanes_data)
- if(!is.data.table(nhanes_key))  setDT(nhanes_data)
+ if(!is.data.table(nhanes_key))  setDT(nhanes_key)
 
  time_variable <- nhanes_key$variable[nhanes_key$type == 'time']
  time_values <- levels(nhanes_data[[time_variable]])

### 2022-11-05 R/replication_tables.R

@@ -1,0 +1,36 @@
+#' Results from Prior Studies of NHANES Data
+#'
+#' These datasets contain values from various publications that have been used
+#'   to validate the `cardioStatsUSA` shiny application (see [app_run]).
+#'
+#' @format Data frames with varying numbers of rows and variables.
+#' \describe{
+#' \item{svy_year}{NHANES cycle.}
+#' \item{variable}{variable name.}
+#' \item{group}{sub-group defined by variable.}
+#' \item{estimate}{point estimate.}
+#' \item{std_error}{standard error of the point estimate.}
+#' \item{ci_lower}{lower bound of 95% confidence interval for the point estimate.}
+#' \item{ci_upper}{upper bound of 95% confidence interval for the point estimate.}
+#' }
+#'
+#' @aliases muntner_jama_2020_etable_1, muntner_hypertension_2022_table_s2,
+#'   cdc_db289_figure_2, cdc_db289_figure_2
+#' @name replication
+"muntner_jama_2020_table_2"
+
+#' @rdname replication
+#' @format NULL
+"muntner_jama_2020_etable_1"
+
+#' @rdname replication
+#' @format NULL
+"muntner_hypertension_2022_table_s2"
+
+#' @rdname replication
+#' @format NULL
+"cdc_db289_figure_2"
+
+#' @rdname replication
+#' @format NULL
+"cdc_db289_figure_2"

### 2022-11-05 data-raw/cdc_db289_figure_1.R

@@ -13,4 +13,6 @@ cdc_db289_figure_1 <- tibble::tribble(
  "Women",         "60+",         66.8,      2.6
 )
 
-usethis::use_data(cdc_db289_figure_1)
+setDT(cdc_db289_figure_1)
+usethis::use_data(cdc_db289_figure_1, overwrite = TRUE)
+

### 2022-11-05 data-raw/muntner_hypertension_2022_table_s2.R

@@ -39,4 +39,5 @@ muntner_hypertension_2022_table_s2 <- tibble::tribble(
  "2017-2020", "Sex", "Men",	49.0, 44.8, 53.3
 )
 
-usethis::use_data(muntner_hypertension_2022_table_s2)
+setDT(muntner_hypertension_2022_table_s2)
+usethis::use_data(muntner_hypertension_2022_table_s2, overwrite = TRUE)

### 2022-11-05 data-raw/muntner_jama_2020_table_2.R

@@ -53,4 +53,6 @@ muntner_jama_2020_table_2 <- tibble::tribble(
  "2017-2018", "gt_160_100",   14.6     ,  11.7    ,  17.5
 )
 
-usethis::use_data(muntner_jama_2020_table_2)
+setDT(muntner_jama_2020_table_2)
+
+usethis::use_data(muntner_jama_2020_table_2, overwrite = TRUE)

### 2022-11-05 data-raw/nhanes_data.R

@@ -20,15 +20,18 @@ nhanes_data <-
  nhanes_recode() %>%
  nhanes_rename()
 
-
 setDT(nhanes_data)
 
 nhanes_key <- cardioStatsUSA::nhanes_key
 
-for(i in names(nhanes_data)){
+for(i in intersect(names(nhanes_data), nhanes_key$variable)){
+
  if(is.null(attr(nhanes_data[[i]], 'label'))){
-  attr(nhanes_data[[i]], 'label') <- nhanes_key$variables[[i]]$label
+  attr(nhanes_data[[i]], 'label') <- nhanes_key %>%
+   .[variable == i] %>%
+   .[['label']]
  }
+
 }
 
 usethis::use_data(nhanes_data, overwrite = TRUE)

### 2022-11-05 data-raw/nhanes_load.R

@@ -12,23 +12,11 @@
 nhanes_load <- function(fpath = file.path(here::here(), 'data-raw'),
                         as = 'tibble'){
 
- data_bp <- haven::read_sas(
-  file.path(fpath, 'htn-raw.sas7bdat')
- )
-
- data_lipids <- haven::read_sas(
-  file.path(fpath, 'lipids-raw.sas7bdat')
- )
+ data_bp <- haven::read_sas(file.path(fpath, 'htn-raw.sas7bdat'))
+ data_lipids <- haven::read_sas(file.path(fpath, 'lipids-raw.sas7bdat'))
 
  out <- left_join(data_bp, data_lipids, by = c("SEQN", "WTSAF2YR")) %>%
   mutate(subpopldl = replace(subpopldl, is.na(subpopldl), 0))
-
- # data_lipids %>% filter(SEQN == 31720)
- # temporary fix
- # out <- out %>%
- #  dplyr::group_by(SEQN) %>%
- #  dplyr::slice(1)
-
 
  switch(
   as,

### 2022-11-05 tests/testthat/test-reliability_mean.R

@@ -1,4 +1,5 @@
 
+# A case where the number of observations is low
 
 smry <- nhanes_summarize(
  data = nhanes_data,
@@ -10,4 +11,8 @@ smry <- nhanes_summarize(
  stratify_variable = "demo_race"
 )
 
-smry[unreliable_status==TRUE]
+reasons <- smry[unreliable_status==TRUE, unreliable_reason]
+
+expect_true(
+ "Sample size < 30" %in% reasons
+)

### 2022-11-05 rmd/nhanes_data.Rmd
stats shown; stringr dropped

@@
  .[order(class)] %>% 
- droplevels()
+ droplevels() %>% 
+ .[, description := gsub('\u2265', '>=', x = description, fixed = T)] %>% 
+ .[, label := gsub('\u2265', '>=', x = label, fixed = T)]
 
-data_to_doc$description %<>% stringr::str_replace_all('\u2265', '>=')
-data_to_doc$label %<>% stringr::str_replace_all('\u2265', '>=')
-
-key_classes <- split(data_to_doc, f = data_to_doc$class)
+key_classes <- split(data_to_doc, by = 'class')
 
@@
   cat("##", key_classes[[i]]$variable[j], "\n")
-  cat("- Label: ", key_classes[[i]]$label[j], "\n")
-  cat("- Description: ", key_classes[[i]]$description[j], "\n\n")
+  cat("- **Label**: ", key_classes[[i]]$label[j], "\n")
+  cat("- **Description**: ", key_classes[[i]]$description[j], "\n")
+
+  if(names(key_classes)[i] != 'Survey'){
+   cat(get_variable_summary(key_classes[[i]]$variable[j],
+                            data = nhanes_data[svy_subpop_htn == 1]),
+       "\n\n")
+  } else {
+   cat("\n")
+  }
+
  }

### 2022-11-05 R/plotly_viz_make_hover.R

@@ -32,43 +32,9 @@ plotly_viz_make_hover <- function(data, stat_all, group, group_label) {
         "review_reason",
         "review_needed",
         "stat_show") := NULL] %>%
-  dcast(data = ., formula = ... ~ statistic, value.var = "stat_label") %>%
+  dcast(formula = ... ~ statistic, value.var = "stat_label") %>%
   .[, hover := do.call(paste, c(.SD, sep="<br>")), .SDcols = stat_cols] %>%
   .[, hover := paste0(group_label, label_sep, hover)]
-
- # data %>%
- #  as_tibble() %>%
- #  mutate(across(all_of(numeric_cols), table_value)) %>%
- #  mutate(
- #   stat_show = if_else(
- #    review_needed,
- #    paste0(" -Review needed-<br>  Reasons: ", review_reason),
- #    ""
- #   ),
- #   stat_show = if_else(
- #    unreliable_status,
- #    paste0(" -Unreliable-<br>  Reasons: ", unreliable_reason),
- #    stat_show
- #   ),
- #   stat_show = if_else(
- #    stat_show == "",
- #    as.character(glue("{estimate} (95% CI {ci_lower}, {ci_upper})")),
- #    stat_show
- #   ),
- #   stat_label = glue(
- #    "{stat_recode(statistic)}: {stat_show}"
- #   )
- #  ) %>%
- #  select(-all_of(numeric_cols),
- #         -c(unreliable_reason,
- #            unreliable_status,
- #            review_reason,
- #            review_needed,
- #            stat_show)) %>%
- #  pivot_wider(names_from = statistic, values_from = stat_label) %>%
- #  unite(col = 'hover', !!!stat_cols, sep = '<br>') %>%
- #  mutate(hover = glue("{group_label}{label_sep}{hover}")) %>%
- #  as.data.table()
 
 }
 
@@ -76,8 +42,13 @@ plotly_viz_make_hover <- function(data, stat_all, group, group_label) {
 stat_recode <- function(stat){
 
  fcase(
-  stat == 'percentage_kg', 'Percentage (Korn & Graubard CI)',
-  rep(TRUE, length(stat)), stringr::str_to_title(stat)
+  stat == "percentage_kg", "Percentage (Korn & Graubard CI)",
+  stat == "mean", "Mean",
+  stat == "percentage", "Percentage",
+  stat == "count", "Count",
+  stat == "q25", "25th %",
+  stat == "q50", "50th %",
+  stat == "q75", "75th %"
  )
 
 }

### 2022-11-05 R/replication_tables.R

@@ -29,8 +29,9 @@
 
 #' @rdname replication
 #' @format NULL
-"cdc_db289_figure_2"
+"cdc_db289_figure_1"
 
 #' @rdname replication
 #' @format NULL
 "cdc_db289_figure_2"
+

### 2022-11-05 R/svy_statby_tidy_quantile.R

@@ -2,13 +2,21 @@
 svy_statby_tidy_quantile <- function(x, outcome, by_vars, ...) {
 
  out_names <- names(x) %>%
-  str_replace(pattern = paste0("^", outcome),
-              replacement = paste0('estimate.', outcome)) %>%
-  str_replace(pattern = outcome,
-              replacement = paste0('_z.z.z_.', outcome, '._z.z.z_')) %>%
-  str_replace(pattern = '^se\\.', replacement = 'std_error.') %>%
-  str_replace(pattern = '^ci_l\\.', replacement = 'ci_lower.') %>%
-  str_replace(pattern = '^ci_u\\.', replacement = 'ci_upper.')
+  gsub(pattern = paste0("^", outcome),
+       replacement = paste0('estimate.', outcome),
+       x = .) %>%
+  gsub(pattern = outcome,
+       replacement = paste0('_z.z.z_.', outcome, '._z.z.z_'),
+       x = .) %>%
+  gsub(pattern = '^se\\.',
+       replacement = 'std_error.',
+       x = .) %>%
+  gsub(pattern = '^ci_l\\.',
+       replacement = 'ci_lower.',
+       x = .) %>%
+  gsub(pattern = '^ci_u\\.',
+       replacement = 'ci_upper.',
+       x = .)
 
  out <- as.data.table(x) %>%
   setnames(new = out_names) %>%

### 2022-11-06 rmd/nhanes_data.Rmd
explain data size

@@
 ```{r child = 'aaa.Rmd'}
+
+```
+
+```{r, echo = FALSE}
+
+n_htn <- table.glue::table_value(sum(nhanes_data$svy_subpop_htn))
+n_chol <- table.glue::table_value(sum(nhanes_data$svy_subpop_chol))
+
+```
+
+Currently, two sub-populations are included in `nhanes_data`:
+
+- blood pressure and hypertension (n = `r n_htn`)
+
+- lipids and cholesterol (n = `r n_chol`)
+
+To access a sub-population, simply filter `nhanes_data` to contain rows where the corresponding sub-population indicator is equal to 1, e.g.,
+
+```{r, results='hide'}
+
+# for blood pressure and hypertension sub-population
+nhanes_data[svy_subpop_htn == 1]
+
+# for lipids and cholesterol sub-population
+nhanes_data[svy_subpop_chol == 1]
 

### 2022-11-11 paper/manuscript.Rmd
LC comment; source code

@@
 Ashley Christenson, BS,^2^ 
-Lisandro D. Colantonio, PhD,^2^
+Lisandro D. Colantonio, MD, PhD,^2^
 Paul Muntner, PhD,^2^
@@
 ^1^Department of Biostatistics and Data Science, Wake Forest School of Medicine, Winston-Salem, NC. `r single_par`
-^2^Department of Epidemiology, University of Alabama at Birmingham. `r single_par`
+^2^Department of Epidemiology, University of Alabama at Birmingham, Birmingham, AL. `r single_par`
 ^3^Informatics, Decision-Enhancement, and Analytic Sciences (IDEAS) Center, Veterans Affairs, Salt Lake City Health Care System, Salt Lake City, UT. `r single_par`
@@
 **Methods**: We developed and validated a web-based application that provides hypertension statistics for US adults using 10 cycles of NHANES data, 1999-2000 through 2017-2020, without requiring users to conduct statistical programming.
-The application’s interface allows users to estimate crude and age-adjusted means, quantiles, and proportions.
+The application's interface allows users to estimate crude and age-adjusted means, quantiles, and proportions.
 Population counts can also be estimated.
@@
 The age-adjusted prevalence of controlled BP, defined by the 2017 American College of Cardiology/American Heart Association BP guideline, among US adults with hypertension was 9.7% in 1999-2000, 25.0% in 2013-2014, and 21.9% in 2017-2020.
-After age-adjustment and among US adults who self-reported taking antihypertensive medication, 27.6%, 48.5%, and 43.0% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively.
+After age-adjustment and among non-pregnant US adults who self-reported taking antihypertensive medication, 27.6%, 48.5%, and 43.0% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively.
 
@@
 We developed an open-source web-based application that provides nationally representative BP and hypertension statistics for non-institutionalized US adults using NHANES data without requiring users to conduct statistical programming.
-The goal of the application is to increase the use of NHANES data for research and policy decision making with a focus on ensuring results are valid, transparent, and reproducible.
+The goal of the application is to increase the use of NHANES data for research and policy decision making with a focus on ensuring results are valid and reproducible.
+The application is also "open-source" as it's source code is publicly available.
 In this manuscript, we review the design, development and validation of the application and present BP and hypertension statistics for US adults that were generated using the application.

### 2022-11-11 paper/manuscript.Rmd
LC comment;
Note that the application can give various formats. We should be
specific about the formats available but there arent enough words in the
Methods section to list them. I will open an issue to specify file types
in the readme.
LC comment; calibration
noting that the calibration occurs in specific groups, consistent
with prior publications.

@@
 Users can restrict analyses to subsets of participants (e.g., participants who self-reported taking antihypertensive medication).
-When population count estimates are requested, survey weights are calibrated to account for missing information on SBP, DBP or self-reported antihypertensive medication use.
+When population count estimates are requested, survey weights are calibrated within race, sex, and age groups to account for missing information on SBP, DBP or self-reported antihypertensive medication use.
 The results can be presented in tables or figures and for the overall population or in subgroups.
-All tables and figures created with the web application can be downloaded and saved.
+All tables and figures created with the web application can be downloaded and saved in standard formats.
 Following CDC recommendations, unreliable statistical estimates are automatically suppressed.[@parker2017national] To increase precision and reliability of estimates, contiguous NHANES cycles can be combined.[@nhanes_tutorial_weights]

### 2022-11-11 paper/manuscript.Rmd
LC comment; drop March
LC text edits

@@
 Antihypertensive medication classes were defined using those listed in the 2017 ACC/AHA BP guideline.[@whelton20182017] During the study examination, height, weight and BP were measured and blood and spot urine samples were collected.
-Of relevance to the application, blood samples were used to measure total and high-density lipoprotein cholesterol, glycated hemoglobin and serum creatinine, and the urine sample was used to measure albumin and creatinine and to conduct a pregnancy test.
+Of relevance to the current application, blood samples were used to measure total and high-density lipoprotein cholesterol, glycated hemoglobin and serum creatinine, and the urine sample was used to measure albumin and creatinine and to conduct a pregnancy test.
 The protocol for measuring BP is available online.[@ostchega_national_2003] In brief, systolic and diastolic BP (SBP and DBP, respectively) were measured three times by trained and certified physicians.
@@
 This exclusion was applied because statistics for BP levels and the prevalence of hypertension in children and adolescents are markedly different than for adults.[@muntner2004trends] We further restricted the population to participants who completed the in-home interview and study examination, with one or more SBP and DBP measurement, and who had data on self-reported antihypertensive medication use.
-After these exclusions were applied, the population in this application included `r n_included` participants (**Figure S1**).
+After these exclusions were applied, the population in the current application included `r n_included` participants (**Figure S1**).
 
@@
 Among US adults with hypertension who self-reported taking antihypertensive medication, 27.6%, 48.5%, and 43.0% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively (**Figure 4; panel B**).
-The age-adjusted prevalence of resistant hypertension among non-pregnant US adults with hypertension, who self-reported taking antihypertensive medication and had ≥1 antihypertensive medication class identified during the pill-bottle review was lowest in 1999-2000 (14.3%), highest in 2005-2006 (21.4%), and 16.6% in 2017-2020 (**Figure S4; panel A**).
+The age-adjusted prevalence of resistant hypertension among non-pregnant US adults with hypertension who self-reported taking antihypertensive medication and had ≥1 antihypertensive medication class identified during the medication inventory was lowest in 1999-2000 (14.3%), highest in 2005-2006 (21.4%), and 16.6% in 2017-2020 (**Figure S4; panel A**).
 Further restricting this analysis to those who were taking three or more classes of antihypertensive medication, the prevalence of resistant hypertension was lowest in 2009-2010 (64.1%), highest in 2003-2004 (78.3%), and 67.9% in 2017-2020 (**Figure S4; panel B**).
@@
 SBP and DBP were measured following a standardized protocol by trained and certified physicians.
-NHANES data are publicly available to download but working with these data requires understanding variable definitions and advanced programming and statistical knowledge.
+NHANES data are publicly available to download, but working with these data requires understanding variable definitions and advanced programming and statistical knowledge.
 Challenges that users encounter when analyzing NHANES data include the need to download and merge multiple data files, even for a single NHANES cycle; combining multiple variables to create outcome definitions, which may require dealing with missing data and questionnaire skip patterns; harmonizing variables across multiple NHANES cycles; and the analysis of complex survey design data.
@@
 Combining the variables, tens of thousands of unique statistics can be estimated using the application.
-In addition, users can generate statistics for US adults from 1999 to March 2020, pooling results or stratifying the data.
+In addition, users can generate statistics for US adults from 1999-2000 to 2017-2020, pooling results or stratifying the data.
 All results from the application can be saved as an image and included in scientific proposals or presentations.
@@
 
-We developed a web-based application for analysis of hypertension outcomes among non-institutionalized adults living in the US from 1999-2000 through 2017-2020.
+We developed a web-based application for analysis of hypertension statistics among non-institutionalized adults living in the US from 1999-2000 through 2017-2020.
 The application is publicly available and produces valid, transparent, and reproducible results.

### 2022-11-11 paper/manuscript.Rmd
AB text edits

@@
 
-In the current manuscript, we present a web application that allows the calculation of nationally representative estimates for BP, hypertension and antihypertensive medication-related outcomes using publicly available NHANES data.
+In the current manuscript, we present a web application with a user-friendly point and click interface that allows the calculation of nationally representative estimates for BP, hypertension and antihypertensive medication-related statistics using publicly available NHANES data.
 Using this application, we generated crude and age-adjusted BP and hypertension statistics.
 Also, we generated statistics stratified by characteristics of US adults and pooling multiple NHANES cycles to obtain more precise estimates when working with the small sub-group of pregnant women.
-Following its validation, version 0.0.1 of the application was released and deployed on a publicly available server on **DATE TBD (it's almost ready)**.[@app_cardioStatsUSA] Researchers, clinicians, and the public can use the application to generate customized BP and hypertension statistics for US adults.
+Following its validation, version 0.0.1 of the application was released and deployed on a publicly available server on **DATE TBD (it's almost ready)**.[@app_cardioStatsUSA] Researchers, clinicians, and the public can use the application to generate customized BP and hypertension statistics for US adults to inform their research and policy decisions.
 
@@
 NHANES data are publicly available to download, but working with these data requires understanding variable definitions and advanced programming and statistical knowledge.
-Challenges that users encounter when analyzing NHANES data include the need to download and merge multiple data files, even for a single NHANES cycle; combining multiple variables to create outcome definitions, which may require dealing with missing data and questionnaire skip patterns; harmonizing variables across multiple NHANES cycles; and the analysis of complex survey design data.
+Challenges of analyzing NHANES data may also include downloading and merging multiple data files, even for a single NHANES cycle; combining multiple variables to create outcome definitions, which may require dealing with missing data and questionnaire skip patterns; harmonizing variables across multiple NHANES cycles; and the analysis of complex survey design data.
 The web application that we present in the current manuscript addresses these challenges, allowing NHANES data to more fully inform public health decisions and future research related to hypertension.
@@
 
-We were able to replicate several prior manuscripts using the application.[@muntner2020htn; @muntner2022blood; @fryar_hypertension_2017] However, we were not able to replicate some results.[@carey_prevalence_2019] For example, a prior manuscript reported the prevalence of resistant hypertension defined by the 2017 ACC/AHA BP guideline to be 19.7% in 2009-2014.[@carey_prevalence_2019] When estimated by the application, the prevalence of resistant hypertension over this time period was 17.1%.
-The difference in the prevalence estimates can be attributed to the approach used to categorize medication into classes.
+Using the application, we replicated results from several prior manuscripts.[@muntner2020htn; @muntner2022blood; @fryar_hypertension_2017] However, we were not able to replicate some results.[@carey_prevalence_2019] For example, a prior manuscript reported the prevalence of resistant hypertension defined by the 2017 ACC/AHA BP guideline to be 19.7% in 2009-2014.[@carey_prevalence_2019] When estimated by the application, the prevalence of resistant hypertension over this time period was 17.1%.
+The difference in the prevalence estimates can be attributed to the approach used to categorize antiyhpertensive medications into their medication classes.
 The prior publication used Lexicon Plus®, which counted spironolactone as two drug classes, a potassium-sparing diuretic and an aldosterone antagonist.
 We included spironolactone as a single drug class, an aldosterone antagonist.
-Additionally, Lexicon Plus® includes Sotalol as a beta blocker and nitroglycerine as a direct vasodilator and we did not include these drugs as antihypertensive medications as they are not listed in the 2017 ACC/AHA BP guideline.[@whelton20182017] The differences in results between the application and this previously published manuscript emphasize that it may not be possible to always generate results that are identical to prior publications.
+Additionally, Lexicon Plus® includes sotalol as a beta blocker and nitroglycerine as a direct vasodilator and we did not include these drugs as antihypertensive medications as they are not listed in the 2017 ACC/AHA BP guideline.[@whelton20182017] The differences in results between the application and this previously published manuscript emphasize that it may not be possible to always generate results that are identical to prior publications.
 Users should be aware of the choices made in defining variables as this may affect the statistical estimates generated using the application.

### 2022-11-12 paper/manuscript.Rmd
bookmark me!

@@
 Ligong Chen, PhD,^2^
-Kathryn Foti,^2^
+Kathryn Foti, PhD, MPH^2^
 Shakia T. Hardy,^2^
@@
 In 2017-2020, 115.3 million US adults had hypertension.
-The age-adjusted prevalence of controlled BP, defined by the 2017 American College of Cardiology/American Heart Association BP guideline, among US adults with hypertension was 9.7% in 1999-2000, 25.0% in 2013-2014, and 21.9% in 2017-2020.
+The age-adjusted prevalence of controlled BP, defined by the 2017 American College of Cardiology/American Heart Association BP guideline, among non-pregnant US adults with hypertension was 9.7% in 1999-2000, 25.0% in 2013-2014, and 21.9% in 2017-2020.
 After age-adjustment and among non-pregnant US adults who self-reported taking antihypertensive medication, 27.6%, 48.5%, and 43.0% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively.

### 2022-11-12 R/app_run.R

@@ -37,10 +37,16 @@ app_run <- function(nhanes_data = cardioStatsUSA::nhanes_data,
   stratify = nhanes_key[stratify == TRUE, .(class, label, variable)]
  ) %>%
   purrr::map(
-  .f = ~ .x %>%
-   split(by = 'class', keep.by = FALSE) %>%
-   purrr::map(deframe)
- )
+   .f = ~ .x %>%
+    split(by = 'class', keep.by = FALSE) %>%
+    purrr::map(
+     .f = function(data_split){
+      data_split %>%
+       getElement('variable') %>%
+       purrr::set_names(data_split$label)
+     }
+    )
+  )
 
  ctns_variables <- nhanes_key[type == 'ctns', variable]
 
@@ -69,444 +75,546 @@ app_run <- function(nhanes_data = cardioStatsUSA::nhanes_data,
    input.statistic.length > 0
   )"
 
+
  # User interface (UI) -----------------------------------------------------
+ ui <- function(request){
 
- ui <- fluidPage(
+  fluidPage(
 
-  introjsUI(),
+   introjsUI(),
 
-  introBox(
-   titlePanel("Cardiometabolic statistics for US adults"),
-   data.step = 1,
-   data.intro = paste(
-    "Welcome! These instructions will teach you",
-    "how to use the \"Cardiometabolic statistics for US adults\" website,",
-    "an open-source application for analysis and visualization of",
-    "National Health and Nutrition Examination Survey (NHANES) data."
-   )
-  ),
-
-  sidebarLayout(
-
-   # Sidebar -----------------------------------------------------------------
-   sidebarPanel(
-
-    actionButton(inputId = "help",
-                 "Press for instructions",
-                 icon = icon("question"),
-                 width = "100%"),
-
-    HTML('<br>'), HTML('<br>'),
-
-    introBox(
-     conditionalPanel(
-      condition = compute_ready,
-  actionButton(
-   inputId =  "run",
-       label = "Compute results",
-       icon = icon("cog"),
-       width = "100%",
-       style = "color: #fff; background-color: #337ab7; border-color: #2e6da4"
-      )
-     ),
-
-     conditionalPanel(
-      condition = paste("!(", compute_ready, ")", sep = ''),
-      actionButton(
-       inputId =  "wont_do_computation",
-       label = "Compute results",
-       icon = icon("cog"),
-       width = "100%",
-       style = "color: #fff; background-color: #808080; border-color: #2e6da4"
-      )
-     ),
[…]

### 2022-11-12 paper/manuscript.Rmd
KF comments

@@
 ^7^ University of Colorado Skaggs School of Pharmacy and Pharmaceutical Sciences, Aurora, CO, USA
+
 \newpage
@@
 NHANES data are currently collected in two-year periods, referred to as cycles, and each cycle includes over 50 data files and accompanying informational material.
-NHANES data collection protocols have changed over time for some variables, and these differences should be taken into account when comparing or aggregating data from multiple cycles.
-Also, tests need to be performed to ensure the statistical estimates are reliable, and unstable estimates should be suppressed.[@parker2017national]
+NHANES data collection protocols have changed over time for some variables, and these differences should be considered when comparing or aggregating data from multiple cycles.
+For example, while BP was measured using a mercury sphygmomanometer until 2016,[@centers2015national] a validated oscillometric device was used in 2017 through 2020.[@ostchega2010assessing] Also, tests need to be performed to ensure the statistical estimates are reliable, and unstable estimates should be suppressed.[@parker2017national]
 
@@
 Challenges of analyzing NHANES data may also include downloading and merging multiple data files, even for a single NHANES cycle; combining multiple variables to create outcome definitions, which may require dealing with missing data and questionnaire skip patterns; harmonizing variables across multiple NHANES cycles; and the analysis of complex survey design data.
+Although multiple reports based on the data are publicly available, it may be difficult to customize reports based on user requests.
 The web application that we present in the current manuscript addresses these challenges, allowing NHANES data to more fully inform public health decisions and future research related to hypertension.
@@
 
-Drs Muntner reported receiving grant funding and consulting fees from Amgen Inc. Dr Colantonio reported receiving grant funding from Amgen Inc. No other disclosures were reported.
+Dr Muntner reported receiving grant funding and consulting fees from Amgen Inc. Dr Colantonio reported receiving grant funding from Amgen Inc. No other disclosures were reported.
 

### 2022-11-13 paper/manuscript.Rmd
CD comments+edits

@@
 
-knitr::opts_chunk$set(
- echo = FALSE,
- message = FALSE,
- warning = FALSE,
- fig.width = 10.5,
- fig.height = 4, 
- dpi = 300,
- ft.keepnext = FALSE
-)
-
 suppressPackageStartupMessages({
@@
 })
+
+knitr::opts_chunk$set(
+ echo = FALSE,
+ message = FALSE,
+ warning = FALSE,
+ fig.width = 10.5,
+ fig.height = 4, 
+ dpi = 300,
+ ft.keepnext = FALSE
+)
 
@@
 Jennifer S. Herrick, MS,^3,6^
-Catherine Derington, PharmD, MS,^7^ 
+Catherine G. Derington, PharmD, MS,^7^ 
 Bharat Poudel, MSPH,^2^
@@
 
-^1^Department of Biostatistics and Data Science, Wake Forest School of Medicine, Winston-Salem, NC. `r single_par`
+^1^Department of Biostatistics and Data Science, Wake Forest University School of Medicine, Winston-Salem, NC. `r single_par`
 ^2^Department of Epidemiology, University of Alabama at Birmingham, Birmingham, AL. `r single_par`
-^3^Informatics, Decision-Enhancement, and Analytic Sciences (IDEAS) Center, Veterans Affairs, Salt Lake City Health Care System, Salt Lake City, UT. `r single_par`
-^4^Department of Population Health Sciences, University of Utah School of Medicine, Salt Lake City, UT. `r single_par`
-^5^Department of Pharmacy Practice, Rosalind Franklin University of Medicine and Science, North Chicago, IL, USA
-^6^Department of Internal Medicine, University of Utah School of Medicine, Salt Lake City, UT
-^7^ University of Colorado Skaggs School of Pharmacy and Pharmaceutical Sciences, Aurora, CO, USA
+^3^Informatics, Decision-Enhancement, and Analytic Sciences (IDEAS) Center, Veterans Affairs, Salt Lake City Health Care System, Salt Lake City, UT.`r single_par`
+^4^ Intermountain Healthcare Department of Population Health Sciences, Spencer Fox Eccles School of Medicine, University of Utah, Salt Lake City, UT. `r single_par`
+^5^Department of Pharmacy Practice, Rosalind Franklin University of Medicine and Science, North Chicago, IL, USA `r single_par`
+^6^Department of Internal Medicine, Spencer Fox Eccles School of Medicine, University of Utah, Salt Lake City, UT `r single_par`
 
@@
 **Methods**: We developed and validated a web-based application that provides hypertension statistics for US adults using 10 cycles of NHANES data, 1999-2000 through 2017-2020, without requiring users to conduct statistical programming.
-The application's interface allows users to estimate crude and age-adjusted means, quantiles, and proportions.
-Population counts can also be estimated.
-To demonstrate the application’s capabilities, we estimated hypertension statistics for US adults.The application was validated by replicating statistics from three publications.
+The application's interface allows users to estimate population counts in addition to crude and age-adjusted summary statistics.
+To demonstrate the application's capabilities, we estimated hypertension statistics for US adults.
+The application was validated by replicating statistics from three publications.
 
@@
 Specific statistical techniques are required to account for the multi-stage sampling design used to select NHANES participants and analyses need to be weighted to produce nationally representative estimates.
-NHANES data are currently collected in two-year periods, referred to as cycles, and each cycle includes over 50 data files and accompanying informational material.
+Prior studies have found analytic errors related to survey design and estimation methods are an important component of total survey error.[@west2017analytic, @west2016big] NHANES data are currently collected in two-year periods, referred to as cycles, and each cycle includes over 50 data files and accompanying informational material.
 NHANES data collection protocols have changed over time for some variables, and these differences should be considered when comparing or aggregating data from multiple cycles.
@@
 During the interview, the labels of medications that participants reported taking in the preceding 30 days were recorded.
-Antihypertensive medication classes were defined using those listed in the 2017 ACC/AHA BP guideline.[@whelton20182017] During the study examination, height, weight and BP were measured and blood and spot urine samples were collected.
+In the current study, antihypertensive medication classes were defined using those listed in the 2017 ACC/AHA BP guideline (Section 8.1.4; Table 18).[@whelton20182017] During the study examination, height, weight and BP were measured and blood and spot urine samples were collected.
 Of relevance to the current application, blood samples were used to measure total and high-density lipoprotein cholesterol, glycated hemoglobin and serum creatinine, and the urine sample was used to measure albumin and creatinine and to conduct a pregnancy test.
@@
 The age-adjusted prevalence of hypertension was highest in 1999-2000 (47.9%), lowest in 2009-2010 (43.0%), and 44.7% in 2017-2020 (**Figure 2**).
-In 1999-2000, there were an estimated 89.8 million US adults with hypertension (**Figure 3**).
-The number of US adults with hypertension increased to 115.3 million in 2017-2020.
+In 1999-2000, there were an estimated 89.8 (95% CI 77.9, 101.7) million US adults with hypertension (**Figure 3**).
+The number of US adults with hypertension increased to 115.3 (95% CI 107.4, 123.2) million in 2017-2020.
 In each NHANES cycle, the estimated prevalence of hypertension was higher among US adults with versus without CKD (**Figure S2**).
@@
 Using this application, we generated crude and age-adjusted BP and hypertension statistics.
-Also, we generated statistics stratified by characteristics of US adults and pooling multiple NHANES cycles to obtain more precise estimates when working with the small sub-group of pregnant women.
+Also, we generated statistics stratified by characteristics of US adults and demonstrated how multiple NHANES cycles can be pooled to obtain more precise estimates when working with small sub-groups (e.g., pregnant women).
 Following its validation, version 0.0.1 of the application was released and deployed on a publicly available server on **DATE TBD (it's almost ready)**.[@app_cardioStatsUSA] Researchers, clinicians, and the public can use the application to generate customized BP and hypertension statistics for US adults to inform their research and policy decisions.
@@
 
-NHANES was designed to obtain nationally representative estimates of the health and nutrition status of non-institutionalized US adults and is an ideal data source to obtain statistics related to hypertension.
+NHANES is an ideal data source to obtain statistics related to hypertension.
+It was designed to obtain nationally representative estimates of the health and nutrition status of non-institutionalized US adults.
 SBP and DBP were measured following a standardized protocol by trained and certified physicians.
@@
 We included spironolactone as a single drug class, an aldosterone antagonist.
-Additionally, Lexicon Plus® includes sotalol as a beta blocker and nitroglycerine as a direct vasodilator and we did not include these drugs as antihypertensive medications as they are not listed in the 2017 ACC/AHA BP guideline.[@whelton20182017] The differences in results between the application and this previously published manuscript emphasize that it may not be possible to always generate results that are identical to prior publications.
+Additionally, Lexicon Plus® includes sotalol as a beta blocker and nitroglycerine as a direct vasodilator and we did not include these drugs as antihypertensive medications as they are not listed in the 2017 ACC/AHA BP guideline and are not used for the treatment of hypertension.[@whelton20182017] The differences in results between the application and this previously published manuscript emphasize that it may not be possible to always generate results that are identical to prior publications.
 Users should be aware of the choices made in defining variables as this may affect the statistical estimates generated using the application.
@@
 
+footer_hyperlink <- function(url){
+ glue::glue(
+  "This commands used to generate this graph are [available online]({url})"
+ )
+}
+
 ```
@@
 
+`r footer_hyperlink(url = "https://bcjaeger.shinyapps.io/nhanesShinyBP/?_inputs_&standard_weights_2=33.6&help=0&stratify_class=null&group=null&wont_do_computation=0&standard_weights_1=49.3&year_stratify=%5B%221999-2000%22%2C%222001-2002%22%2C%222003-2004%22%2C%222005-2006%22%2C%222007-2008%22%2C%222009-2010%22%2C%222011-2012%22%2C%222013-2014%22%2C%222015-2016%22%2C%222017-2020%22%5D&shinyjs-delay-09f5dfa7e9684def7b5c54ddf5175adf=200&geom=%22scatter%22&outcome=%22bp_sys_mean%22&statistic_primary=%22mean%22&standard_weights_4=7&shinyjs-delay-9c5ec1c3f7219bb7bf0d2e2dae255e33=200&shinyjs-delay-c2a3e79393f36660786d425d151ef85e=500&n_exposure_group=null&pool=%22no%22&stratify=null&exposure_cut_type=%22interval%22&group_class=null&statistic=%22mean%22&subset_n=%220%22&outcome_class=%22Blood%20pressure%22&standard_weights_3=10.1&.clientValue-default-plotlyCrosstalkOpts=%7B%22on%22%3A%22plotly_click%22%2C%22persistent%22%3Afalse%2C%22dynamic%22%3Afalse%2C%22selectize%22%3Afalse%2C%22opacityDim%22%3A0.2%2C%22selected%22%3A%7B%22opacity%22%3A1%7D%2C%22debounce%22%3A0%2C%22color%22%3A%5B%5D%7D&do=%22figure%22&year_pool=%5B%221999-2000%22%2C%222017-2020%22%5D&plotly_afterplot-A=%22%5C%22htmlwidget-b15366fdef4727083aad%5C%22%22&age_standardize=false&plotly_hover-A=null")` `r single_par`
+
 \newpage
@@
 
+`r footer_hyperlink(url = "https://bcjaeger.shinyapps.io/nhanesShinyBP/?_inputs_&standard_weights_2=33.6&help=0&stratify_class=null&group=null&wont_do_computation=0&standard_weights_1=49.3&year_stratify=%5B%221999-2000%22%2C%222001-2002%22%2C%222003-2004%22%2C%222005-2006%22%2C%222007-2008%22%2C%222009-2010%22%2C%222011-2012%22%2C%222013-2014%22%2C%222015-2016%22%2C%222017-2020%22%5D&geom=%22bar%22&outcome=%22htn_accaha%22&statistic_primary=%22percentage%22&standard_weights_4=7&outcome_open=false&plotly_hover-A=null&n_exposure_group=null&pool=%22no%22&stratify=null&exposure_cut_type=%22interval%22&group_class=null&statistic=%22percentage%22&subset_n=%220%22&outcome_class=%22Hypertension%22&standard_weights_3=10.1&.clientValue-default-plotlyCrosstalkOpts=%7B%22on%22%3A%22plotly_click%22%2C%22persistent%22%3Afalse%2C%22dynamic%22%3Afalse%2C%22selectize%22%3Afalse%2C%22opacityDim%22%3A0.2%2C%22selected%22%3A%7B%22opacity%22%3A1%7D%2C%22debounce%22%3A0%2C%22color%22%3A%5B%5D%7D&do=%22figure%22&year_pool=%5B%221999-2000%22%2C%222017-2020%22%5D&plotly_afterplot-A=%22%5C%22htmlwidget-9eb170c07bf326bb924d%5C%22%22&age_standardize=true&outcome_class_open=false")` `r single_par`
+
 \newpage
@@
 `r footer_hover(parenthetical = "estimated count and the upper and lower limits of the 95% confidence interval", hover_over = "the bars")` `r single_par`
+
+`r footer_hyperlink(url = "https://bcjaeger.shinyapps.io/nhanesShinyBP/?_inputs_&standard_weights_2=33.6&help=0&stratify_class=null&group=null&wont_do_computation=0&standard_weights_1=49.3&year_stratify=%5B%221999-2000%22%2C%222001-2002%22%2C%222003-2004%22%2C%222005-2006%22%2C%222007-2008%22%2C%222009-2010%22%2C%222011-2012%22%2C%222013-2014%22%2C%222015-2016%22%2C%222017-2020%22%5D&geom=%22bar%22&outcome=%22htn_accaha%22&statistic_primary=%22count%22&standard_weights_4=7&outcome_open=false&plotly_hover-A=null&n_exposure_group=null&pool=%22no%22&stratify=null&exposure_cut_type=%22interval%22&group_class=null&statistic=%22count%22&subset_n=%220%22&outcome_class=%22Hypertension%22&standard_weights_3=10.1&.clientValue-default-plotlyCrosstalkOpts=%7B%22on%22%3A%22plotly_click%22%2C%22persistent%22%3Afalse%2C%22dynamic%22%3Afalse%2C%22selectize%22%3Afalse%2C%22opacityDim%22%3A0.2%2C%22selected%22%3A%7B%22opacity%22%3A1%7D%2C%22debounce%22%3A0%2C%22color%22%3A%5B%5D%7D&do=%22figure%22&year_pool=%5B%221999-2000%22%2C%222017-2020%22%5D&plotly_afterplot-A=%22%5C%22htmlwidget-72fc5230e0cbdfdb5676%5C%22%22&age_standardize=false&outcome_class_open=false")` `r single_par`
 
@@
 
+`r footer_hyperlink(url = "https://bcjaeger.shinyapps.io/nhanesShinyBP/?_inputs_&standard_weights_2=43.5&subset_variable_2=%22demo_pregnant%22&help=0&stratify_class=null&subset_value_1_ctns=%5B0%2C0%5D&subset_value_2_catg=%5B%22No%22%2C%22Missing%22%5D&subset_variable_1=%22htn_accaha%22&group=null&wont_do_computation=0&standard_weights_1=26.4&subset_variable_2_open=false&year_stratify=%5B%221999-2000%22%2C%222001-2002%22%2C%222003-2004%22%2C%222005-2006%22%2C%222007-2008%22%2C%222009-2010%22%2C%222011-2012%22%2C%222013-2014%22%2C%222015-2016%22%2C%222017-2020%22%5D&subset_value_1_catg=%22Yes%22&geom=%22bar%22&outcome=%22bp_control_accaha%22&statistic_primary=%22percentage%22&subset_n_open=false&standard_weights_4=13.2&outcome_open=false&plotly_hover-A=null&n_exposure_group=null&pool=%22no%22&stratify=null&exposure_cut_type=%22interval%22&group_class=null&statistic=%22percentage%22&subset_n=%222%22&outcome_class=%22Blood%20pressure%22&standard_weights_3=17&.clientValue-default-plotlyCrosstalkOpts=%7B%22on%22%3A%22plotly_click%22%2C%22persistent%22%3Afalse%2C%22dynamic%22%3Afalse%2C%22selectize%22%3Afalse%2C%22opacityDim%22%3A0.2%2C%22selected%22%3A%7B%22opacity%22%3A1%7D%2C%22debounce%22%3A0%2C%22color%22%3A%5B%5D%7D&do=%22figure%22&year_pool=%5B%221999-2000%22%2C%222017-2020%22%5D&plotly_afterplot-A=%22%5C%22htmlwidget-1d7ba8c29733e6d67fa7%5C%22%22&age_standardize=true&statistic_primary_open=false&subset_value_2_ctns=%5B0%2C0%5D&subset_variable_1_open=false&outcome_class_open=false")` `r single_par`
+
 \newpage
@@
 Survey participants with missing values for pregnancy status were assumed to be non-pregnant `r single_par`
+
+`r footer_hyperlink(url = "https://bcjaeger.shinyapps.io/nhanesShinyBP/?_inputs_&standard_weights_2=43.5&subset_variable_2=%22demo_pregnant%22&subset_variable_3=%22bp_med_use%22&help=0&stratify_class=null&subset_value_1_ctns=%5B0%2C0%5D&subset_value_2_catg=%5B%22No%22%2C%22Missing%22%5D&subset_variable_1=%22htn_accaha%22&group=null&wont_do_computation=0&standard_weights_1=26.4&subset_variable_2_open=false&subset_value_3_ctns=%5B0%2C0%5D&year_stratify=%5B%221999-2000%22%2C%222001-2002%22%2C%222003-2004%22%2C%222005-2006%22%2C%222007-2008%22%2C%222009-2010%22%2C%222011-2012%22%2C%222013-2014%22%2C%222015-2016%22%2C%222017-2020%22%5D&subset_value_1_catg=%22Yes%22&geom=%22bar%22&outcome=%22bp_control_accaha%22&statistic_primary=%22percentage%22&subset_n_open=false&standard_weights_4=13.2&outcome_open=false&plotly_hover-A=null&n_exposure_group=null&pool=%22no%22&stratify=null&exposure_cut_type=%22interval%22&group_class=null&statistic=%22percentage%22&subset_n=%223%22&outcome_class=%22Blood%20pressure%22&standard_weights_3=17&.clientValue-default-plotlyCrosstalkOpts=%7B%22on%22%3A%22plotly_click%22%2C%22persistent%22%3Afalse%2C%22dynamic%22%3Afalse%2C%22selectize%22%3Afalse%2C%22opacityDim%22%3A0.2%2C%22selected%22%3A%7B%22opacity%22%3A1%7D%2C%22debounce%22%3A0%2C%22color%22%3A%5B%5D%7D&do=%22figure%22&year_pool=%5B%221999-2000%22%2C%222017-2020%22%5D&plotly_afterplot-A=%22%5C%22htmlwidget-49855de1c73b4f93593a%5C%22%22&age_standardize=true&statistic_primary_open=false&subset_value_2_ctns=%5B0%2C0%5D&subset_variable_3_open=false&subset_value_3_catg=%22Yes%22&subset_variable_1_open=false&outcome_class_open=false")` `r single_par`
 

### 2022-11-14 paper/manuscript.Rmd
light edits

@@
 Specific statistical techniques are required to account for the multi-stage sampling design used to select NHANES participants and analyses need to be weighted to produce nationally representative estimates.
-Prior studies have found analytic errors related to survey design and estimation methods are an important component of total survey error.[@west2017analytic, @west2016big] NHANES data are currently collected in two-year periods, referred to as cycles, and each cycle includes over 50 data files and accompanying informational material.
+Prior studies have found errors related to survey design and estimation be an important component of total survey error.[@west2017analytic; @west2016big].
+NHANES data are currently collected in two-year periods, referred to as cycles, and each cycle includes over 50 data files and accompanying informational material.
 NHANES data collection protocols have changed over time for some variables, and these differences should be considered when comparing or aggregating data from multiple cycles.
@@
 The goal of the application is to increase the use of NHANES data for research and policy decision making with a focus on ensuring results are valid and reproducible.
-The application is also "open-source" as it's source code is publicly available.
-In this manuscript, we review the design, development and validation of the application and present BP and hypertension statistics for US adults that were generated using the application.
+The application is also "open-source" as it's source code is publicly available.[@rpack_cardioStatsUSA] In this manuscript, we review the design, development and validation of the application and present BP and hypertension statistics for US adults that were generated using the application.
 
@@
 
-The web application was created using Shiny,[@rpack_shiny; @rpack_shinyWidgets; @rpack_shinyalert] an open-source software package that translates code from the R programming language into HTML, CSS, or JavaScript commands and creates a website interface.[@r_language] We created the “cardioStatsUSA” R package to provide additional details on the web application’s design and comprehensive documentation of its components.[@rpack_cardioStatsUSA] We validated the web application by using it to reproduce statistics reported in two prior studies and one CDC report.[@muntner2020htn; @muntner2022blood; @fryar_hypertension_2017]
+The web application was created using Shiny,[@rpack_shiny; @rpack_shinyWidgets; @rpack_shinyalert] an open-source software package that translates code from the R programming language into HTML, CSS, or JavaScript commands and creates a website interface.[@r_language] We created the “cardioStatsUSA” R package to provide additional details on the web application’s design and comprehensive documentation of its components.[@rpack_cardioStatsUSA] We validated the web application by using it to reproduce statistics reported in two prior studies and one CDC report.[@muntner2020htn; @muntner2022blood; @fryar_hypertension_2017] Details on the validation are available online.
 
@@
   values = abbrvs_write(
-   abbrvs[c('BP', 'SBP', 'DBP', 'ACC', 'AHA', 'JNC7')]
+   abbrvs[c('BP', 'SBP', 'DBP', 'ACC', 'ACE', 'AHA', 'JNC7')]
   )

### 2022-11-15 paper/manuscript.Rmd
working through v9 edits

@@
  library(plotly)
+ library(table.glue)
 })
@@
 standard_weights_overall <- c(49.3, 33.6, 10.1, 7)
-standard_weights_htn <- c(26.4, 43.5, 17.0, 13.2)
+standard_weights_htn <- c(26.4, 43.4, 17.0, 13.2)
 
@@
 
+fig_print <- function(x){
+ x$fig_object[[1]] %>% 
+  layout(font = list(size = 40))
+}
+
+est_ci <- function(data){
+ with(data, table_glue("{estimate} (95% CI {ci_lower}, {ci_upper})"))
+}
+
+est_cols <- c("estimate", "ci_lower", "ci_upper")
+
+# Main tables ----
+
+# table 1 is created in a later code chunk
+
+# table 2
+
+# round to 1 decimal for consistency with the app
+rspec <- round_spec() %>% 
+ round_using_decimal(digits = 1)
+
+tbl_main_2_data <- nhanes_data %>% 
+ nhanes_summarize(key = nhanes_key,
+                  outcome_variable = 'demo_race',
+                  group_variable = 'htn_accaha',
+                  time_values = 'most_recent') %>%
+ mutate(
+  across(where(is.numeric), .fns = table_value, rspec = rspec)
+ ) %>% 
+ filter(statistic == 'percentage')
+
+# inline results for table 2
+
+htn_black_2017 <- tbl_main_2_data %>% 
+ .[htn_accaha == 'Yes'] %>% 
+ .[demo_race == 'Non-Hispanic Black'] %>% 
+ est_ci()
+
+nrm_black_2017 <- tbl_main_2_data %>% 
+ .[htn_accaha == 'No'] %>% 
+ .[demo_race == 'Non-Hispanic Black'] %>% 
+ est_ci()
+
+htn_hispanic_2017 <- tbl_main_2_data %>% 
+ .[htn_accaha == 'Yes'] %>% 
+ .[demo_race == 'Hispanic'] %>% 
+ est_ci()
+
+nrm_hispanic_2017 <- tbl_main_2_data %>% 
+ .[htn_accaha == 'No'] %>% 
+ .[demo_race == 'Hispanic'] %>% 
+ est_ci()
+
+# putting table 2 into flextable format
+
+tbl_main_2_flex <- flextable(tbl_main_2_data) %>% 
+ theme_box() %>% 
+ width(width = 0.875) %>% 
+ width(j = ~demo_race, width = 1) %>% 
+ width(j = ~statistic, width = 1) %>% 
+ width(j = ~unreliable_reason, width = 0.7) %>% 
+ width(j = ~review_reason, width = 0.7) %>% 
+ align(align = 'center', part = 'all') %>% 
+ align(j = 1, align = 'center', part = 'all') %>% 
+ add_footer_lines(
+  values = "Table format is identical to the web application's output."
+ ) %>% 
+ add_footer_lines(
+  values = "svy_year indicates the NHANES cycle"
+ ) %>% 
+ add_footer_lines(
+  values = "htn_accaha indicates hypertension status." 
+ ) %>% 
+ add_footer_lines(
+  values = "demo_race indicates the race subgroup."
+ ) %>% 
+ add_footer_lines(
+  values = "statistic indicates the type of statistic computed."
+ ) %>% 
+ add_footer_lines(
+  values = "estimate is the point estimate of the given statistic"
+ ) %>% 
+ add_footer_lines(
+  values = "std_error is the standard error for estimate"
+ ) %>% 
+ add_footer_lines(
+  values = "ci_lower is the lower bound of a 95% confidence interval for estimate"
+ ) %>% 
+ add_footer_lines(
+  values = "ci_upper is the upper bound of a 95% confidence interval for estimate"
+ ) %>% 
+ add_footer_lines(
+  values = "n_obs is the number of observations (unweighted) used for computations"
+ ) %>% 
+ add_footer_lines(
+  values = "unreliable_status is TRUE if the result is unreliable, FALSE otherwise"
+ ) %>% 
+ add_footer_lines(
+  values = "unreliable_reason is blank if the result is reliable, and lists reason(s) for unreliability otherwise"
+ ) %>% 
+ add_footer_lines(
+  values = "review_needed is TRUE if the result should be reviewed, FALSE otherwise"
+ ) %>% 
+ add_footer_lines(
+  values = "review_reason is blank if the result does not require review, and lists reason(s) for requiring review otherwise"
+ )
+
+# Main figures ----
+
+# figure 1
+fig_main_1 <- nhanes_data %>% 
+ nhanes_visualize(key = nhanes_key,
+                  outcome_variable = 'bp_sys_mean',
+                  time_values = 'all', 
+                  geom = 'points',
+                  size_point = 25,
+                  size_error = 1)
+
+# inline results for figure 1
+sbp_1999 <- est_ci(fig_main_1$fig_data[svy_year == '1999-2000'])
+sbp_2009 <- est_ci(fig_main_1$fig_data[svy_year == '2009-2010'])
+sbp_2017 <- est_ci(fig_main_1$fig_data[svy_year == '2017-2020'])
+
+# figure 2
+fig_main_2 <- nhanes_data %>% 
+ nhanes_visualize(key = nhanes_key,
+                  outcome_variable = 'htn_accaha',
+                  time_values = 'all', 
+                  standard_weights = standard_weights_overall, 
+                  geom = 'bar')
+
+# inline results for figure 2
+htn_1999 <- fig_main_2 %>% 
+ getElement("fig_data") %>% 
+ .[svy_year == '1999-2000' & htn_accaha == 'Yes'] %>% 
+ est_ci()
+
+htn_2009 <- fig_main_2 %>% 
+ getElement("fig_data") %>% 
+ .[svy_year == '2009-2010' & htn_accaha == 'Yes'] %>% 
+ est_ci()
+
+htn_2017 <- fig_main_2 %>% 
+ getElement("fig_data") %>% 
+ .[svy_year == '2017-2020' & htn_accaha == 'Yes'] %>% 
+ est_ci()
+
+# figure 3
+fig_main_3 <- nhanes_data %>% 
+ nhanes_visualize(key = nhanes_key,
+                  outcome_variable = 'htn_accaha',
+                  time_values = 'all',
+                  statistic_primary = 'count',
+                  geom = 'bar')
+
+# inline results for figure 3
+n_htn_1999 <- fig_main_3 %>% 
+ getElement("fig_data") %>% 
+ .[svy_year == '1999-2000' & htn_accaha == 'Yes'] %>% 
+ .[, (est_cols) := lapply(.SD, function(x) table_value(x/1e6)), 
+   .SDcols = est_cols] %>% 
+ est_ci()
+
+n_htn_2017 <- fig_main_3 %>% 
+ getElement("fig_data") %>% 
+ .[svy_year == '2017-2020' & htn_accaha == 'Yes'] %>% 
+ .[, (est_cols) := lapply(.SD, function(x) table_value(x/1e6)), 
+   .SDcols = est_cols] %>% 
+ est_ci()
+
+# figure 4, panel a
+fig_main_4_a <- nhanes_data %>% 
+ nhanes_visualize(
+  key = nhanes_key,
+  outcome_variable = 'bp_control_accaha',
+  subset_calls = list("htn_accaha" = "Yes",
+                      "demo_pregnant" = c("No", "Missing")),
+  standard_weights = standard_weights_htn,
+  time_values = 'all', 
+  geom = 'bar')
+
+# inline results for figure 4 panel a
+bp_control_1999a <- fig_main_4_a %>% 
+ getElement("fig_data") %>% 
+ .[svy_year == '1999-2000' & bp_control_accaha == 'Yes'] %>% 
+ est_ci()
+
+bp_control_2013a <- fig_main_4_a %>% 
+ getElement("fig_data") %>% 
+ .[svy_year == '2013-2014' & bp_control_accaha == 'Yes'] %>% 
+ est_ci()
+
+bp_control_2017a <- fig_main_4_a %>% 
+ getElement("fig_data") %>% 
+ .[svy_year == '2017-2020' & bp_control_accaha == 'Yes'] %>% 
+ est_ci()
+
+# figure 4, panel b
+fig_main_4_b <- nhanes_data %>% 
+ nhanes_visualize(
+  key = nhanes_key,
+  outcome_variable = 'bp_control_accaha',
+  subset_calls = list("htn_accaha" = "Yes",
+                      "bp_med_use" = "Yes",
+                      "demo_pregnant" = c("No", "Missing")),
+  standard_weights = standard_weights_htn,
+  time_values = 'all', 
+  geom = 'bar'
+ )
+
+bp_control_1999b <- fig_main_4_b %>% 
+ getElement("fig_data") %>% 
+ .[svy_year == '1999-2000' & bp_control_accaha == 'Yes'] %>% 
+ est_ci()
+
+bp_control_2013b <- fig_main_4_b %>% 
+ getElement("fig_data") %>% 
+ .[svy_year == '2013-2014' & bp_control_accaha == 'Yes'] %>% 
+ est_ci()
+
+bp_control_2017b <- fig_main_4_b %>% 
+ getElement("fig_data") %>% 
+ .[svy_year == '2017-2020' & bp_control_accaha == 'Yes'] %>% 
+ est_ci()
+
+# figure S4, panel a
+fig_supp_4_a <- nhanes_data %>% 
+ nhanes_visualize(
+  key = nhanes_key,
+  outcome_variable = 'htn_resistant_accaha',
+  subset_calls = list(
+   "demo_pregnant" = c("No", "Missing"),
+   "bp_med_use" = "Yes",
+   "bp_med_n_class" = c('One', 'Two', 'Three', 'Four or more')
+  ),
+  standard_weights = standard_weights_htn,
+  time_values = 'all', 
+  geom = 'bar'
+ )
+
+htn_resistant_1999a <- fig_supp_4_a %>% 
+ getElement("fig_data") %>% 
+ .[svy_year == '1999-2000' & htn_resistant_accaha == 'Yes'] %>% 
+ est_ci()
+
+htn_resistant_2005a <- fig_supp_4_a %>% 
+ getElement("fig_data") %>% 
+ .[svy_year == '2005-2006' & htn_resistant_accaha == 'Yes'] %>% 
+ est_ci()
+
+htn_resistant_2017a <- fig_supp_4_a %>% 
+ getElement("fig_data") %>% 
+ .[svy_year == '2017-2020' & htn_resistant_accaha == 'Yes'] %>% 
+ est_ci()
+
+# figure S4, panel b
+fig_supp_4_b <- nhanes_data %>% 
+ nhanes_visualize(
+  key = nhanes_key,
+  outcome_variable = 'htn_resistant_accaha',
+  subset_calls = list(
+   "demo_pregnant" = c("No", "Missing"),
+   "bp_med_use" = "Yes",
+   "bp_med_n_class" = c('Three', 'Four or more')
+  ),
+  standard_weights = standard_weights_htn,
+  time_values = 'all', 
+  geom = 'bar'
+ )
+
+htn_resistant_2009b <- fig_supp_4_b %>% 
+ getElement("fig_data") %>% 
+ .[svy_year == '2009-2010' & htn_resistant_accaha == 'Yes'] %>% 
+ est_ci()
+
+htn_resistant_2003b <- fig_supp_4_b %>% 
+ getElement("fig_data") %>% 
+ .[svy_year == '2003-2004' & htn_resistant_accaha == 'Yes'] %>% 
+ est_ci()
+
+htn_resistant_2017b <- fig_supp_4_b %>% 
+ getElement("fig_data") %>% 
+ .[svy_year == '2017-2020' & htn_resistant_accaha == 'Yes'] %>% 
+ est_ci()
+
 ```
@@
 **Methods**: We developed and validated a web-based application that provides hypertension statistics for US adults using 10 cycles of NHANES data, 1999-2000 through 2017-2020, without requiring users to conduct statistical programming.
-The application's interface allows users to estimate population counts in addition to crude and age-adjusted summary statistics.
-To demonstrate the application's capabilities, we estimated hypertension statistics for US adults.
-The application was validated by replicating statistics from three publications.
+The application’s interface allows users to estimate crude and age-adjusted means, quantiles and proportions.
+Population counts can also be estimated.
+To demonstrate the application’s capabilities, we estimated hypertension statistics for US adults.
 
-**Results**: The estimated mean systolic blood pressure (BP) declined from 123 mmHg in 1999-2000 to 120 mmHg in 2009-2010 and increased to 123 mmHg in 2017-2020.
-The age-adjusted prevalence of hypertension (i.e., systolic BP ≥130 mmHg, diastolic BP ≥80 mmHg or self-reported antihypertensive medication use) was 47.9% in 1999-2000, 43.0% in 2009-2010, and 44.7% in 2017-2020.
-In 2017-2020, 115.3 million US adults had hypertension.
-The age-adjusted prevalence of controlled BP, defined by the 2017 American College of Cardiology/American Heart Association BP guideline, among non-pregnant US adults with hypertension was 9.7% in 1999-2000, 25.0% in 2013-2014, and 21.9% in 2017-2020.
-After age-adjustment and among non-pregnant US adults who self-reported taking antihypertensive medication, 27.6%, 48.5%, and 43.0% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively.
+**Results**: The estimated mean systolic blood pressure (BP) declined from `r bracket_drop(sbp_1999)` mmHg in 1999-2000 to `r bracket_drop(sbp_2009)` mmHg in 2009-2010 and increased to `r bracket_drop(sbp_2017)` mmHg in 2017-2020.
+The age-adjusted prevalence of hypertension (i.e., systolic BP ≥130 mmHg, diastolic BP ≥80 mmHg or self-reported antihypertensive medication use) was `r bracket_drop(htn_1999)`% in 1999-2000, `r bracket_drop(htn_2009)`% in 2009-2010, and `r bracket_drop(htn_2017)`% in 2017-2020.
+In 2017-2020, an estimated `r bracket_drop(n_htn_2017)` million US adults had hypertension.
+The age-adjusted prevalence of controlled BP, defined by the 2017 American College of Cardiology/American Heart Association BP guideline, among non-pregnant US adults with hypertension was `r bracket_drop(bp_control_1999a)`% in 1999-2000, `r bracket_drop(bp_control_2013a)`% in 2013-2014, and `r bracket_drop(bp_control_2017a)`% in 2017-2020.
+After age-adjustment and among non-pregnant US adults who self-reported taking antihypertensive medication, `r bracket_drop(bp_control_1999b)`%, `r bracket_drop(bp_control_2013b)`%, and `r bracket_drop(bp_control_2017b)`% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively.
 
@@
 
-The National Health and Nutrition Examination Survey (NHANES) is a program conducted by the US National Center for Health Statistics (NCHS) of the Centers for Disease Control and Prevention (CDC) and is designed to assess the ongoing health and nutritional status of the non-institutionalized US population.[@nhanes_about] NHANES data have been analyzed to provide hypertension statistics for non-institutionalized US adults with important policy and public health implications.
-For example, NHANES data have been used to estimate the impact of the lower blood pressure (BP) levels that define hypertension and controlled BP in the 2017 American College of Cardiology/American Heart Association (ACC/AHA) BP guideline versus the Seventh Report of the Joint National Committee on Prevention, Detection, Evaluation, and Treatment of High BP (JNC7) guideline.[@muntner_potential_2018] Additionally, NHANES data have been used to track the proportion of US adults with hypertension that have controlled BP, which were included in a Call-to-Action to Control BP from the US Surgeon General.[@us_surgeon_2020]
+The National Health and Nutrition Examination Survey (NHANES) is a program conducted by the US National Center for Health Statistics (NCHS) of the Centers for Disease Control and Prevention (CDC) and is designed to assess the health and nutritional status of the non-institutionalized US population.[@nhanes_about] NHANES data have been analyzed to provide hypertension statistics for non-institutionalized US adults with important policy and public health implications.
+For example, NHANES data have been used to estimate the impact of the lower blood pressure (BP) levels that define hypertension and controlled BP in the 2017 American College of Cardiology/American Heart Association (ACC/AHA) BP guideline compared with the Seventh Report of the Joint National Committee on Prevention, Detection, Evaluation, and Treatment of High BP (JNC7) guideline.[@muntner_potential_2018] Additionally, NHANES data have been used to track the proportion of US adults with hypertension that have controlled BP, which were included in a Call-to-Action to Control BP from the US Surgeon General.[@us_surgeon_2020]
 
@@
 
-NHANES data are publicly available and accessible through the CDC website.[@nhanes_home; @nhanes_data] However, analyzing NHANES data may be challenging for a number of reasons.
-Specific statistical techniques are required to account for the multi-stage sampling design used to select NHANES participants and analyses need to be weighted to produce nationally representative estimates.
-Prior studies have found errors related to survey design and estimation be an important component of total survey error.[@west2017analytic; @west2016big].
-NHANES data are currently collected in two-year periods, referred to as cycles, and each cycle includes over 50 data files and accompanying informational material.
-NHANES data collection protocols have changed over time for some variables, and these differences should be considered when comparing or aggregating data from multiple cycles.
-For example, while BP was measured using a mercury sphygmomanometer until 2016,[@centers2015national] a validated oscillometric device was used in 2017 through 2020.[@ostchega2010assessing] Also, tests need to be performed to ensure the statistical estimates are reliable, and unstable estimates should be suppressed.[@parker2017national]
+NHANES data are publicly available and accessible through the CDC website.[@nhanes_home; @nhanes_data] However, analyzing NHANES data requires specific statistical techniques to account for the multi-stage sampling design used to select participants and analyses need to be weighted to produce nationally representative estimates.
+Also, NHANES data are collected in two-year periods referred to as cycles and protocols have changed over time for some variables.
+These changes need to be incorporated when estimating statistics from multiple NHANES cycles.
 
@@
 
-We developed an open-source web-based application that provides nationally representative BP and hypertension statistics for non-institutionalized US adults using NHANES data without requiring users to conduct statistical programming.
+We developed a web-based application that provides nationally representative BP and hypertension statistics for non-institutionalized US adults using NHANES data without requiring users to conduct statistical programming.
 The goal of the application is to increase the use of NHANES data for research and policy decision making with a focus on ensuring results are valid and reproducible.
-The application is also "open-source" as it's source code is publicly available.[@rpack_cardioStatsUSA] In this manuscript, we review the design, development and validation of the application and present BP and hypertension statistics for US adults that were generated using the application.
-
-<br>
+The application is open-source,[@perens1999open] which means the code supporting it is publicly available and may be re-distributed or modified by anyone.[@rpack_cardioStatsUSA] In this manuscript, we review the design, development and validation of the application and present BP and hypertension statistics for US adults that were generated using the application.
 
@@
 
-NHANES data were collected through an in-home interview and study examination conducted at a mobile examination center.
+NHANES data were collected through an in-home interview and a study examination conducted at a mobile examination center.
 The interview included questions about demographics, health behaviors, medical history, and medication use.
 During the interview, the labels of medications that participants reported taking in the preceding 30 days were recorded.
-In the current study, antihypertensive medication classes were defined using those listed in the 2017 ACC/AHA BP guideline (Section 8.1.4; Table 18).[@whelton20182017] During the study examination, height, weight and BP were measured and blood and spot urine samples were collected.
+We defined antihypertensive medication classes using those listed in the 2017 ACC/AHA BP guideline (Section 8.1.4; Table 18).[@whelton20182017] During the study examination, height, weight and BP were measured and blood and spot urine samples were collected.
 Of relevance to the current application, blood samples were used to measure total and high-density lipoprotein cholesterol, glycated hemoglobin and serum creatinine, and the urine sample was used to measure albumin and creatinine and to conduct a pregnancy test.
@@
 The mean SBP and DBP levels were computed over all available measurements for each participant.
-The oscillometric SBP and DBP values were calibrated to the mercury device.[@muntner2022blood; @ostchega2021comparing] For the current analysis, we defined hypertension, BP control, and resistant hypertension according to the 2017 ACC/AHA BP guideline.[@whelton20182017] The application also has these variables defined according to the JNC7 definitions.[@chobanian2003seventh] Chronic kidney disease (CKD) was defined by estimated glomerular filtration rate < 60 ml/min/1.73 m^2^ calculated using the 2021 serum creatinine-based equation or albumin-to-creatinine ratio > 30 mg/g.[@inker_new_2021] A list of BP, hypertension, and antihypertensive medication variables is provided in **Table 1** with full definitions for all variables provided in **Table S1**.
+The oscillometric SBP and DBP values were calibrated to the mercury device.[@muntner2022blood; @ostchega2021comparing] For the current analysis, we defined hypertension, BP control, and resistant hypertension according to the 2017 ACC/AHA BP guideline.[@whelton20182017] The application also has these variables defined according to the JNC7 definitions.[@chobanian2003seventh] A list of BP, hypertension, and antihypertensive medication variables available in the application is provided in **Table 1** with full definitions for all variables included in the application provided in **Table S1**.
 
@@
 
-A full summary of this application’s features and associated tutorials are available online.[@rpack_cardioStatsUSA] Briefly, users can select NHANES cycles from 1999-2000 to 2017-2020 to be analyzed.
-Estimates are weighted to represent the non-institutionalized US population and users may incorporate age-adjustment through direct standardization.
+A full summary of the application’s features and associated tutorials are available online.[@rpack_cardioStatsUSA] Briefly, users can select NHANES cycles from 1999-2000 to 2017-2020 to be analyzed.
+Estimates are weighted to represent the non-institutionalized US population and users can incorporate age-adjustment through direct standardization.
 Users can restrict analyses to subsets of participants (e.g., participants who self-reported taking antihypertensive medication).
-When population count estimates are requested, survey weights are calibrated within race, sex, and age groups to account for missing information on SBP, DBP or self-reported antihypertensive medication use.
-The results can be presented in tables or figures and for the overall population or in subgroups.
-All tables and figures created with the web application can be downloaded and saved in standard formats.
+When population count estimates are requested, survey weights are calibrated within race, sex, and age groups to account for missing information on SBP, DBP or self-reported antihypertensive medication use.[@coresh_prevalence_2003] The results can be presented in tables or figures and for the overall population or in subgroups.
+All tables and figures created with the web application can be downloaded and saved.
 Following CDC recommendations, unreliable statistical estimates are automatically suppressed.[@parker2017national] To increase precision and reliability of estimates, contiguous NHANES cycles can be combined.[@nhanes_tutorial_weights]
@@
 
-The web application was created using Shiny,[@rpack_shiny; @rpack_shinyWidgets; @rpack_shinyalert] an open-source software package that translates code from the R programming language into HTML, CSS, or JavaScript commands and creates a website interface.[@r_language] We created the “cardioStatsUSA” R package to provide additional details on the web application’s design and comprehensive documentation of its components.[@rpack_cardioStatsUSA] We validated the web application by using it to reproduce statistics reported in two prior studies and one CDC report.[@muntner2020htn; @muntner2022blood; @fryar_hypertension_2017] Details on the validation are available online.
+The web application was created using Shiny,[@rpack_shiny; @rpack_shinyWidgets; @rpack_shinyalert] an open-source software package that translates code from the R programming language into HTML, CSS, or JavaScript commands and creates a website interface.[@r_language] We validated the web application by using it to reproduce statistics reported in two prior studies and one CDC report.[@muntner2020htn; @muntner2022blood; @fryar_hypertension_2017] We created the “cardioStatsUSA” R package to provide additional details on the web application’s design,  documentation of its components, and further details on its validation.[@rpack_cardioStatsUSA]
 
@@
 Specifically, we estimated the age-adjusted proportion of US adults with controlled BP by NHANES cycle among non-pregnant US adults with hypertension, overall and among those who self-reported taking antihypertensive medication.
-We also estimated the age-adjusted prevalence of resistant hypertension by NHANES cycle for non-pregnant US adults with hypertension who self-reported taking antihypertensive medication and had ≥ 1 classes of antihypertensive medication identified during the medication inventory and among those with ≥ 3 classes of antihypertensive medication identified during the medication inventory.
+We also estimated the age-adjusted prevalence of resistant hypertension by NHANES cycle for non-pregnant US adults with hypertension who self-reported taking antihypertensive medication and had ≥ 1 classes of antihypertensive medication identified during the medication inventory.
+This was repeated among those taking ≥ 3 classes of antihypertensive medication.
 For age adjustment in the analysis of BP control and resistant hypertension, we set the age distribution for the standard population to represent US adults with hypertension from 1999 to 2020: 26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥ 75 years of age, respectively.
@@
 
-Among non-institutionalized US adults ≥ 18 years of age, the estimated mean SBP was 123 (95% CI 121, 124) mm Hg in 1999-2000, 120 (95% CI 120, 121) mm Hg in 2009-2010 and 123 (95% CI 122, 124) mm Hg in 2017-2020 (**Figure 1**).
-The age-adjusted prevalence of hypertension was highest in 1999-2000 (47.9%), lowest in 2009-2010 (43.0%), and 44.7% in 2017-2020 (**Figure 2**).
-In 1999-2000, there were an estimated 89.8 (95% CI 77.9, 101.7) million US adults with hypertension (**Figure 3**).
-The number of US adults with hypertension increased to 115.3 (95% CI 107.4, 123.2) million in 2017-2020.
+Among non-institutionalized US adults ≥ 18 years of age, the estimated mean SBP was `r sbp_1999` mm Hg in 1999-2000, `r sbp_2009` mm Hg in 2009-2010 and `r sbp_2017` mm Hg in 2017-2020 (**Figure 1**).
+The age-adjusted prevalence of hypertension defined as in the 2017 ACC/AHA BP guideline was highest in 1999-2000 (`r bracket_drop(htn_1999)`%), lowest in 2009-2010 (`r bracket_drop(htn_2009)` ), and `r bracket_drop(htn_2017)`% in 2017-2020 (**Figure 2**).
+In 1999-2000, there were an estimated `r n_htn_1999` million US adults with hypertension (**Figure 3**).
+The number of US adults with hypertension increased to `r n_htn_2017` million in 2017-2020.
 In each NHANES cycle, the estimated prevalence of hypertension was higher among US adults with versus without CKD (**Figure S2**).
-In 2017-2020, a higher percentage of US adults with versus without hypertension were non-Hispanic Black (13.5% versus 9.3%) while a lower percentage of US adults with versus without hypertension were Hispanic (12.3% versus 18.1%) (**Table 2**).
+In 2017-2020, a higher percentage of US adults with versus without hypertension were non-Hispanic Black (`r bracket_drop(htn_black_2017)`% versus `r bracket_drop(nrm_black_2017)`%) while a lower percentage of US adults with versus without hypertension were Hispanic (`r bracket_drop(htn_hispanic_2017)`% versus `r bracket_drop(nrm_hispanic_2017)`%) (**Table 2**).
 Among pregnant women, the distribution of BP categories could not be estimated reliably in 2017-2020 (**Figure S3; Panel A**), but it could be estimated reliably after pooling NHANES cycles from 2009-2010 through 2017-2020 (**Figure S3; Panel B**)
@@
 
-The age-adjusted prevalence of BP control among non-pregnant US adults with hypertension was lowest in 1999-2000 (9.7%), highest in 2013-2014 (25.0%), and 21.9% in 2017-2020 (**Figure 4; panel A**).
-Among US adults with hypertension who self-reported taking antihypertensive medication, 27.6%, 48.5%, and 43.0% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively (**Figure 4; panel B**).
-The age-adjusted prevalence of resistant hypertension among non-pregnant US adults with hypertension who self-reported taking antihypertensive medication and had ≥1 antihypertensive medication class identified during the medication inventory was lowest in 1999-2000 (14.3%), highest in 2005-2006 (21.4%), and 16.6% in 2017-2020 (**Figure S4; panel A**).
-Further restricting this analysis to those who were taking three or more classes of antihypertensive medication, the prevalence of resistant hypertension was lowest in 2009-2010 (64.1%), highest in 2003-2004 (78.3%), and 67.9% in 2017-2020 (**Figure S4; panel B**).
+The age-adjusted prevalence of BP control defined as in the 2017 ACC/AHA BP guideline among non-pregnant US adults with hypertension was lowest in 1999-2000 (`r bracket_drop(bp_control_1999a)`%), highest in 2013-2014 (`r bracket_drop(bp_control_2013a)`%), and `r bracket_drop(bp_control_2017a)`% in 2017-2020 (**Figure 4; Panel A**).
+Among US adults with hypertension who self-reported taking antihypertensive medication, `r bracket_drop(bp_control_1999b)`%, `r bracket_drop(bp_control_2013b)`%, and `r bracket_drop(bp_control_2017b)`% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively (**Figure 4; Panel B**).
+The age-adjusted prevalence of resistant hypertension defined as in the 2017 ACC/AHA BP guideline among non-pregnant US adults with hypertension who self-reported taking antihypertensive medication and had ≥1 antihypertensive medication class identified during the medication inventory was lowest in 1999-2000 (`r bracket_drop(htn_resistant_1999a)`%), highest in 2005-2006 (21.4%), and 16.6% in 2017-2020 (**Figure S4; Panel A**).
+Further restricting this analysis to those who were taking three or more classes of antihypertensive medication, the prevalence of resistant hypertension was lowest in 2009-2010 (64.1%), highest in 2003-2004 (78.3%), and 67.9% in 2017-2020 (**Figure S4; Panel B**).
 
@@
 
-In the current manuscript, we present a web application with a user-friendly point and click interface that allows the calculation of nationally representative estimates for BP, hypertension and antihypertensive medication-related statistics using publicly available NHANES data.
+In the current manuscript, we present a web application with a user-friendly point and click interface that allows the calculation of nationally representative estimates for BP, hypertension and antihypertensive medication-related statistics using NHANES data.
 Using this application, we generated crude and age-adjusted BP and hypertension statistics.
@@
 Although multiple reports based on the data are publicly available, it may be difficult to customize reports based on user requests.
-The web application that we present in the current manuscript addresses these challenges, allowing NHANES data to more fully inform public health decisions and future research related to hypertension.
+The web application that we present in the current manuscript addresses these challenges, increasing the accessibility of NHANES data to clinicians, researchers, and decision-makers.
 
@@
 We included spironolactone as a single drug class, an aldosterone antagonist.
-Additionally, Lexicon Plus® includes sotalol as a beta blocker and nitroglycerine as a direct vasodilator and we did not include these drugs as antihypertensive medications as they are not listed in the 2017 ACC/AHA BP guideline and are not used for the treatment of hypertension.[@whelton20182017] The differences in results between the application and this previously published manuscript emphasize that it may not be possible to always generate results that are identical to prior publications.
+Additionally, Lexicon Plus® includes sotalol as a beta blocker and nitroglycerine as a direct vasodilator, and although these are mechanisms of actions for these drugs, we did not include these drugs as antihypertensive medications as they are not listed in the 2017 ACC/AHA BP guideline and are not approved by the Food and Drug Administration with an indication for treating essential hypertension.[@whelton20182017] The differences in results between the application and this previously published manuscript emphasize that it may not be possible to always generate results that are identical to prior publications.
 Users should be aware of the choices made in defining variables as this may affect the statistical estimates generated using the application.
@@
 This study also has several limitations.
-NHANES participants had their mean BP measured during a single visit, and BP guidelines recommend obtaining mean BP using at least two BP values measured on separate days.[@whelton20182017] The response rate for NHANES has declined from 1999-2000 through 2017-2020.
+NHANES participants had their mean BP measured during a single visit, and BP guidelines recommend obtaining mean BP using at least two BP values measured on separate days.[@whelton20182017] Many estimates, including medication use, are derived from participant report, and we cannot exclude the possibilities of reporting or recall biases.
+The response rate for NHANES has declined from 1999-2000 through 2017-2020.
 Effects of this decline are unclear.
@@
 
-library(table.glue)
-
-rspec <- round_spec() %>% 
- round_using_decimal(digits = 1)
-
-nhanes_data %>% 
- nhanes_summarize(key = nhanes_key,
-                  outcome_variable = 'demo_race',
-                  group_variable = 'htn_accaha',
-                  time_values = 'most_recent') %>%
- mutate(
-  across(where(is.numeric), 
-         .fns = table_value, 
-         rspec = rspec)
- ) %>% 
- filter(statistic == 'percentage') %>% 
- flextable() %>% 
- theme_box() %>% 
- width(width = 0.875) %>% 
- width(j = ~demo_race, width = 1) %>% 
- width(j = ~statistic, width = 1) %>% 
- width(j = ~unreliable_reason, width = 0.7) %>% 
- width(j = ~review_reason, width = 0.7) %>% 
- align(align = 'center', part = 'all') %>% 
- align(j = 1, align = 'center', part = 'all') %>% 
- add_footer_lines(
-  values = "Table format is identical to the web application's output."
- ) %>% 
- add_footer_lines(
-  values = "svy_year indicates the NHANES cycle"
- ) %>% 
- add_footer_lines(
-  values = "htn_accaha indicates hypertension status." 
- ) %>% 
- add_footer_lines(
-  values = "demo_race indicates the race subgroup."
- ) %>% 
- add_footer_lines(
-  values = "statistic indicates the type of statistic computed."
- ) %>% 
- add_footer_lines(
-  values = "estimate is the point estimate of the given statistic"
- ) %>% 
- add_footer_lines(
-  values = "std_error is the standard error for estimate"
- ) %>% 
- add_footer_lines(
-  values = "ci_lower is the lower bound of a 95% confidence interval for estimate"
- ) %>% 
- add_footer_lines(
-  values = "ci_upper is the upper bound of a 95% confidence interval for estimate"
- ) %>% 
- add_footer_lines(
-  values = "n_obs is the number of observations (unweighted) used for computations"
- ) %>% 
- add_footer_lines(
-  values = "unreliable_status is TRUE if the result is unreliable, FALSE otherwise"
- ) %>% 
- add_footer_lines(
-  values = "unreliable_reason is blank if the result is reliable, and lists reason(s) for unreliability otherwise"
- ) %>% 
- add_footer_lines(
-  values = "review_needed is TRUE if the result should be reviewed, FALSE otherwise"
- ) %>% 
- add_footer_lines(
-  values = "review_reason is blank if the result does not require review, and lists reason(s) for requiring review otherwise"
- )
+tbl_main_2_flex
 
@@
 
-fig_print <- function(x){
- x[[1]] %>% 
-  layout(font = list(size = 40))
-}
-
-nhanes_visualize(data = nhanes_data,
-                 key = nhanes_key,
-                 outcome_variable = 'bp_sys_mean',
-                 time_values = 'all', 
-                 geom = 'points',
-                 size_point = 25,
-                 size_error = 1) %>% 
- fig_print()
+fig_print(fig_main_1)
 
@@
 
-nhanes_visualize(data = nhanes_data,
-                 key = nhanes_key,
-                 outcome_variable = 'htn_accaha',
-                 time_values = 'all', 
-                 standard_weights = standard_weights_overall, 
-                 geom = 'bar') %>% 
- fig_print()
+fig_print(fig_main_2)
 
@@
 
-nhanes_visualize(data = nhanes_data,
-                 key = nhanes_key,
-                 outcome_variable = 'htn_accaha',
-                 time_values = 'all',
-                 statistic_primary = 'count',
-                 geom = 'bar') %>% 
- fig_print()
+fig_print(fig_main_3)
 
@@
 
-p1 <- 
- nhanes_visualize(
-  data = nhanes_data,
-  key = nhanes_key,
-  outcome_variable = 'bp_control_accaha',
-  subset_calls = list("htn_accaha" = "Yes",
-                      "demo_pregnant" = c("No", "Missing")),
-  standard_weights = standard_weights_htn,
-  time_values = 'all', 
-  geom = 'bar')
-
-p2 <- 
- nhanes_visualize(
-  data = nhanes_data,
-  key = nhanes_key,
-  outcome_variable = 'bp_control_accaha',
-  subset_calls = list("htn_accaha" = "Yes",
-                      "bp_med_use" = "Yes",
-                      "demo_pregnant" = c("No", "Missing")),
-  standard_weights = standard_weights_htn,
-  time_values = 'all', 
-  geom = 'bar'
- )
-
-fig_print(p1)
+fig_print(fig_main_4_a)
 
@@
 ```{r, fig.height=3.75}
-fig_print(p2)
+fig_print(fig_main_4_b)
 ```
@@
 
-p1 <- nhanes_visualize(
- data = nhanes_data,
- key = nhanes_key,
- outcome_variable = 'htn_resistant_accaha',
- subset_calls = list(
-  "demo_pregnant" = c("No", "Missing"),
-  "bp_med_use" = "Yes",
-  "bp_med_n_class" = c('One', 'Two', 'Three', 'Four or more')
- ),
- time_values = 'all', 
- geom = 'bar'
-)
-
-p2 <- nhanes_visualize(
- data = nhanes_data,
- key = nhanes_key,
- outcome_variable = 'htn_resistant_accaha',
- subset_calls = list(
-  "demo_pregnant" = c("No", "Missing"),
-  "bp_med_use" = "Yes",
-  "bp_med_n_class" = c('Three', 'Four or more')
- ),
- time_values = 'all', 
- geom = 'bar'
-)
-
-fig_print(p1)
+fig_print(fig_supp_4_a)
 
@@
 ```{r, fig.height=3.75}
-fig_print(p2)
+fig_print(fig_supp_4_b)
 ```

### 2022-11-15 R/nhanes_visualize.R

@@ -61,9 +61,7 @@ nhanes_visualize <- function(data,
                           standard_weights = standard_weights,
                           simplify_output = FALSE)
 
- # browser()
-
- nhanes_design_viz(
+ fig <- nhanes_design_viz(
   smry,
   statistic_primary = statistic_primary,
   title = title,
@@ -74,5 +72,8 @@ nhanes_visualize <- function(data,
   size_point = size_point,
   size_error = size_error
  )
+
+ list(fig_data = smry$results,
+      fig_object = fig)
 
 }

### 2022-11-15 R/roxygen_document.R

@@ -23,6 +23,18 @@ describe_module_sample <- function(type = 'htn'){
  )
 }
 
+describe_replicate_with_code <- function(){
+
+ "The code in this section does exactly what the videos above do, and provides a glimpse of the R code that is used on the back-end of the shiny application."
+
+}
+
+describe_replicate_intro <- function(replicate_from){
+
+ glue::glue("This article shows how `cardioStatsUSA` can replicate results from {replicate_from}. The results can be replicated with our shiny application or with code.")
+
+}
+
 tabulate_module_sample <- function(type = 'htn',
                                    subset_calls = NULL,
                                    subset_labs = NULL){

### 2022-11-22 paper/manuscript.Rmd
Help arrives
finishing touches to close issue #46

@@
 We restricted the dataset to adults ≥ 18 years of age.
-This exclusion was applied because statistics for BP levels and the prevalence of hypertension in children and adolescents are markedly different than for adults.[@muntner2004trends] We further restricted the population to participants who completed the in-home interview and study examination, with one or more SBP and DBP measurement, and who had data on self-reported antihypertensive medication use.
+This exclusion was applied because hypertension is defined differently for children and adolescents compared to adults.[@flynn_clinical_2017] We further restricted the population to participants who completed the in-home interview and study examination, with one or more SBP and DBP measurement, and who had data on self-reported antihypertensive medication use.
 After these exclusions were applied, the population in the current application included `r n_included` participants (**Figure S1**).
@@
 For age adjustment in the analysis of BP control and resistant hypertension, we set the age distribution for the standard population to represent US adults with hypertension from 1999 to 2020: 26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥ 75 years of age, respectively.
+Participants with missing pregnancy status were assumed to be non-pregnant for these analyses.
 

### 2022-11-24 README.Rmd
progressing validation + tutorials
videos and dynamic results

@@
 output: github_document
+always_allow_html: true
 ---
@@
 Our goal is to provide a platform for exploration of the National Health and Nutrition Examination Survey (NHANES) data.
-We created this website to give context and documentation for users of the application, which can be accessed here: https://bcjaeger.shinyapps.io/nhanesShinyBP/
+
+## Web application
+
+- Live version here: https://bcjaeger.shinyapps.io/nhanesShinyBP/
+
+- This website provides context and documentation for our web application
 
@@
 
-**Systolic blood pressure by age group in the last three NHANES cycles.**
+**Figure 1: Systolic blood pressure by age group in the last three NHANES cycles.**
 
@@
                  width = 800, 
-                 height = 600)[[1]]
+                 height = 600)$fig_object[[1]]
 
@@
 
-**Prevalence of blood pressure control by sex among adults with hypertension.**
+**Figure 2: Prevalence of blood pressure control by sex among adults with hypertension.**
 
@@
                  width = 800, 
-                 height = 600)[[1]]
+                 height = 600)$fig_object[[1]]
 
+```
+
+## Make your own results
+
+You can use the `cardioStatsUSA` web application to run customized analyses of NHANES data. The videos below show how I used it to make Figures 1 and 2.
+
+- This video shows how **Figure 1** was made
+
+```{r, out.width='100%', echo = FALSE}
+vembedr::embed_url("https://www.youtube.com/watch?v=SjuzOp4ZMww")
+```
+
+- This video shows how **Figure 2** was made
+
+```{r, out.width='100%', echo = FALSE}
+vembedr::embed_url("https://www.youtube.com/watch?v=oMporQuiodo")
 ```

### 2022-11-24 paper/manuscript.Rmd
videos and dynamic results

@@
 
-htn_resistant_2009b <- fig_supp_4_b %>% 
+htn_resistant_1999b <- fig_supp_4_b %>% 
  getElement("fig_data") %>% 
- .[svy_year == '2009-2010' & htn_resistant_accaha == 'Yes'] %>% 
+ .[svy_year == '1999-2000' & htn_resistant_accaha == 'Yes'] %>% 
  est_ci()
 
-htn_resistant_2003b <- fig_supp_4_b %>% 
+htn_resistant_2005b <- fig_supp_4_b %>% 
  getElement("fig_data") %>% 
- .[svy_year == '2003-2004' & htn_resistant_accaha == 'Yes'] %>% 
+ .[svy_year == '2005-2006' & htn_resistant_accaha == 'Yes'] %>% 
  est_ci()
@@
 Among US adults with hypertension who self-reported taking antihypertensive medication, `r bracket_drop(bp_control_1999b)`%, `r bracket_drop(bp_control_2013b)`%, and `r bracket_drop(bp_control_2017b)`% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively (**Figure 4; Panel B**).
-The age-adjusted prevalence of resistant hypertension defined as in the 2017 ACC/AHA BP guideline among non-pregnant US adults with hypertension who self-reported taking antihypertensive medication and had ≥1 antihypertensive medication class identified during the medication inventory was lowest in 1999-2000 (`r bracket_drop(htn_resistant_1999a)`%), highest in 2005-2006 (21.4%), and 16.6% in 2017-2020 (**Figure S4; Panel A**).
-Further restricting this analysis to those who were taking three or more classes of antihypertensive medication, the prevalence of resistant hypertension was lowest in 2009-2010 (64.1%), highest in 2003-2004 (78.3%), and 67.9% in 2017-2020 (**Figure S4; Panel B**).
+The age-adjusted prevalence of resistant hypertension defined as in the 2017 ACC/AHA BP guideline among non-pregnant US adults with hypertension who self-reported taking antihypertensive medication and had ≥1 antihypertensive medication class identified during the medication inventory was lowest in 1999-2000 (`r bracket_drop(htn_resistant_1999a)`%), highest in 2005-2006 (`r bracket_drop(htn_resistant_2005a)`%), and `r bracket_drop(htn_resistant_2017a)`% in 2017-2020 (**Figure S4; Panel A**).
+Further restricting this analysis to those who were taking three or more classes of antihypertensive medication, the prevalence of resistant hypertension was `r bracket_drop(htn_resistant_1999b)`%,  `r bracket_drop(htn_resistant_2005b)`%, and `r bracket_drop(htn_resistant_2017b)`% in 1999-2000, 2005-2006, and 2017-2020, respectively (**Figure S4; Panel B**).
 

### 2022-12-12 paper/manuscript.Rmd
medication data update

@@
  add_footer_lines(
-  values = "htn_accaha indicates hypertension status." 
- ) %>% 
- add_footer_lines(
-  values = "demo_race indicates the race subgroup."
+  values = "according the 2017 American College of Cardiology/American Heart Association blood pressure guideline." 
+ ) %>% 
+ add_footer_lines(
+  values = "demo_race indicates the race/ethnicity subgroup."
  ) %>% 
@@
  add_footer_lines(
-  values = "n_obs is the number of observations (unweighted) used for computations"
- ) %>% 
- add_footer_lines(
-  values = "unreliable_status is TRUE if the result is unreliable, FALSE otherwise"
- ) %>% 
- add_footer_lines(
-  values = "unreliable_reason is blank if the result is reliable, and lists reason(s) for unreliability otherwise"
- ) %>% 
- add_footer_lines(
-  values = "review_needed is TRUE if the result should be reviewed, FALSE otherwise"
+  values = "n_obs is the number of observations (unweighted) used for the computation"
+ ) %>% 
+ add_footer_lines(
+  values = "unreliable_status is TRUE if the result is unreliable, FALSE if the estimate is reliable"
+ ) %>% 
+ add_footer_lines(
+  values = "unreliable_reason is blank if the result is reliable, and lists reason(s) for not being reliable otherwise"
+ ) %>% 
+ add_footer_lines(
+  values = "review_needed is TRUE if the result should be reviewed by a statistical expert, FALSE if the estimate does not need to be reviewed.
+The estimate may or may not need to be suppressed depending on the outcome of the review."
  ) %>% 
@@
 
+r1 <- round_spec() %>% 
+ round_using_decimal(digits = 1)
+
 # inline results for figure 3
@@
  .[svy_year == '1999-2000' & htn_accaha == 'Yes'] %>% 
- .[, (est_cols) := lapply(.SD, function(x) table_value(x/1e6)), 
+ .[, (est_cols) := lapply(.SD, function(x) table_value(x/1e6, rspec = r1)), 
    .SDcols = est_cols] %>% 
@@
  .[svy_year == '2017-2020' & htn_accaha == 'Yes'] %>% 
- .[, (est_cols) := lapply(.SD, function(x) table_value(x/1e6)), 
+ .[, (est_cols) := lapply(.SD, function(x) table_value(x/1e6, rspec = r1)), 
    .SDcols = est_cols] %>% 
@@
 Sean P. Kane, PharmD, BCPS,^5^
+Lei Huang, MS,^2^
 Jennifer S. Herrick, MS,^3,6^
@@
 Lisandro D. Colantonio, MD, PhD,^2^
-Paul Muntner, PhD,^2^
-
-**AUTHOR ORDER NOT FINALIZED**
+Paul Muntner, PhD^2^
 
@@
 These changes need to be incorporated when estimating statistics from multiple NHANES cycles.
+These analytic challenges may present barriers to the analysis of NHANES data.
 
@@
 We developed a web-based application that provides nationally representative BP and hypertension statistics for non-institutionalized US adults using NHANES data without requiring users to conduct statistical programming.
-The goal of the application is to increase the use of NHANES data for research and policy decision making with a focus on ensuring results are valid and reproducible.
-The application is open-source,[@perens1999open] which means the code supporting it is publicly available and may be re-distributed or modified by anyone.[@rpack_cardioStatsUSA] In this manuscript, we review the design, development and validation of the application and present BP and hypertension statistics for US adults that were generated using the application.
+The goal of the application is to make NHANES data available for research and policy decision-making with a focus on ensuring results are valid and reproducible.
+The application is open-source,[@perens1999open] which means its code is publicly available and may be re-distributed or modified by anyone.[@rpack_cardioStatsUSA] In this manuscript, we review the design, development and validation of the application and present BP and hypertension statistics for US adults that were generated using the application.
 
@@
 During the interview, the labels of medications that participants reported taking in the preceding 30 days were recorded.
-We defined antihypertensive medication classes using those listed in the 2017 ACC/AHA BP guideline (Section 8.1.4; Table 18).[@whelton20182017] During the study examination, height, weight and BP were measured and blood and spot urine samples were collected.
+Antihypertensive medication classes were defined based on being commercially available according to the US Food and Drug Administration’s Orange Book, Drugs\@FDA, and Lexicon Plus® databases, and either: 1) Being listed as an antihypertensive medication in the 2017 ACC/AHA BP guideline or 2) Having a labeled indication for the treatment of essential (primary) hypertension.[@whelton20182017] Only orally-administered medications were included.
+During the study examination, height, weight and BP were measured and blood and spot urine samples were collected.
 Of relevance to the current application, blood samples were used to measure total and high-density lipoprotein cholesterol, glycated hemoglobin and serum creatinine, and the urine sample was used to measure albumin and creatinine and to conduct a pregnancy test.
-The protocol for measuring BP is available online.[@ostchega_national_2003] In brief, systolic and diastolic BP (SBP and DBP, respectively) were measured three times by trained and certified physicians.
+The protocol for measuring BP is available online.[@ostchega_national_2003] In brief, systolic and diastolic BP (SBP and DBP, respectively) were measured up to three times by trained and certified physicians.
 BP was measured using a mercury sphygmomanometer from 1999-2000 through 2015-2016 and using an oscillometric device in 2017-2020.
 The mean SBP and DBP levels were computed over all available measurements for each participant.
-The oscillometric SBP and DBP values were calibrated to the mercury device.[@muntner2022blood; @ostchega2021comparing] For the current analysis, we defined hypertension, BP control, and resistant hypertension according to the 2017 ACC/AHA BP guideline.[@whelton20182017] The application also has these variables defined according to the JNC7 definitions.[@chobanian2003seventh] A list of BP, hypertension, and antihypertensive medication variables available in the application is provided in **Table 1** with full definitions for all variables included in the application provided in **Table S1**.
+The oscillometric SBP and DBP values were calibrated to the mercury device.[@muntner2022blood; @ostchega2021comparing] For the current analysis, we defined hypertension, BP control, and resistant hypertension according to the 2017 ACC/AHA BP guideline.[@whelton20182017] The application also has these variables defined according to the JNC7 guideline definitions.[@chobanian2003seventh] A list of BP, hypertension, and antihypertensive medication variables available in the application is provided in **Table 1** with full definitions for all variables included in the application provided in **Table S1**.
 
@@
 
-The web application was created using Shiny,[@rpack_shiny; @rpack_shinyWidgets; @rpack_shinyalert] an open-source software package that translates code from the R programming language into HTML, CSS, or JavaScript commands and creates a website interface.[@r_language] We validated the web application by using it to reproduce statistics reported in two prior studies and one CDC report.[@muntner2020htn; @muntner2022blood; @fryar_hypertension_2017] We created the “cardioStatsUSA” R package to provide additional details on the web application’s design,  documentation of its components, and further details on its validation.[@rpack_cardioStatsUSA]
+The web application was created using Shiny,[@rpack_shiny; @rpack_shinyWidgets; @rpack_shinyalert] an open-source software package that translates code from the R programming language into HTML, CSS, or JavaScript commands and creates a website interface.[@r_language] We validated the web application by using it to reproduce statistics reported in two prior studies and one CDC report.[@muntner2020htn; @muntner2022blood; @fryar_hypertension_2017] We created the “cardioStatsUSA” R package to provide additional details on the web application’s design,  documentation of its components, and validation.[@rpack_cardioStatsUSA]
 
@@
 Among non-institutionalized US adults ≥ 18 years of age, the estimated mean SBP was `r sbp_1999` mm Hg in 1999-2000, `r sbp_2009` mm Hg in 2009-2010 and `r sbp_2017` mm Hg in 2017-2020 (**Figure 1**).
-The age-adjusted prevalence of hypertension defined as in the 2017 ACC/AHA BP guideline was highest in 1999-2000 (`r bracket_drop(htn_1999)`%), lowest in 2009-2010 (`r bracket_drop(htn_2009)` ), and `r bracket_drop(htn_2017)`% in 2017-2020 (**Figure 2**).
+The age-adjusted prevalence of hypertension defined according to the 2017 ACC/AHA BP guideline was highest in 1999-2000 (`r bracket_drop(htn_1999)`%), lowest in 2009-2010 and 2013-2014 (`r bracket_drop(htn_2009)`%), and `r bracket_drop(htn_2017)`% in 2017-2020 (**Figure 2**).
 In 1999-2000, there were an estimated `r n_htn_1999` million US adults with hypertension (**Figure 3**).
@@
 
-The age-adjusted prevalence of BP control defined as in the 2017 ACC/AHA BP guideline among non-pregnant US adults with hypertension was lowest in 1999-2000 (`r bracket_drop(bp_control_1999a)`%), highest in 2013-2014 (`r bracket_drop(bp_control_2013a)`%), and `r bracket_drop(bp_control_2017a)`% in 2017-2020 (**Figure 4; Panel A**).
+The age-adjusted prevalence of BP control defined according to the 2017 ACC/AHA BP guideline among non-pregnant US adults with hypertension was lowest in 1999-2000 (`r bracket_drop(bp_control_1999a)`%), highest in 2013-2014 (`r bracket_drop(bp_control_2013a)`%), and `r bracket_drop(bp_control_2017a)`% in 2017-2020 (**Figure 4; Panel A**).
 Among US adults with hypertension who self-reported taking antihypertensive medication, `r bracket_drop(bp_control_1999b)`%, `r bracket_drop(bp_control_2013b)`%, and `r bracket_drop(bp_control_2017b)`% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively (**Figure 4; Panel B**).
-The age-adjusted prevalence of resistant hypertension defined as in the 2017 ACC/AHA BP guideline among non-pregnant US adults with hypertension who self-reported taking antihypertensive medication and had ≥1 antihypertensive medication class identified during the medication inventory was lowest in 1999-2000 (`r bracket_drop(htn_resistant_1999a)`%), highest in 2005-2006 (`r bracket_drop(htn_resistant_2005a)`%), and `r bracket_drop(htn_resistant_2017a)`% in 2017-2020 (**Figure S4; Panel A**).
+The age-adjusted prevalence of resistant hypertension defined according to the 2017 ACC/AHA BP guideline among non-pregnant US adults with hypertension who self-reported taking antihypertensive medication and had ≥1 antihypertensive medication class identified during the medication inventory was lowest in 1999-2000 (`r bracket_drop(htn_resistant_1999a)`%), highest in 2005-2006 (`r bracket_drop(htn_resistant_2005a)`%), and `r bracket_drop(htn_resistant_2017a)`% in 2017-2020 (**Figure S4; Panel A**).
 Further restricting this analysis to those who were taking three or more classes of antihypertensive medication, the prevalence of resistant hypertension was `r bracket_drop(htn_resistant_1999b)`%,  `r bracket_drop(htn_resistant_2005b)`%, and `r bracket_drop(htn_resistant_2017b)`% in 1999-2000, 2005-2006, and 2017-2020, respectively (**Figure S4; Panel B**).
@@
 
-In the current manuscript, we present a web application with a user-friendly point and click interface that allows the calculation of nationally representative estimates for BP, hypertension and antihypertensive medication-related statistics using NHANES data.
+In the current manuscript, we present a web application with a point and click interface that allows the calculation of nationally representative estimates for BP, hypertension and antihypertensive medication-related statistics using NHANES data.
 Using this application, we generated crude and age-adjusted BP and hypertension statistics.
 Also, we generated statistics stratified by characteristics of US adults and demonstrated how multiple NHANES cycles can be pooled to obtain more precise estimates when working with small sub-groups (e.g., pregnant women).
-Following its validation, version 0.0.1 of the application was released and deployed on a publicly available server on **DATE TBD (it's almost ready)**.[@app_cardioStatsUSA] Researchers, clinicians, and the public can use the application to generate customized BP and hypertension statistics for US adults to inform their research and policy decisions.
+Following its validation, version 0.0.1 of the application was released and deployed on a publicly available server on **DATE TBD (it's almost ready)**.[@app_cardioStatsUSA] Anyone can use the application to generate customized BP and hypertension statistics for US adults to inform their research and policy decisions.
 
@@
 Challenges of analyzing NHANES data may also include downloading and merging multiple data files, even for a single NHANES cycle; combining multiple variables to create outcome definitions, which may require dealing with missing data and questionnaire skip patterns; harmonizing variables across multiple NHANES cycles; and the analysis of complex survey design data.
-Although multiple reports based on the data are publicly available, it may be difficult to customize reports based on user requests.
-The web application that we present in the current manuscript addresses these challenges, increasing the accessibility of NHANES data to clinicians, researchers, and decision-makers.
+Although multiple reports based on the data are publicly available, it may be difficult to generate customized reports.
+The web application that we present in the current manuscript addresses these challenges, increasing the accessibility of NHANES data to clinicians, researchers, and the public.
 
@@
 Several design decisions have been incorporated into the application.
-We required participants to have a single SBP and DBP to be included while some prior analyses required three SBP and DBP measurements.[@muntner_potential_2018]  We chose an approach that is consistent with several analyses conducted by CDC investigators.[@yoon_hypertension_2015; @fryar_hypertension_2017] However, mean BP and the prevalence of high BP may be lower if we required multiple BP measurements.[@handler_impact_2012] Any bias resulting from this decision is likely to be small as over 95% of adult NHANES participants with at least one SBP and DBP measurement had three SBP and DBP measurements.
+We required participants to have a single SBP and DBP to be included while some prior analyses required three SBP and DBP measurements.[@muntner_potential_2018] We chose this approach as it is consistent with several analyses conducted by CDC scientists.[@yoon_hypertension_2015; @fryar_hypertension_2017] However, mean BP and the prevalence of high BP may be lower if we required multiple BP measurements.[@handler_impact_2012] Any bias resulting from this decision is likely to be small as over 95% of adult NHANES participants with at least one SBP and DBP measurement had three SBP and DBP measurements.
 The application re-calibrates the NHANES weights for the estimation of population counts.[@coresh_prevalence_2003] This was done because participants missing data on SBP, DBP or antihypertensive medication use cannot have BP or hypertension-related outcomes.
 Weights were not re-calibrated when estimating proportions as participants missing data are removed from the numerator and denominator.
-Medication classes were coded using generic names and the drug classes in the 2017 ACC/AHA BP guideline.
-We recognize the NCHS recommends using Lexicon Plus®, a proprietary database, to categorize medication classes.[@nhanes_rxq_drug] While the categorization of most medications is identical using generic drug names and Lexicon Plus®, some differences exist.
+Medication classes were coded using generic names and the drug classes after review by pharmacists.
+We recognize the NCHS recommends using Lexicon Plus®, a proprietary database, to categorize medication classes.[@nhanes_rxq_drug] While the categorization of most medications is identical using generic drug names and Lexicon Plus®, some differences existed.
 Many additional decisions were made regarding the definitions of variables, inclusion of study participants, and analytic approach.
@@
 
-Using the application, we replicated results from several prior manuscripts.[@muntner2020htn; @muntner2022blood; @fryar_hypertension_2017] However, we were not able to replicate some results.[@carey_prevalence_2019] For example, a prior manuscript reported the prevalence of resistant hypertension defined by the 2017 ACC/AHA BP guideline to be 19.7% in 2009-2014.[@carey_prevalence_2019] When estimated by the application, the prevalence of resistant hypertension over this time period was 17.1%.
-The difference in the prevalence estimates can be attributed to the approach used to categorize antiyhpertensive medications into their medication classes.
+```{r}
+
+htn_resistant_prev <- 
+ nhanes_summarize(outcome_variable = 'htn_resistant_accaha', 
+                  time_values = c("2009-2010", "2011-2012", "2013-2014"),
+                  pool = TRUE,
+                  subset_calls = list("htn_accaha" = "Yes",
+                                      "bp_med_n_class" = c("One", 
+                                                           "Two", 
+                                                           "Three",
+                                                           "Four or more"),
+                                      "bp_med_use" = "Yes")) %>% 
+ .[htn_resistant_accaha == 'Yes'] %>% 
+ est_ci()
+
+```
+
+Using the application, we replicated results from several prior manuscripts.[@muntner2020htn; @muntner2022blood; @fryar_hypertension_2017] However, we were not able to replicate some results.[@carey_prevalence_2019] For example, a prior manuscript reported the prevalence of resistant hypertension defined by the 2017 ACC/AHA BP guideline to be 19.7% in 2009-2014.[@carey_prevalence_2019] When estimated by the application, the prevalence of resistant hypertension over this time period was `r bracket_drop(htn_resistant_prev)`%.
+The difference in the prevalence estimates can be attributed to the approach used to categorize antihypertensive medications into their medication classes.
 The prior publication used Lexicon Plus®, which counted spironolactone as two drug classes, a potassium-sparing diuretic and an aldosterone antagonist.
@@
 This study also has several limitations.
-NHANES participants had their mean BP measured during a single visit, and BP guidelines recommend obtaining mean BP using at least two BP values measured on separate days.[@whelton20182017] Many estimates, including medication use, are derived from participant report, and we cannot exclude the possibilities of reporting or recall biases.
+NHANES participants had their mean BP measured during a single visit, and BP guidelines recommend obtaining mean BP using at least two BP values measured on separate days.[@whelton20182017] Many estimates, including medication use, were derived from participant self-report, and we cannot exclude the possibilities of reporting or recall biases.
 The response rate for NHANES has declined from 1999-2000 through 2017-2020.
@@
 
-## Conclusions
-
-We developed a web-based application for analysis of hypertension statistics among non-institutionalized adults living in the US from 1999-2000 through 2017-2020.
+## Perspectives
+
+We developed a web-based application for the  analysis of hypertension statistics among non-institutionalized adults living in the US from 1999-2000 through 2017-2020.
 The application is publicly available and produces valid, transparent, and reproducible results.
@@
 
-### Table 2:	Race/ethnicity distribution of US adults with and without hypertension in 2017-2020
-
-```{r}
-
-tbl_main_2_flex
-
-```
-
-\newpage
-
 ```{r footer_functions}
@@
 footer_hover <- function(parenthetical,
+                         estimate,
                          hover_over){
@@
  "The graph is identical to the web application's output. \\
- All values ({parenthetical}) can be obtained in the app \\
- by hovering over {hover_over}."
+ The estimated {estimate} and the upper and lower limits of \\
+ the 95% confidence interval can be displayed in the \\
+ application by hovering the cursor over the {hover_over}."
  )
@@
  glue::glue(
-  "This commands used to generate this graph are [available online]({url})"
+  "The inputs used to generate this graph are [available online]({url})"
  )
@@
 
+### Table 2:	Race/ethnicity distribution of US adults with and without hypertension in 2017-2020
+
+```{r}
+
+tbl_main_2_flex
+
+```
+
+<!-- `r footer_hyperlink(url = "https://bcjaeger.shinyapps.io/nhanesShinyBP/?_inputs_&standard_weights_2=33.6&group_class_open=false&help=0&stratify_class=null&do_open=false&group=%22htn_accaha%22&wont_do_computation=0&standard_weights_1=49.3&explore_output_cells_selected=%5B%5D&explore_output_search_columns=%5B%22%22%2C%22%22%2C%22%22%2C%22%22%2C%22%22%2C%22%22%2C%22%22%2C%22%22%2C%22%22%2C%22%22%2C%22%22%2C%22%22%2C%22%22%5D&year_stratify=%222017-2020%22&explore_output_search=%22%22&geom=%22bar%22&outcome=%22demo_race%22&statistic_primary=%22percentage%22&standard_weights_4=7&outcome_open=false&explore_output_columns_selected=null&n_exposure_group=null&pool=%22no%22&stratify=null&exposure_cut_type=%22interval%22&group_class=%22Hypertension%22&statistic=%22percentage%22&subset_n=%220%22&outcome_class=%22Demographics%22&standard_weights_3=10.1&explore_output_rows_current=%5B1%2C2%2C3%2C4%2C5%2C6%2C7%2C8%2C9%2C10%5D&do=%22data%22&year_pool=%5B%221999-2000%22%2C%222017-2020%22%5D&explore_output_rows_all=%5B1%2C2%2C3%2C4%2C5%2C6%2C7%2C8%2C9%2C10%5D&age_standardize=false&group_open=false&explore_output_rows_selected=null&explore_output_state=%7B%22time%22%3A1670012255861%2C%22start%22%3A0%2C%22length%22%3A10%2C%22order%22%3A%5B%5D%2C%22search%22%3A%7B%22search%22%3A%22%22%2C%22smart%22%3Atrue%2C%22regex%22%3Afalse%2C%22caseInsensitive%22%3Atrue%7D%2C%22columns%22%3A%5B%7B%22visible%22%3Atrue%2C%22search%22%3A%7B%22search%22%3A%22%22%2C%22smart%22%3Atrue%2C%22regex%22%3Afalse%2C%22caseInsensitive%22%3Atrue%7D%7D%2C%7B%22visible%22%3Atrue%2C%22search%22%3A%7B%22search%22%3A%22%22%2C%22smart%22%3Atrue%2C%22regex%22%3Afalse%2C%22caseInsensitive%22%3Atrue%7D%7D%2C%7B%22visible%22%3Atrue%2C%22search%22%3A%7B%22search%22%3A%22%22%2C%22smart%22%3Atrue%2C%22regex%22%3Afalse%2C%22caseInsensitive%22%3Atrue%7D%7D%2C%7B%22visible%22%3Atrue%2C%22search%22%3A%7B%22search%22%3A%22%22%2C%22smart%22%3Atrue%2C%22regex%22%3Afalse%2C%22caseInsensitive%22%3Atrue%7D%7D%2C%7B%22visible%22%3Atrue%2C%22search%22%3A%7B%22search%22%3A%22%22%2C%22smart%22%3Atrue%2C%22regex%22%3Afalse%2C%22caseInsensitive%22%3Atrue%7D%7D%2C%7B%22visible%22%3Atrue%2C%22search%22%3A%7B%22search%22%3A%22%22%2C%22smart%22%3Atrue%2C%22regex%22%3Afalse%2C%22caseInsensitive%22%3Atrue%7D%7D%2C%7B%22visible%22%3Atrue%2C%22search%22%3A%7B%22search%22%3A%22%22%2C%22smart%22%3Atrue%2C%22regex%22%3Afalse%2C%22caseInsensitive%22%3Atrue%7D%7D%2C%7B%22visible%22%3Atrue%2C%22search%22%3A%7B%22search%22%3A%22%22%2C%22smart%22%3Atrue%2C%22regex%22%3Afalse%2C%22caseInsensitive%22%3Atrue%7D%7D%2C%7B%22visible%22%3Atrue%2C%22search%22%3A%7B%22search%22%3A%22%22%2C%22smart%22%3Atrue%2C%22regex%22%3Afalse%2C%22caseInsensitive%22%3Atrue%7D%7D%2C%7B%22visible%22%3Atrue%2C%22search%22%3A%7B%22search%22%3A%22%22%2C%22smart%22%3Atrue%2C%22regex%22%3Afalse%2C%22caseInsensitive%22%3Atrue%7D%7D%2C%7B%22visible%22%3Atrue%2C%22search%22%3A%7B%22search%22%3A%22%22%2C%22smart%22%3Atrue%2C%22regex%22%3Afalse%2C%22caseInsensitive%22%3Atrue%7D%7D%2C%7B%22visible%22%3Atrue%2C%22search%22%3A%7B%22search%22%3A%22%22%2C%22smart%22%3Atrue%2C%22regex%22%3Afalse%2C%22caseInsensitive%22%3Atrue%7D%7D%2C%7B%22visible%22%3Atrue%2C%22search%22%3A%7B%22search%22%3A%22%22%2C%22smart%22%3Atrue%2C%22regex%22%3Afalse%2C%22caseInsensitive%22%3Atrue%7D%7D%5D%7D&pool_open=false&explore_output_cell_clicked=%7B%7D&outcome_class_open=false")` `r single_par` -->
+
+\newpage
+
 ### Figure 1: Mean systolic blood pressure for US adults by calendar year
@@
 
-`r footer_hover(parenthetical = "estimated mean and the upper and lower limits of the 95% confidence interval", hover_over = "the mean systolic blood pressure")` `r single_par`
+`r footer_hover(estimate = "prevalence", hover_over = "dot")` `r single_par`
 
@@
 
-`r footer_hover(parenthetical = "estimated prevalence and the upper and lower limits of the 95% confidence interval", hover_over = "the bars")` `r single_par`
+`r footer_hover(estimate = "prevalence", hover_over = "bar")` `r single_par`
 
@@
 ```
-`r footer_hover(parenthetical = "estimated count and the upper and lower limits of the 95% confidence interval", hover_over = "the bars")` `r single_par`
+
+`r footer_hover(estimate = "prevalence", hover_over = "bar")` `r single_par`
 
@@
 
-`r footer_hover(parenthetical = "estimated prevalence and the upper and lower limits of the 95% confidence interval", hover_over = "the bars")` `r single_par`
-
-Survey participants with missing values for pregnancy status were assumed to be non-pregnant `r single_par`
-
-`r footer_hyperlink(url = "https://bcjaeger.shinyapps.io/nhanesShinyBP/?_inputs_&standard_weights_2=43.5&subset_variable_2=%22demo_pregnant%22&help=0&stratify_class=null&subset_value_1_ctns=%5B0%2C0%5D&subset_value_2_catg=%5B%22No%22%2C%22Missing%22%5D&subset_variable_1=%22htn_accaha%22&group=null&wont_do_computation=0&standard_weights_1=26.4&subset_variable_2_open=false&year_stratify=%5B%221999-2000%22%2C%222001-2002%22%2C%222003-2004%22%2C%222005-2006%22%2C%222007-2008%22%2C%222009-2010%22%2C%222011-2012%22%2C%222013-2014%22%2C%222015-2016%22%2C%222017-2020%22%5D&subset_value_1_catg=%22Yes%22&geom=%22bar%22&outcome=%22bp_control_accaha%22&statistic_primary=%22percentage%22&subset_n_open=false&standard_weights_4=13.2&outcome_open=false&plotly_hover-A=null&n_exposure_group=null&pool=%22no%22&stratify=null&exposure_cut_type=%22interval%22&group_class=null&statistic=%22percentage%22&subset_n=%222%22&outcome_class=%22Blood%20pressure%22&standard_weights_3=17&.clientValue-default-plotlyCrosstalkOpts=%7B%22on%22%3A%22plotly_click%22%2C%22persistent%22%3Afalse%2C%22dynamic%22%3Afalse%2C%22selectize%22%3Afalse%2C%22opacityDim%22%3A0.2%2C%22selected%22%3A%7B%22opacity%22%3A1%7D%2C%22debounce%22%3A0%2C%22color%22%3A%5B%5D%7D&do=%22figure%22&year_pool=%5B%221999-2000%22%2C%222017-2020%22%5D&plotly_afterplot-A=%22%5C%22htmlwidget-1d7ba8c29733e6d67fa7%5C%22%22&age_standardize=true&statistic_primary_open=false&subset_value_2_ctns=%5B0%2C0%5D&subset_variable_1_open=false&outcome_class_open=false")` `r single_par`
+`r footer_hover(estimate = "prevalence", hover_over = "bar")` `r single_par`
+
+Participants with missing values for pregnancy status were assumed to be non-pregnant in this analysis `r single_par`
+
+`r footer_hyperlink(url = "https://bcjaeger.shinyapps.io/nhanesShinyBP/?_inputs_&standard_weights_2=43.4&subset_variable_2=%22demo_pregnant%22&shinyjs-delay-83b9d0e407c7920ca132675a5f5a2420=200&help=0&stratify_class=null&subset_value_1_ctns=%5B0%2C0%5D&subset_value_2_catg=%5B%22No%22%2C%22Missing%22%5D&subset_variable_1=%22htn_accaha%22&group=null&wont_do_computation=0&standard_weights_1=26.4&shinyjs-delay-64fac92b188a33b6e29b5c1462d93c5f=500&year_stratify=%5B%221999-2000%22%2C%222001-2002%22%2C%222003-2004%22%2C%222005-2006%22%2C%222007-2008%22%2C%222009-2010%22%2C%222011-2012%22%2C%222013-2014%22%2C%222015-2016%22%2C%222017-2020%22%5D&subset_value_1_catg=%22Yes%22&geom=%22bar%22&shinyjs-delay-f68f72a7dc73646d6d53de72584d1314=200&outcome=%22bp_control_accaha%22&statistic_primary=%22percentage%22&standard_weights_4=13.2&shinyjs-delay-4cde34965ec4abc6815eda510975a7a8=200&n_exposure_group=null&pool=%22no%22&stratify=null&exposure_cut_type=%22interval%22&group_class=null&statistic=%22percentage%22&subset_n=%222%22&outcome_class=%22Blood%20pressure%22&standard_weights_3=17&subset_value_2_ctns=%5B0%2C0%5D&do=%22figure%22&year_pool=%5B%221999-2000%22%2C%222017-2020%22%5D&.clientValue-default-plotlyCrosstalkOpts=%7B%22on%22%3A%22plotly_click%22%2C%22persistent%22%3Afalse%2C%22dynamic%22%3Afalse%2C%22selectize%22%3Afalse%2C%22opacityDim%22%3A0.2%2C%22selected%22%3A%7B%22opacity%22%3A1%7D%2C%22debounce%22%3A0%2C%22color%22%3A%5B%5D%7D&age_standardize=true&plotly_afterplot-A=%22%5C%22htmlwidget-81fa8fc4f6f4e307e35b%5C%22%22")` `r single_par`
 
@@
 
-`r footer_hover(parenthetical = "estimated prevalence and the upper and lower limits of the 95% confidence interval", hover_over = "the bars")` `r single_par`
-
-Survey participants with missing values for pregnancy status were assumed to be non-pregnant `r single_par`
-
-`r footer_hyperlink(url = "https://bcjaeger.shinyapps.io/nhanesShinyBP/?_inputs_&standard_weights_2=43.5&subset_variable_2=%22demo_pregnant%22&subset_variable_3=%22bp_med_use%22&help=0&stratify_class=null&subset_value_1_ctns=%5B0%2C0%5D&subset_value_2_catg=%5B%22No%22%2C%22Missing%22%5D&subset_variable_1=%22htn_accaha%22&group=null&wont_do_computation=0&standard_weights_1=26.4&subset_variable_2_open=false&subset_value_3_ctns=%5B0%2C0%5D&year_stratify=%5B%221999-2000%22%2C%222001-2002%22%2C%222003-2004%22%2C%222005-2006%22%2C%222007-2008%22%2C%222009-2010%22%2C%222011-2012%22%2C%222013-2014%22%2C%222015-2016%22%2C%222017-2020%22%5D&subset_value_1_catg=%22Yes%22&geom=%22bar%22&outcome=%22bp_control_accaha%22&statistic_primary=%22percentage%22&subset_n_open=false&standard_weights_4=13.2&outcome_open=false&plotly_hover-A=null&n_exposure_group=null&pool=%22no%22&stratify=null&exposure_cut_type=%22interval%22&group_class=null&statistic=%22percentage%22&subset_n=%223%22&outcome_class=%22Blood%20pressure%22&standard_weights_3=17&.clientValue-default-plotlyCrosstalkOpts=%7B%22on%22%3A%22plotly_click%22%2C%22persistent%22%3Afalse%2C%22dynamic%22%3Afalse%2C%22selectize%22%3Afalse%2C%22opacityDim%22%3A0.2%2C%22selected%22%3A%7B%22opacity%22%3A1%7D%2C%22debounce%22%3A0%2C%22color%22%3A%5B%5D%7D&do=%22figure%22&year_pool=%5B%221999-2000%22%2C%222017-2020%22%5D&plotly_afterplot-A=%22%5C%22htmlwidget-49855de1c73b4f93593a%5C%22%22&age_standardize=true&statistic_primary_open=false&subset_value_2_ctns=%5B0%2C0%5D&subset_variable_3_open=false&subset_value_3_catg=%22Yes%22&subset_variable_1_open=false&outcome_class_open=false")` `r single_par`
+`r footer_hover(estimate = "prevalence", hover_over = "bar")` `r single_par`
+
+Participants with missing values for pregnancy status were assumed to be non-pregnant in this analysis `r single_par`
+
+`r footer_hyperlink(url = "https://bcjaeger.shinyapps.io/nhanesShinyBP/?_inputs_&standard_weights_2=43.4&subset_variable_2=%22demo_pregnant%22&subset_variable_3=%22bp_med_use%22&help=0&stratify_class=null&subset_value_1_ctns=%5B0%2C0%5D&subset_value_2_catg=%5B%22No%22%2C%22Missing%22%5D&subset_variable_1=%22htn_accaha%22&group=null&wont_do_computation=0&standard_weights_1=26.4&subset_value_3_ctns=%5B0%2C0%5D&shinyjs-delay-8692c04cdef53922346e48054dbd93a1=200&year_stratify=%5B%221999-2000%22%2C%222001-2002%22%2C%222003-2004%22%2C%222005-2006%22%2C%222007-2008%22%2C%222009-2010%22%2C%222011-2012%22%2C%222013-2014%22%2C%222015-2016%22%2C%222017-2020%22%5D&subset_value_1_catg=%22Yes%22&shinyjs-delay-facdd7b8cab855f088d44290099dfbef=200&geom=%22bar%22&outcome=%22bp_control_accaha%22&statistic_primary=%22percentage%22&shinyjs-delay-a1efa56454e54123cf7630b981e03ad9=500&standard_weights_4=13.2&plotly_hover-A=null&n_exposure_group=null&pool=%22no%22&stratify=null&exposure_cut_type=%22interval%22&group_class=null&statistic=%22percentage%22&subset_n=%223%22&outcome_class=%22Blood%20pressure%22&standard_weights_3=17&subset_value_2_ctns=%5B0%2C0%5D&do=%22figure%22&year_pool=%5B%221999-2000%22%2C%222017-2020%22%5D&shinyjs-delay-8010218891a07c8626e99b56bc03ac48=200&age_standardize=true&.clientValue-default-plotlyCrosstalkOpts=%7B%22on%22%3A%22plotly_click%22%2C%22persistent%22%3Afalse%2C%22dynamic%22%3Afalse%2C%22selectize%22%3Afalse%2C%22opacityDim%22%3A0.2%2C%22selected%22%3A%7B%22opacity%22%3A1%7D%2C%22debounce%22%3A0%2C%22color%22%3A%5B%5D%7D&plotly_afterplot-A=%22%5C%22htmlwidget-ff33b629debbee169b28%5C%22%22&subset_value_3_catg=%22Yes%22")` `r single_par`
 
@@
               col_keys = c('class', 'label', 'description')) %>% 
- set_header_labels(label = "All variables included in module",
+ set_header_labels(label = "Variables included in the module",
                    description = "Variable definition") %>% 
@@
  padding(i = indent_meds, j = 1, padding.left = 20) %>% 
- width(j = 1, width = 2) %>%
+ width(j = 1, width = 3.5) %>%
  width(j = 2, width = 7) %>% 
@@
 ```
-`r footer_hover(parenthetical = "estimated prevalence and the upper and lower limits of the 95% confidence interval", hover_over = "the bars")` `r single_par`
+
+`r footer_hover(estimate = "prevalence", hover_over = "bar")` `r single_par`
+
+`r footer_hyperlink(url = "https://bcjaeger.shinyapps.io/nhanesShinyBP/?_inputs_&standard_weights_2=33.6&shinyjs-delay-379cecbbd3bd0926bf37d5396e79af6f=200&group_class_open=false&help=0&stratify_class=null&do_open=false&group=%22cc_ckd%22&wont_do_computation=0&standard_weights_1=49.3&shinyjs-delay-1523a834e5f76f6c6878c89805380cb9=500&explore_output_cells_selected=%5B%5D&year_stratify=%5B%221999-2000%22%2C%222001-2002%22%2C%222003-2004%22%2C%222005-2006%22%2C%222007-2008%22%2C%222009-2010%22%2C%222011-2012%22%2C%222013-2014%22%2C%222015-2016%22%2C%222017-2020%22%5D&explore_output_search=null&explore_output_search_columns=null&geom=%22bar%22&outcome=%22htn_accaha%22&statistic_primary=%22percentage%22&standard_weights_4=7&outcome_open=false&explore_output_columns_selected=null&n_exposure_group=null&pool=%22no%22&stratify=null&exposure_cut_type=%22interval%22&group_class=%22Comorbidities%22&statistic=%22percentage%22&subset_n=%220%22&outcome_class=%22Hypertension%22&standard_weights_3=10.1&explore_output_rows_current=null&do=%22figure%22&year_pool=%5B%221999-2000%22%2C%222017-2020%22%5D&explore_output_rows_all=null&age_standardize=false&.clientValue-default-plotlyCrosstalkOpts=%7B%22on%22%3A%22plotly_click%22%2C%22persistent%22%3Afalse%2C%22dynamic%22%3Afalse%2C%22selectize%22%3Afalse%2C%22opacityDim%22%3A0.2%2C%22selected%22%3A%7B%22opacity%22%3A1%7D%2C%22debounce%22%3A0%2C%22color%22%3A%5B%5D%7D&shinyjs-delay-53ed67400615637051b4a98ee01f93c5=200&plotly_afterplot-A=%22%5C%22htmlwidget-e51ff4a9a0ba21276c1f%5C%22%22&group_open=false&plotly_hover-A=null&explore_output_rows_selected=null&explore_output_state=null&explore_output_cell_clicked=null&outcome_class_open=false")` `r single_par`
 
@@
 ```
-The `r paste("--")` symbol indicates that an estimate is not reliable. `r single_par`
+The `r paste("--")` symbol indicates that the estimate was not reliable. `r single_par`
 
@@
 
-`r footer_hover(parenthetical = "estimated prevalence, upper and lower limits of the 95% confidence interval, or reasons for the suppression of data if applicable", hover_over = "the bars")` `r single_par`
-
-There were no pregnant women in the two highest blood pressure categories, so the figure does not include segments for these categories. `r single_par`
+The graph is identical to the web application’s output.
+The estimated prevalence, upper and lower limits of the 95% confidence interval, and reasons for the suppression of data if applicable can be obtained in the application by hovering the mouse over the bars.
+
+There were no pregnant women in the two highest blood pressure categories, so the figure does not include red and purple segments for these categories. `r single_par`
+
+`r footer_hyperlink(url = "https://bcjaeger.shinyapps.io/nhanesShinyBP/?_inputs_&standard_weights_2=33.6&shinyjs-delay-379cecbbd3bd0926bf37d5396e79af6f=200&group_class_open=false&help=0&stratify_class=null&subset_value_1_ctns=%5B0%2C0%5D&do_open=false&group=null&wont_do_computation=0&standard_weights_1=49.3&shinyjs-delay-1523a834e5f76f6c6878c89805380cb9=500&explore_output_cells_selected=%5B%5D&year_stratify=%222017-2020%22&explore_output_search=null&explore_output_search_columns=null&geom=%22bar%22&subset_variable_1=%22demo_pregnant%22&outcome=%22bp_cat_meds_excluded%22&statistic_primary=%22percentage%22&subset_value_1_catg=%22Yes%22&subset_n_open=false&standard_weights_4=7&outcome_open=false&explore_output_columns_selected=null&n_exposure_group=null&pool=%22no%22&stratify=null&exposure_cut_type=%22interval%22&group_class=null&statistic=%22percentage%22&subset_n=%221%22&outcome_class=%22Blood%20pressure%22&standard_weights_3=10.1&explore_output_rows_current=null&do=%22figure%22&year_pool=%5B%221999-2000%22%2C%222017-2020%22%5D&explore_output_rows_all=null&age_standardize=false&.clientValue-default-plotlyCrosstalkOpts=%7B%22on%22%3A%22plotly_click%22%2C%22persistent%22%3Afalse%2C%22dynamic%22%3Afalse%2C%22selectize%22%3Afalse%2C%22opacityDim%22%3A0.2%2C%22selected%22%3A%7B%22opacity%22%3A1%7D%2C%22debounce%22%3A0%2C%22color%22%3A%5B%5D%7D&shinyjs-delay-53ed67400615637051b4a98ee01f93c5=200&plotly_afterplot-A=%22%5C%22htmlwidget-7efd8162074e261eeb99%5C%22%22&group_open=false&plotly_hover-A=null&explore_output_rows_selected=null&subset_variable_1_open=false&explore_output_state=null&explore_output_cell_clicked=null&outcome_class_open=false")` `r single_par`
 
@@
 ```
-`r footer_hover(parenthetical = "estimated prevalence and the upper and lower limits of the 95% confidence interval", hover_over = "the bars")` `r single_par`
+
+`r footer_hover(estimate = "prevalence", hover_over = "bar")` `r single_par`
 
 The estimate for systolic blood pressure of 140 to < 160 or diastolic blood pressure of 90 to < 100 mm Hg is 1.1%.
-This can be seen on the application by hovering over the red bar with your mouse.
+This can be seen on the application by hovering the mouse over the red bar with your mouse.
 `r single_par`
+
+`r footer_hyperlink(url = "https://bcjaeger.shinyapps.io/nhanesShinyBP/?_inputs_&standard_weights_2=33.6&shinyjs-delay-379cecbbd3bd0926bf37d5396e79af6f=200&group_class_open=false&help=0&stratify_class=null&subset_value_1_ctns=%5B0%2C0%5D&do_open=false&group=null&wont_do_computation=0&standard_weights_1=49.3&shinyjs-delay-1523a834e5f76f6c6878c89805380cb9=500&explore_output_cells_selected=%5B%5D&year_stratify=%5B%222009-2010%22%2C%222011-2012%22%2C%222013-2014%22%2C%222015-2016%22%2C%222017-2020%22%5D&explore_output_search=null&explore_output_search_columns=null&geom=%22bar%22&subset_variable_1=%22demo_pregnant%22&outcome=%22bp_cat_meds_excluded%22&statistic_primary=%22percentage%22&subset_value_1_catg=%22Yes%22&subset_n_open=false&standard_weights_4=7&outcome_open=false&explore_output_columns_selected=null&n_exposure_group=null&pool=%22yes%22&stratify=null&exposure_cut_type=%22interval%22&group_class=null&statistic=%22percentage%22&subset_n=%221%22&outcome_class=%22Blood%20pressure%22&standard_weights_3=10.1&explore_output_rows_current=null&do=%22figure%22&year_pool=%5B%222009-2010%22%2C%222017-2020%22%5D&explore_output_rows_all=null&age_standardize=false&.clientValue-default-plotlyCrosstalkOpts=%7B%22on%22%3A%22plotly_click%22%2C%22persistent%22%3Afalse%2C%22dynamic%22%3Afalse%2C%22selectize%22%3Afalse%2C%22opacityDim%22%3A0.2%2C%22selected%22%3A%7B%22opacity%22%3A1%7D%2C%22debounce%22%3A0%2C%22color%22%3A%5B%5D%7D&shinyjs-delay-53ed67400615637051b4a98ee01f93c5=200&plotly_afterplot-A=%22%5C%22htmlwidget-b1f957b964c8c871b2ea%5C%22%22&group_open=false&plotly_hover-A=null&explore_output_rows_selected=null&subset_variable_1_open=false&explore_output_state=null&pool_open=false&explore_output_cell_clicked=null&outcome_class_open=false")` `r single_par`
 
@@
 
-`r footer_hover(parenthetical = "estimated prevalence and the upper and lower limits of the 95% confidence interval", hover_over = "the bars")` `r single_par`
-
-Survey participants with missing values for pregnancy status were assumed to be non-pregnant `r single_par`
+`r footer_hover(estimate = "prevalence", hover_over = "bar")` `r single_par`
+
+Participants with missing values for pregnancy status were assumed to be non-pregnant in this analysis `r single_par`
+
+`r footer_hyperlink(url = "https://bcjaeger.shinyapps.io/nhanesShinyBP/?_inputs_&standard_weights_2=43.4&subset_variable_2=%22bp_med_use%22&subset_variable_3=%22bp_med_n_class%22&help=0&stratify_class=null&subset_value_1_ctns=%5B0%2C0%5D&subset_value_2_catg=%22Yes%22&subset_variable_1=%22demo_pregnant%22&group=null&wont_do_computation=0&standard_weights_1=26.4&subset_value_3_ctns=%5B0%2C0%5D&subset_variable_2_open=false&year_stratify=%5B%221999-2000%22%2C%222001-2002%22%2C%222003-2004%22%2C%222005-2006%22%2C%222007-2008%22%2C%222009-2010%22%2C%222011-2012%22%2C%222013-2014%22%2C%222015-2016%22%2C%222017-2020%22%5D&subset_value_1_catg=%5B%22No%22%2C%22Missing%22%5D&geom=%22bar%22&outcome=%22htn_resistant_accaha%22&statistic_primary=%22percentage%22&subset_n_open=false&standard_weights_4=13.2&outcome_open=false&plotly_hover-A=null&n_exposure_group=null&pool=%22no%22&stratify=null&exposure_cut_type=%22interval%22&group_class=null&statistic=%22percentage%22&subset_n=%223%22&outcome_class=%22Hypertension%22&standard_weights_3=17&subset_value_2_ctns=%5B0%2C0%5D&do=%22figure%22&year_pool=%5B%221999-2000%22%2C%222017-2020%22%5D&subset_variable_3_open=false&age_standardize=true&.clientValue-default-plotlyCrosstalkOpts=%7B%22on%22%3A%22plotly_click%22%2C%22persistent%22%3Afalse%2C%22dynamic%22%3Afalse%2C%22selectize%22%3Afalse%2C%22opacityDim%22%3A0.2%2C%22selected%22%3A%7B%22opacity%22%3A1%7D%2C%22debounce%22%3A0%2C%22color%22%3A%5B%5D%7D&plotly_afterplot-A=%22%5C%22htmlwidget-dbbceadbcdd51ccb28f9%5C%22%22&subset_value_3_catg=%5B%22One%22%2C%22Two%22%2C%22Three%22%2C%22Four%20or%20more%22%5D&subset_variable_1_open=false&outcome_class_open=false")` `r single_par`
 
@@
 
-`r footer_hover(parenthetical = "estimated prevalence and the upper and lower limits of the 95% confidence interval", hover_over = "the bars")` `r single_par`
-
-Survey participants with missing values for pregnancy status were assumed to be non-pregnant `r single_par`
+`r footer_hover(estimate = "prevalence", hover_over = "bar")` `r single_par`
+
+Participants with missing values for pregnancy status were assumed to be non-pregnant in this analysis `r single_par`
+
+`r footer_hyperlink(url = "https://bcjaeger.shinyapps.io/nhanesShinyBP/?_inputs_&standard_weights_2=43.4&subset_variable_2=%22bp_med_use%22&subset_variable_3=%22bp_med_n_class%22&help=0&stratify_class=null&subset_value_1_ctns=%5B0%2C0%5D&subset_value_2_catg=%22Yes%22&subset_variable_1=%22demo_pregnant%22&group=null&wont_do_computation=0&standard_weights_1=26.4&subset_value_3_ctns=%5B0%2C0%5D&subset_variable_2_open=false&year_stratify=%5B%221999-2000%22%2C%222001-2002%22%2C%222003-2004%22%2C%222005-2006%22%2C%222007-2008%22%2C%222009-2010%22%2C%222011-2012%22%2C%222013-2014%22%2C%222015-2016%22%2C%222017-2020%22%5D&subset_value_1_catg=%5B%22No%22%2C%22Missing%22%5D&plotly_relayout-A=%22%7B%5C%22width%5C%22%3A1123%2C%5C%22height%5C%22%3A600%7D%22&geom=%22bar%22&outcome=%22htn_resistant_accaha%22&statistic_primary=%22percentage%22&subset_n_open=false&standard_weights_4=13.2&outcome_open=false&plotly_hover-A=%22%5B%7B%5C%22curveNumber%5C%22%3A0%2C%5C%22pointNumber%5C%22%3A0%2C%5C%22x%5C%22%3A%5C%221999-2000%5C%22%2C%5C%22y%5C%22%3A13.5556325036742%7D%5D%22&n_exposure_group=null&pool=%22no%22&stratify=null&exposure_cut_type=%22interval%22&group_class=null&statistic=%22percentage%22&subset_n=%223%22&outcome_class=%22Hypertension%22&standard_weights_3=17&subset_value_2_ctns=%5B0%2C0%5D&do=%22figure%22&year_pool=%5B%221999-2000%22%2C%222017-2020%22%5D&subset_variable_3_open=false&age_standardize=true&.clientValue-default-plotlyCrosstalkOpts=%7B%22on%22%3A%22plotly_click%22%2C%22persistent%22%3Afalse%2C%22dynamic%22%3Afalse%2C%22selectize%22%3Afalse%2C%22opacityDim%22%3A0.2%2C%22selected%22%3A%7B%22opacity%22%3A1%7D%2C%22debounce%22%3A0%2C%22color%22%3A%5B%5D%7D&plotly_afterplot-A=%22%5C%22htmlwidget-0a02a3f59338fc876b7b%5C%22%22&subset_value_3_catg=%5B%22Three%22%2C%22Four%20or%20more%22%5D&subset_variable_1_open=false&outcome_class_open=false")` `r single_par`
 

### 2022-12-12 data-raw/nhanes_load.R

@@ -13,9 +13,12 @@ nhanes_load <- function(fpath = file.path(here::here(), 'data-raw'),
                         as = 'tibble'){
 
  data_bp <- haven::read_sas(file.path(fpath, 'htn-raw.sas7bdat'))
- data_lipids <- haven::read_sas(file.path(fpath, 'lipids-raw.sas7bdat'))
 
- out <- left_join(data_bp, data_lipids, by = c("SEQN", "WTSAF2YR")) %>%
+ data_lipids <-
+  haven::read_sas(file.path(fpath, 'lipids-raw.sas7bdat')) %>%
+  select(-highriskcond)
+
+ out <- left_join(data_bp, data_lipids, by = c("SEQN")) %>%
   mutate(subpopldl = replace(subpopldl, is.na(subpopldl), 0))
 
  switch(

### 2022-12-12 data-raw/nhanes_recode.R

@@ -198,22 +198,22 @@ nhanes_recode <- function(data = NULL){
     )
    ),
 
-   highriskcond = factor(
-    highriskcond,
-    levels = 0:9,
-    labels = c(
-     "0",
-     "1",
-     "2",
-     "3 or more",
-     "3 or more",
-     "3 or more",
-     "3 or more",
-     "3 or more",
-     "3 or more",
-     "3 or more"
-    )
-   ),
+   # highriskcond = factor(
+   #  highriskcond,
+   #  levels = 0:9,
+   #  labels = c(
+   #   "0",
+   #   "1",
+   #   "2",
+   #   "3 or more",
+   #   "3 or more",
+   #   "3 or more",
+   #   "3 or more",
+   #   "3 or more",
+   #   "3 or more",
+   #   "3 or more"
+   #  )
+   # ),
 
    # 0/1 indicators ----
 
@@ -262,6 +262,7 @@ nhanes_recode <- function(data = NULL){
               total_c240pl,
               persistent_ldl,
               vhr,
+              combination,
               addontherapy,
               recommended_ill,
               llt_rec_2018,

### 2024-02-09 paper/manuscript.Rmd
finalized changes for pub, committed long after they were made

@@
 
-```{r, fig.height=3.75}
+```{r fig_1, fig.height=3.75}
 
@@
 
-```{r, fig.height=3.75}
+```{r fig_2, fig.height=3.75}
 
@@
 
-```{r}
+```{r fig_3}
 
@@
 A. Among non-pregnant US adults with hypertension `r single_par`
-```{r, fig.height=3.75}
+```{r fig_4a, fig.height=3.75}
 
@@
 B. Among non-pregnant US adults who self-report taking antihypertensive medication `r single_par`
-```{r, fig.height=3.75}
+```{r fig_4b, fig.height=3.75}
 fig_print(fig_main_4_b)
